---
title: "03--resazurin-20260624-mgig-33C"
author: "Sam White"
date: "2026-06-24"
output:
  github_document:
    toc: true
    number_sections: true
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
editor_options: 
  chunk_output_type: console
---

# Background

Adult oysters from nine USDA families were placed in plastic cups and submerged
in ~100 mL of freshly prepared resazurin working solution at room temperature
(~22°C). Cups were incubated in a water table at 33°C. Three rounds of
measurements were performed, with different individuals from each family used in
each round. At designated timepoints, 200 µL of resazurin was sampled from each
cup and placed in a 96-well plate. Fluorescence was measured using a Synergy HTX
(Agilent) plate reader.

See `Resazurin/data/20260624-mgig-33C/README.md` for full experimental notes.

## Expected inputs

| Path | Description |
|------|-------------|
| `Resazurin/data/20260624-mgig-33C/plate-*-T*.txt` | Plate reader fluorescence exports (one file per plate per timepoint) |
| `Resazurin/data/20260624-mgig-33C/layout.csv` | Well metadata: plate ID, well ID, blank flag, family groups, cup ID (`cup_ID.group`), measurement round (`round.group`), unique individual ID (`sample_ID.group`), area measurements (mm², from ImageJ) |

## Expected outputs

All outputs are written to `Resazurin/outputs/03--resazurin-20260624-mgig-33C/`.

| File | Description |
|------|-------------|
| `figures/` | All plots generated by this script |
| `auc_all_metrics.csv` | Per-individual AUC values for every active measurement metric |
| `auc_summary.csv` | Group-level AUC summary statistics (mean, SD, SE, median) |
| `metabolism.csv` | Full per-well per-timepoint metabolism data frame |
| `pairwise_stats.csv` | Tukey-adjusted pairwise comparisons from AUC linear models |
| `curve_features.csv` | Per-individual curve traits describing metabolic capacity, responsiveness, and stress-like trajectory shape |
| `curve_feature_family_summary.csv` | Family-level summary statistics for every curve trait |
| `curve_feature_family_stats.csv` | Ranked tests identifying which curve traits best distinguish families |
| `curve_feature_pairwise_stats.csv` | Tukey-adjusted family pairwise comparisons for each curve trait |

# Setup

## Knitr options
```{r setup}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = "",         # Prevents appending '##' to beginning of lines in code output
  results = 'hold'     # Holds output so it's all printed together after code chunk
)
```

## Load libraries
```{r libraries}
library(tidyverse)
library(colorspace)   # qualitative_hcl() for large palettes

has_lme4 <- requireNamespace("lme4", quietly = TRUE)
has_lmerTest <- requireNamespace("lmerTest", quietly = TRUE)
has_emmeans <- requireNamespace("emmeans", quietly = TRUE)

if (!has_lme4 || !has_lmerTest || !has_emmeans) {
  message(
    "Optional mixed-model/post-hoc packages are not all installed; ",
    "time-series mixed models and emmeans pairwise tests will be skipped ",
    "or replaced with base R TukeyHSD fallbacks where possible."
  )
}
```

# Helper Functions

```{r helper-fns}
normalize_well_id <- function(x) {
  x <- toupper(trimws(x))
  valid <- str_detect(x, "^[A-Z]+[0-9]+$")
  out <- rep(NA_character_, length(x))
  if (!any(valid)) return(out)
  m <- str_match(x[valid], "^([A-Z]+)([0-9]+)$")
  out[valid] <- paste0(m[, 2], as.integer(m[, 3]))
  out
}

parse_time_hr <- function(path) {
  hit <- str_match(basename(path),
                   "(?i)[_-]T([0-9]+(?:\\.[0-9]+)?)\\.txt$")
  as.numeric(hit[, 2])
}

parse_plate_id <- function(path) {
  hit <- str_match(basename(path),
    "(?i)^[0-9]{8}_(round[0-9]+)_T[0-9]+(?:\\.[0-9]+)?\\.txt$")
  id <- hit[, 2]
  ifelse(is.na(id), "unknown", id)
}

extract_results_block <- function(lines) {
  results_idx <- which(trimws(lines) == "Results")
  if (length(results_idx) == 0) stop("No Results section found")
  idx <- results_idx[1]
  header_tokens <- str_split(lines[idx + 1], "\\t")[[1]] |> trimws()
  col_ids <- header_tokens[
    header_tokens != "" & str_detect(header_tokens, "^[0-9]+$")]
  j <- idx + 2
  data_lines <- character()
  while (j <= length(lines)) {
    line <- lines[j]
    if (trimws(line) == "") break
    if (!str_detect(line, "^[A-Za-z]\\t")) break
    data_lines <- c(data_lines, line)
    j <- j + 1
  }
  list(col_ids = col_ids, data_lines = data_lines)
}

parse_plate_export <- function(path) {
  lines <- readLines(path, warn = FALSE)
  res <- extract_results_block(lines)

  map_dfr(res$data_lines, function(line) {
    tokens <- str_split(line, "\\t")[[1]] |> trimws()
    tokens <- tokens[tokens != ""]
    row_letter <- tokens[1]
    nums <- suppressWarnings(as.numeric(tokens[-1]))
    valid_idx <- which(!is.na(nums))
    if (length(valid_idx) == 0) return(tibble())
    vals <- nums[valid_idx]
    n <- min(length(vals), length(res$col_ids))
    tibble(
      row_id  = toupper(row_letter),
      col_id  = as.integer(res$col_ids[seq_len(n)]),
      well_id = normalize_well_id(
        paste0(toupper(row_letter), res$col_ids[seq_len(n)])),
      value   = vals[seq_len(n)]
    )
  }) %>%
    mutate(
      plate_id = str_to_lower(parse_plate_id(path)),
      time_hr  = parse_time_hr(path)
    )
}

trapezoid_auc <- function(time_hr, value) {
  ok <- is.finite(time_hr) & is.finite(value)
  t <- time_hr[ok]
  v <- value[ok]
  if (length(t) < 2) return(NA_real_)
  ord <- order(t)
  t <- t[ord]; v <- v[ord]
  sum(diff(t) * (head(v, -1) + tail(v, -1)) / 2)
}

# Shared helper: extract display unit string from a measurement column name.
# e.g. "area_mm2_measurement" -> "mm²", "weight_mg_measurement" -> "mg"
parse_meas_unit <- function(col_name) {
  unit_raw <- col_name |>
    str_remove("^metabolism_per_") |>
    str_remove("_measurement$") |>
    str_extract("[^_]+$")
  case_when(
    unit_raw == "mm2" ~ "mm²",
    unit_raw == "cm2" ~ "cm²",
    unit_raw == "mm3" ~ "mm³",
    unit_raw == "cm3" ~ "cm³",
    TRUE              ~ unit_raw
  )
}

# y-axis label for metabolism line plots: "fold change/mm²"
metabolism_y_label <- function(col_name) {
  paste0("Metabolism (fold change/", parse_meas_unit(col_name), ")")
}

# y-axis label for AUC box plots: "Metabolism (AUC; mm²)"
auc_y_label <- function(metric_name) {
  paste0("Metabolism (AUC; ", parse_meas_unit(metric_name), ")")
}
```

# Load Data

## Plate export files

```{r load-plates}
proj_root <- rprojroot::find_rstudio_root_file()
data_dir  <- file.path(proj_root, "Resazurin", "data", "20260624-mgig-33C")
out_dir   <- file.path(proj_root, "Resazurin", "outputs",
                        "03--resazurin-20260624-mgig-33C")
fig_dir   <- file.path(out_dir, "figures")

dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

plate_files <- list.files(
  data_dir,
  pattern = "(?i)^[0-9]{8}_round[0-9]+_T[0-9]+(?:\\.[0-9]+)?\\.txt$",
  full.names = TRUE
)

plate_raw <- map_dfr(plate_files, function(path) {
  tryCatch(parse_plate_export(path),
           error = function(e) {
             message("Parse error in ", basename(path), ": ", e$message)
             tibble()
           })
})

str(plate_raw)
```

## Plate consistency check

Checks that every plate has the same number of wells at every timepoint.
The expected well count is the mode across all plate × timepoint reads.
Any plate with at least one deviating read is flagged and dropped entirely
before any further analysis — removing only the aberrant timepoint would
break the fold-change baseline calculation.

```{r plate-consistency-check, results='asis'}
well_counts <- plate_raw %>%
  group_by(plate_id, time_hr) %>%
  summarise(n_wells = n_distinct(well_id), .groups = "drop")

expected_n_wells <- as.integer(
  names(which.max(table(well_counts$n_wells)))
)

inconsistent_reads <- well_counts %>%
  filter(n_wells != expected_n_wells) %>%
  arrange(plate_id, time_hr)

inconsistent_plate_ids <- unique(inconsistent_reads$plate_id)

if (nrow(inconsistent_reads) > 0) {
  cat("**Plate consistency check FAILED.**",
      "Expected", expected_n_wells, "wells per plate-timepoint read.",
      length(inconsistent_plate_ids),
      "plate(s) have at least one deviating read and are excluded",
      "from all analyses:\n\n")
  cat(knitr::kable(
    inconsistent_reads,
    col.names = c("Plate", "Time (h)", "Wells read"),
    caption   = paste("Expected:", expected_n_wells, "wells per read")
  ), sep = "\n")
  cat("\n")
  plate_raw <- plate_raw %>%
    filter(!plate_id %in% inconsistent_plate_ids)
  message(length(inconsistent_plate_ids),
          " plate(s) removed from plate_raw: ",
          paste(inconsistent_plate_ids, collapse = ", "))
} else {
  cat("Plate consistency check passed: all",
      n_distinct(well_counts$plate_id), "plates have",
      expected_n_wells, "wells at every timepoint.\n")
}
```

## Layout file

```{r load-layout}
layout_path <- file.path(data_dir, "layout.csv")

layout_raw <- read_csv(layout_path,
                       col_types = cols(.default = "c"),
                       show_col_types = FALSE)

# Standardise column names to snake_case
names(layout_raw) <- names(layout_raw) |>
  str_to_lower() |>
  str_replace_all("[^a-z0-9]+", "_") |>
  str_replace_all("_+", "_") |>
  str_replace("_$", "")

# Normalise plate_id to match plate file ids (strip "plate-" prefix)
layout_clean <- layout_raw %>%
  mutate(
    plate_id = str_remove(str_to_lower(plate_id), "^plate-"),
    well_id  = normalize_well_id(plate_well),
    is_blank = if ("is_blank" %in% names(layout_raw))
      toupper(trimws(is_blank)) %in% c("TRUE", "T", "1", "YES", "Y")
    else
      FALSE
  )

found_exclude_col <- intersect(
  c("exclude_from_analysis", "exclude", "omit", "not_analyzed"),
  names(layout_clean)
)[1]
layout_clean <- layout_clean %>%
  mutate(
    exclude_from_analysis = if (!is.na(found_exclude_col))
      toupper(trimws(.data[[found_exclude_col]])) %in%
        c("TRUE", "T", "1", "YES", "Y")
    else
      FALSE
  )

# Identify measurement columns and group columns
measurement_cols <- names(layout_clean)[
  str_detect(names(layout_clean), "_measurement$")]
group_cols <- names(layout_clean)[
  str_detect(names(layout_clean), "_group$")]

# Cast measurement columns to numeric
layout_clean <- layout_clean %>%
  mutate(across(all_of(measurement_cols),
                ~ suppressWarnings(as.numeric(.x))))

# Determine which measurement columns actually contain finite data
active_meas_cols <- measurement_cols[
  sapply(measurement_cols, function(col)
    any(is.finite(layout_clean[[col]]), na.rm = TRUE))]

# Normalise group values to lowercase so they match colour scale definitions
layout_clean <- layout_clean %>%
  mutate(across(all_of(group_cols),
                ~ str_to_lower(trimws(as.character(.x)))))

message("Group columns: ", paste(group_cols, collapse = ", "))
message("Active measurement columns: ",
        paste(active_meas_cols, collapse = ", "))

str(layout_clean)
```

# Merge Plate Data with Layout

```{r merge-data}
dat <- plate_raw %>%
  left_join(
    layout_clean %>%
      select(plate_id, well_id, is_blank, exclude_from_analysis,
             any_of("exclude_reason"),
             all_of(group_cols), all_of(measurement_cols)),
    by = c("plate_id", "well_id")
  ) %>%
  mutate(
    is_blank = replace_na(is_blank, FALSE),
    exclude_from_analysis = replace_na(exclude_from_analysis, FALSE)
  )

str(dat)
```

# Raw Fluorescence

## Data frame

```{r raw-fluor-df}
# Wells in the plate reader output that have no layout entry get all-NA group
# columns after the join. Keep only wells assigned to at least one group.
active_gc <- intersect(group_cols, names(dat))

raw_df <- dat %>%
  filter(
    !is_blank,
    if (length(active_gc) > 0)
      if_any(all_of(active_gc), ~ !is.na(.))
    else
      TRUE
  ) %>%
  mutate(
    trace_id = if_else(
      !is.na(sample_id_group) & trimws(as.character(sample_id_group)) != "",
      as.character(sample_id_group),
      paste(plate_id, well_id, sep = "_")
    )
  )

families   <- str_sort(unique(na.omit(raw_df$family_id_group)), numeric = TRUE)
treatments <- sort(unique(na.omit(raw_df$treatment_group)))

n_fam <- length(families)
n_trt <- length(treatments)

# Palette strategy:
#   <= 7 groups : Okabe-Ito (gold standard for colorblind-safe figures).
#   >  7 groups : colorspace::qualitative_hcl("Dynamic") scales to any N
#                 using perceptually uniform HCL space — no colour collisions.
# Black (#000000) is excluded from both and reserved for blank wells.
okabe_ito_7 <- c(
  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
  "#0072B2", "#D55E00", "#CC79A7"
)
make_palette <- function(n) {
  if (n == 0L) return(character(0))
  if (n <= length(okabe_ito_7)) return(okabe_ito_7[seq_len(n)])
  colorspace::qualitative_hcl(n, palette = "Dynamic")
}

all_colours   <- make_palette(n_fam + n_trt)
fam_colours   <- setNames(all_colours[seq_len(n_fam)], families)
trt_colours   <- setNames(all_colours[n_fam + seq_len(n_trt)], treatments)

lty_pool <- c("solid", "dashed", "dotted", "dotdash", "longdash")
trt_linetypes <- setNames(
  lty_pool[(seq_len(n_trt) - 1L) %% length(lty_pool) + 1L],
  treatments
)
plate_well_colours <- c(blank = "black", fam_colours)

has_trt <- n_trt > 0

str(raw_df)
```

## Raw fluorescence by plate (including blanks)

```{r raw-fluor-by-plate, fig.width=10, fig.height=8}
p_raw_plates <- dat %>%
  filter(is.finite(time_hr), is.finite(value)) %>%
  mutate(
    colour_group = if_else(is_blank, "blank",
                           coalesce(family_id_group, "sample")),
    trace_id     = paste(plate_id, well_id, sep = "_")
  ) %>%
  ggplot(aes(x = time_hr, y = value,
             group = trace_id, colour = colour_group)) +
  geom_line(alpha = 0.6) +
  geom_point(size = 1, alpha = 0.7) +
  facet_wrap(~ plate_id) +
  scale_colour_manual(
    values   = plate_well_colours,
    name     = "Group",
    breaks   = names(plate_well_colours),
    na.value = "grey80"
  ) +
  labs(x = "Time (h)", y = "Raw fluorescence (RFU)") +
  theme_classic(base_size = 12) +
  theme(strip.background = element_blank(),
        strip.text       = element_text(face = "bold"))

p_raw_plates
ggsave(file.path(fig_dir, "raw_fluor_by_plate.png"),
       p_raw_plates, width = 10, height = 8)
```

## Mean raw fluorescence by family

```{r raw-mean-by-family, fig.width=8, fig.height=5}
raw_family_summary <- raw_df %>%
  filter(!is.na(family_id_group), !exclude_from_analysis) %>%
  group_by(family_id_group, treatment_group, time_hr) %>%
  summarise(
    mean_fluor = mean(value, na.rm = TRUE),
    se_fluor   = sd(value, na.rm = TRUE) /
      sqrt(sum(!is.na(value))),
    n          = sum(!is.na(value)),
    .groups    = "drop"
  ) %>%
  mutate(group_var = if (has_trt)
    paste(family_id_group, treatment_group, sep = ".")
  else
    family_id_group)

p_raw_mean <- ggplot(raw_family_summary,
    aes(x = time_hr, y = mean_fluor,
        colour = family_id_group,
        group = group_var)) +
  geom_ribbon(aes(ymin = mean_fluor - se_fluor,
                  ymax = mean_fluor + se_fluor,
                  fill = family_id_group),
              alpha = 0.15, colour = NA) +
  geom_line(
    mapping   = if (has_trt) aes(linetype = treatment_group) else NULL,
    linewidth = 1) +
  geom_point(size = 2) +
  scale_colour_manual(values = fam_colours, name = "Family",
                      breaks = families) +
  scale_fill_manual(values = fam_colours, name = "Family",
                    breaks = families) +
  labs(x = "Time (h)", y = "Mean raw fluorescence (RFU ± SE)") +
  theme_classic(base_size = 13) +
  if (has_trt) scale_linetype_manual(values = trt_linetypes, name = "Treatment") else NULL

p_raw_mean
ggsave(file.path(fig_dir, "raw_mean_by_family.png"),
       p_raw_mean, width = 8, height = 5)
```

## Individual raw fluorescence traces by family

```{r raw-individual-by-family, fig.width=10, fig.height=5}
p_raw_by_family <- raw_df %>%
  filter(!is.na(family_id_group)) %>%
  ggplot(aes(x = time_hr, y = value, group = trace_id,
             colour = .data[[if (has_trt) "treatment_group" else "family_id_group"]])) +
  geom_line(alpha = 0.6) +
  geom_point(size = 1.2, alpha = 0.7) +
  facet_wrap(~ family_id_group) +
  scale_colour_manual(
    values = if (has_trt) trt_colours else fam_colours,
    name   = if (has_trt) "Treatment" else "Family",
    breaks = if (has_trt) treatments else families) +
  labs(x = "Time (h)", y = "Raw fluorescence (RFU)") +
  theme_classic(base_size = 12) +
  theme(strip.background = element_blank(),
        strip.text       = element_text(face = "bold"))

p_raw_by_family
ggsave(file.path(fig_dir, "raw_individual_by_family.png"),
       p_raw_by_family, width = 10, height = 5)
```

## Individual raw fluorescence traces by treatment

```{r raw-individual-by-treatment, fig.width=10, fig.height=5}
if (has_trt) {
  p_raw_by_treatment <- raw_df %>%
    ggplot(aes(x = time_hr, y = value,
               group = trace_id, colour = family_id_group)) +
    geom_line(alpha = 0.6) +
    geom_point(size = 1.2, alpha = 0.7) +
    facet_wrap(~ treatment_group) +
    scale_colour_manual(values = fam_colours, name = "Family",
                        breaks = families) +
    labs(x = "Time (h)", y = "Raw fluorescence (RFU)") +
    theme_classic(base_size = 12) +
    theme(strip.background = element_blank(),
          strip.text       = element_text(face = "bold"))

  p_raw_by_treatment
  ggsave(file.path(fig_dir, "raw_individual_by_treatment.png"),
         p_raw_by_treatment, width = 10, height = 5)
}
```

## Excluded samples

Wells flagged `exclude_from_analysis = TRUE` appear in the raw fluorescence
plots above but are omitted from all analyses that follow.

```{r excluded-wells, results='asis'}
excluded_wells <- dat %>%
  filter(!is_blank, exclude_from_analysis) %>%
  mutate(
    sample = if_else(
      !is.na(sample_id_group) & trimws(as.character(sample_id_group)) != "",
      as.character(sample_id_group),
      paste(plate_id, well_id, sep = "_")
    )
  ) %>%
  select(plate_id, well_id, sample, family_id_group, treatment_group,
         any_of("exclude_reason")) %>%
  distinct() %>%
  arrange(plate_id, well_id)

if (nrow(excluded_wells) > 0) {
  col_names <- c("Plate", "Well", "Sample", "Family", "Treatment")
  if ("exclude_reason" %in% names(excluded_wells))
    col_names <- c(col_names, "Reason")
  cat(knitr::kable(excluded_wells, col.names = col_names), sep = "\n")
} else {
  cat("No wells are excluded from analysis.\n")
}
```

# Blank Correction via Fold-Change Normalization

T0 is the earliest timepoint present in the dataset (not necessarily 0 hr).
Sample fold-change is expressed relative to each individual's T0 reading,
resolved by `sample_id_group` when that column is populated — allowing the
same animal to be tracked across plates — or by `plate_id + well_id` when
no sample IDs exist (backward-compatible with single-plate, multi-timepoint
designs). Blank fold-change is the per-plate mean blank RFU at each timepoint
divided by the pooled mean blank RFU at T0. Subtracting blank fold-change from
sample fold-change removes background fluorescence drift; all samples start at
exactly 0 at T0 by construction.

## Step 1 – Identify T0 and compute per-sample fold-change

```{r fold-change-all}
# T0 = earliest timepoint present in the dataset
t0_time <- min(dat$time_hr[is.finite(dat$time_hr)], na.rm = TRUE)
message("T0 timepoint: ", t0_time, " hr")

# T0 reference value per individual.
# Resolved by sample_id_group (cross-plate tracking) when available;
# falls back to plate+well for layouts without explicit sample IDs.
t0_all <- dat %>%
  filter(time_hr == t0_time, !is_blank, is.finite(value)) %>%
  mutate(sample_key = if_else(
    !is.na(sample_id_group) & trimws(as.character(sample_id_group)) != "",
    as.character(sample_id_group),
    paste(plate_id, well_id, sep = "_")
  )) %>%
  group_by(sample_key) %>%
  summarise(value_t0 = mean(value, na.rm = TRUE), .groups = "drop")

dat_fc <- dat %>%
  mutate(sample_key = if_else(
    !is_blank &
      !is.na(sample_id_group) & trimws(as.character(sample_id_group)) != "",
    as.character(sample_id_group),
    paste(plate_id, well_id, sep = "_")
  )) %>%
  left_join(t0_all, by = "sample_key") %>%
  mutate(fold_change = if_else(
    !is_blank & is.finite(value_t0) & value_t0 > 0,
    value / value_t0,
    NA_real_
  ))

str(dat_fc)
```

## Step 2 – Blank fold-change reference per plate per timepoint

```{r blank-fc-ref}
# Pooled mean blank RFU at T0 across all T0 plates
mean_blank_t0 <- dat %>%
  filter(is_blank, time_hr == t0_time, is.finite(value)) %>%
  pull(value) %>%
  mean(na.rm = TRUE)

if (!is.finite(mean_blank_t0))
  message("No blank readings found at T0 (", t0_time,
          " hr); blank correction will produce NA.")

# Per-plate per-timepoint mean blank expressed as fold-change relative to T0
blank_fc_ref <- dat %>%
  filter(is_blank, is.finite(value)) %>%
  group_by(plate_id, time_hr) %>%
  summarise(mean_blank_rfu = mean(value, na.rm = TRUE), .groups = "drop") %>%
  mutate(mean_blank_fc = mean_blank_rfu / mean_blank_t0)

str(blank_fc_ref)
```

## Step 3 – Subtract blank fold-change from sample fold-change

```{r corrected-fc}
samples <- dat_fc %>%
  filter(!is_blank, !exclude_from_analysis) %>%
  mutate(
    trace_id = if_else(
      !is.na(sample_id_group) & trimws(as.character(sample_id_group)) != "",
      as.character(sample_id_group),
      paste(plate_id, well_id, sep = "_")
    )
  ) %>%
  left_join(blank_fc_ref, by = c("plate_id", "time_hr")) %>%
  mutate(corrected_fc = fold_change - mean_blank_fc)

str(samples)
```

# Blank-Corrected Fold-Change

## Mean by family

```{r bc-fc-mean-by-family, fig.width=8, fig.height=5}
bc_fc_summary <- samples %>%
  filter(!is.na(family_id_group), !exclude_from_analysis) %>%
  group_by(family_id_group, treatment_group, time_hr) %>%
  summarise(
    mean_val = mean(corrected_fc, na.rm = TRUE),
    se_val   = sd(corrected_fc, na.rm = TRUE) /
      sqrt(sum(!is.na(corrected_fc))),
    n        = sum(!is.na(corrected_fc)),
    .groups  = "drop"
  ) %>%
  mutate(group_var = if (has_trt)
    paste(family_id_group, treatment_group, sep = ".")
  else
    family_id_group)

p_bc_fc_mean <- ggplot(bc_fc_summary,
    aes(x = time_hr, y = mean_val,
        colour = family_id_group,
        group = group_var)) +
  geom_ribbon(aes(ymin = mean_val - se_val,
                  ymax = mean_val + se_val,
                  fill = family_id_group),
              alpha = 0.15, colour = NA) +
  geom_line(
    mapping   = if (has_trt) aes(linetype = treatment_group) else NULL,
    linewidth = 1) +
  geom_point(size = 2) +
  scale_colour_manual(values = fam_colours, name = "Family",
                      breaks = families) +
  scale_fill_manual(values = fam_colours, name = "Family",
                    breaks = families) +
  labs(x = "Time (h)",
       y = "Mean blank-corrected fold-change (± SE)") +
  theme_classic(base_size = 13) +
  if (has_trt) scale_linetype_manual(values = trt_linetypes, name = "Treatment") else NULL

p_bc_fc_mean
ggsave(file.path(fig_dir, "blank_corrected_fc_mean_by_family.png"),
       p_bc_fc_mean, width = 8, height = 5)
```

## Individual traces by family

```{r bc-fc-individual-by-family, fig.width=10, fig.height=5}
p_bc_fc_by_family <- samples %>%
  filter(!is.na(family_id_group)) %>%
  ggplot(aes(x = time_hr, y = corrected_fc, group = trace_id,
             colour = .data[[if (has_trt) "treatment_group" else "family_id_group"]])) +
  geom_line(alpha = 0.6) +
  geom_point(size = 1.2, alpha = 0.7) +
  facet_wrap(~ family_id_group) +
  scale_colour_manual(
    values = if (has_trt) trt_colours else fam_colours,
    name   = if (has_trt) "Treatment" else "Family",
    breaks = if (has_trt) treatments else families) +
  labs(x = "Time (h)", y = "Blank-corrected fold-change") +
  theme_classic(base_size = 12) +
  theme(strip.background = element_blank(),
        strip.text       = element_text(face = "bold"))

p_bc_fc_by_family
ggsave(file.path(fig_dir, "blank_corrected_fc_by_family.png"),
       p_bc_fc_by_family, width = 10, height = 5)
```

## Individual blank-corrected fold-change traces by treatment

```{r bc-fc-individual-by-treatment, fig.width=10, fig.height=5}
if (has_trt) {
  p_bc_fc_by_treatment <- samples %>%
    ggplot(aes(x = time_hr, y = corrected_fc,
               group = trace_id, colour = family_id_group)) +
    geom_line(alpha = 0.6) +
    geom_point(size = 1.2, alpha = 0.7) +
    facet_wrap(~ treatment_group) +
    scale_colour_manual(values = fam_colours, name = "Family",
                        breaks = families) +
    labs(x = "Time (h)", y = "Blank-corrected fold-change") +
    theme_classic(base_size = 12) +
    theme(strip.background = element_blank(),
          strip.text       = element_text(face = "bold"))

  p_bc_fc_by_treatment
  ggsave(file.path(fig_dir, "blank_corrected_fc_by_treatment.png"),
         p_bc_fc_by_treatment, width = 10, height = 5)
}
```

# Metabolism (Size-Normalised Fold-Change)

Blank-corrected fold-change divided by each active measurement column.
This is "metabolism" as defined in Huffmyer et al.

```{r metabolism-df}
if (length(active_meas_cols) == 0) {
  message("No active measurement columns: skipping metabolism calculation.")
  metabolism_df <- tibble()
} else {
  metabolism_df <- samples
  for (mc in active_meas_cols) {
    out_col <- paste0("metabolism_per_", mc)
    metabolism_df <- metabolism_df %>%
      mutate(!!out_col := if_else(
        is.finite(.data[[mc]]) & .data[[mc]] > 0 &
          is.finite(corrected_fc),
        corrected_fc / .data[[mc]],
        NA_real_
      ))
  }
}

str(metabolism_df)
```

## Mean metabolism by family

```{r metabolism-mean-by-family, fig.width=8, fig.height=5}
if (nrow(metabolism_df) > 0) {

  metab_cols <- paste0("metabolism_per_", active_meas_cols)

  for (col in metab_cols) {
    if (!col %in% names(metabolism_df)) next
    mc_label <- str_remove(col, "^metabolism_per_")

    metab_summary <- metabolism_df %>%
      filter(!is.na(family_id_group), !exclude_from_analysis) %>%
      group_by(family_id_group, treatment_group, time_hr) %>%
      summarise(
        mean_val = mean(.data[[col]], na.rm = TRUE),
        se_val   = sd(.data[[col]], na.rm = TRUE) /
          sqrt(sum(!is.na(.data[[col]]))),
        .groups  = "drop"
      ) %>%
      mutate(group_var = if (has_trt)
        paste(family_id_group, treatment_group, sep = ".")
      else
        family_id_group)

    p_metab_mean <- ggplot(metab_summary,
        aes(x = time_hr, y = mean_val,
            colour = family_id_group,
            group = group_var)) +
      geom_ribbon(aes(ymin = mean_val - se_val,
                      ymax = mean_val + se_val,
                      fill = family_id_group),
                  alpha = 0.15, colour = NA) +
      geom_line(
        mapping   = if (has_trt) aes(linetype = treatment_group) else NULL,
        linewidth = 1) +
      geom_point(size = 2) +
      scale_colour_manual(values = fam_colours, name = "Family",
                          breaks = families) +
      scale_fill_manual(values = fam_colours, name = "Family",
                        breaks = families) +
      labs(x = "Time (h)",
           y = paste0(metabolism_y_label(col), " (± SE)")) +
      theme_classic(base_size = 13) +
      if (has_trt) scale_linetype_manual(values = trt_linetypes, name = "Treatment") else NULL

    print(p_metab_mean)
    ggsave(
      file.path(fig_dir,
                paste0("metabolism_mean_", mc_label, ".png")),
      p_metab_mean, width = 8, height = 5)
  }
}
```

## Individual metabolism traces by family

```{r metabolism-individual-by-family, fig.width=10, fig.height=5}
if (nrow(metabolism_df) > 0) {

  for (col in metab_cols) {
    if (!col %in% names(metabolism_df)) next
    mc_label <- str_remove(col, "^metabolism_per_")

    p_metab_by_family <- metabolism_df %>%
      filter(!is.na(family_id_group)) %>%
      ggplot(aes(x = time_hr, y = .data[[col]], group = trace_id,
                 colour = .data[[if (has_trt) "treatment_group" else "family_id_group"]])) +
      geom_line(alpha = 0.6) +
      geom_point(size = 1.2, alpha = 0.7) +
      facet_wrap(~ family_id_group) +
      scale_colour_manual(
        values = if (has_trt) trt_colours else fam_colours,
        name   = if (has_trt) "Treatment" else "Family",
        breaks = if (has_trt) treatments else families) +
      labs(x = "Time (h)", y = metabolism_y_label(col)) +
      theme_classic(base_size = 12) +
      theme(strip.background = element_blank(),
            strip.text       = element_text(face = "bold"))

    print(p_metab_by_family)
    ggsave(
      file.path(fig_dir,
                paste0("metabolism_individual_", mc_label, "_by_family.png")),
      p_metab_by_family, width = 10, height = 5)

    if (has_trt) {
      p_metab_by_treatment <- ggplot(metabolism_df,
          aes(x = time_hr, y = .data[[col]],
              group = trace_id, colour = family_id_group)) +
        geom_line(alpha = 0.6) +
        geom_point(size = 1.2, alpha = 0.7) +
        facet_wrap(~ treatment_group) +
        scale_colour_manual(values = fam_colours, name = "Family",
                            breaks = families) +
        labs(x = "Time (h)", y = metabolism_y_label(col)) +
        theme_classic(base_size = 12) +
        theme(strip.background = element_blank(),
              strip.text       = element_text(face = "bold"))

      print(p_metab_by_treatment)
      ggsave(
        file.path(fig_dir,
                  paste0("metabolism_individual_", mc_label, "_by_treatment.png")),
        p_metab_by_treatment, width = 10, height = 5)
    }
  }
}
```

# Time-Series Statistical Analysis

Linear mixed effects models test the effect of experimental variables on
metabolism over time. Individual (`sample_id_group`) is included as a random
intercept to account for repeated measures across timepoints. Type III ANOVA
with Satterthwaite's approximation (lmerTest) assesses significance; post-hoc
pairwise comparisons use estimated marginal means (emmeans, Tukey adjustment).

```{r ts-model}
run_ts_stats <- function(df, value_col) {
  if (!has_lme4 || !has_lmerTest || !has_emmeans) {
    message("Skipping time-series mixed model for ", value_col,
            ": lme4, lmerTest, and emmeans are required.")
    return(NULL)
  }

  has_family    <- "family_id_group" %in% names(df) &&
    length(unique(na.omit(df$family_id_group))) > 1
  has_treatment <- "treatment_group" %in% names(df) &&
    length(unique(na.omit(df$treatment_group))) > 1

  if (!has_family && !has_treatment) return(NULL)

  df <- df %>%
    filter(is.finite(.data[[value_col]]), is.finite(time_hr)) %>%
    mutate(
      time_f     = factor(time_hr),
      individual = factor(trace_id)
    )

  if (nrow(df) == 0) return(NULL)

  if (has_family)    df <- df %>% mutate(family    = factor(family_id_group))
  if (has_treatment) df <- df %>% mutate(treatment = factor(treatment_group))

  if (has_family    && length(unique(na.omit(df$family)))    < 2) return(NULL)
  if (has_treatment && length(unique(na.omit(df$treatment))) < 2) return(NULL)

  fixed <- if (has_family && has_treatment) {
    paste0(value_col, " ~ time_f * family * treatment")
  } else if (has_family) {
    paste0(value_col, " ~ time_f * family")
  } else {
    paste0(value_col, " ~ time_f * treatment")
  }

  model <- lme4::lmer(
    as.formula(paste0(fixed, " + (1 | individual)")),
    data = df
  )

  anova_res <- lmerTest::anova(model, type = 3, ddf = "Satterthwaite")

  # Pairwise comparisons of group combinations at each timepoint
  emm_spec <- if (has_family && has_treatment) {
    ~ family * treatment | time_f
  } else if (has_family) {
    ~ family | time_f
  } else {
    ~ treatment | time_f
  }

  emm       <- emmeans::emmeans(model, emm_spec)
  pairs_res <- as.data.frame(pairs(emm, adjust = "tukey"))

  # Main-effect marginal means (collapsed across time)
  emm_main <- if (has_family && has_treatment) {
    emmeans::emmeans(model, ~ family * treatment)
  } else if (has_family) {
    emmeans::emmeans(model, ~ family)
  } else {
    emmeans::emmeans(model, ~ treatment)
  }

  pairs_main <- as.data.frame(pairs(emm_main, adjust = "tukey"))

  list(
    model         = model,
    anova         = anova_res,
    pairs_by_time = pairs_res,
    pairs_main    = pairs_main,
    has_family    = has_family,
    has_treatment = has_treatment
  )
}

ts_stats <- list()
if (nrow(metabolism_df) > 0) {
  for (mc in active_meas_cols) {
    col <- paste0("metabolism_per_", mc)
    if (col %in% names(metabolism_df))
      ts_stats[[col]] <- run_ts_stats(metabolism_df, col)
  }
}
```

## Results

```{r ts-stats-print, results='asis'}
for (col in names(ts_stats)) {
  res <- ts_stats[[col]]
  if (is.null(res)) next

  cat("\n\n### Metric:", col, "\n\n")

  cat("**Type III ANOVA (Satterthwaite approximation):**\n\n")
  cat(knitr::kable(as.data.frame(res$anova), digits = 4, format = "pipe"), sep = "\n")
  cat("\n")

  cat("**Marginal means – main effects (collapsed across time):**\n\n")
  cat(knitr::kable(as.data.frame(res$pairs_main), digits = 4, format = "pipe"), sep = "\n")
  cat("\n")

  cat("**Pairwise comparisons by timepoint (Tukey):**\n\n")
  cat(knitr::kable(as.data.frame(res$pairs_by_time), digits = 4, format = "pipe"), sep = "\n")
  cat("\n")
}
```

# Area Under the Curve (AUC)

AUC computed per individual via the trapezoid rule across all timepoints.
`metabolism_per_*` is the primary metric matching the paper; `corrected_fc`
and `raw_fluorescence` are retained for reference.

```{r auc-calc}
compute_auc <- function(df, value_col, group_vars) {
  df %>%
    filter(is.finite(time_hr), is.finite(.data[[value_col]])) %>%
    group_by(across(all_of(group_vars))) %>%
    summarise(
      AUC          = trapezoid_auc(time_hr, .data[[value_col]]),
      n_timepoints = n(),
      .groups      = "drop"
    ) %>%
    filter(is.finite(AUC))
}

# Only include grouping columns that are actually present in the data
individual_vars <- intersect(
  c("trace_id", "family_id_group", "treatment_group"),
  names(metabolism_df)
)

auc_metab_list <- list()
if (nrow(metabolism_df) > 0) {
  for (mc in active_meas_cols) {
    col <- paste0("metabolism_per_", mc)
    if (col %in% names(metabolism_df)) {
      auc_metab_list[[col]] <-
        compute_auc(metabolism_df, col, individual_vars) %>%
        mutate(metric = col)
    }
  }
}

auc_all <- bind_rows(auc_metab_list)

str(auc_all)
```

## AUC summary tables

```{r auc-summary}
sum_vars <- intersect(
  c("metric", "family_id_group", "treatment_group"),
  names(auc_all)
)
auc_summary <- auc_all %>%
  group_by(across(all_of(sum_vars))) %>%
  summarise(
    n      = n(),
    mean   = mean(AUC, na.rm = TRUE),
    sd     = sd(AUC, na.rm = TRUE),
    se     = sd / sqrt(n),
    median = median(AUC, na.rm = TRUE),
    .groups = "drop"
  )

print(auc_summary)
```

# Statistical Analysis

Each individual oyster (`sample_id_group`) is the observational unit. The
model is built from whichever grouping factors are present: both family and
treatment (with interaction) when both exist, or a one-way model when only
one factor is available. Each plate maps to a unique family × treatment
combination, so plate-level and group-level variance are confounded;
interpret accordingly.

```{r stats-models}
tidy_tukey <- function(model, term) {
  tk <- tryCatch(TukeyHSD(aov(model), which = term)[[term]],
                 error = function(e) NULL)
  if (is.null(tk)) return(tibble())
  as_tibble(tk, rownames = "contrast") %>%
    rename(
      estimate = diff,
      p.value = `p adj`
    ) %>%
    mutate(SE = NA_real_, df = NA_real_, t.ratio = NA_real_) %>%
    select(contrast, estimate, SE, df, t.ratio, p.value)
}

run_auc_stats <- function(auc_df) {
  empty <- tibble()

  has_family    <- "family_id_group" %in% names(auc_df) &&
    length(unique(na.omit(auc_df$family_id_group))) > 1
  has_treatment <- "treatment_group" %in% names(auc_df) &&
    length(unique(na.omit(auc_df$treatment_group))) > 1

  if (!has_family && !has_treatment) {
    return(list(model = NULL, anova = NULL,
                pairs_full = empty, pairs_family = empty,
                pairs_trt = empty,
                has_family = FALSE, has_treatment = FALSE))
  }

  if (has_family)    auc_df <- auc_df %>% mutate(family    = factor(family_id_group))
  if (has_treatment) auc_df <- auc_df %>% mutate(treatment = factor(treatment_group))

  formula_str <- if (has_family && has_treatment) {
    "AUC ~ family * treatment"
  } else if (has_family) {
    "AUC ~ family"
  } else {
    "AUC ~ treatment"
  }
  model     <- lm(as.formula(formula_str), data = auc_df)
  anova_res <- anova(model)

  if (has_family && has_treatment) {
    if (has_emmeans) {
      pairs_full   <- as.data.frame(pairs(emmeans::emmeans(model, ~ family * treatment),
                                          adjust = "tukey"))
      pairs_family <- as.data.frame(pairs(emmeans::emmeans(model, ~ family),
                                          adjust = "tukey"))
      pairs_trt    <- as.data.frame(pairs(emmeans::emmeans(model, ~ treatment),
                                          adjust = "tukey"))
    } else {
      pairs_full <- empty
      pairs_family <- tidy_tukey(model, "family")
      pairs_trt <- tidy_tukey(model, "treatment")
    }
  } else if (has_family) {
    pairs_family <- if (has_emmeans) {
      as.data.frame(pairs(emmeans::emmeans(model, ~ family),
                          adjust = "tukey"))
    } else {
      tidy_tukey(model, "family")
    }
    pairs_full   <- pairs_family
    pairs_trt    <- empty
  } else {
    pairs_trt    <- if (has_emmeans) {
      as.data.frame(pairs(emmeans::emmeans(model, ~ treatment),
                          adjust = "tukey"))
    } else {
      tidy_tukey(model, "treatment")
    }
    pairs_full   <- pairs_trt
    pairs_family <- empty
  }

  list(
    model         = model,
    anova         = anova_res,
    pairs_full    = pairs_full,
    pairs_family  = pairs_family,
    pairs_trt     = pairs_trt,
    has_family    = has_family,
    has_treatment = has_treatment
  )
}

metrics_to_test <- unique(auc_all$metric)
stats_results   <- map(
  set_names(metrics_to_test),
  ~ run_auc_stats(auc_all %>% filter(metric == .x))
)
```

## Results by metric

```{r stats-print, results='asis'}
for (met in metrics_to_test) {
  stats <- stats_results[[met]]
  cat("\n\n### Metric:", met, "\n\n")
  cat("**ANOVA:**\n\n")
  cat(knitr::kable(as.data.frame(stats$anova), digits = 4, format = "pipe"), sep = "\n")
  cat("\n")
  if (stats$has_family && stats$has_treatment) {
    cat("**Pairwise: family × treatment (Tukey):**\n\n")
    cat(knitr::kable(as.data.frame(stats$pairs_full), digits = 4, format = "pipe"), sep = "\n")
  cat("\n")
    cat("**Pairwise: family main effect:**\n\n")
    cat(knitr::kable(as.data.frame(stats$pairs_family), digits = 4, format = "pipe"), sep = "\n")
  cat("\n")
    cat("**Pairwise: treatment main effect:**\n\n")
    cat(knitr::kable(as.data.frame(stats$pairs_trt), digits = 4, format = "pipe"), sep = "\n")
  cat("\n")
  } else if (stats$has_family) {
    cat("**Pairwise: family (Tukey):**\n\n")
    cat(knitr::kable(as.data.frame(stats$pairs_family), digits = 4, format = "pipe"), sep = "\n")
  cat("\n")
  } else if (stats$has_treatment) {
    cat("**Pairwise: treatment (Tukey):**\n\n")
    cat(knitr::kable(as.data.frame(stats$pairs_trt), digits = 4, format = "pipe"), sep = "\n")
  cat("\n")
  }
}
```

# AUC Box Plots with Statistical Annotations

Significance labels: `***` p < 0.001, `**` p < 0.01, `*` p < 0.05.
Brackets are drawn only for significant pairs (p < 0.05).
Plots are generated for whichever grouping factors are present:
treatment-only, family-only, all-groups, within-family, and within-treatment.

```{r sig-label-fn}
sig_label <- function(p) {
  case_when(p < 0.001 ~ "***", p < 0.01 ~ "**", p < 0.05 ~ "*",
            TRUE ~ "ns")
}

# Add significance brackets to an existing ggplot.
# pairs_df   : data frame with $contrast and $p.value columns
# group_levels: ordered character vector matching x-axis factor levels
# y_vals     : numeric vector of AUC values used to set bracket heights
add_sig_brackets <- function(p, pairs_df, group_levels, y_vals) {
  sig_pairs <- pairs_df %>%
    mutate(label = sig_label(p.value)) %>%
    filter(label != "ns")
  if (nrow(sig_pairs) == 0) return(p)

  y_max   <- max(y_vals, na.rm = TRUE)
  y_range <- diff(range(y_vals, na.rm = TRUE))
  step    <- y_range * 0.12

  for (i in seq_len(nrow(sig_pairs))) {
    parts <- str_split(as.character(sig_pairs$contrast[i]), " - ", 2)[[1]]
    g1 <- trimws(parts[1])
    g2 <- trimws(parts[2])
    x1 <- match(g1, group_levels)
    x2 <- match(g2, group_levels)
    if (is.na(x1) || is.na(x2)) next
    bar_y <- y_max + i * step
    p <- p +
      annotate("segment", x = x1, xend = x2,
               y = bar_y, yend = bar_y,
               colour = "black", linewidth = 0.6) +
      annotate("segment", x = x1, xend = x1,
               y = bar_y, yend = bar_y - step * 0.3,
               colour = "black", linewidth = 0.6) +
      annotate("segment", x = x2, xend = x2,
               y = bar_y, yend = bar_y - step * 0.3,
               colour = "black", linewidth = 0.6) +
      annotate("text", x = (x1 + x2) / 2,
               y = bar_y + step * 0.15,
               label = sig_pairs$label[i], size = 4.5)
  }
  p
}
```

```{r auc-boxplots, fig.width=6, fig.height=5}
for (met in metrics_to_test) {
  df      <- auc_all %>% filter(metric == met)
  stats   <- stats_results[[met]]
  y_lab   <- auc_y_label(met)
  has_fam <- stats$has_family
  has_trt <- stats$has_treatment

  # ── Treatment main effect (x = treatment, tick = treatment name) ───────
  if (has_trt) {
    df_p <- df %>%
      mutate(x = factor(treatment_group, levels = sort(unique(treatment_group))))
    grps <- levels(df_p$x)
    p <- ggplot(df_p, aes(x = x, y = AUC, fill = x)) +
      geom_boxplot(alpha = 0.6, outlier.shape = NA) +
      geom_jitter(width = 0.15, alpha = 0.4, size = 1.5) +
      scale_fill_manual(values = trt_colours[grps], guide = "none") +
      labs(x = "Treatment", y = y_lab) +
      theme_classic(base_size = 13)
    p <- add_sig_brackets(p, stats$pairs_trt, grps, df_p$AUC)
    print(p)
    ggsave(file.path(fig_dir, paste0("auc_treatment_", met, ".png")),
           p, width = 5, height = 5)
  }

  # ── Family main effect (x = family, tick = family name) ───────────────
  if (has_fam) {
    df_p <- df %>%
      mutate(x = factor(family_id_group,
                        levels = str_sort(unique(family_id_group), numeric = TRUE)))
    grps <- levels(df_p$x)
    p <- ggplot(df_p, aes(x = x, y = AUC, fill = x)) +
      geom_boxplot(alpha = 0.6, outlier.shape = NA) +
      geom_jitter(width = 0.15, alpha = 0.4, size = 1.5) +
      scale_fill_manual(values = fam_colours[grps], guide = "none") +
      labs(x = "Family", y = y_lab) +
      theme_classic(base_size = 13)
    p <- add_sig_brackets(p, stats$pairs_family, grps, df_p$AUC)
    print(p)
    ggsave(file.path(fig_dir, paste0("auc_family_", met, ".png")),
           p, width = 5, height = 5)
  }

  # Remaining plots require both factors
  if (!has_fam || !has_trt) next

  # ── All family:treatment groups (x = family:treatment) ─────────────────
  # emmeans contrasts use spaces; convert to colon to match tick labels
  pairs_fc <- stats$pairs_full %>%
    mutate(contrast = str_replace_all(
      contrast,
      "([a-z]+) ([a-z]+)( - )([a-z]+) ([a-z]+)",
      "\\1:\\2\\3\\4:\\5"
    ))
  df_p <- df %>%
    mutate(x = factor(
      paste(family_id_group, treatment_group, sep = ":"),
      levels = str_sort(unique(paste(family_id_group, treatment_group, sep = ":")),
                        numeric = TRUE)
    ))
  grps     <- levels(df_p$x)
  fill_map <- setNames(make_palette(length(grps)), grps)
  p <- ggplot(df_p, aes(x = x, y = AUC, fill = x)) +
    geom_boxplot(alpha = 0.6, outlier.shape = NA) +
    geom_jitter(width = 0.15, alpha = 0.4, size = 1.5) +
    scale_fill_manual(values = fill_map, guide = "none") +
    labs(x = "Family : Treatment", y = y_lab) +
    theme_classic(base_size = 13) +
    theme(axis.text.x = element_text(angle = 20, hjust = 1))
  p <- add_sig_brackets(p, pairs_fc, grps, df_p$AUC)
  print(p)
  ggsave(file.path(fig_dir, paste0("auc_all_groups_", met, ".png")),
         p, width = 6, height = 5)

  # ── Within each family: treatment comparison (x = family:treatment) ────
  # Tick labels are family:treatment so these plots are visually distinct
  # from the treatment main-effect plot above.
  for (fam in str_sort(unique(df$family_id_group), numeric = TRUE)) {
    df_p <- df %>%
      filter(family_id_group == fam) %>%
      mutate(x = factor(
        paste(family_id_group, treatment_group, sep = ":"),
        levels = str_sort(unique(paste(family_id_group, treatment_group, sep = ":")),
                          numeric = TRUE)
      ))
    grps     <- levels(df_p$x)
    pairs_sub <- pairs_fc %>%
      filter(str_count(contrast, paste0(fam, ":")) == 2)
    p <- ggplot(df_p, aes(x = x, y = AUC, fill = x)) +
      geom_boxplot(alpha = 0.6, outlier.shape = NA) +
      geom_jitter(width = 0.15, alpha = 0.4, size = 1.5) +
      scale_fill_manual(values = fill_map[grps], guide = "none") +
      labs(x = "Family : Treatment", y = y_lab) +
      theme_classic(base_size = 13)
    p <- add_sig_brackets(p, pairs_sub, grps, df_p$AUC)
    print(p)
    ggsave(file.path(fig_dir, paste0("auc_", fam, "_trt_", met, ".png")),
           p, width = 5, height = 5)
  }

  # ── Within each treatment: family comparison (x = family:treatment) ────
  # Tick labels are family:treatment so these plots are visually distinct
  # from the family main-effect plot above.
  for (trt in sort(unique(df$treatment_group))) {
    df_p <- df %>%
      filter(treatment_group == trt) %>%
      mutate(x = factor(
        paste(family_id_group, treatment_group, sep = ":"),
        levels = str_sort(unique(paste(family_id_group, treatment_group, sep = ":")),
                          numeric = TRUE)
      ))
    grps      <- levels(df_p$x)
    pairs_sub <- pairs_fc %>%
      filter(str_count(contrast, paste0(":", trt)) == 2)
    p <- ggplot(df_p, aes(x = x, y = AUC, fill = x)) +
      geom_boxplot(alpha = 0.6, outlier.shape = NA) +
      geom_jitter(width = 0.15, alpha = 0.4, size = 1.5) +
      scale_fill_manual(values = fill_map[grps], guide = "none") +
      labs(x = "Family : Treatment", y = y_lab) +
      theme_classic(base_size = 13)
    p <- add_sig_brackets(p, pairs_sub, grps, df_p$AUC)
    print(p)
    ggsave(file.path(fig_dir, paste0("auc_", trt, "_fam_", met, ".png")),
           p, width = 5, height = 5)
  }
}
```

# Curve Feature Analysis: Which Aspects Best Distinguish Families?

This section converts each individual's resazurin trajectory into a suite of
curve traits. These include metabolic capacity (`auc_total`,
`metabolic_scope`), responsiveness (`initial_slope`, `vmax`,
`time_to_vmax`, `time_to_peak`, `inflection_time`), trajectory/stress-like
features (`late_early_auc_ratio`, `delta_auc_late_minus_early`,
`metabolic_depression_index`, `final_delta`, `trajectory_class`), and
metabolic stability (`stability_cv`). Traits are ranked by family effect size
and FDR-adjusted significance to identify the most distinguishing aspects of
the curves.

```{r curve-feature-helper-fns}
safe_divide <- function(num, den) {
  if_else(is.finite(num) & is.finite(den) & abs(den) > .Machine$double.eps,
          num / den, NA_real_)
}

segment_auc <- function(time_hr, value) {
  ok <- is.finite(time_hr) & is.finite(value)
  t <- time_hr[ok]
  v <- value[ok]
  if (length(t) < 2) {
    return(tibble(t_start = numeric(), t_end = numeric(),
                  t_mid = numeric(), auc = numeric()))
  }
  ord <- order(t)
  t <- t[ord]
  v <- v[ord]
  tibble(
    t_start = head(t, -1),
    t_end   = tail(t, -1),
    t_mid   = (head(t, -1) + tail(t, -1)) / 2,
    auc     = diff(t) * (head(v, -1) + tail(v, -1)) / 2
  )
}

classify_trajectory <- function(slopes, slope_tol = 1e-12) {
  if (length(slopes) == 0 || all(!is.finite(slopes))) return(NA_character_)
  s <- slopes[is.finite(slopes)]
  signs <- case_when(
    s >  slope_tol ~  1L,
    s < -slope_tol ~ -1L,
    TRUE           ~  0L
  )
  nonzero <- signs[signs != 0L]
  if (length(nonzero) == 0) return("stable")
  if (all(nonzero > 0)) return("monotonic_increase")
  if (all(nonzero < 0)) return("monotonic_decrease")
  if (any(nonzero > 0) && any(nonzero < 0)) {
    first_neg <- which(nonzero < 0)[1]
    first_pos <- which(nonzero > 0)[1]
    if (first_pos < first_neg) return("rise_then_depress")
    if (first_neg < first_pos) return("depress_then_recover")
  }
  "complex"
}

extract_curve_features_one <- function(df, value_col, metric_label) {
  df <- df %>%
    filter(is.finite(time_hr), is.finite(.data[[value_col]])) %>%
    arrange(time_hr)

  if (nrow(df) < 2) return(tibble())

  t <- df$time_hr
  v <- df[[value_col]]
  dt <- diff(t)
  dv <- diff(v)
  valid_slope <- is.finite(dt) & dt > 0 & is.finite(dv)
  slopes <- dv[valid_slope] / dt[valid_slope]
  slope_midpoints <- (head(t, -1) + tail(t, -1)) / 2
  slope_midpoints <- slope_midpoints[valid_slope]
  seg_auc <- segment_auc(t, v)
  split_time <- median(seg_auc$t_mid, na.rm = TRUE)
  early_auc <- seg_auc %>%
    filter(t_mid <= split_time) %>%
    summarise(x = sum(auc, na.rm = TRUE)) %>%
    pull(x)
  late_auc <- seg_auc %>%
    filter(t_mid > split_time) %>%
    summarise(x = sum(auc, na.rm = TRUE)) %>%
    pull(x)

  if (length(early_auc) == 0) early_auc <- NA_real_
  if (length(late_auc) == 0) late_auc <- NA_real_

  baseline_value <- v[1]
  final_value <- v[length(v)]
  peak_idx <- which.max(v)
  trough_idx <- which.min(v)
  peak_value <- v[peak_idx]
  trough_value <- v[trough_idx]
  peak_after_baseline <- peak_value - baseline_value
  depression_abs <- if (peak_idx < length(v)) {
    max(peak_value - final_value, 0, na.rm = TRUE)
  } else {
    0
  }

  vmax_idx <- if (length(slopes) > 0) which.max(slopes) else NA_integer_
  min_slope_idx <- if (length(slopes) > 0) which.min(slopes) else NA_integer_
  curvature <- diff(slopes)
  inflection_idx <- if (length(curvature) > 0) which.max(abs(curvature)) + 1L else NA_integer_

  tibble(
    value_metric = metric_label,
    baseline_value = baseline_value,
    final_value = final_value,
    peak_value = peak_value,
    trough_value = trough_value,
    time_to_peak = t[peak_idx],
    time_to_trough = t[trough_idx],
    auc_total = trapezoid_auc(t, v),
    auc_early = early_auc,
    auc_late = late_auc,
    delta_auc_late_minus_early = late_auc - early_auc,
    late_early_auc_ratio = safe_divide(late_auc, early_auc),
    initial_slope = slopes[1],
    vmax = if (length(slopes) > 0) max(slopes, na.rm = TRUE) else NA_real_,
    time_to_vmax = if (!is.na(vmax_idx)) slope_midpoints[vmax_idx] else NA_real_,
    min_slope = if (length(slopes) > 0) min(slopes, na.rm = TRUE) else NA_real_,
    time_to_min_slope = if (!is.na(min_slope_idx)) slope_midpoints[min_slope_idx] else NA_real_,
    inflection_time = if (!is.na(inflection_idx)) slope_midpoints[inflection_idx] else NA_real_,
    metabolic_scope = peak_after_baseline,
    final_delta = final_value - baseline_value,
    depression_abs = depression_abs,
    metabolic_depression_index = safe_divide(depression_abs, abs(peak_after_baseline)),
    resilience_ratio = safe_divide(final_value, peak_value),
    stability_cv = safe_divide(sd(v, na.rm = TRUE), abs(mean(v, na.rm = TRUE))),
    trajectory_class = classify_trajectory(slopes),
    n_timepoints = nrow(df)
  )
}

compute_curve_features <- function(df, value_col, metric_label) {
  id_vars <- intersect(
    c("trace_id", "family_id_group", "treatment_group", "round_group",
      "cup_id_group", "sample_id_group"),
    names(df)
  )

  df %>%
    filter(!is.na(family_id_group), is.finite(.data[[value_col]])) %>%
    group_by(across(all_of(id_vars))) %>%
    group_modify(~ extract_curve_features_one(.x, value_col, metric_label)) %>%
    ungroup()
}
```

## Per-individual Curve Features

```{r curve-features-calc}
feature_metric_cols <- c(
  corrected_fc = "corrected_fc",
  setNames(paste0("metabolism_per_", active_meas_cols),
           paste0("metabolism_per_", active_meas_cols))
)
feature_metric_cols <- feature_metric_cols[
  feature_metric_cols %in% names(metabolism_df)
]

curve_features <- imap_dfr(
  feature_metric_cols,
  ~ compute_curve_features(metabolism_df, .x, .y)
)

curve_feature_long <- curve_features %>%
  pivot_longer(
    cols = where(is.numeric) & !any_of(c("n_timepoints")),
    names_to = "feature",
    values_to = "feature_value"
  ) %>%
  filter(is.finite(feature_value))

trajectory_features <- curve_features %>%
  count(value_metric, family_id_group, trajectory_class, name = "n") %>%
  group_by(value_metric, family_id_group) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

str(curve_features)
```

## Family Summaries for Curve Features

```{r curve-feature-family-summary}
curve_feature_family_summary <- curve_feature_long %>%
  group_by(value_metric, feature, family_id_group) %>%
  summarise(
    n = n(),
    mean = mean(feature_value, na.rm = TRUE),
    sd = sd(feature_value, na.rm = TRUE),
    se = sd / sqrt(n),
    median = median(feature_value, na.rm = TRUE),
    iqr = IQR(feature_value, na.rm = TRUE),
    .groups = "drop"
  )

print(curve_feature_family_summary)
```

## Rank Features by Family Separation

For each numeric feature, a one-way family model is fit using individual oysters
as observational units. Features are ranked by adjusted p-value and eta-squared
family effect size. `max_abs_pairwise_d` is the largest absolute pairwise
standardized family mean difference and helps identify features with strong
family separation even when sample sizes are modest.

```{r curve-feature-stats}
run_feature_family_stats <- function(df) {
  df <- df %>%
    filter(is.finite(feature_value), !is.na(family_id_group)) %>%
    mutate(family = factor(family_id_group))

  if (nrow(df) < 3 || n_distinct(df$family) < 2) return(tibble())

  model <- lm(feature_value ~ family, data = df)
  aov_tab <- anova(model)

  ss_family <- aov_tab["family", "Sum Sq"]
  df_family <- aov_tab["family", "Df"]
  ms_error <- aov_tab["Residuals", "Mean Sq"]
  ss_total <- sum(aov_tab$`Sum Sq`, na.rm = TRUE)
  eta_sq <- ss_family / ss_total
  omega_sq <- (ss_family - df_family * ms_error) / (ss_total + ms_error)

  fam_means <- df %>%
    group_by(family) %>%
    summarise(mean = mean(feature_value), .groups = "drop")

  pair_grid <- tidyr::crossing(fam1 = fam_means$family,
                               fam2 = fam_means$family) %>%
    filter(as.character(fam1) < as.character(fam2)) %>%
    left_join(fam_means, by = c("fam1" = "family")) %>%
    rename(mean1 = mean) %>%
    left_join(fam_means, by = c("fam2" = "family")) %>%
    rename(mean2 = mean) %>%
    mutate(abs_d = abs(mean1 - mean2) / sqrt(ms_error))

  kw <- kruskal.test(feature_value ~ family, data = df)

  tibble(
    n = nrow(df),
    n_family = n_distinct(df$family),
    f_statistic = aov_tab["family", "F value"],
    p_value = aov_tab["family", "Pr(>F)"],
    kruskal_p_value = kw$p.value,
    eta_sq = eta_sq,
    omega_sq = max(omega_sq, 0, na.rm = TRUE),
    max_abs_pairwise_d = max(pair_grid$abs_d, na.rm = TRUE)
  )
}

curve_feature_family_stats <- curve_feature_long %>%
  group_by(value_metric, feature) %>%
  group_modify(~ run_feature_family_stats(.x)) %>%
  ungroup() %>%
  group_by(value_metric) %>%
  mutate(
    p_adj_bh = p.adjust(p_value, method = "BH"),
    rank_within_metric = min_rank(p_adj_bh),
    rank_score = dense_rank(desc(eta_sq)) + dense_rank(p_adj_bh)
  ) %>%
  ungroup() %>%
  arrange(p_adj_bh, desc(eta_sq), desc(max_abs_pairwise_d))

top_curve_features <- curve_feature_family_stats %>%
  group_by(value_metric) %>%
  slice_min(order_by = rank_score, n = 5, with_ties = FALSE) %>%
  ungroup() %>%
  arrange(value_metric, rank_score)

best_curve_feature <- curve_feature_family_stats %>%
  slice_min(order_by = rank_score, n = 1, with_ties = FALSE)

print(curve_feature_family_stats)
```

## Pairwise Family Comparisons for Curve Features

```{r curve-feature-pairwise}
run_feature_pairwise <- function(df) {
  df <- df %>%
    filter(is.finite(feature_value), !is.na(family_id_group)) %>%
    mutate(family = factor(family_id_group))

  if (nrow(df) < 3 || n_distinct(df$family) < 2) return(tibble())

  model <- lm(feature_value ~ family, data = df)
  if (has_emmeans) {
    as.data.frame(pairs(emmeans::emmeans(model, ~ family), adjust = "tukey")) %>%
      as_tibble()
  } else {
    tidy_tukey(model, "family")
  }
}

curve_feature_pairwise_stats <- curve_feature_long %>%
  semi_join(top_curve_features, by = c("value_metric", "feature")) %>%
  group_by(value_metric, feature) %>%
  group_modify(~ run_feature_pairwise(.x)) %>%
  ungroup()

print(curve_feature_pairwise_stats)
```

## Most Distinguishing Curve Aspects

```{r curve-feature-ranking-print, results='asis'}
cat("### Best overall numeric curve discriminator\n\n")
cat(knitr::kable(best_curve_feature, digits = 5, format = "pipe"), sep = "\n")
cat("\n\n### Top numeric curve discriminators by metric\n\n")
cat(knitr::kable(top_curve_features, digits = 5, format = "pipe"), sep = "\n")
```

```{r curve-feature-rank-plot, fig.width=9, fig.height=6}
rank_plot_df <- curve_feature_family_stats %>%
  group_by(value_metric) %>%
  slice_max(order_by = eta_sq, n = 12, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(feature = fct_reorder(feature, eta_sq))

p_feature_rank <- ggplot(rank_plot_df,
                         aes(x = eta_sq, y = feature, fill = p_adj_bh < 0.05)) +
  geom_col(width = 0.75) +
  facet_wrap(~ value_metric, scales = "free_y") +
  scale_fill_manual(values = c("FALSE" = "grey70", "TRUE" = "#0072B2"),
                    name = "BH-adjusted p < 0.05") +
  labs(x = "Family effect size (eta-squared)", y = "Curve feature") +
  theme_classic(base_size = 12)

p_feature_rank
ggsave(file.path(fig_dir, "curve_feature_family_effect_ranking.png"),
       p_feature_rank, width = 9, height = 6)
```

```{r curve-feature-best-boxplot, fig.width=8, fig.height=5}
if (nrow(best_curve_feature) > 0) {
  best_metric <- best_curve_feature$value_metric[1]
  best_feature <- best_curve_feature$feature[1]

  best_plot_df <- curve_feature_long %>%
    filter(value_metric == best_metric, feature == best_feature) %>%
    mutate(family_id_group = factor(
      family_id_group,
      levels = str_sort(unique(family_id_group), numeric = TRUE)
    ))

  p_best_feature <- ggplot(best_plot_df,
      aes(x = family_id_group, y = feature_value, fill = family_id_group)) +
    geom_boxplot(alpha = 0.65, outlier.shape = NA) +
    geom_jitter(width = 0.15, alpha = 0.55, size = 1.8) +
    scale_fill_manual(values = fam_colours[levels(best_plot_df$family_id_group)],
                      guide = "none") +
    labs(x = "Family",
         y = best_feature,
         title = paste("Most distinguishing curve feature:", best_feature),
         subtitle = best_metric) +
    theme_classic(base_size = 13)

  p_best_feature
  ggsave(file.path(fig_dir, "curve_feature_best_family_boxplot.png"),
         p_best_feature, width = 8, height = 5)
}
```

```{r curve-feature-family-heatmap, fig.width=10, fig.height=7}
heatmap_features <- curve_feature_family_stats %>%
  group_by(value_metric) %>%
  slice_max(order_by = eta_sq, n = 10, with_ties = FALSE) %>%
  ungroup() %>%
  select(value_metric, feature)

curve_feature_heatmap_df <- curve_feature_family_summary %>%
  semi_join(heatmap_features, by = c("value_metric", "feature")) %>%
  group_by(value_metric, feature) %>%
  mutate(mean_z = as.numeric(scale(mean))) %>%
  ungroup() %>%
  mutate(
    feature_label = paste(value_metric, feature, sep = " | "),
    family_id_group = factor(
      family_id_group,
      levels = str_sort(unique(family_id_group), numeric = TRUE)
    )
  )

p_feature_heatmap <- ggplot(curve_feature_heatmap_df,
    aes(x = family_id_group, y = feature_label, fill = mean_z)) +
  geom_tile(colour = "white", linewidth = 0.25) +
  scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#B2182B",
                       midpoint = 0, name = "Family mean z") +
  labs(x = "Family", y = "Curve feature") +
  theme_classic(base_size = 11) +
  theme(axis.text.y = element_text(size = 8))

p_feature_heatmap
ggsave(file.path(fig_dir, "curve_feature_family_mean_heatmap.png"),
       p_feature_heatmap, width = 10, height = 7)
```

```{r trajectory-class-summary, fig.width=8, fig.height=5}
p_trajectory <- ggplot(trajectory_features,
    aes(x = family_id_group, y = prop, fill = trajectory_class)) +
  geom_col(position = "stack", colour = "white", linewidth = 0.2) +
  facet_wrap(~ value_metric) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Family", y = "Individuals", fill = "Trajectory class") +
  theme_classic(base_size = 12)

p_trajectory
ggsave(file.path(fig_dir, "curve_feature_trajectory_classes.png"),
       p_trajectory, width = 8, height = 5)
```

# Save Output Data

```{r save-outputs}
write_csv(auc_all,      file.path(out_dir, "auc_all_metrics.csv"))
write_csv(auc_summary,  file.path(out_dir, "auc_summary.csv"))
write_csv(curve_features,
          file.path(out_dir, "curve_features.csv"))
write_csv(curve_feature_family_summary,
          file.path(out_dir, "curve_feature_family_summary.csv"))
write_csv(curve_feature_family_stats,
          file.path(out_dir, "curve_feature_family_stats.csv"))
write_csv(curve_feature_pairwise_stats,
          file.path(out_dir, "curve_feature_pairwise_stats.csv"))
write_csv(trajectory_features,
          file.path(out_dir, "curve_feature_trajectory_classes.csv"))

if (nrow(metabolism_df) > 0)
  write_csv(metabolism_df,
            file.path(out_dir, "metabolism.csv"))

stats_compiled <- map_dfr(metrics_to_test, function(met) {
  bind_rows(
    stats_results[[met]]$pairs_full %>%
      mutate(comparison = "family:treatment"),
    stats_results[[met]]$pairs_family %>%
      mutate(comparison = "family"),
    stats_results[[met]]$pairs_trt %>%
      mutate(comparison = "treatment")
  ) %>% mutate(metric = met)
})

write_csv(stats_compiled,
          file.path(out_dir, "pairwise_stats.csv"))

message("Output files written to: ", out_dir)
```
