---
title: "00.00-mgig-heat-survivorship.Rmd"
author: "Sam White"
date: "2026-05-07"
output: 
  github_document:
    toc: true
    number_sections: true
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
---

# BACKGROUND

This is a Kaplan-Meier survival analysis of 33<sup>o</sup>C heat stress of _Magallana gigas_ oysters, [conducted on 20260427](https://github.com/RobertsLab/sormi-assay-development/tree/main/heat-survivorship/20260427-33C-USDA-families) (GitHub repo). It compares 9 different families bred/selected by the USDA.

These oysters had been [previously heat stressed @ 36<sup>o</sup>C and assessed by the resazurin assay on 20260415](https://github.com/RobertsLab/sormi-assay-development/blob/main/Resazurin/20260415-36C/README.md) (GitHub README)

## SETUP

### Load packages

```{r setup, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(survival)
library(ggplot2)

knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = "",         # Prevents appending '##' to beginning of lines in code output
  results = 'hold',    # Holds output so it's all printed together after code chunk
  fig.path = "../outputs/00.00-mgig-heat-survivorship/"
)
```


### Read in and view raw data

```{r read-raw-data}
survivorship_raw <- read_csv(
	"../data/20260427-33C-USDA-families/survivorship.csv",
	show_col_types = FALSE
)

# Preview newly created object
cat("\n=== survivorship_raw: str() ===\n")
str(survivorship_raw)
```


## ANALYSIS

### Clean and standardize columns used for survival analysis

```{r clean-columns, message=FALSE, warning=FALSE}
survivorship_clean <- survivorship_raw %>%
	mutate(
		individual_id = as.character(individual_id),
		family_group = as.character(`familly_id.group`),
		timepoint_hrs = as.numeric(timepoint_hrs),
		alive_chr = toupper(trimws(as.character(`alive.measurement`))),
		alive = case_when(
			alive_chr %in% c("TRUE", "T", "1", "YES", "Y") ~ TRUE,
			alive_chr %in% c("FALSE", "F", "0", "NO", "N") ~ FALSE,
			TRUE ~ NA
		)
	) %>%
	select(
		individual_id,
		family_group,
		timepoint_count,
		timepoint_hrs,
		alive,
		date,
		time,
		everything()
	)

# Preview newly created object
cat("\n=== survivorship_clean: str() ===\n")
str(survivorship_clean)
cat("\n=== survivorship_clean: summary(alive) ===\n")
summary(survivorship_clean$alive)
```

### Reduce repeated observations to one survival row per individual

```{r build-individual-survival, message=FALSE, warning=FALSE}
individual_survival <- survivorship_clean %>%
	filter(!is.na(individual_id), !is.na(timepoint_hrs), !is.na(alive)) %>%
	arrange(individual_id, timepoint_hrs) %>%
	group_by(individual_id) %>%
	summarise(
		family_group = first(family_group[!is.na(family_group)]),
		event = if_else(any(!alive), 1L, 0L),
		time_to_event = {
			if (any(!alive)) {
				min(timepoint_hrs[!alive])
			} else {
				max(timepoint_hrs)
			}
		},
		n_observations = n(),
		.groups = "drop"
	)

# Preview newly created object
cat("\n=== individual_survival: str() ===\n")
str(individual_survival)
cat("\n=== individual_survival: summary(time_to_event) ===\n")
summary(individual_survival$time_to_event)
cat("\n=== individual_survival: table(event) ===\n")
table(individual_survival$event)
```

### Create Surv object and fit Kaplan-Meier models

```{r fit-km-models, message=FALSE, warning=FALSE}
surv_object <- with(individual_survival, Surv(time = time_to_event, event = event))

km_fit_overall <- survfit(surv_object ~ 1, data = individual_survival)
km_fit_by_family <- survfit(surv_object ~ family_group, data = individual_survival)

# Preview newly created objects
cat("\n=== km_fit_overall ===\n")
print(km_fit_overall)
cat("\n=== km_fit_by_family ===\n")
print(km_fit_by_family)
cat("\n=== km_fit_overall: str() ===\n")
str(km_fit_overall)
cat("\n=== km_fit_by_family: str() ===\n")
str(km_fit_by_family)
```

### Perform log-rank test to compare survival between families

```{r log-rank-test, message=FALSE, warning=FALSE}
logrank_test <- survdiff(surv_object ~ family_group, data = individual_survival)

# Display test results
cat("\n=== Log-rank test: survdiff() output ===\n")
print(logrank_test)

# Extract and report key statistics
cat("\n=== Log-Rank Test Summary ===\n")
cat("Chi-square statistic:", logrank_test$chisq, "\n")
cat("Degrees of freedom:", length(logrank_test$n) - 1, "\n")
cat("p-value:", 1 - pchisq(logrank_test$chisq, df = length(logrank_test$n) - 1), "\n")
cat("\nInterpretation:\n")
p_val <- 1 - pchisq(logrank_test$chisq, df = length(logrank_test$n) - 1)
if (p_val < 0.05) {
  cat("p < 0.05: Family groups show SIGNIFICANTLY DIFFERENT survival (reject null hypothesis)\n")
} else {
  cat("p >= 0.05: No significant difference in survival detected between family groups\n")
}
```

### Convert Kaplan-Meier fit to plotting data frames

```{r build-plot-data, message=FALSE, warning=FALSE}
km_overall_df <- bind_rows(
	data.frame(
		time = 0,
		surv = 1,
		n_risk = km_fit_overall$n,
		n_event = 0,
		n_censor = 0
	),
	summary(km_fit_overall) %>%
		with(
			data.frame(
				time = time,
				surv = surv,
				n_risk = n.risk,
				n_event = n.event,
				n_censor = n.censor
			)
		)
) %>%
	arrange(time)

km_family_df <- bind_rows(
	data.frame(
		time = 0,
		surv = 1,
		n_risk = as.numeric(km_fit_by_family$n),
		n_event = 0,
		n_censor = 0,
		strata = names(km_fit_by_family$strata)
	),
	summary(km_fit_by_family) %>%
		with(
			data.frame(
				time = time,
				surv = surv,
				n_risk = n.risk,
				n_event = n.event,
				n_censor = n.censor,
				strata = strata
			)
		)
) %>%
	mutate(family_group = sub("^family_group=", "", strata)) %>%
	arrange(family_group, time)

# Preview newly created objects
cat("\n=== km_overall_df: head() ===\n")
print(head(km_overall_df))
cat("\n=== km_overall_df: str() ===\n")
str(km_overall_df)

cat("\n=== km_family_df: head() ===\n")
print(head(km_family_df))
cat("\n=== km_family_df: str() ===\n")
str(km_family_df)
```

## PLOTS

### Plot Kaplan-Meier curves

#### Overall

```{r plot-km-overall, fig.width=8, fig.height=5, message=FALSE, warning=FALSE}
ggplot(km_overall_df, aes(x = time, y = surv)) +
	geom_step(linewidth = 1.1) +
	labs(
		title = "Kaplan-Meier Survival Curve (Overall)",
		x = "Time (hours)",
		y = "Survival probability"
	) +
	ylim(0, 1) +
	theme_minimal(base_size = 12)
```

#### Families

```{r plot-km-by-family, fig.width=9, fig.height=6, message=FALSE, warning=FALSE}
ggplot(km_family_df, aes(x = time, y = surv, color = family_group)) +
	geom_step(linewidth = 1.0) +
	labs(
		title = "Kaplan-Meier Survival Curves by Family Group",
		x = "Time (hours)",
		y = "Survival probability",
		color = "Family group"
	) +
	ylim(0, 1) +
	theme_minimal(base_size = 12)
```
