---
title: "Kaplan-Meier Survivorship Analysis — 20260609 (36 °C)"
author: "Steven Roberts"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
editor: visual
---

BACKGROUND - Re-run of the 20260609 36 °C *Magallana gigas* USDA-family heat-stress survivorship experiment through the reusable Kaplan-Meier template (`code/01-mgig-heat-survivorship-20260604.qmd`), so all dates are analyzed with an identical pipeline (KM curves via `ggsurvfit`, global log-rank test, and a Cox Proportional Hazards model with hazard ratios).

Original (un-templated) analysis: `code/00.00-mgig-heat-survivorship-20260609.Rmd`.

```{=html}
<!--
================================================================================
REUSABLE KAPLAN-MEIER SURVIVORSHIP TEMPLATE (long-format input variant)
================================================================================
Adapted from code/01-mgig-heat-survivorship-20260604.qmd (which itself adapts
A. Huffmyer's Manchester stress-survival analysis). The downstream KM / log-rank
/ Cox sections are kept identical to that template so results are directly
comparable across experiments. ONLY the data read + reshape is changed to accept
the LONG-format survivorship.csv used by the 20260427/0511/0609 experiments:

  one row per individual x time point, with columns
    individual_id, familly_id.group, timepoint_hrs, alive.measurement
  (plus, in later experiments, plate_ID / plate_well / is_blank).

  alive.measurement TRUE  = alive  -> status 0
  alive.measurement FALSE = dead   -> status 1 (event)
================================================================================
-->
```

## Set up

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(survival)
library(ggsurvfit)
library(gtsummary)
library(cardx)
library(cowplot)
```

```{r config}
# >>> EDIT: analysis configuration ---------------------------------------------

# Path to the LONG-format mortality csv for this experiment
data_path  <- "../data/20260609-36C-USDA-families/survivorship.csv"

# Output directory for saved figures (new directory for this re-run)
out_dir    <- "../outputs/04-mgig-heat-survivorship-20260609/"

# Grouping / factor column(s) to compare survival across.
group_vars <- c("Family")

# Individual ID column produced in the reshape below.
id_var     <- "uid"
# ------------------------------------------------------------------------------
```

## Read in data

```{r read-data, message=FALSE}
raw <- readr::read_csv(data_path, show_col_types = FALSE)
glimpse(raw)
```

## Reshape to one (time, status) row per individual

```{r reshape}
# Standardize the long-format columns this experiment family uses.
raw <- raw %>%
  rename(Family = `familly_id.group`, alive_raw = `alive.measurement`) %>%
  mutate(
    Family        = as.character(Family),
    timepoint_hrs = as.numeric(timepoint_hrs),
    alive = case_when(
      toupper(trimws(as.character(alive_raw))) %in% c("TRUE","T","1","YES","Y") ~ TRUE,
      toupper(trimws(as.character(alive_raw))) %in% c("FALSE","F","0","NO","N") ~ FALSE,
      TRUE ~ NA
    )
  )

# Drop blank wells when the experiment flags them.
if ("is_blank" %in% names(raw)) {
  raw <- raw %>%
    filter(!(toupper(trimws(as.character(is_blank))) %in% c("TRUE","T","1","YES","Y")))
}

# Build a TRULY unique individual key. When plate_ID + plate_well exist they are
# the physical, guaranteed-unique well identity; otherwise individual_id is unique.
if (all(c("plate_ID","plate_well") %in% names(raw))) {
  raw <- raw %>% mutate(uid = paste(plate_ID, plate_well, sep = "_"))
} else {
  raw <- raw %>% mutate(uid = as.character(individual_id))
}

# Collapse to ONE (time, status) row per individual:
#   - if it ever died (alive == FALSE), keep the FIRST death time (event)
#   - otherwise keep the LAST observed time point (right-censored)
data <- raw %>%
  filter(!is.na(uid), !is.na(timepoint_hrs), !is.na(alive)) %>%
  group_by(uid) %>%
  summarise(
    Family = first(Family[!is.na(Family)]),
    status = as.integer(any(!alive)),
    time   = if (any(!alive)) min(timepoint_hrs[!alive]) else max(timepoint_hrs),
    .groups = "drop"
  ) %>%
  mutate(
    Family = factor(Family, levels = unique(Family[order(readr::parse_number(Family), Family)]))
  )

cat("n individuals:", nrow(data), " | events (deaths):", sum(data$status), "\n")
head(data)
```

## Generate Kaplan-Meier survival curves

```{r km-fit}
km_formula <- as.formula(
  paste("Surv(time, status) ~", paste(group_vars, collapse = " + "))
)

s1 <- survival::survfit(km_formula, data = data)
print(s1)
summary(s1)
```

```{r km-plot-combined, warning=FALSE}
ggsurvfit::survfit2(km_formula, data = data) %>%
  ggsurvfit::ggsurvfit() +
  labs(x = "Time (hours)", y = "Survival probability") +
  ylim(0, 1)
```

## Global log-rank test (family comparison)

```{r log-rank}
logrank <- survival::survdiff(km_formula, data = data)
print(logrank)
p_val <- 1 - pchisq(logrank$chisq, df = length(logrank$n) - 1)
cat("Log-rank chi-square:", round(logrank$chisq, 4),
    "| df:", length(logrank$n) - 1,
    "| p-value:", signif(p_val, 4), "\n")
```

## Cox Proportional Hazards model

```{r cox-model}
cox_formula <- as.formula(
  paste("Surv(time, status) ~", paste(group_vars, collapse = " * "))
)
cox_fit <- survival::coxph(cox_formula, data = data)
cox_fit
cox_fit %>% tbl_regression(exp = TRUE)
```

## Publication plot

```{r theme}
my_theme <- theme_classic() +
  theme(
    axis.title.y = element_text(size = 14, color = "black"),
    axis.text.y  = element_text(size = 12, color = "black"),
    axis.text.x  = element_text(size = 12, color = "black"),
    axis.title.x = element_text(size = 14, color = "black")
  )
```

```{r plot-by-family, warning=FALSE}
g1 <- group_vars[1]
f1 <- as.formula(paste("Surv(time, status) ~", g1))
n_groups1     <- nlevels(factor(data[[g1]]))
group_colors1 <- scales::hue_pal()(n_groups1)

plot1 <- ggsurvfit::survfit2(f1, data = data) %>%
  ggsurvfit::ggsurvfit() +
  labs(x = "Time (hours)", y = "Survival probability", title = "") +
  ylim(0, 1) +
  my_theme +
  scale_colour_manual(values = group_colors1) +
  theme(legend.position = "right")

plot1
```

```{r save-plots}
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
ggsave(file.path(out_dir, "survival_by_family.png"), plot1, width = 7, height = 5, dpi = 300)
```

## Take homes

```{r take-homes}
med <- summary(survival::survfit(Surv(time, status) ~ 1, data = data))$table["median"]
cat("Experiment: 20260609, 36 C\n")
cat("Individuals:", nrow(data), "| Deaths:", sum(data$status),
    sprintf("(%.0f%%)\n", 100 * mean(data$status)))
cat("Overall median survival (hrs):", ifelse(is.na(med), "not reached (>50% survived)", med), "\n")
cat("Family log-rank p-value:", signif(p_val, 4),
    ifelse(p_val < 0.05, "-> families differ significantly\n",
                         "-> no significant family difference\n"))
```
