---
title: "Heat Survivorship Cross-Experiment Summary"
author: "Steven Roberts"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
editor: visual
---

## Background

This report summarizes the heat survivorship experiments in
`heat-survivorship/data` and compares survival patterns across experiment dates,
temperatures, and USDA families.

The experiments were not all run with identical input formats or protocols. The
20260604 data are stored in wide format, while the other experiments are stored
as one observation per individual and time point. This report standardizes both
formats into one survival-analysis table with one row per individual:

- `time`: first observed death time, or final observed time if still alive
- `status`: death event indicator, where `1 = dead` and `0 = censored/alive`
- `Experiment`: experiment date and temperature label
- `Family`: observed USDA family/group label
- `Family_identity_note`: notes ambiguous labels that should not be interpreted
  as confirmed biological family identity

Cross-experiment comparisons should be interpreted as descriptive because
temperature, container format, prior handling, and time-point schedules differ
among experiments.

Size data are also not available uniformly across experiments. The 20260427
experiment includes ImageJ area measurements (`area_mm2.measurement` in
`metadata.csv` and `area_mm2.measurment` in `survivorship.csv`), but the
20260511, 20260604, and 20260609 survivorship files do not include comparable
oyster size measurements. Therefore, size differences can be summarized for
20260427 only and cannot be evaluated directly across all experiments with the
current data files.

Briefly, the family coverage also differs among experiments. The 20260427
experiment examined families 1, 2, 3, 5, 6, 7, 8, 9, and 10; 20260511 examined
families 1, 3, 5, 6, 7, 8, 9, 9B, and 10; 20260604 examined families 1, 2, 3,
5, 6, 8, 9, and 10; and 20260609 examined families 1, 3, 5, 6, 7, 8, 9, 9B,
and 10. In the 20260511 and 20260609 experiments, two groups were both
originally labelled family 9. One of those groups was actually family 2 and the
other was family 9, but the true identity of each group could not be resolved;
therefore one group is retained as `9` and the other as `9B`. In this report,
`9` and `9B` for those two experiments are treated as observed groups with
ambiguous 2/9 identity, not as two confirmed biological families. Families 1, 3,
5, 6, 8, and 10 are the clearly shared families across all experiments; family
7 is absent from 20260604; and family 2 can only be interpreted directly in
20260427 and 20260604.

## Set up

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(survival)
library(ggplot2)
```

```{r config}
experiments <- tribble(
  ~experiment_id, ~experiment_label, ~temperature_C, ~format, ~data_path,
  "20260427", "20260427 - 33 C", 33, "long",
    "../data/20260427-33C-USDA-families/survivorship.csv",
  "20260511", "20260511 - 33 C", 33, "long",
    "../data/20260511-33C-USDA-families/survivorship.csv",
  "20260604", "20260604 - 36 C", 36, "wide",
    "../data/20260604-36C-USDA-families/survivorship.csv",
  "20260609", "20260609 - 36 C", 36, "long",
    "../data/20260609-36C-USDA-families/survivorship.csv"
)

out_dir <- "../outputs/05-mgig-heat-survivorship-experiment-summary/"
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
```

## Standardize data

```{r helper-functions}
parse_alive <- function(x) {
  case_when(
    toupper(trimws(as.character(x))) %in% c("TRUE", "T", "1", "YES", "Y") ~ TRUE,
    toupper(trimws(as.character(x))) %in% c("FALSE", "F", "0", "NO", "N") ~ FALSE,
    TRUE ~ NA
  )
}

clean_family <- function(x) {
  x <- as.character(x)
  x <- str_remove(x, "^f")
  x <- str_replace(x, "^0+", "")
  str_to_upper(x)
}

family_identity_note <- function(family, experiment_id) {
  case_when(
    experiment_id %in% c("20260511", "20260609") & family %in% c("9", "9B") ~
      "Ambiguous duplicate 9 label: one group is family 2 and one is family 9",
    TRUE ~ "Observed label treated as family identity"
  )
}

km_curve_data <- function(data, groups) {
  data %>%
    group_by(across(all_of(groups))) %>%
    group_modify(\(.x, .y) {
      fit <- survfit(Surv(time, status) ~ 1, data = .x)
      fit_summary <- summary(fit)

      tibble(
        time = c(0, fit_summary$time),
        survival = c(1, fit_summary$surv),
        n_risk = c(nrow(.x), fit_summary$n.risk),
        n_event = c(0, fit_summary$n.event)
      )
    }) %>%
    ungroup()
}

read_long_survival <- function(path, experiment_id, experiment_label, temperature_C) {
  raw <- readr::read_csv(path, show_col_types = FALSE) %>%
    rename(Family = `familly_id.group`, alive_raw = `alive.measurement`) %>%
    mutate(
      Family = clean_family(Family),
      timepoint_hrs = as.numeric(timepoint_hrs),
      alive = parse_alive(alive_raw)
    )

  if ("is_blank" %in% names(raw)) {
    raw <- raw %>%
      filter(!parse_alive(is_blank))
  }

  if (all(c("plate_ID", "plate_well") %in% names(raw))) {
    raw <- raw %>% mutate(uid = paste(plate_ID, plate_well, sep = "_"))
  } else {
    raw <- raw %>% mutate(uid = as.character(individual_id))
  }

  raw %>%
    filter(!is.na(uid), !is.na(timepoint_hrs), !is.na(alive), !is.na(Family)) %>%
    group_by(uid) %>%
    summarise(
      Family = first(Family[!is.na(Family)]),
      time = if (any(!alive)) min(timepoint_hrs[!alive]) else max(timepoint_hrs),
      status = as.integer(any(!alive)),
      n_observations = n(),
      .groups = "drop"
    ) %>%
    mutate(
      experiment_id = experiment_id,
      Experiment = experiment_label,
      temperature_C = temperature_C,
      Family_identity_note = family_identity_note(Family, experiment_id)
    )
}

read_wide_survival <- function(path, experiment_id, experiment_label, temperature_C) {
  raw <- readr::read_csv(path, show_col_types = FALSE)

  time_cols <- names(raw)[!is.na(suppressWarnings(as.numeric(names(raw))))]

  raw %>%
    mutate(
      uid = as.character(ID),
      Family = clean_family(Family)
    ) %>%
    pivot_longer(
      cols = all_of(time_cols),
      names_to = "time",
      values_to = "status_at_time"
    ) %>%
    mutate(
      time = as.numeric(time),
      status_at_time = as.integer(status_at_time)
    ) %>%
    filter(!is.na(uid), !is.na(time), !is.na(status_at_time), !is.na(Family)) %>%
    group_by(uid, Family) %>%
    summarise(
      time = if (any(status_at_time == 1)) min(time[status_at_time == 1]) else max(time),
      status = as.integer(any(status_at_time == 1)),
      n_observations = n(),
      .groups = "drop"
    ) %>%
    mutate(
      experiment_id = experiment_id,
      Experiment = experiment_label,
      temperature_C = temperature_C,
      Family_identity_note = family_identity_note(Family, experiment_id)
    )
}

read_experiment <- function(experiment_id, experiment_label, temperature_C, format, data_path) {
  if (format == "long") {
    read_long_survival(data_path, experiment_id, experiment_label, temperature_C)
  } else if (format == "wide") {
    read_wide_survival(data_path, experiment_id, experiment_label, temperature_C)
  } else {
    stop("Unsupported experiment format: ", format)
  }
}
```

```{r build-standardized-data}
survival_data <- pmap_dfr(experiments, read_experiment) %>%
  mutate(
    Experiment = factor(Experiment, levels = experiments$experiment_label),
    temperature_C = factor(temperature_C),
    Family = factor(Family, levels = unique(Family[order(readr::parse_number(Family), Family)]))
  )

readr::write_csv(survival_data, file.path(out_dir, "standardized_survival_data.csv"))

glimpse(survival_data)
```

## Experiment summaries

```{r experiment-summary}
overall_summary <- survival_data %>%
  group_by(Experiment, temperature_C) %>%
  summarise(
    individuals = n(),
    observed_family_groups = n_distinct(Family),
    deaths = sum(status),
    death_percent = 100 * mean(status),
    first_event_hr = min(time[status == 1], na.rm = TRUE),
    last_observed_hr = max(time, na.rm = TRUE),
    median_survival_hr = summary(survfit(Surv(time, status) ~ 1))$table["median"],
    .groups = "drop"
  ) %>%
  mutate(
    first_event_hr = if_else(is.infinite(first_event_hr), NA_real_, first_event_hr),
    median_survival_hr = if_else(is.na(median_survival_hr), NA_real_, median_survival_hr)
  )

overall_summary %>%
  mutate(across(
    c(death_percent, first_event_hr, last_observed_hr, median_survival_hr),
    \(x) round(x, 1)
  )) %>%
  knitr::kable()
```

```{r family-summary}
family_summary <- survival_data %>%
  group_by(Experiment, temperature_C, Family, Family_identity_note) %>%
  summarise(
    individuals = n(),
    deaths = sum(status),
    death_percent = 100 * mean(status),
    median_survival_hr = summary(survfit(Surv(time, status) ~ 1))$table["median"],
    .groups = "drop"
  ) %>%
  mutate(median_survival_hr = if_else(is.na(median_survival_hr), NA_real_, median_survival_hr))

readr::write_csv(family_summary, file.path(out_dir, "family_survival_summary.csv"))

family_summary %>%
  arrange(Experiment, readr::parse_number(as.character(Family))) %>%
  mutate(across(c(death_percent, median_survival_hr), \(x) round(x, 1))) %>%
  knitr::kable()
```

## Size data availability

```{r size-data-availability}
size_metadata_20260427 <- readr::read_csv(
  "../data/20260427-33C-USDA-families/metadata.csv",
  show_col_types = FALSE
) %>%
  mutate(Family = clean_family(`familly_id.group`))

size_overall_20260427 <- size_metadata_20260427 %>%
  summarise(
    Experiment = "20260427 - 33 C",
    individuals = n(),
    individuals_with_area = sum(!is.na(area_mm2.measurement)),
    mean_area_mm2 = mean(area_mm2.measurement, na.rm = TRUE),
    median_area_mm2 = median(area_mm2.measurement, na.rm = TRUE),
    min_area_mm2 = min(area_mm2.measurement, na.rm = TRUE),
    max_area_mm2 = max(area_mm2.measurement, na.rm = TRUE)
  )

size_by_family_20260427 <- size_metadata_20260427 %>%
  group_by(Family) %>%
  summarise(
    individuals = n(),
    individuals_with_area = sum(!is.na(area_mm2.measurement)),
    mean_area_mm2 = mean(area_mm2.measurement, na.rm = TRUE),
    median_area_mm2 = median(area_mm2.measurement, na.rm = TRUE),
    min_area_mm2 = min(area_mm2.measurement, na.rm = TRUE),
    max_area_mm2 = max(area_mm2.measurement, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(readr::parse_number(Family), Family)

readr::write_csv(size_overall_20260427, file.path(out_dir, "size_summary_20260427.csv"))
readr::write_csv(size_by_family_20260427, file.path(out_dir, "size_by_family_20260427.csv"))

size_overall_20260427 %>%
  mutate(across(ends_with("_mm2"), \(x) round(x, 1))) %>%
  knitr::kable()
```

```{r size-by-family-20260427}
size_by_family_20260427 %>%
  mutate(across(ends_with("_mm2"), \(x) round(x, 1))) %>%
  knitr::kable()
```

Only 20260427 has explicit oyster size measurements in the files currently in
this directory. Across that experiment, all 128 individuals have ImageJ area
measurements, with a mean area of 779.1 mm2 and median area of 741.6 mm2. Family
means range from 674.6 mm2 in family 5 to 878.1 mm2 in family 10. Because the
other experiments lack comparable size columns, these values should not be used
to infer size differences across experiment dates.

## Speculative identity of 9 and 9B

Because `9` and `9B` in 20260511 and 20260609 represent two unresolved groups,
one truly family 2 and one truly family 9, the analysis below compares their
survival summaries with experiments where families 2 and 9 were labelled
directly. This is a hypothesis-generating comparison only; it should not be used
to recode the data.

```{r speculative-identity}
two_nine_summary <- family_summary %>%
  filter(Family %in% c("2", "9", "9B")) %>%
  mutate(
    label_status = case_when(
      Family %in% c("9", "9B") &
        as.character(Experiment) %in% c("20260511 - 33 C", "20260609 - 36 C") ~
        "ambiguous 2/9 group",
      TRUE ~ "known family label"
    )
  ) %>%
  arrange(temperature_C, Experiment, Family)

two_nine_summary %>%
  select(
    Experiment,
    temperature_C,
    Family,
    label_status,
    individuals,
    deaths,
    death_percent,
    median_survival_hr
  ) %>%
  mutate(across(c(death_percent, median_survival_hr), \(x) round(x, 1))) %>%
  knitr::kable()
```

Known-label experiments provide limited guidance. In the 20260427 33 C
experiment, known families 2 and 9 were similar, with family 2 at 37.5% deaths
and family 9 at 33.3% deaths, and neither reached median survival. In the
20260604 36 C experiment, known family 2 appeared more susceptible than known
family 9: family 2 had 100% deaths and median survival of 17 h, while family 9
had 80% deaths and median survival of 42 h.

Using that 36 C pattern as the strongest available clue, the 20260609 ambiguous
groups weakly suggest that `9B` could be family 2 and `9` could be family 9,
because `9B` had slightly higher mortality than `9` (72.7% vs. 63.6%). This is
weak evidence because both groups had the same median survival estimate (48 h)
and the experiment differs from 20260604 in design and timing. For 20260511,
there is no useful survival-based distinction: both ambiguous groups had 100%
deaths and median survival of 43 h, so the identities cannot be inferred from
these survivorship data.

## Survival curves

```{r survival-by-experiment, warning=FALSE}
plot_by_experiment <- km_curve_data(survival_data, "Experiment") %>%
  ggplot(aes(x = time, y = survival, color = Experiment)) +
  geom_step(linewidth = 0.9) +
  labs(x = "Time (hours)", y = "Survival probability", color = "Experiment") +
  ylim(0, 1) +
  theme_classic()

plot_by_experiment
```

```{r survival-by-temperature, warning=FALSE}
plot_by_temperature <- km_curve_data(survival_data, "temperature_C") %>%
  ggplot(aes(x = time, y = survival, color = temperature_C)) +
  geom_step(linewidth = 0.9) +
  labs(x = "Time (hours)", y = "Survival probability", color = "Temperature (C)") +
  ylim(0, 1) +
  theme_classic()

plot_by_temperature
```

```{r survival-family-facets, fig.width=9, fig.height=7, warning=FALSE}
plot_family_facets <- km_curve_data(survival_data, c("Experiment", "Family")) %>%
  ggplot(aes(x = time, y = survival, color = Family)) +
  geom_step(linewidth = 0.7) +
  labs(x = "Time (hours)", y = "Survival probability", color = "Family/group") +
  ylim(0, 1) +
  facet_wrap(~Experiment, scales = "free_x") +
  theme_classic() +
  theme(legend.position = "right")

plot_family_facets
```

```{r save-plots}
ggsave(file.path(out_dir, "survival_by_experiment.png"),
       plot_by_experiment, width = 7, height = 5, dpi = 300)
ggsave(file.path(out_dir, "survival_by_temperature.png"),
       plot_by_temperature, width = 7, height = 5, dpi = 300)
ggsave(file.path(out_dir, "survival_by_family_faceted_by_experiment.png"),
       plot_family_facets, width = 9, height = 7, dpi = 300)
```

## Statistical comparisons

```{r log-rank-tests}
logrank_experiment <- survdiff(Surv(time, status) ~ Experiment, data = survival_data)
logrank_temperature <- survdiff(Surv(time, status) ~ temperature_C, data = survival_data)

logrank_by_experiment <- survival_data %>%
  group_by(Experiment) %>%
  group_modify(\(.x, .y) {
    fit <- survdiff(Surv(time, status) ~ Family, data = .x)
    tibble(
      chisq = unname(fit$chisq),
      df = length(fit$n) - 1,
      p_value = 1 - pchisq(fit$chisq, df = length(fit$n) - 1)
    )
  }) %>%
  ungroup()

comparison_summary <- tribble(
  ~comparison, ~chisq, ~df, ~p_value,
  "Across experiments", unname(logrank_experiment$chisq),
    length(logrank_experiment$n) - 1,
    1 - pchisq(logrank_experiment$chisq, df = length(logrank_experiment$n) - 1),
  "Across temperatures", unname(logrank_temperature$chisq),
    length(logrank_temperature$n) - 1,
    1 - pchisq(logrank_temperature$chisq, df = length(logrank_temperature$n) - 1)
)

comparison_summary %>%
  bind_rows(
    logrank_by_experiment %>%
      transmute(
        comparison = paste("Across observed family/groups within", Experiment),
        chisq,
        df,
        p_value
      )
  ) %>%
  mutate(p_value = signif(p_value, 4)) %>%
  mutate(chisq = round(chisq, 3)) %>%
  knitr::kable()
```

```{r cox-model}
cox_experiment <- coxph(Surv(time, status) ~ Experiment, data = survival_data)
cox_temperature <- coxph(Surv(time, status) ~ temperature_C, data = survival_data)

summary(cox_experiment)
summary(cox_temperature)
```

## Take homes

```{r take-homes}
ranked_experiments <- overall_summary %>%
  arrange(desc(death_percent), median_survival_hr)

most_severe <- ranked_experiments %>% slice(1)
least_severe <- ranked_experiments %>% slice_tail(n = 1)

cat("Experiments summarized:", n_distinct(survival_data$Experiment), "\n")
cat("Total individuals:", nrow(survival_data), "\n")
cat("Most severe by death percent:", as.character(most_severe$Experiment),
    sprintf("(%.1f%% deaths)\n", most_severe$death_percent))
cat("Least severe by death percent:", as.character(least_severe$Experiment),
    sprintf("(%.1f%% deaths)\n", least_severe$death_percent))
cat("Cross-experiment log-rank p-value:",
    signif(comparison_summary$p_value[comparison_summary$comparison == "Across experiments"], 4),
    "\n")
cat("Temperature log-rank p-value:",
    signif(comparison_summary$p_value[comparison_summary$comparison == "Across temperatures"], 4),
    "\n")
```
