---
title: "Kaplan-Meier Survivorship Analysis — 20260706 (35 °C)"
author: "Steven Roberts"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
editor: visual
---

BACKGROUND - Kaplan-Meier survivorship analysis of the 20260706 35 °C
*Magallana gigas* USDA-family heat-stress experiment, run through the reusable
Kaplan-Meier pipeline (`code/01-mgig-heat-survivorship-20260604.qmd`), so results
are directly comparable to the 20260427 / 0511 / 0604 / 0609 dates.

Oysters (9 USDA families) were placed in a 35 °C tank at 3 pm on 2026-07-06;
mortality was scored at 20, 43, 67, and 92 hours post-immersion (see
`data/20260706-35C-USDA-families/README.md`).

The downstream KM / log-rank / Cox sections are kept IDENTICAL to the template so
results are directly comparable across experiments. ONLY the data read + reshape
is changed, because this date uses a different input layout than the earlier
long-format files:

```{=html}
<!--
================================================================================
INPUT FORMAT FOR THIS DATE (WIDE, mortality-coded)
================================================================================
survivorship.csv is WIDE: one row per individual, with columns

    ID, Family, 20, 43, 67, 92

where 20/43/67/92 are the timepoint hours and each cell is a MORTALITY flag:

    1 = dead   -> status 1 (event)
    0 = alive  -> status 0

(NB: this is the OPPOSITE polarity of the long-format `alive.measurement`
column used by the 20260427/0511/0609 files, where TRUE = alive.)

The flags are monotone non-decreasing across time (once dead, stays dead), so an
individual's death TIME is the first timepoint at which the flag == 1; an
individual that is never flagged dead is right-censored at the last timepoint.
================================================================================
-->
```

## Set up

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(survival)
library(ggsurvfit)
library(gtsummary)
library(cardx)
library(cowplot)
```

```{r config}
# >>> EDIT: analysis configuration ---------------------------------------------

# Path to the WIDE-format mortality csv for this experiment
data_path  <- "../data/20260706-35C-USDA-families/survivorship.csv"

# Output directory for saved figures
out_dir    <- "../outputs/06-mgig-heat-survivorship-20260706/"

# Grouping / factor column(s) to compare survival across.
group_vars <- c("Family")

# Individual ID column produced in the reshape below.
id_var     <- "uid"
# ------------------------------------------------------------------------------
```

## Read in data

```{r read-data, message=FALSE}
raw <- readr::read_csv(data_path, show_col_types = FALSE)
glimpse(raw)
```

## Reshape to one (time, status) row per individual

```{r reshape}
# Timepoint columns are the numeric-named columns (hours). Everything else
# (ID, Family) is metadata.
time_cols <- names(raw)[grepl("^[0-9]+$", names(raw))]
stopifnot(length(time_cols) > 0)

long <- raw %>%
  # keep only rows that actually have a family assignment (drop blank/footer rows)
  filter(!is.na(Family), trimws(as.character(Family)) != "") %>%
  # one wide row == one individual; give each a guaranteed-unique key so that a
  # missing ID does not silently drop a real (e.g. all-survived) oyster.
  mutate(
    row_uid = paste0("row", row_number()),
    uid     = ifelse(is.na(ID) | trimws(as.character(ID)) == "",
                     row_uid, as.character(ID)),
    Family  = as.character(Family)
  ) %>%
  pivot_longer(
    cols      = all_of(time_cols),
    names_to  = "timepoint_hrs",
    values_to = "dead_raw"
  ) %>%
  mutate(
    timepoint_hrs = as.numeric(timepoint_hrs),
    # 1 = dead (event), 0 = alive; be tolerant of stray whitespace / text
    dead = case_when(
      toupper(trimws(as.character(dead_raw))) %in% c("1","TRUE","T","DEAD","YES","Y") ~ TRUE,
      toupper(trimws(as.character(dead_raw))) %in% c("0","FALSE","F","ALIVE","NO","N") ~ FALSE,
      TRUE ~ NA
    )
  )

# Collapse to ONE (time, status) row per individual:
#   - if it ever died (dead == TRUE), keep the FIRST death time (event)
#   - otherwise keep the LAST observed time point (right-censored)
data <- long %>%
  filter(!is.na(uid), !is.na(timepoint_hrs), !is.na(dead)) %>%
  group_by(uid) %>%
  summarise(
    Family = first(Family[!is.na(Family)]),
    status = as.integer(any(dead)),
    time   = if (any(dead)) min(timepoint_hrs[dead]) else max(timepoint_hrs),
    .groups = "drop"
  ) %>%
  mutate(
    Family = factor(Family, levels = unique(Family[order(readr::parse_number(Family), Family)]))
  )

cat("n individuals:", nrow(data), " | events (deaths):", sum(data$status), "\n")
cat("families:", paste(levels(data$Family), collapse = ", "), "\n")
head(data)
```

## Generate Kaplan-Meier survival curves

```{r km-fit}
km_formula <- as.formula(
  paste("Surv(time, status) ~", paste(group_vars, collapse = " + "))
)

s1 <- survival::survfit(km_formula, data = data)
print(s1)
summary(s1)
```

```{r km-plot-combined, warning=FALSE}
ggsurvfit::survfit2(km_formula, data = data) %>%
  ggsurvfit::ggsurvfit() +
  labs(x = "Time (hours)", y = "Survival probability") +
  ylim(0, 1)
```

## Global log-rank test (family comparison)

```{r log-rank}
logrank <- survival::survdiff(km_formula, data = data)
print(logrank)
p_val <- 1 - pchisq(logrank$chisq, df = length(logrank$n) - 1)
cat("Log-rank chi-square:", round(logrank$chisq, 4),
    "| df:", length(logrank$n) - 1,
    "| p-value:", signif(p_val, 4), "\n")
```

## Cox Proportional Hazards model

```{r cox-model}
cox_formula <- as.formula(
  paste("Surv(time, status) ~", paste(group_vars, collapse = " * "))
)
cox_fit <- survival::coxph(cox_formula, data = data)
cox_fit
cox_fit %>% tbl_regression(exp = TRUE)
```

## Publication plot

```{r theme}
my_theme <- theme_classic() +
  theme(
    axis.title.y = element_text(size = 14, color = "black"),
    axis.text.y  = element_text(size = 12, color = "black"),
    axis.text.x  = element_text(size = 12, color = "black"),
    axis.title.x = element_text(size = 14, color = "black")
  )
```

```{r plot-by-family, warning=FALSE}
g1 <- group_vars[1]
f1 <- as.formula(paste("Surv(time, status) ~", g1))
n_groups1     <- nlevels(factor(data[[g1]]))
group_colors1 <- scales::hue_pal()(n_groups1)

plot1 <- ggsurvfit::survfit2(f1, data = data) %>%
  ggsurvfit::ggsurvfit() +
  labs(x = "Time (hours)", y = "Survival probability", title = "") +
  ylim(0, 1) +
  my_theme +
  scale_colour_manual(values = group_colors1) +
  theme(legend.position = "right")

plot1
```

```{r save-plots}
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
ggsave(file.path(out_dir, "survival_by_family.png"), plot1, width = 7, height = 5, dpi = 300)
```

## Take homes

```{r take-homes}
med <- summary(survival::survfit(Surv(time, status) ~ 1, data = data))$table["median"]
cat("Experiment: 20260706, 35 C\n")
cat("Individuals:", nrow(data), "| Deaths:", sum(data$status),
    sprintf("(%.0f%%)\n", 100 * mean(data$status)))
cat("Overall median survival (hrs):", ifelse(is.na(med), "not reached (>50% survived)", med), "\n")
cat("Family log-rank p-value:", signif(p_val, 4),
    ifelse(p_val < 0.05, "-> families differ significantly\n",
                         "-> no significant family difference\n"))
```
