---
title: "Cross-Experiment Survivorship Summary & Family Ranking"
author: "Steven Roberts"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    df-print: paged
editor: visual
---

BACKGROUND - This document pools EVERY *Magallana gigas* USDA-family
survivorship experiment run to date into a single analysis, harmonising the two
different raw data layouts used across dates, and then ranks families by as many
survival phenotypes as the data support.

Experiments included (see each `data/<date>-.../README.md` for full protocols):

| exp_id | date | stressor | raw format |
|---|---|---|---|
| 20260427-33C | 2026-04-27 | 33 °C | long (`alive.measurement`) |
| 20260511-33C | 2026-05-11 | 33 °C | long (`alive.measurement`) |
| 20260604-36C | 2026-06-04 | 36 °C | wide (mortality 0/1) |
| 20260609-36C | 2026-06-09 | 36 °C | long (`alive.measurement`) |
| 20260706-35C | 2026-07-06 | 35 °C | wide (mortality 0/1) |
| 20260706-35C09ppt | 2026-07-06 | 35 °C + 9 ppt | wide (mortality 0/1) |
| 20260707-FW | 2026-07-07 | freshwater → 36 °C | wide (mortality 0/1) |

The 20260714-H2O2FW experiment has a protocol README but no scored
`survivorship.csv` yet, so it is not included.

**Note on the 9 / 9b label ambiguity:** the 20260511 and 20260609 experiments
each contain both a `9` and a `9b` family and are missing family `2`. One of the
two 9-labels in those experiments is really family 2 and the other is really
family 9, but the correct assignment was lost. Because any guess would be a 50/50
that corrupts *both* family 2 and family 9, all `9`/`9b` individuals are
**excluded** from those two experiments (family 9 is retained wherever it appears
unambiguously; family `9b` drops out entirely).

```{=html}
<!--
================================================================================
HOW FAMILIES ARE COMPARED ACROSS EXPERIMENTS
================================================================================
Experiments differ in severity (33 vs 35 vs 36 C, salinity, duration), so raw
survival is NOT directly comparable across dates. To compare families fairly we
score each family WITHIN each experiment and then aggregate:

  surv_prop      = fraction of the family alive at the LAST timepoint
  rel_surv_time  = mean restricted lifetime / experiment duration (0-1)
                   (= RMST normalised to the observation window; because every
                    individual is observed at every timepoint until it dies or
                    the experiment ends, this equals the restricted mean
                    survival time exactly.)

Both are 0-1, higher = better. Within each experiment each family's two metrics
are converted to a PERCENTILE (0-100) relative to the other families in that same
experiment, which removes the experiment-severity effect. A family's composite
score is the mean of those percentiles across all experiments it appears in.
================================================================================
-->
```

## Set up

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(survival)
library(ggsurvfit)
library(knitr)

out_dir <- "../outputs/09-mgig-survivorship-cross-experiment-summary/"
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
```

## Experiment registry

```{r registry}
experiments <- tibble::tribble(
  ~exp_id,               ~date,        ~stressor,             ~format, ~path,
  "20260427-33C",        "2026-04-27", "33 C",                "long",  "../data/20260427-33C-USDA-families/survivorship.csv",
  "20260511-33C",        "2026-05-11", "33 C",                "long",  "../data/20260511-33C-USDA-families/survivorship.csv",
  "20260604-36C",        "2026-06-04", "36 C",                "wide",  "../data/20260604-36C-USDA-families/survivorship.csv",
  "20260609-36C",        "2026-06-09", "36 C",                "long",  "../data/20260609-36C-USDA-families/survivorship.csv",
  "20260706-35C",        "2026-07-06", "35 C",                "wide",  "../data/20260706-35C-USDA-families/survivorship.csv",
  "20260706-35C09ppt",   "2026-07-06", "35 C / 9 ppt",        "wide",  "../data/20260706-35C09ppt-USDA-families/survivorship.csv",
  "20260707-FW",         "2026-07-07", "Freshwater -> 36 C",  "wide",  "../data/20260707-FW-USDA-families/survivorship.csv"
)
knitr::kable(experiments |> dplyr::select(-path))
```

## Harmonise all experiments to one (family, individual, time, status) table

```{r readers}
# Normalise family labels: strip a leading "f" and zero-padding so "f01","01","1"
# all become "1"; non-numeric labels such as "9b" are kept verbatim (lower-cased).
norm_family <- function(x) {
  x <- tolower(trimws(as.character(x)))
  x <- sub("^f", "", x)          # f01 -> 01, f2 -> 2, 9b unchanged
  x <- ifelse(grepl("^[0-9]+$", x), as.character(as.integer(x)), x)  # 01 -> 1
  x
}

# ---- LONG format: alive.measurement TRUE = alive, one row per indiv x timepoint
read_long <- function(path) {
  df <- readr::read_csv(path, show_col_types = FALSE, name_repair = "minimal")
  names(df) <- trimws(names(df))
  if ("is_blank" %in% names(df)) {
    df <- df %>% filter(is.na(is_blank) | tolower(as.character(is_blank)) != "true")
  }
  df %>%
    transmute(
      uid        = as.character(individual_id),
      family_raw = as.character(familly_id.group),
      time       = as.numeric(timepoint_hrs),
      dead       = !as.logical(alive.measurement)
    ) %>%
    filter(!is.na(uid), !is.na(time), !is.na(dead))
}

# ---- WIDE format: numeric-named timepoint columns, cell 1 = dead
read_wide <- function(path) {
  df <- readr::read_csv(path, show_col_types = FALSE, name_repair = "minimal")
  names(df) <- trimws(gsub("﻿", "", names(df)))          # drop BOM/whitespace
  time_cols <- names(df)[grepl("^[0-9]+$", names(df))]
  fam_col   <- names(df)[grepl("^family$", names(df), ignore.case = TRUE)][1]
  df %>%
    mutate(uid = paste0("row", row_number())) %>%             # one row = one indiv
    rename(family_raw = !!fam_col) %>%
    filter(!is.na(family_raw), trimws(as.character(family_raw)) != "") %>%
    pivot_longer(all_of(time_cols), names_to = "time", values_to = "dead_raw") %>%
    transmute(
      uid        = as.character(uid),
      family_raw = as.character(family_raw),
      time = as.numeric(time),
      dead = toupper(trimws(as.character(dead_raw))) %in% c("1","TRUE","T","DEAD","YES","Y")
    ) %>%
    filter(!is.na(time))
}

read_exp <- function(fmt, path) if (fmt == "long") read_long(path) else read_wide(path)
```

```{r build-individual-table}
# Collapse each individual to ONE (time, status) row within its experiment:
#   status = 1 if ever dead (event); time = first death time, else last timepoint.
indiv0 <- experiments %>%
  mutate(raw = purrr::map2(format, path, read_exp)) %>%
  select(exp_id, date, stressor, raw) %>%
  unnest(raw) %>%
  mutate(family = norm_family(family_raw)) %>%
  group_by(exp_id, date, stressor, family, uid) %>%
  summarise(
    status = as.integer(any(dead, na.rm = TRUE)),
    time   = if (any(dead, na.rm = TRUE)) min(time[dead], na.rm = TRUE) else max(time, na.rm = TRUE),
    .groups = "drop"
  )

# ---- Resolve the 9 / 9b mislabelling -----------------------------------------
# In experiments where BOTH "9" and "9b" were run, family 2 is ABSENT: one of the
# two 9-labels is really family 2 and the other is really family 9, but which is
# which was lost at labelling. Rather than guess (a 50/50 that would corrupt BOTH
# family 2 and family 9), we EXCLUDE every "9" and "9b" individual from those
# experiments. In experiments where only "9" appears it is unambiguous family 9
# and is kept; family "9b" therefore drops out of the analysis entirely.
amb_exps <- indiv0 %>%
  group_by(exp_id) %>%
  summarise(both = all(c("9", "9b") %in% family), .groups = "drop") %>%
  filter(both) %>% pull(exp_id)

n_dropped <- indiv0 %>%
  filter(exp_id %in% amb_exps, family %in% c("9", "9b")) %>% nrow()

indiv <- indiv0 %>%
  filter(!(exp_id %in% amb_exps & family %in% c("9", "9b"))) %>%
  group_by(exp_id) %>%
  mutate(tau = max(time)) %>%          # experiment observation window (hrs)
  ungroup()

cat("Experiments with unresolved 9/9b labels (9 & 9b excluded there):",
    ifelse(length(amb_exps) == 0, "none", paste(amb_exps, collapse = ", ")), "\n")
cat("Individuals dropped due to 9/9b ambiguity:", n_dropped, "\n")
cat("Total individuals pooled:", nrow(indiv),
    "| experiments:", dplyr::n_distinct(indiv$exp_id),
    "| families:", dplyr::n_distinct(indiv$family), "\n")
cat("Family labels seen:", paste(sort(unique(indiv$family)), collapse = ", "), "\n")
```

## Per-experiment performance summary

```{r exp-summary}
exp_summary <- indiv %>%
  group_by(exp_id, date, stressor) %>%
  group_modify(function(d, key) {
    fit_all <- survival::survfit(Surv(time, status) ~ 1, data = d)
    med <- summary(fit_all)$table["median"]
    # family log-rank (only if >1 family with events possible)
    p_lr <- NA_real_
    if (dplyr::n_distinct(d$family) > 1) {
      lr <- survival::survdiff(Surv(time, status) ~ family, data = d)
      p_lr <- 1 - pchisq(lr$chisq, df = length(lr$n) - 1)
    }
    tibble(
      n            = nrow(d),
      n_families   = dplyr::n_distinct(d$family),
      deaths       = sum(d$status),
      mortality_pct = round(100 * mean(d$status), 1),
      tau_hrs      = max(d$time),
      median_surv_hrs = ifelse(is.na(med), NA, unname(med)),
      logrank_p    = signif(p_lr, 3)
    )
  }) %>%
  ungroup() %>%
  arrange(date)

knitr::kable(exp_summary, caption = "Per-experiment survivorship summary")
readr::write_csv(exp_summary, file.path(out_dir, "experiment_summary.csv"))
```

```{r km-overall-by-exp, warning=FALSE, fig.width=8, fig.height=5}
# Overall KM curve per experiment on a common (normalised) time axis so severity
# is directly comparable regardless of experiment duration.
indiv_norm <- indiv %>% mutate(time_frac = time / tau)

km_overall <- survival::survfit(Surv(time_frac, status) ~ exp_id, data = indiv_norm)
p_overall <- ggsurvfit::survfit2(Surv(time_frac, status) ~ stressor, data = indiv_norm) %>%
  ggsurvfit::ggsurvfit(linewidth = 1) +
  labs(x = "Fraction of experiment duration", y = "Survival probability",
       title = "Overall survival by experiment (time normalised to each experiment's duration)") +
  ylim(0, 1) +
  theme_classic() +
  theme(legend.position = "right")
p_overall
ggsave(file.path(out_dir, "km_overall_by_experiment.png"), p_overall, width = 8, height = 5, dpi = 300)
```

## Family phenotypes within each experiment

```{r family-metrics}
# Two phenotypes per family x experiment (both 0-1, higher = better):
#   surv_prop     = fraction alive at end
#   rel_surv_time = mean restricted lifetime / tau  (= normalised RMST)
fam_exp <- indiv %>%
  group_by(exp_id, date, stressor, family, tau) %>%
  summarise(
    n             = dplyr::n(),
    deaths        = sum(status),
    surv_prop     = mean(status == 0),
    rel_surv_time = mean(ifelse(status == 1, time, tau)) / dplyr::first(tau),
    .groups = "drop"
  ) %>%
  # within-experiment percentile (0-100) for each phenotype; ties share mean rank
  group_by(exp_id) %>%
  mutate(
    pct_surv = 100 * (rank(surv_prop,     ties.method = "average") - 0.5) / dplyr::n(),
    pct_rst  = 100 * (rank(rel_surv_time, ties.method = "average") - 0.5) / dplyr::n()
  ) %>%
  ungroup()

readr::write_csv(fam_exp, file.path(out_dir, "family_by_experiment.csv"))
```

```{r family-surv-heatmap, fig.width=9, fig.height=5}
hm <- fam_exp %>%
  mutate(
    family = factor(family, levels = unique(family[order(readr::parse_number(family), family)])),
    exp_lab = paste0(stressor, "\n", date)
  )
p_hm <- ggplot(hm, aes(x = exp_lab, y = family, fill = surv_prop)) +
  geom_tile(color = "white") +
  geom_text(aes(label = sprintf("%.0f%%", 100 * surv_prop)), size = 3) +
  scale_fill_viridis_c(name = "Survival", limits = c(0, 1), labels = scales::percent) +
  labs(x = NULL, y = "Family",
       title = "Final survival proportion by family and experiment") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 8))
p_hm
ggsave(file.path(out_dir, "family_survival_heatmap.png"), p_hm, width = 9, height = 5, dpi = 300)
```

## Family ranking across all experiments

```{r family-ranking}
family_rank <- fam_exp %>%
  group_by(family) %>%
  summarise(
    n_experiments   = dplyr::n(),
    total_n         = sum(n),
    total_deaths    = sum(deaths),
    overall_mort_pct = round(100 * sum(deaths) / sum(n), 1),
    mean_surv_prop  = round(mean(surv_prop), 3),
    mean_rel_surv   = round(mean(rel_surv_time), 3),
    composite_score = round(mean((pct_surv + pct_rst) / 2), 1),  # 0-100, higher = better
    .groups = "drop"
  ) %>%
  arrange(desc(composite_score)) %>%
  mutate(rank = dplyr::row_number()) %>%
  select(rank, family, n_experiments, total_n, overall_mort_pct,
         mean_surv_prop, mean_rel_surv, composite_score)

knitr::kable(family_rank,
             caption = "Family ranking (composite = mean within-experiment percentile across survival phenotypes; higher = hardier)")
readr::write_csv(family_rank, file.path(out_dir, "family_ranking.csv"))
```

```{r family-rank-plot, fig.width=8, fig.height=5}
p_rank <- family_rank %>%
  mutate(family = forcats::fct_reorder(family, composite_score)) %>%
  ggplot(aes(x = composite_score, y = family, fill = composite_score)) +
  geom_col() +
  geom_text(aes(label = paste0("n=", n_experiments, " exp")), hjust = -0.1, size = 3) +
  scale_fill_viridis_c(guide = "none") +
  xlim(0, 105) +
  labs(x = "Composite survival score (0-100, higher = hardier)", y = "Family",
       title = "Family survival ranking across all experiments") +
  theme_classic()
p_rank
ggsave(file.path(out_dir, "family_ranking.png"), p_rank, width = 8, height = 5, dpi = 300)
```

## Consistency check: within-experiment survival percentile by family

```{r consistency, fig.width=9, fig.height=5}
p_cons <- fam_exp %>%
  mutate(family = factor(family, levels = family_rank$family)) %>%
  ggplot(aes(x = family, y = pct_surv)) +
  geom_boxplot(outlier.shape = NA, fill = "grey90") +
  geom_jitter(aes(color = stressor), width = 0.15, height = 0, size = 2) +
  labs(x = "Family (ordered by composite rank)", y = "Within-experiment survival percentile",
       title = "Ranking consistency across experiments",
       color = "Experiment") +
  theme_classic()
p_cons
ggsave(file.path(out_dir, "family_consistency.png"), p_cons, width = 9, height = 5, dpi = 300)
```

## Take homes

```{r take-homes}
best  <- family_rank %>% slice_head(n = 3)
worst <- family_rank %>% slice_tail(n = 3)

cat("Experiments pooled:", nrow(exp_summary), "\n")
cat("Total individuals:", sum(exp_summary$n),
    "| overall deaths:", sum(exp_summary$deaths), "\n\n")

cat("Most heat/stress tolerant families (top composite score):\n")
for (i in seq_len(nrow(best))) {
  cat(sprintf("  %d. Family %s  (score %.1f, mean survival %.0f%%, %d experiments)\n",
              best$rank[i], best$family[i], best$composite_score[i],
              100 * best$mean_surv_prop[i], best$n_experiments[i]))
}
cat("\nLeast tolerant families (bottom composite score):\n")
for (i in seq_len(nrow(worst))) {
  cat(sprintf("  %d. Family %s  (score %.1f, mean survival %.0f%%, %d experiments)\n",
              worst$rank[i], worst$family[i], worst$composite_score[i],
              100 * worst$mean_surv_prop[i], worst$n_experiments[i]))
}
```

<!--
Interpretation notes:
- composite_score ranks families on within-experiment percentile so it is not
  distorted by how mild/severe any single experiment was.
- Families appearing in fewer experiments (n_experiments column) have less robust
  scores; weight accordingly.
- The 9/9b label ambiguity on 0511 & 0609 is handled by exclusion (see note at
  top), so families 2 and 9 are scored only from experiments where their identity
  is certain, and "9b" does not appear in the ranking.
-->
