---
title: "Gen5-20260824-mgig-sormi-citrate_synthase-F07-ambient_and_36C"
author: "Sam White"
date: "2026-08-24"
output:
  github_document:
    toc: true
    number_sections: true
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
---

# BACKGROUND

Citrate synthase (CS) activity in ctenidia of sixteen *Magallana gigas*
(Pacific oyster) individuals from **family F07** -- eight held at
**ambient** temperature and eight at **36 &deg;C** -- assayed together on a
single plate 2026-08-24 with the
[Abcam Citrate Synthase Assay Kit (ab239712), v4a](https://github.com/RobertsLab/resources/blob/master/protocols/Commercial_Protocols/ABCAM-Citrate-Synthase-Assay-v4a-ab239712.pdf).
Both temperature exposures for `F07` are combined on this one plate.
**This plate also has a confirmed reagent mix-up affecting its background
control (note 4) -- background-corrected rates on this plate equal the raw
rates.**

The kit is a **coupled kinetic** assay. CS condenses acetyl-CoA and
oxaloacetate, releasing free CoA-SH; the liberated thiol reduces DTNB to
TNB<sup>2-</sup>, which absorbs at 412 nm. The **rate** of A412 increase is
proportional to CS activity, and absolute nmol of thiol are assigned from a
**GSH (reduced glutathione) standard curve** read on the same plate.

Absorbance was read at 412 nm in kinetic mode, 25 &deg;C, every 2 min for 20
min (11 reads) on a Synergy HTX. See note 1.

Total extracted protein per homogenate (used for normalization, note 2) was
not measured on this plate; it comes from two BCA/Bradford-style
protein-quantification assays run 2026-08-13:
[`Gen5-20260813-mgig-sormi-BSA-F07-protein.Rmd`](https://github.com/RobertsLab/sormi-assay-development/blob/main/Citrate_synthase/code/Gen5-20260813-mgig-sormi-BSA-F07-protein.Rmd)
(12 of the 16 samples) and a follow-up re-assay plate
(the remaining 4 -- `F07_05_36C`, `F07_01_ambient`, `F07_04_ambient`,
`F07_06_ambient` -- which were re-assayed at dilution after failing QC on
the first pass). Homogenization used 350 &micro;L of Assay Buffer 7 per
sample (reported by the bench operator; see the lab notebook homogenization
post for this assay).

This document is fully self-contained: layout parsing, kinetic QC,
standard-curve fitting, rate extraction, CV computation, protein
normalization, and results assembly are all defined directly below (see
[`## Pipeline functions`](#pipeline-functions) under SETUP) rather than sourced from a separate
script, so the whole analysis can be reproduced from this one file.

## Sample naming convention

Well labels follow `<family>_<individual>_<temperature>-<assay_type>-<weight>-df.<n>`
for sample wells (e.g. `F07_01_ambient-citrate_synthase-7.3-df.0`,
`F07_01_36C-citrate_synthase-9.0-df.0`), `STD-<assay_type>-<nmol_per_well>`
for GSH standards, `POS-<assay_type>` for the positive control, and
`BG-<assay_type>` for the pooled background control (see note 3) -- the
background label carries no weight or dilution suffix, since it is not tied
to any single individual's tissue. Because both temperature exposures are
present on this one plate, `<temperature>` genuinely varies (`ambient` or
`36C`) within this document; nothing in the parsing below hardcodes a
temperature.
The trailing number in the sample label (e.g. `7.3`) is tissue weight in mg,
which is **not** used for normalization here (note 2); it is parsed and
validated but plays no role in the activity calculation.

## Notes

1. **Read duration for this plate is 20 minutes** (11 reads at 2-min
   intervals). The assay parameters below set `read_duration_min = 20` to
   match. Rate extraction uses a fixed-width sliding window
   (`rate_window_n = 5` points = 8 min) rather than the full trace; see the
   KINETIC TRACES section for whether the assay is still in its linear phase
   over this window.
2. **Activity is normalized to total extracted protein, not tissue weight**,
   because protein concentration was measured directly (BCA-style assay),
   and controls for extraction efficiency in a way tissue weight cannot. No
   `Sample` string in either protein-concentration source needed correction
   for this plate's samples.
3. **Background correction uses a single pooled background control**,
   rather than a paired per-sample background for each individual. A pool
   was made from 5 &micro;L of each sample homogenate, and
   2 &micro;L of that pool (matching `sample_volume_uL` used for every
   reaction well) was run in triplicate with Background Control Mix (no CS
   Substrate Mix) to produce wells `BG-citrate_synthase` (F10-F12). Because
   there is only one background estimate for the whole plate, the same
   background-corrected rate would normally be subtracted from every sample
   -- but see note 4.
4. **Reaction Mix (containing CS Substrate) appears to have been added to
   all three pooled-background wells on this plate instead of Background
   Control Mix.** All three `BG-citrate_synthase` replicates rise
   substantially over the run (net change 0.067-0.119 A412, vs. <0.05 for a
   flat background well) rather than staying flat, and this has been
   confirmed as a pipetting mix-up rather than genuine background signal.
   Because 0 of 3 replicates are flat, `compute_background_correction()`
   cannot estimate a background rate (`bg_rate_flat` is `NA`) and, per its
   existing fallback, the background correction used for every sample on
   this plate is 0 -- i.e., **corrected rate = raw rate** throughout. See
   BACKGROUND CORRECTION and the SUMMARY for what this means for this
   plate's results.

```{r setup, include=TRUE}
library(knitr)
library(ggplot2)
library(dplyr)
library(tidyr)

knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,         # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = "",        # Prevents appending '##' to beginning of lines in code output
  results = 'hold'     # Holds output so it's all printed together after code chunk
)

```

## Pipeline functions

All layout parsing, kinetic QC, standard-curve fitting, rate extraction,
CV computation, protein normalization, and results assembly are defined
directly in this document (rather than sourced from a separate script) so
that it is fully self-contained and reproducible from this single file.
The functions are grouped below by pipeline stage; each is called in the
DATA / KINETIC TRACES / ANOMALY DETECTION / GSH STANDARD CURVE / RATE
EXTRACTION / BACKGROUND CORRECTION / TECHNICAL REPLICATE PRECISION /
CITRATE SYNTHASE ACTIVITY / QC SUMMARY sections that follow. Nothing
plate-specific (typo fixes, disqualified samples, thresholds) is
hardcoded in any function -- it is always passed in as an argument, from
the `assay_params` list defined further below or from a plate-specific
fix list.

### Data import and layout parsing

Reads the raw Gen5 exports (kinetic absorbance CSV, full-report text backup, plate-layout CSV), reconciles them against each other, and attaches well-type/sample-id/replicate metadata plus measured protein concentrations to every well.

```{r pipeline-fns-io}
## -----------------------------------------------------------------------
## elapsed_to_min()
## Purpose : Convert Gen5 "H:MM:SS" elapsed-time strings to decimal minutes.
## Inputs  : x - character vector of "H:MM:SS" strings
## Outputs : numeric vector of elapsed minutes, same length as x
## -----------------------------------------------------------------------
elapsed_to_min <- function(x) {
  vapply(strsplit(x, ":"), function(p) {
    p <- as.numeric(p)
    p[1] * 60 + p[2] + p[3] / 60
  }, numeric(1))
}

## -----------------------------------------------------------------------
## read_absorbance_csv()
## Purpose : Parse a Gen5 "absorbance-*.csv" kinetic export: wavelength
##           header, a "Time," row, a temperature row, then one row per well
##           (well ID in column 1, one OD reading per timepoint thereafter).
## Inputs  : path - path to the absorbance-*.csv file
## Outputs : long-format data.frame, one row per well x timepoint:
##           well (chr), time_min (dbl), od (dbl)
## -----------------------------------------------------------------------
read_absorbance_csv <- function(path) {
  ln <- iconv(readLines(path, warn = FALSE, encoding = "latin1"), "latin1", "UTF-8")
  time_i   <- grep("^Time,", ln)[1]
  time_str <- strsplit(ln[time_i], ",")[[1]]
  time_str <- time_str[!time_str %in% c("", "Time")]
  tmin     <- elapsed_to_min(time_str)

  well_ln <- ln[grepl("^[A-H][0-9]{1,2},", ln)]
  bind_rows(lapply(strsplit(well_ln, ","), function(p) {
    data.frame(well = p[1], time_min = tmin,
               od = as.numeric(p[seq_along(tmin) + 1]),
               stringsAsFactors = FALSE)
  }))
}

## -----------------------------------------------------------------------
## read_full_report()
## Purpose : Parse a Gen5 "full_report-*.txt" export: tab-delimited, wells
##           across columns, timepoints down rows. Used only as an
##           independent cross-check against the absorbance CSV, never as a
##           primary data source.
## Inputs  : path - path to the full_report-*.txt file
## Outputs : long-format data.frame, one row per well x timepoint:
##           well (chr), time_min (dbl), od (dbl)
## -----------------------------------------------------------------------
read_full_report <- function(path) {
  ln  <- iconv(readLines(path, warn = FALSE, encoding = "latin1"), "latin1", "UTF-8")
  hdr_i <- grep("^Time\t.*\tA1\t", ln)[1]
  hdr   <- trimws(strsplit(ln[hdr_i], "\t")[[1]])
  rows  <- ln[grepl("^[0-9]:[0-9]{2}:[0-9]{2}\t", ln)]
  m     <- do.call(rbind, strsplit(rows, "\t"))
  tmin  <- elapsed_to_min(m[, 1])

  well_cols <- which(grepl("^[A-H][0-9]{1,2}$", hdr))
  bind_rows(lapply(well_cols, function(j) {
    data.frame(well = hdr[j], time_min = tmin,
               od = as.numeric(m[, j]), stringsAsFactors = FALSE)
  }))
}

## -----------------------------------------------------------------------
## parse_plate_layout()
## Purpose : Convert a raw Gen5 plate-layout CSV export into one row per
##           occupied well with its descriptive label. Auto-detects two
##           export formats seen in practice:
##             "single"  - one row per plate row (row 1 = column numbers,
##                         column 1 = row letter, last column = "Name" tag);
##                         each cell already holds the full descriptive label.
##             "double"  - each plate row is exported as TWO rows: a
##                         "Well ID" row (SPL codes) immediately followed by
##                         a "Name" row (the descriptive label actually used
##                         downstream).
##           The format is detected from column 1: if a value repeats twice
##           in a row (once for the well-ID row, once for the label row)
##           immediately below a numeric header, it is treated as "double".
## Inputs  : path - path to the layout-*.csv file
## Outputs : data.frame, one row per occupied well:
##           well (chr, e.g. "A1"), plate_row (chr), plate_col (int),
##           label (chr, trimmed descriptive label)
## -----------------------------------------------------------------------
parse_plate_layout <- function(path) {
  plate_layout <- read.csv(path, header = FALSE, stringsAsFactors = FALSE)

  col1 <- trimws(plate_layout$V1)
  # "double" format: row letters appear twice in a row in column 1
  # (once tagging the Well-ID row, once tagging the Name row), e.g.
  # "A", "A", "B", "B", ... Detect by checking whether non-empty,
  # non-header values in col1 are each immediately followed by a repeat.
  data_rows   <- which(col1 != "" & !is.na(col1))[-1]  # drop header row (row 1)
  is_double <- length(data_rows) >= 2 &&
    all(col1[data_rows[c(TRUE, FALSE)]] == col1[data_rows[c(FALSE, TRUE)]][seq_len(floor(length(data_rows) / 2))])

  if (is_double) {
    # Pair each row-letter row with the following label row.
    row_letter_rows <- data_rows[c(TRUE, FALSE)]
    label_rows      <- data_rows[c(FALSE, TRUE)]
    stopifnot(length(row_letter_rows) == length(label_rows))

    out <- expand.grid(pair_idx = seq_along(row_letter_rows),
                        col_idx  = 2:(ncol(plate_layout) - 1)) %>%
      mutate(
        plate_row = col1[row_letter_rows[pair_idx]],
        plate_col = as.integer(plate_layout[1, col_idx]),
        well      = paste0(plate_row, plate_col),
        label     = trimws(mapply(function(i, j) plate_layout[i, j],
                                   label_rows[pair_idx], col_idx))
      ) %>%
      filter(label != "") %>%
      select(well, plate_row, plate_col, label) %>%
      arrange(plate_row, plate_col)
  } else {
    # "single" format: row 1 = column numbers, column 1 = row letters,
    # last column = "Name" tag; each cell already holds the full label.
    out <- expand.grid(row_idx = 2:nrow(plate_layout),
                        col_idx = 2:(ncol(plate_layout) - 1)) %>%
      mutate(
        plate_row = plate_layout$V1[row_idx],
        plate_col = as.integer(plate_layout[1, col_idx]),
        well      = paste0(plate_row, plate_col),
        label     = trimws(mapply(function(i, j) plate_layout[i, j], row_idx, col_idx))
      ) %>%
      filter(label != "") %>%
      select(well, plate_row, plate_col, label) %>%
      arrange(plate_row, plate_col)
  }

  attr(out, "layout_format") <- ifelse(is_double, "double", "single")
  out
}

## -----------------------------------------------------------------------
## reconcile_kinetic_sources()
## Purpose : Cross-check the absorbance CSV (primary data source) against
##           the full_report text export (independent instrument export)
##           and the plate layout (expected well set). Fails loudly
##           (stopifnot) if any layout well is missing from the CSV, or if
##           the two kinetic sources disagree on any shared reading.
## Inputs  : absorbance_csv - long-format data.frame from read_absorbance_csv()
##           full_report    - long-format data.frame from read_full_report()
##           layout_wells   - data.frame from parse_plate_layout()
##           tolerance      - max allowed |CSV - report| absolute difference
##                            (default 1e-9, i.e. exact agreement)
## Outputs : list(plate_readings, overlap_check, summary), where
##           plate_readings = absorbance_csv with a `source` column added,
##           summary        = named list of counts for reporting
## -----------------------------------------------------------------------
reconcile_kinetic_sources <- function(absorbance_csv, full_report, layout_wells,
                                       tolerance = 1e-9) {
  wells_csv    <- unique(absorbance_csv$well)
  wells_report <- unique(full_report$well)

  wells_report_not_csv <- setdiff(wells_report, wells_csv)
  wells_layout_not_csv <- setdiff(layout_wells$well, wells_csv)
  wells_csv_not_layout <- setdiff(wells_csv, layout_wells$well)

  overlap_check <- inner_join(absorbance_csv, full_report,
                              by = c("well", "time_min"), suffix = c("_csv", "_report")) %>%
    mutate(abs_diff = abs(od_csv - od_report))

  max_disagreement <- if (nrow(overlap_check) > 0) max(overlap_check$abs_diff) else NA_real_

  stopifnot(
    length(wells_layout_not_csv) == 0,
    is.na(max_disagreement) || max_disagreement < tolerance
  )

  plate_readings <- absorbance_csv %>% mutate(source = "absorbance_csv")

  list(
    plate_readings = plate_readings,
    overlap_check  = overlap_check,
    summary = list(
      n_wells_csv               = length(wells_csv),
      n_wells_report            = length(wells_report),
      n_wells_report_not_csv    = length(wells_report_not_csv),
      n_wells_layout_not_csv    = length(wells_layout_not_csv),
      n_wells_csv_not_layout    = length(wells_csv_not_layout),
      n_shared_readings         = nrow(overlap_check),
      max_disagreement          = max_disagreement
    )
  )
}

## -----------------------------------------------------------------------
## annotate_wells()
## Purpose : Join layout labels onto the kinetic readings and parse each
##           label into well_type, sample_id, family, individual,
##           temperature, tissue weight, and standard concentration.
##           Expected label grammar:
##             sample              : <family>_<individual>_<temperature>-<assay_type>-<weight>-df.<n>
##             standard            : STD-<assay_type>-<nmol_per_well>
##             positive control    : POS-<assay_type>
##             background (pooled) : BG-<assay_type>
##           Unlike the paired per-sample background design used on earlier
##           plates (`_BG` appended to a sample's own label), this plate's
##           background wells are a single pooled control with no sample_id,
##           family, individual, or weight of their own (note 3) -- so
##           sample_id/family/individual/temperature/weight_mg parsing is
##           restricted to well_type == "sample" only.
##           Label typos (e.g. a misspelled assay_type) are corrected before
##           parsing via `label_fixes`, a named character vector of
##           c(pattern = replacement) pairs applied in order with gsub().
## Inputs  : plate_readings - data.frame from reconcile_kinetic_sources()$plate_readings
##           layout_wells   - data.frame from parse_plate_layout()
##           label_fixes    - optional named character vector, e.g.
##                            c("citrate_synthasse" = "citrate_synthase")
## Outputs : plate_long - fully annotated long-format data.frame, one row
##           per well x timepoint, with well_type, sample_id, family,
##           individual, temperature, weight_mg, std_nmol columns added
## -----------------------------------------------------------------------
annotate_wells <- function(plate_readings, layout_wells, label_fixes = NULL) {
  plate_long <- plate_readings %>%
    left_join(layout_wells, by = "well") %>%
    mutate(
      well_type = case_when(
        grepl("^STD-", label) ~ "standard",
        grepl("^POS-", label) ~ "positive_control",
        grepl("^BG-",  label) ~ "background",
        TRUE                  ~ "sample"
      ),
      label_clean = label
    )

  if (!is.null(label_fixes)) {
    for (pat in names(label_fixes)) {
      plate_long$label_clean <- gsub(pat, label_fixes[[pat]], plate_long$label_clean)
    }
  }

  plate_long <- plate_long %>%
    mutate(
      sample_id   = ifelse(well_type == "sample",
                           sub("-[a-zA-Z_]+-.*$", "", label_clean), NA_character_),
      family      = ifelse(!is.na(sample_id), sub("^(F[0-9]+)_.*$", "\\1", sample_id), NA_character_),
      individual  = ifelse(!is.na(sample_id), sub("^F[0-9]+_([0-9]+)_.*$", "\\1", sample_id), NA_character_),
      temperature = ifelse(!is.na(sample_id), sub("^F[0-9]+_[0-9]+_(.*)$", "\\1", sample_id), NA_character_),
      weight_mg   = ifelse(well_type == "sample",
                           as.numeric(sub("^.*-[a-zA-Z_]+-([0-9.]+)-df\\..*$", "\\1", label_clean)),
                           NA_real_),
      std_nmol    = ifelse(well_type == "standard",
                           as.numeric(sub("^STD-[a-zA-Z_]+-", "", label_clean)), NA_real_)
    ) %>%
    arrange(well_type, sample_id, well, time_min)

  # Fail loudly rather than silently dropping malformed labels
  stopifnot(
    !any(is.na(plate_long$well_type)),
    !any(is.na(plate_long$weight_mg[plate_long$well_type == "sample"])),
    !any(is.na(plate_long$std_nmol[plate_long$well_type == "standard"]))
  )

  plate_long
}

## -----------------------------------------------------------------------
## load_protein_concentrations()
## Purpose : Load one or more BCA/Bradford-style protein-concentration CSV
##           exports, concatenate them, and match each plate sample to
##           exactly one protein-concentration record. Fails loudly if any
##           plate sample has zero or more than one match. Converts
##           concentration (ug/mL) to total protein mass (mg) in the full
##           homogenate using the measured homogenization volume.
## Inputs  : protein_files        - character vector of CSV paths, each with
##                                   columns `Sample` and
##                                   `Calculated concentration (ug/mL)`
##           plate_sample_ids     - character vector of this plate's sample IDs
##           homogenate_volume_uL - measured homogenization volume (uL),
##                                   applied identically to every sample
##           sample_id_fixes      - optional named character vector of
##                                   c(pattern = replacement) label-typo
##                                   corrections applied to the protein
##                                   file's Sample column before matching
## Outputs : data.frame, one row per plate sample:
##           sample_id, source_file, conc_ug_mL, total_protein_mg
## -----------------------------------------------------------------------
load_protein_concentrations <- function(protein_files, plate_sample_ids,
                                         homogenate_volume_uL,
                                         sample_id_fixes = NULL) {
  protein_all <- protein_files %>%
    purrr::map_dfr(~ read.csv(.x, check.names = FALSE) %>%
                      mutate(source_file = basename(.x))) %>%
    rename(sample_id_protein = Sample, conc_ug_mL = `Calculated concentration (ug/mL)`) %>%
    select(sample_id_protein, source_file, conc_ug_mL)

  if (!is.null(sample_id_fixes)) {
    for (pat in names(sample_id_fixes)) {
      protein_all$sample_id_protein <- gsub(pat, sample_id_fixes[[pat]],
                                             protein_all$sample_id_protein)
    }
  }

  protein_matches <- protein_all %>% filter(sample_id_protein %in% plate_sample_ids)
  match_counts    <- protein_matches %>% count(sample_id_protein, name = "n_matches")

  stopifnot(
    all(plate_sample_ids %in% protein_matches$sample_id_protein),
    all(match_counts$n_matches == 1)
  )

  protein_matches %>%
    rename(sample_id = sample_id_protein) %>%
    mutate(total_protein_mg = conc_ug_mL * homogenate_volume_uL / 1e6) %>%
    arrange(sample_id)
}
```

### Well-level kinetic QC

Flags read glitches, non-rising ('decreasing') traces, elevated t0 baselines, and other well-level anomalies before any rate is trusted.

```{r pipeline-fns-qc}
## -----------------------------------------------------------------------
## local_step_excess()
## Purpose : For a kinetic trace, measure how much each read-to-read step
##           departs from the AVERAGE of its two immediate neighbouring
##           steps. A large local excess flags a read glitch; comparing to
##           local neighbours (rather than the trace's overall median step)
##           tolerates traces that smoothly decelerate throughout.
## Inputs  : od - numeric vector of OD readings in time order
## Outputs : numeric vector of length length(od) - 1 (one per step)
## -----------------------------------------------------------------------
local_step_excess <- function(od) {
  d <- diff(od)
  n <- length(d)
  local_expect <- vapply(seq_len(n), function(i) {
    mean(d[intersect(c(i - 1L, i + 1L), seq_len(n))])
  }, numeric(1))
  d - local_expect
}

## -----------------------------------------------------------------------
## compute_well_diagnostics()
## Purpose : Compute four independent per-well trace-shape diagnostics from
##           the raw kinetic trace, before any rate fitting: direction
##           (net change), monotonicity (fraction of rising steps),
##           over-range (any read above the photometric linearity ceiling),
##           and discontinuity (a read-to-read glitch via local_step_excess).
##           Also flags an elevated starting absorbance for sample wells,
##           and background wells that behave like active reactions.
## Inputs  : plate_long   - data.frame from annotate_wells()
##           assay_params - list with od_linear_max, glitch_excess_od,
##                          sample_baseline_od, bg_flat_od_max,
##                          bg_flat_drift_max
## Outputs : well_diagnostics - one row per well, with od_first, od_last,
##           od_max, net_change, frac_rising, max_step, typical_step,
##           step_excess, glitch_at_min, and flag_* / n_flags columns
## -----------------------------------------------------------------------
compute_well_diagnostics <- function(plate_long, assay_params) {
  plate_long %>%
    arrange(well, time_min) %>%
    group_by(well, plate_row, plate_col, well_type, sample_id, label, std_nmol, source) %>%
    summarise(
      od_first        = first(od),
      od_last         = last(od),
      od_max          = max(od),
      net_change      = last(od) - first(od),
      frac_rising     = mean(diff(od) > 0),
      max_step        = diff(od)[which.max(abs(diff(od)))],
      typical_step    = median(abs(diff(od))),
      step_excess     = local_step_excess(od)[which.max(abs(local_step_excess(od)))],
      glitch_at_min   = time_min[which.max(abs(local_step_excess(od))) + 1L],
      .groups = "drop"
    ) %>%
    mutate(
      step_ratio      = abs(max_step) / pmax(typical_step, 0.001),
      flag_decreasing = net_change <= 0,
      flag_over_range = od_max > assay_params$od_linear_max,
      flag_glitch     = abs(step_excess) > assay_params$glitch_excess_od,
      flag_high_baseline = well_type == "sample" & od_first > assay_params$sample_baseline_od,
      flag_bg_active  = well_type == "background" &
                        (od_first > assay_params$bg_flat_od_max |
                         abs(net_change) > assay_params$bg_flat_drift_max),
      n_flags         = flag_decreasing + flag_over_range + flag_glitch +
                        flag_bg_active + flag_high_baseline
    )
}

## -----------------------------------------------------------------------
## compute_baseline_diagnostics()
## Purpose : Classify sample/background wells by starting absorbance
##           (elevated vs. normal) and summarize, per sample, how many of
##           its replicates are baseline-compromised. A sample with all
##           replicates elevated has no clean replicate at all.
## Inputs  : well_diagnostics - data.frame from compute_well_diagnostics()
##           assay_params     - list with sample_baseline_od
## Outputs : list(baseline_check, baseline_per_sample)
## -----------------------------------------------------------------------
compute_baseline_diagnostics <- function(well_diagnostics, assay_params) {
  baseline_check <- well_diagnostics %>%
    filter(well_type %in% c("sample", "background")) %>%
    mutate(baseline = ifelse(od_first > assay_params$sample_baseline_od,
                             "elevated", "normal")) %>%
    arrange(desc(od_first))

  baseline_per_sample <- baseline_check %>%
    filter(well_type == "sample") %>%
    group_by(sample_id) %>%
    summarise(n_elevated = sum(baseline == "elevated"), n = n(),
              median_baseline = median(od_first), .groups = "drop") %>%
    arrange(desc(n_elevated))

  list(baseline_check = baseline_check, baseline_per_sample = baseline_per_sample)
}
```

### Standard curve and rate extraction

Fits the GSH standard curve, extracts each well's rate via a sliding-window linear regression gated on R-squared, and summarizes the positive control.

```{r pipeline-fns-rates}
## -----------------------------------------------------------------------
## fit_standard_curve()
## Purpose : Extract t=0 GSH standard readings, quantify signal drift over
##           the run, compute per-concentration replicate statistics, flag
##           outlier wells by deviation from their triplicate median, and
##           fit three candidate calibration lines (all wells, per-
##           concentration means, outlier-excluded). The outlier-excluded
##           fit is the calibration used downstream.
## Inputs  : plate_long   - data.frame from annotate_wells()
##           assay_params - list with std_outlier_od
## Outputs : list with standards_t0, standard_drift, standard_summary,
##           standards_flagged, fit_all_wells, fit_means, fit_no_outliers,
##           fit_comparison, std_slope, std_intercept, std_r2, std_nmol_max
## -----------------------------------------------------------------------
fit_standard_curve <- function(plate_long, assay_params) {
  standards_all <- plate_long %>%
    filter(well_type == "standard") %>%
    select(well, std_nmol, time_min, od, source)

  standards_t0 <- standards_all %>%
    filter(time_min == 0) %>%
    select(well, std_nmol, od, source) %>%
    arrange(std_nmol, well)

  standard_drift <- standards_all %>%
    group_by(well, std_nmol) %>%
    summarise(od_t0 = od[time_min == 0], od_t40 = od[time_min == max(time_min)],
              drift = od_t40 - od_t0, .groups = "drop") %>%
    arrange(std_nmol, well)

  standard_summary <- standards_t0 %>%
    group_by(std_nmol) %>%
    summarise(n = n(), mean_od = mean(od), sd_od = sd(od),
              se_od = sd(od) / sqrt(n()), cv_pct = 100 * sd(od) / mean(od),
              median_od = median(od), .groups = "drop") %>%
    arrange(std_nmol) %>%
    mutate(net_od = mean_od - mean_od[std_nmol == 0],
           od_per_nmol = ifelse(std_nmol > 0, net_od / std_nmol, NA_real_))

  standards_flagged <- standards_t0 %>%
    group_by(std_nmol) %>%
    mutate(triplicate_median = median(od),
           deviation = od - triplicate_median,
           is_outlier = abs(deviation) > assay_params$std_outlier_od) %>%
    ungroup() %>%
    arrange(desc(abs(deviation)))

  fit_all_wells   <- lm(od ~ std_nmol, data = standards_flagged)
  fit_means       <- lm(mean_od ~ std_nmol, data = standard_summary)
  fit_no_outliers <- lm(od ~ std_nmol, data = standards_flagged %>% filter(!is_outlier))

  fit_comparison <- bind_rows(
    data.frame(fit = "all wells",          n = nobs(fit_all_wells),
               slope = coef(fit_all_wells)[2],   intercept = coef(fit_all_wells)[1],
               r_squared = summary(fit_all_wells)$r.squared),
    data.frame(fit = "concentration means", n = nobs(fit_means),
               slope = coef(fit_means)[2],       intercept = coef(fit_means)[1],
               r_squared = summary(fit_means)$r.squared),
    data.frame(fit = "outlier-excluded",   n = nobs(fit_no_outliers),
               slope = coef(fit_no_outliers)[2], intercept = coef(fit_no_outliers)[1],
               r_squared = summary(fit_no_outliers)$r.squared)
  )

  list(
    standards_all      = standards_all,
    standards_t0       = standards_t0,
    standard_drift     = standard_drift,
    standard_summary   = standard_summary,
    standards_flagged  = standards_flagged,
    fit_all_wells      = fit_all_wells,
    fit_means          = fit_means,
    fit_no_outliers    = fit_no_outliers,
    fit_comparison     = fit_comparison,
    std_slope          = unname(coef(fit_no_outliers)[2]),
    std_intercept      = unname(coef(fit_no_outliers)[1]),
    std_r2             = summary(fit_no_outliers)$r.squared,
    std_nmol_max       = max(standard_summary$std_nmol)
  )
}

## -----------------------------------------------------------------------
## window_slopes()
## Purpose : Fit every sliding window of `w` consecutive timepoints on one
##           kinetic trace and return the slope (mOD/min) and R^2 of each.
## Inputs  : t - numeric vector of timepoints (min)
##           y - numeric vector of OD readings, same length as t
##           w - window width in number of points
## Outputs : data.frame, one row per window: t_start, t_end,
##           slope_mOD_min, r2
## -----------------------------------------------------------------------
window_slopes <- function(t, y, w) {
  n <- length(y)
  bind_rows(lapply(seq_len(n - w + 1), function(s) {
    i  <- s:(s + w - 1)
    f  <- lm(y[i] ~ t[i])
    r2 <- summary(f)$r.squared
    data.frame(t_start = t[i[1]], t_end = t[i[w]],
               slope_mOD_min = unname(coef(f)[2]) * 1000,
               r2 = ifelse(is.na(r2), 0, r2))
  }))
}

## -----------------------------------------------------------------------
## compute_well_rates()
## Purpose : For every well, scan all sliding windows of assay_params$
##           rate_window_n consecutive reads and record the max-increasing
##           window (used for activity) and the max-absolute window
##           (diagnostic only -- matches Gen5's own `Max V` convention,
##           which can report large negative "rates" for degrading wells).
##           A rate is `rate_usable` only if its R^2 clears the floor, the
##           well's net change is positive, and no read glitch falls inside
##           the fitted window.
## Inputs  : plate_long       - data.frame from annotate_wells()
##           well_diagnostics - data.frame from compute_well_diagnostics()
##           assay_params     - list with rate_window_n, rate_min_r2
## Outputs : well_rates - one row per well: t_start, t_end, slope_mOD_min,
##           r2, max_abs_slope_mOD_min, abs_window_is_negative,
##           glitch_in_window, rate_usable (plus diagnostic columns joined in)
## -----------------------------------------------------------------------
compute_well_rates <- function(plate_long, well_diagnostics, assay_params) {
  plate_long %>%
    arrange(well, time_min) %>%
    group_by(well, well_type, sample_id, std_nmol) %>%
    group_modify(~ {
      w   <- window_slopes(.x$time_min, .x$od, assay_params$rate_window_n)
      inc <- w[which.max(w$slope_mOD_min), ]
      abs_ <- w[which.max(abs(w$slope_mOD_min)), ]
      data.frame(
        t_start = inc$t_start, t_end = inc$t_end,
        slope_mOD_min = inc$slope_mOD_min, r2 = inc$r2,
        max_abs_slope_mOD_min = abs_$slope_mOD_min,
        abs_window_is_negative = abs_$slope_mOD_min < 0
      )
    }) %>%
    ungroup() %>%
    left_join(well_diagnostics %>% select(well, net_change, frac_rising, od_max,
                                          flag_decreasing, flag_over_range,
                                          flag_glitch, glitch_at_min,
                                          flag_bg_active, n_flags),
              by = "well") %>%
    mutate(
      glitch_in_window = flag_glitch & glitch_at_min > t_start & glitch_at_min <= t_end,
      rate_usable      = r2 >= assay_params$rate_min_r2 & net_change > 0 & !glitch_in_window
    )
}

## -----------------------------------------------------------------------
## compute_positive_control()
## Purpose : Compute the CS Positive Control's rate and activity, and check
##           that its replicates rose linearly -- the plate's proof that the
##           reaction chemistry (DTNB, coupling, reader) worked, independent
##           of any sample-specific problem.
## Inputs  : well_rates   - data.frame from compute_well_rates()
##           std_slope    - calibration slope (A412/nmol), from fit_standard_curve()
##           assay_params - list with sample_volume_uL, dilution_factor
## Outputs : pos_control - one row per positive-control well: t_start, t_end,
##           slope_mOD_min, r2, net_change, activity_mU_uL
## -----------------------------------------------------------------------
compute_positive_control <- function(well_rates, std_slope, assay_params) {
  well_rates %>%
    filter(well_type == "positive_control") %>%
    mutate(activity_mU_uL = (slope_mOD_min / 1000) / std_slope /
                            assay_params$sample_volume_uL * assay_params$dilution_factor)
}
```

### Background correction, replicate CV, and activity calculation

Estimates a single pooled background rate from the flat (well-behaved) background replicates, computes technical-replicate coefficients of variation, and converts corrected rates into protein-normalized CS activity.

```{r pipeline-fns-activity}
## -----------------------------------------------------------------------
## compute_background_correction()
## Purpose : Estimate a single, plate-wide background rate from the flat
##           (well-behaved) replicates of the pooled background control
##           (note 3), since anomalous background wells do not measure
##           background. This plate has one background triplicate for the
##           whole plate, so the estimate is a single row rather than one
##           row per sample_id.
## Inputs  : well_rates - data.frame from compute_well_rates()
## Outputs : background_correction - one-row data.frame: n_bg_total,
##           n_bg_flat, bg_rate_flat (NA if no flat replicate exists),
##           bg_rate_all
## -----------------------------------------------------------------------
compute_background_correction <- function(well_rates) {
  bg <- well_rates %>% filter(well_type == "background")
  data.frame(
    n_bg_total   = nrow(bg),
    n_bg_flat    = sum(!bg$flag_bg_active),
    bg_rate_flat = ifelse(sum(!bg$flag_bg_active) > 0,
                          mean(bg$slope_mOD_min[!bg$flag_bg_active]), NA_real_),
    bg_rate_all  = mean(bg$slope_mOD_min)
  )
}

## -----------------------------------------------------------------------
## compute_replicate_cv()
## Purpose : Compute technical-replicate coefficient of variation (CV%) on
##           the extracted rate, both across ALL three replicates and
##           across USABLE replicates only (rate_usable == TRUE). Flags any
##           sample whose CV exceeds assay_params$cv_threshold_pct.
## Inputs  : well_rates       - data.frame from compute_well_rates()
##           sample_id_order  - character vector giving the row order /
##                              complete sample-ID set to report (typically
##                              the plate's sample IDs, e.g. from
##                              protein_by_sample$sample_id)
##           assay_params     - list with cv_threshold_pct
## Outputs : cv_summary - one row per sample: n_all, mean_all, sd_all,
##           cv_all, n_usable, mean_usable, sd_usable, cv_usable,
##           excluded_wells, fails_cv_all, fails_cv_usable
## -----------------------------------------------------------------------
compute_replicate_cv <- function(well_rates, sample_id_order, assay_params) {
  sample_rates <- well_rates %>% filter(well_type == "sample") %>% arrange(sample_id, well)

  cv_all <- sample_rates %>%
    group_by(sample_id) %>%
    summarise(n_all = n(), mean_all = mean(slope_mOD_min), sd_all = sd(slope_mOD_min),
              cv_all = 100 * sd(slope_mOD_min) / mean(slope_mOD_min), .groups = "drop")

  cv_usable <- sample_rates %>%
    filter(rate_usable) %>%
    group_by(sample_id) %>%
    summarise(n_usable = n(), mean_usable = mean(slope_mOD_min),
              sd_usable = sd(slope_mOD_min),
              cv_usable = 100 * sd(slope_mOD_min) / mean(slope_mOD_min), .groups = "drop")

  data.frame(sample_id = sample_id_order, stringsAsFactors = FALSE) %>%
    left_join(cv_all, by = "sample_id") %>%
    left_join(cv_usable, by = "sample_id") %>%
    mutate(
      excluded_wells = vapply(sample_id, function(s) {
        w <- sample_rates$well[sample_rates$sample_id == s & !sample_rates$rate_usable]
        if (length(w) == 0) "-" else paste(w, collapse = ", ")
      }, character(1)),
      fails_cv_all    = cv_all > assay_params$cv_threshold_pct,
      fails_cv_usable = !is.na(cv_usable) & cv_usable > assay_params$cv_threshold_pct
    ) %>%
    arrange(sample_id)
}

## -----------------------------------------------------------------------
## calculate_cs_activity()
## Purpose : Compute per-sample CS activity following Abcam sec 10.3, using
##           only usable replicates, corrected for the single pooled
##           background rate (note 3), scaled through the standard curve,
##           and normalized to total extracted protein (rather than tissue
##           weight).
## Inputs  : well_rates            - data.frame from compute_well_rates()
##           protein_by_sample     - data.frame from load_protein_concentrations()
##           background_correction - one-row data.frame from
##                                    compute_background_correction()
##           std_slope             - calibration slope (A412/nmol), from fit_standard_curve()
##           std_nmol_max          - max calibrated GSH concentration, from fit_standard_curve()
##           assay_params          - list with sample_volume_uL, dilution_factor,
##                                    homogenate_volume_uL, read_duration_min
##           plate_long            - data.frame from annotate_wells(), used to
##                                    attach family/individual/temperature
## Outputs : cs_activity - one row per sample: n_reps_used, mean_rate_mOD_min,
##           sd_rate, cv_rate, family, individual, temperature, conc_ug_mL,
##           total_protein_mg, bg_rate_mOD_min, corrected_mOD_min,
##           activity_mU_per_uL, total_mU_in_homogenate,
##           activity_mU_per_mg_protein, within_std_range
## -----------------------------------------------------------------------
calculate_cs_activity <- function(well_rates, protein_by_sample, background_correction,
                                   std_slope, std_nmol_max, assay_params, plate_long) {
  sample_rates <- well_rates %>% filter(well_type == "sample") %>% arrange(sample_id, well)

  bg_rate <- ifelse(is.na(background_correction$bg_rate_flat), 0, background_correction$bg_rate_flat)

  sample_rates %>%
    filter(rate_usable) %>%
    group_by(sample_id) %>%
    summarise(n_reps_used = n(), mean_rate_mOD_min = mean(slope_mOD_min),
              sd_rate = sd(slope_mOD_min),
              cv_rate = 100 * sd(slope_mOD_min) / mean(slope_mOD_min), .groups = "drop") %>%
    left_join(plate_long %>% filter(well_type == "sample") %>%
                distinct(sample_id, family, individual, temperature),
              by = "sample_id") %>%
    left_join(protein_by_sample %>% select(sample_id, conc_ug_mL, total_protein_mg),
              by = "sample_id") %>%
    mutate(
      bg_rate_mOD_min       = bg_rate,
      corrected_mOD_min     = mean_rate_mOD_min - bg_rate_mOD_min,
      rate_OD_min           = corrected_mOD_min / 1000,
      nmol_per_min          = rate_OD_min / std_slope,
      activity_mU_per_uL    = nmol_per_min / assay_params$sample_volume_uL *
                              assay_params$dilution_factor,
      total_mU_in_homogenate = activity_mU_per_uL * assay_params$homogenate_volume_uL,
      activity_mU_per_mg_protein = total_mU_in_homogenate / total_protein_mg,
      nmol_in_window        = nmol_per_min * (assay_params$read_duration_min),
      within_std_range      = nmol_in_window <= std_nmol_max
    ) %>%
    arrange(sample_id)
}

## -----------------------------------------------------------------------
## build_results_table()
## Purpose : Assemble the final, presentation-ready per-sample results
##           table from cs_activity, cv_summary and baseline diagnostics,
##           with a plain-language Interpretation column.
## Inputs  : cs_activity          - data.frame from calculate_cs_activity()
##           cv_summary           - data.frame from compute_replicate_cv()
##           baseline_per_sample  - data.frame from compute_baseline_diagnostics()
##           assay_params         - list with cv_threshold_pct, homogenate_volume_uL
## Outputs : results_table - formatted data.frame, ready for kable()/write.csv()
## -----------------------------------------------------------------------
build_results_table <- function(cs_activity, cv_summary, baseline_per_sample, assay_params) {
  cs_activity %>%
    left_join(cv_summary %>% select(sample_id, n_all, cv_all), by = "sample_id") %>%
    left_join(baseline_per_sample %>% select(sample_id, n_elevated), by = "sample_id") %>%
    transmute(
      Sample                       = sample_id,
      Family                       = family,
      Individual                   = individual,
      Temperature                  = temperature,
      `Protein conc (ug/mL)`       = round(conc_ug_mL, 1),
      `Total protein (mg)`         = round(total_protein_mg, 3),
      `Reps used`                  = paste0(n_reps_used, "/", n_all),
      `CV all reps (%)`            = round(cv_all, 1),
      `CV used reps (%)`           = round(cv_rate, 1),
      `Rate (mA412/min)`           = round(mean_rate_mOD_min, 2),
      `BG rate (mA412/min)`        = round(bg_rate_mOD_min, 3),
      `Corrected rate (mA412/min)` = round(corrected_mOD_min, 2),
      `Activity (mU/uL)`           = round(activity_mU_per_uL, 4),
      `Activity (mU/mg protein)`   = round(activity_mU_per_mg_protein, 3),
      `CV flag`                    = ifelse(cv_all > assay_params$cv_threshold_pct,
                                            paste0("FAIL >", assay_params$cv_threshold_pct, "%"), "pass"),
      `Elevated baseline reps`     = paste0(n_elevated, "/3"),
      Interpretation               = case_when(
        n_elevated == 3 ~ "DO NOT USE - no clean replicate",
        n_elevated > 0  ~ "caution - some reps compromised",
        cv_rate > assay_params$cv_threshold_pct | is.na(cv_rate) ~ "caution - imprecise",
        TRUE ~ "usable"
      )
    )
}
```

### QC summary and narrative generation

Assembles the per-well QC summary table and auto-generates the plain-language QC findings bullets used in the QC SUMMARY section below.

```{r pipeline-fns-summary}
## -----------------------------------------------------------------------
## build_qc_summary_table()
## Purpose : Consolidate every QC check performed across the pipeline into a
##           single two-column (check, value) table, suitable for kable() and
##           for writing to qc_summary.csv. Mirrors the individual find
##           functions but as compact, at-a-glance figures rather than prose.
## Inputs  : layout_wells, reconcile_summary, std_curve, well_diagnostics,
##           well_rates, cv_summary, baseline_per_sample, pos_control,
##           assay_params
## Outputs : qc_summary - data.frame with columns check, value (both chr)
## -----------------------------------------------------------------------
build_qc_summary_table <- function(layout_wells, reconcile_summary, std_curve,
                                    well_diagnostics, well_rates, cv_summary,
                                    baseline_per_sample, pos_control, assay_params) {
  standard_summary  <- std_curve$standard_summary
  standards_flagged <- std_curve$standards_flagged
  standard_drift    <- std_curve$standard_drift
  sample_rates      <- well_rates %>% filter(well_type == "sample")

  data.frame(
    check = c(
      "Occupied wells in layout",
      "Layout wells missing from absorbance CSV",
      "Max disagreement between absorbance CSV and full report",
      "Standard concentrations with replicate CV > threshold",
      "Standard wells flagged as outliers",
      "Standard curve R^2 (outlier-excluded, replicate level)",
      "Standard curve R^2 (all wells, replicate level)",
      "Standard wells that LOST signal over the run",
      "Standards above photometric linearity ceiling",
      "Background control wells behaving as active reactions",
      "Sample wells with a decreasing trace",
      "Sample wells with an elevated starting A412",
      "Samples with ALL THREE replicates baseline-compromised",
      "Sample wells with a read glitch",
      "Sample wells usable for rate extraction",
      "Samples with technical CV > threshold (all reps)",
      "Samples with technical CV > threshold (usable reps)",
      "Positive control replicates rising and linear"
    ),
    value = c(
      nrow(layout_wells),
      reconcile_summary$n_wells_layout_not_csv,
      sprintf("%.g", reconcile_summary$max_disagreement),
      paste(standard_summary$std_nmol[standard_summary$cv_pct > assay_params$cv_threshold_pct],
            collapse = ", "),
      paste0(sum(standards_flagged$is_outlier), "/", nrow(standards_flagged), " (",
             paste(standards_flagged$well[standards_flagged$is_outlier], collapse = ", "), ")"),
      sprintf("%.4f", std_curve$std_r2),
      sprintf("%.4f", summary(std_curve$fit_all_wells)$r.squared),
      paste0(sum(standard_drift$drift < 0), "/", nrow(standard_drift)),
      paste(standard_summary$std_nmol[standard_summary$mean_od > assay_params$od_linear_max],
            collapse = ", "),
      paste0(sum(well_diagnostics$flag_bg_active), "/",
             sum(well_diagnostics$well_type == "background")),
      paste0(sum(well_diagnostics$flag_decreasing & well_diagnostics$well_type == "sample"), "/",
             sum(well_diagnostics$well_type == "sample")),
      paste0(sum(well_diagnostics$flag_high_baseline), "/",
             sum(well_diagnostics$well_type == "sample"), " (",
             paste(well_diagnostics$well[well_diagnostics$flag_high_baseline], collapse = ", "), ")"),
      paste0(sum(baseline_per_sample$n_elevated == baseline_per_sample$n), "/",
             nrow(baseline_per_sample), " (",
             paste(baseline_per_sample$sample_id[baseline_per_sample$n_elevated == baseline_per_sample$n],
                   collapse = ", "), ")"),
      paste0(sum(well_diagnostics$flag_glitch & well_diagnostics$well_type == "sample"), "/",
             sum(well_diagnostics$well_type == "sample")),
      paste0(sum(sample_rates$rate_usable), "/", nrow(sample_rates)),
      paste0(sum(cv_summary$fails_cv_all), "/", nrow(cv_summary)),
      paste0(sum(cv_summary$fails_cv_usable), "/", nrow(cv_summary)),
      paste0(sum(pos_control$net_change > 0 & pos_control$r2 > 0.99), "/", nrow(pos_control))
    ),
    stringsAsFactors = FALSE
  )
}

## -----------------------------------------------------------------------
## generate_qc_findings()
## Purpose : Walk the QC/CV/anomaly objects for a plate and emit a
##           consistent, generic-language markdown bullet list of findings
##           (decreasing/glitched/high-baseline wells, background-control
##           mix-ups, standard-curve QC failures, CV failures, disqualified
##           samples). Intended to seed the SUMMARY section of a plate's
##           .Rmd; a short hand-written subsection for plate-specific
##           interpretation should follow it.
## Inputs  : well_diagnostics    - data.frame from compute_well_diagnostics()
##           well_rates          - data.frame from compute_well_rates()
##           cv_summary          - data.frame from compute_replicate_cv()
##           baseline_per_sample - data.frame from compute_baseline_diagnostics()
##           std_curve           - list from fit_standard_curve()
##           reconcile_summary   - list from reconcile_kinetic_sources()$summary
##           assay_params        - list with cv_threshold_pct, od_linear_max
## Outputs : character vector, one markdown bullet per finding (empty
##           findings are omitted, so a clean plate returns a short list)
## -----------------------------------------------------------------------
generate_qc_findings <- function(well_diagnostics, well_rates, cv_summary,
                                  baseline_per_sample, std_curve,
                                  reconcile_summary, assay_params) {
  bullets <- character(0)

  # Source reconciliation
  if (reconcile_summary$n_wells_layout_not_csv > 0 ||
      (!is.na(reconcile_summary$max_disagreement) && reconcile_summary$max_disagreement >= 1e-9)) {
    bullets <- c(bullets, sprintf(
      "**Kinetic source disagreement.** %d layout well(s) missing from the absorbance CSV; max disagreement between absorbance CSV and full report was %.4g A412.",
      reconcile_summary$n_wells_layout_not_csv, reconcile_summary$max_disagreement))
  } else {
    bullets <- c(bullets, sprintf(
      "**Kinetic sources agree.** absorbance CSV and full report agree exactly on all %d shared well x timepoint readings; all layout wells are present.",
      reconcile_summary$n_shared_readings))
  }

  # Sample wells with decreasing kinetics
  dec_samples <- well_diagnostics %>% filter(well_type == "sample", flag_decreasing)
  if (nrow(dec_samples) > 0) {
    bullets <- c(bullets, sprintf(
      "**Decreasing kinetics.** %d sample well(s) end lower than (or equal to) their starting A412 and are excluded from rate extraction: %s.",
      nrow(dec_samples), paste(dec_samples$well, collapse = ", ")))
  }

  # Read glitches
  glitch_samples <- well_diagnostics %>% filter(well_type == "sample", flag_glitch)
  if (nrow(glitch_samples) > 0) {
    bullets <- c(bullets, sprintf(
      "**Read glitches.** %d sample well(s) contain a single read-to-read step inconsistent with its neighbours: %s (only disqualifying if it falls inside the fitted rate window).",
      nrow(glitch_samples), paste(glitch_samples$well, collapse = ", ")))
  }

  # Elevated baseline
  high_baseline <- well_diagnostics %>% filter(well_type == "sample", flag_high_baseline)
  if (nrow(high_baseline) > 0) {
    bullets <- c(bullets, sprintf(
      "**Elevated starting absorbance.** %d sample well(s) start above the baseline threshold (A412 > %s), indicating pre-existing thiol/contamination before the reaction began: %s.",
      nrow(high_baseline), assay_params$sample_baseline_od, paste(high_baseline$well, collapse = ", ")))
  }

  # Samples fully disqualified by baseline
  disqualified <- baseline_per_sample %>% filter(n_elevated == n)
  if (nrow(disqualified) > 0) {
    bullets <- c(bullets, sprintf(
      "**Sample(s) with no clean replicate.** Every replicate of %s has an elevated starting A412; these samples have no usable measurement on this plate regardless of CV.",
      paste(disqualified$sample_id, collapse = ", ")))
  }

  # Over-range wells
  over_range <- well_diagnostics %>% filter(flag_over_range)
  if (nrow(over_range) > 0) {
    bullets <- c(bullets, sprintf(
      "**Over-range wells.** %d well(s) exceed the photometric linearity ceiling (A412 > %s): %s.",
      nrow(over_range), assay_params$od_linear_max, paste(over_range$well, collapse = ", ")))
  }

  # Background wells behaving as active reactions
  bg_active <- well_diagnostics %>% filter(well_type == "background", flag_bg_active)
  if (nrow(bg_active) > 0) {
    bullets <- c(bullets, sprintf(
      "**Background control wells behaving as active reactions.** %d of %d background wells show a starting A412 or drift inconsistent with a flat, inactive well: %s.",
      nrow(bg_active), sum(well_diagnostics$well_type == "background"),
      paste(bg_active$well, collapse = ", ")))
  }

  # Standard curve QC
  std_summary <- std_curve$standard_summary
  cv_fail_std <- std_summary %>% filter(cv_pct > assay_params$cv_threshold_pct)
  if (nrow(cv_fail_std) > 0) {
    bullets <- c(bullets, sprintf(
      "**GSH standard curve replicate imprecision.** %d of %d concentrations exceed %s%% CV: %s nmol/well.",
      nrow(cv_fail_std), nrow(std_summary), assay_params$cv_threshold_pct,
      paste(cv_fail_std$std_nmol, collapse = ", ")))
  }
  n_std_outliers <- sum(std_curve$standards_flagged$is_outlier)
  if (n_std_outliers > 0) {
    bullets <- c(bullets, sprintf(
      "**Standard outlier wells.** %d of %d standard wells deviate from their triplicate median by more than %s A412: %s.",
      n_std_outliers, nrow(std_curve$standards_flagged), assay_params$std_outlier_od,
      paste(std_curve$standards_flagged$well[std_curve$standards_flagged$is_outlier], collapse = ", ")))
  }
  over_range_std <- std_summary %>% filter(mean_od > assay_params$od_linear_max)
  if (nrow(over_range_std) > 0) {
    bullets <- c(bullets, sprintf(
      "**Standards above the photometric linear range.** %s nmol/well read above A412 %s; the calibration is trustworthy only below these concentrations.",
      paste(over_range_std$std_nmol, collapse = ", "), assay_params$od_linear_max))
  }
  n_drift_neg <- sum(std_curve$standard_drift$drift < 0)
  if (n_drift_neg > 0) {
    bullets <- c(bullets, sprintf(
      "**Standard signal decay.** %d of %d standard wells lost signal over the run (TNB instability); the standard curve is read at t = 0.",
      n_drift_neg, nrow(std_curve$standard_drift)))
  }
  bullets <- c(bullets, sprintf(
    "**Standard curve fit (outlier-excluded):** slope = %.5f A412/nmol, R^2 = %.4f (all-wells R^2 = %.4f).",
    std_curve$std_slope, std_curve$std_r2, summary(std_curve$fit_all_wells)$r.squared))

  # Technical replicate CV failures
  fail_all <- cv_summary %>% filter(fails_cv_all)
  if (nrow(fail_all) > 0) {
    bullets <- c(bullets, sprintf(
      "**Technical replicate CV > %s%% (all replicates).** %d of %d samples: %s.",
      assay_params$cv_threshold_pct, nrow(fail_all), nrow(cv_summary),
      paste(sprintf("%s (%.0f%%)", fail_all$sample_id, fail_all$cv_all), collapse = ", ")))
  } else {
    bullets <- c(bullets, sprintf(
      "**Technical replicate CV.** All %d samples are within %s%% CV across all three replicates.",
      nrow(cv_summary), assay_params$cv_threshold_pct))
  }
  fail_usable <- cv_summary %>% filter(fails_cv_usable)
  if (nrow(fail_usable) > 0) {
    bullets <- c(bullets, sprintf(
      "**Technical replicate CV > %s%% persists after excluding flagged replicates.** %s.",
      assay_params$cv_threshold_pct,
      paste(sprintf("%s (%.0f%%, n=%d)", fail_usable$sample_id, fail_usable$cv_usable,
                    fail_usable$n_usable), collapse = ", ")))
  }

  bullets
}
```


## Assay parameters

All assay constants and QC thresholds are collected here so nothing
downstream hard-codes a number.

```{r assay-params}
assay_params <- list(
  sample_volume_uL      = 2,     # homogenate volume per reaction well (V in Abcam formula); also the pooled-background volume (note 3)
  buffer_volume_uL      = 48,    # Assay Buffer 7 added to reach 50 uL pre-Reaction Mix
  dilution_factor       = 1,     # D; layout `df.0` = undiluted homogenate
  homogenate_volume_uL  = 350,   # MEASURED buffer volume tissue was homogenized in (see note 2)
  read_wavelength_nm    = 412,
  read_interval_min     = 2,
  read_duration_min     = 20,    # see note 1
  rate_window_n         = 5,     # points per sliding regression window (5 pts = 8 min)
  rate_min_r2           = 0.80,  # min R^2 for a rate window to be trusted
  glitch_excess_od      = 0.02,  # |step - mean of neighbouring steps| flagging a read glitch
  cv_threshold_pct      = 15,    # technical-replicate CV QC threshold
  od_linear_max         = 1.5,   # upper A412 bound for reliable photometry
  std_outlier_od        = 0.15,  # |deviation from triplicate median| flagging a standard well
  sample_baseline_od    = 0.35,  # max acceptable t0 A412 for a sample well
  bg_flat_od_max        = 0.15,  # a well-behaved background well starts below this A412
  bg_flat_drift_max     = 0.05   # ...and drifts less than this over the whole run
)

cat("--- assay_params: assay constants and QC thresholds ---\n\n")
str(assay_params)
```

## Output directory

```{r output-dir}
# Output directory for this analysis (matches this file's name, per ../code/README.md)
output_dir <- "../outputs/Gen5-20260824-mgig-sormi-citrate_synthase-F07-ambient_and_36C"
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

cat("--- output_dir: destination for all figures and tables ---\n")
str(output_dir)
```

# DATA

Data are read from the local repo (`../data/raw_absorbance/`) so this document
renders before/after the files are pushed to GitHub.

## Plate layout

```{r data-layout}
data_dir <- "../data/raw_absorbance"
run_stem <- "Gen5-20260824-mgig-sormi-citrate_synthase-F07-ambient_and_36C"

layout_wells <- parse_plate_layout(file.path(data_dir, paste0("layout-", run_stem, ".csv")))

cat("Layout format detected:", attr(layout_wells, "layout_format"), "\n")
cat("Occupied wells in layout:", nrow(layout_wells), "\n")

cat("\n--- layout_wells: one row per occupied well, with its descriptive label ---\n\n")
str(layout_wells)
```

## Kinetic readers

```{r kinetic-readers}
absorbance_csv <- read_absorbance_csv(file.path(data_dir, paste0("absorbance-", run_stem, ".csv")))
full_report    <- read_full_report(file.path(data_dir, paste0("full_report-", run_stem, ".txt")))

cat("absorbance CSV : ", dplyr::n_distinct(absorbance_csv$well), " wells x ",
    dplyr::n_distinct(absorbance_csv$time_min), " timepoints\n", sep = "")
cat("full report    : ", dplyr::n_distinct(full_report$well), " wells x ",
    dplyr::n_distinct(full_report$time_min), " timepoints\n", sep = "")

cat("\n--- absorbance_csv: long-format readings from absorbance-*.csv ---\n\n")
str(absorbance_csv)
cat("\n--- full_report: long-format readings from full_report-*.txt ---\n\n")
str(full_report)
```

## Cross-check against the full report

`absorbance-*.csv` is used for every well; `full_report-*.txt` is used only
to confirm the CSV is trustworthy: every well it reports should also be in
the CSV, every well the layout expects should be in the CSV, and every
reading the two files share should agree exactly.

```{r reconcile-sources}
recon <- reconcile_kinetic_sources(absorbance_csv, full_report, layout_wells)
plate_readings <- recon$plate_readings

cat("Wells in absorbance CSV          :", recon$summary$n_wells_csv, "\n")
cat("Wells in full report             :", recon$summary$n_wells_report, "\n")
cat("Wells in full report not in CSV  :", recon$summary$n_wells_report_not_csv, "\n")
cat("Layout wells missing from CSV    :", recon$summary$n_wells_layout_not_csv, "\n")
cat("CSV wells not in layout          :", recon$summary$n_wells_csv_not_layout, "\n")
cat("Shared well x timepoints         :", recon$summary$n_shared_readings, "\n")
cat("Max |CSV - report| disagreement  :", recon$summary$max_disagreement, "\n")

cat("\n--- recon$overlap_check: per-reading comparison of the two raw files ---\n\n")
str(recon$overlap_check)
cat("\n--- plate_readings: readings for all occupied wells, from absorbance_csv ---\n\n")
str(plate_readings)
```

This plate's `absorbance-*.csv` covers the full 72 occupied wells (48 sample
+ 18 standard + 3 positive-control + 3 pooled-background, note 3) and the
layout carries no spelling typo, so no label fix is needed here. An earlier
export of this file was missing 6 wells (the `BG` triplicate and the `STD-0`
triplicate) relative to the layout and `full_report`; it was re-exported
from Gen5 before this analysis was run, and the check above now passes
cleanly.

## Annotate wells and parse metadata

```{r annotate-wells}
plate_long <- suppressWarnings(annotate_wells(plate_readings, layout_wells))

cat("Wells per type:\n")
print(plate_long %>% distinct(well, well_type) %>% count(well_type, name = "n_wells") %>%
        as.data.frame(), row.names = FALSE)
cat("\nSamples:", paste(sort(unique(na.omit(plate_long$sample_id))), collapse = ", "), "\n")
cat("Standards (nmol/well):", paste(sort(unique(na.omit(plate_long$std_nmol))), collapse = ", "), "\n")

cat("\n--- plate_long: fully annotated long-format plate, one row per well x timepoint ---\n\n")
str(plate_long)
```

`annotate_wells()` emits a harmless `NAs introduced by coercion` warning
(suppressed above), because its `weight_mg` parser only applies to `sample`
wells and returns `NA` for standards/positive control/background by
construction (background is a pooled control with no individual weight of
its own, note 3).
No `weight_mg` value is actually missing for a sample well;
`annotate_wells()`'s internal `stopifnot()` would halt otherwise.

## Protein concentration (normalization factor)

Activity is normalized to total extracted protein rather than tissue weight
(note 2). Protein concentration for each sample was measured on a separate
BCA/Bradford-style plate (595 nm): 12 of this plate's 16 samples come from
the primary `F07` protein-quantification run, and the remaining 4 (flagged
for re-assay in that run) come from a follow-up re-assay plate.
No `Sample` string typo needed correction for this plate's samples.

```{r protein-concentration}
protein_files <- c(
  "../data/BSA/raw_absorbance/sample_protein_concentrations_2.csv",
  "../outputs/Gen5-20260813-mgig-BSA-F07/sample_protein_concentrations.csv"
)

plate_sample_ids <- sort(unique(na.omit(plate_long$sample_id)))

protein_by_sample <- load_protein_concentrations(protein_files, plate_sample_ids,
                                                  assay_params$homogenate_volume_uL) %>%
  left_join(plate_long %>% filter(well_type == "sample") %>%
              distinct(sample_id, family, individual, temperature),
            by = "sample_id")

cat("Samples matched to a protein concentration record:", nrow(protein_by_sample),
    "/", length(plate_sample_ids), "\n")
cat("Protein concentration range (ug/mL):",
    paste(range(protein_by_sample$conc_ug_mL), collapse = " - "), "\n")
cat("Total protein per homogenate range (mg):",
    paste(round(range(protein_by_sample$total_protein_mg), 3), collapse = " - "), "\n\n")
print(protein_by_sample %>% select(sample_id, source_file, conc_ug_mL, total_protein_mg) %>%
        as.data.frame(), row.names = FALSE)

cat("\n--- protein_by_sample: matched protein concentration and total protein per homogenate ---\n\n")
str(protein_by_sample)
```

# KINETIC TRACES

Every trace is plotted before any rate is extracted. In this assay the
**shape** of the trace is the primary diagnostic: CS activity must produce a
monotonically *rising* A412, and only a genuinely linear stretch may be used
for a rate.

```{r plot-kinetic-traces, fig.width = 11, fig.height = 8}
trace_plot <- ggplot(plate_long, aes(x = time_min, y = od, group = well, colour = well_type)) +
  geom_line(linewidth = 0.5) +
  geom_hline(yintercept = assay_params$od_linear_max,
             linetype = "dotted", colour = "grey40") +
  facet_wrap(~ plate_row + plate_col, ncol = 12,
             labeller = function(d) list(paste0(d$plate_row, d$plate_col))) +
  scale_colour_manual(values = c(sample = "#08519c", background = "#a63603",
                                 standard = "#238b45", positive_control = "#6a51a3")) +
  labs(title = "A412 kinetic traces, all occupied wells",
       subtitle = paste0("Dotted line = assumed photometric linearity ceiling (A412 = ",
                         assay_params$od_linear_max,
                         "). Rising traces only are usable for rate extraction."),
       x = "Time (min)", y = "A412", colour = "Well type") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        strip.text = element_text(size = 6),
        axis.text  = element_text(size = 5),
        legend.position = "bottom")

ggsave(file.path(output_dir, "kinetic_traces_all_wells.png"), trace_plot,
       width = 11, height = 8, dpi = 300)

cat("--- trace_plot: ggplot object structure ---\n\n")
summary(trace_plot)

trace_plot
```

Nearly every sample and positive-control trace on this plate rises across
the full 20-minute window (replicate-level R<sup>2</sup> > 0.99 for the
fitted rate window on 47 of the 48 sample wells, computed below) -- the sole
exception, `A7` (`F07_03_ambient`), is completely flat at A412 0.099 for
all 11 reads, a dead/failed well rather than a slow reaction (see ANOMALY
DETECTION). The three pooled-background traces (`F10`-`F12`) are **not**
flat here -- they rise at a rate closer to a sample well than a background
well, the signature of the confirmed Reaction Mix mix-up (note 4).

# ANOMALY DETECTION

## Per-well trace diagnostics

Four independent checks, computed from the raw trace before any rate
fitting:

1. **Direction** — net A412 change from first to last read. A CS-containing
   well must be positive.
2. **Monotonicity** — fraction of the 10 read-to-read intervals that rise.
3. **Over-range** — any read above the photometric linearity ceiling.
4. **Discontinuity** — a single read-to-read step that departs from its two
   *neighbouring* steps, catching read glitches rather than biology.

Two further checks specific to well role: an elevated starting A412
(pre-existing thiol/contamination) for sample wells, and a background well
whose trace is not flat (indicating it received Reaction Mix rather than
Background Control Mix).

```{r anomaly-diagnostics}
well_diagnostics <- compute_well_diagnostics(plate_long, assay_params)

cat("Flag counts across all", nrow(well_diagnostics), "wells:\n")
print(well_diagnostics %>% summarise(across(starts_with("flag_"), sum)) %>%
        as.data.frame(), row.names = FALSE)

cat("\nWells with at least one flag:", sum(well_diagnostics$n_flags > 0), "/", nrow(well_diagnostics), "\n")

cat("\n--- well_diagnostics: per-well trace diagnostics and anomaly flags ---\n\n")
str(well_diagnostics)
```

## Anomaly table

```{r anomaly-table}
anomaly_table <- well_diagnostics %>%
  filter(n_flags > 0) %>%
  select(well, sample_id, well_type, od_first, net_change, n_flags,
         flag_decreasing, flag_over_range, flag_glitch,
         flag_high_baseline, flag_bg_active) %>%
  arrange(desc(n_flags), well)

kable(anomaly_table, digits = 3,
      caption = "Wells with at least one anomaly flag")

cat("\nTotal flagged wells:", nrow(anomaly_table), "/", nrow(well_diagnostics), "\n")

cat("\n--- anomaly_table: all flagged wells with their specific flags ---\n\n")
str(anomaly_table)
```

19 of the 72 wells are flagged, in three distinct groups:

1. **15 GSH standard wells** are flagged `flag_decreasing` -- all five
   non-zero concentrations (8, 16, 24, 32, 40 nmol) end at or below their
   own starting A412 over the 20-minute window, consistent with
   TNB<sup>2-</sup> signal decay (see GSH STANDARD CURVE below). Because the
   standard curve is calibrated at t = 0 (not from the endpoint), this does
   not compromise the calibration (R<sup>2</sup> = 0.9992, no outlier
   wells).
2. **All 3 pooled-background wells** (`F10`, `F11`, `F12`) are flagged
   `flag_bg_active` -- the confirmed Reaction Mix mix-up (note 4). See
   BACKGROUND CORRECTION for what this means for this plate's activity
   values.
3. **One sample well**, `A7` (`F07_03_ambient`), is flagged `flag_decreasing`
   -- it is completely flat (net change = 0.000, A412 = 0.099 for all 11
   reads), not a slow-rising reaction. Its two sibling replicates (`A8`,
   `A9`) rise normally, so `F07_03_ambient`'s activity is still computed
   from 2 of 3 replicates (see RATE EXTRACTION and TECHNICAL REPLICATE
   PRECISION).

No sample well shows a read glitch or an elevated starting baseline, and no
sample well besides `A7` fails to rise.

## Background wells behaving as active reactions

```{r bg-anomaly}
bg_active_wells <- well_diagnostics %>%
  filter(well_type == "background", flag_bg_active) %>%
  select(well, sample_id, od_first, net_change)

kable(bg_active_wells, digits = 3,
      caption = "Background wells with a non-flat trace")

cat("Background wells behaving as active reactions:", nrow(bg_active_wells),
    "/", sum(well_diagnostics$well_type == "background"), "\n")

cat("\n--- bg_active_wells: background wells excluded from the background-rate estimate ---\n\n")
str(bg_active_wells)
```

**All three pooled-background replicates are flagged.** `F10` and `F11`
start slightly above the 0.15 A412 threshold (0.156, 0.152) and `F12` starts
just under it (0.118), but all three fail on drift regardless: net change
0.067-0.119 A412 over 20 minutes, 1.3-2.4x the 0.05 flat-well threshold and
of the same order of magnitude as a real sample's signal. This is the
confirmed Reaction Mix mix-up (note 4), not genuine background activity.
Because 0 of 3 replicates are flat, none are used in a background-rate
estimate -- see BACKGROUND CORRECTION.

## Baseline diagnostic (elevated starting absorbance)

```{r baseline-diagnostic}
bd <- compute_baseline_diagnostics(well_diagnostics, assay_params)
baseline_check       <- bd$baseline_check
baseline_per_sample  <- bd$baseline_per_sample

cat("Sample/background wells with elevated starting A412 (> ",
    assay_params$sample_baseline_od, "):",
    sum(baseline_check$baseline == "elevated"), "/", nrow(baseline_check), "\n", sep = "")

kable(baseline_per_sample, digits = 3,
      caption = "Elevated-baseline replicate count per sample")

cat("\n--- baseline_check: per-well baseline classification ---\n\n")
str(baseline_check)
cat("\n--- baseline_per_sample: elevated-replicate count per sample ---\n\n")
str(baseline_per_sample)
```

No sample or background well anywhere on this plate starts above A412 0.35
— sample wells start within a 0.099-0.161 range (the low end is `A7`'s dead
well) and the pooled background wells start within 0.118-0.156, both well
below the 0.35 threshold. No sample has any elevated-baseline replicate, and
none is disqualified on this basis -- the background wells' problem is
their *drift*, not their starting absorbance (note 4).

## Plot starting absorbance

```{r plot-baseline, fig.width = 9, fig.height = 6}
baseline_plot <- ggplot(baseline_check, aes(x = reorder(well, od_first), y = od_first,
                                            fill = well_type)) +
  geom_col() +
  geom_hline(yintercept = assay_params$sample_baseline_od,
             linetype = "dashed", colour = "#cb181d") +
  scale_fill_manual(values = c(sample = "#08519c", background = "#a63603")) +
  labs(title = "Starting A412 (t = 0) for every sample/background well",
       subtitle = paste0("Dashed line = elevated-baseline threshold (A412 = ",
                         assay_params$sample_baseline_od, "). No well on this plate exceeds it."),
       x = "Well", y = "A412 at t = 0", fill = "Well type") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        axis.text.x = element_text(angle = 90, size = 5, vjust = 0.5),
        legend.position = "bottom")

ggsave(file.path(output_dir, "starting_absorbance.png"), baseline_plot,
       width = 9, height = 6, dpi = 300)

cat("--- baseline_plot: ggplot object structure ---\n\n")
summary(baseline_plot)

baseline_plot
```

# GSH STANDARD CURVE

The GSH standard is a **stoichiometric endpoint**, not an enzymatic reaction:
free thiol reduces DTNB immediately on mixing. The standard curve is
therefore read from a single timepoint, and t = 0 is used because
TNB<sup>2-</sup> degrades over the run (quantified below).

## Fit the standard curve

Three candidate fits are compared:

- **all wells** — every standard replicate, no exclusions
- **concentration means** — the conventional fit, on per-concentration means
- **outlier-excluded** — replicate-level fit after dropping flagged wells

The outlier-excluded fit is used for activity calculation.

```{r standard-curve-fit}
std_curve <- fit_standard_curve(plate_long, assay_params)

standards_t0       <- std_curve$standards_t0
standard_drift     <- std_curve$standard_drift
standard_summary   <- std_curve$standard_summary
standards_flagged  <- std_curve$standards_flagged
fit_all_wells      <- std_curve$fit_all_wells
fit_comparison     <- std_curve$fit_comparison
std_slope          <- std_curve$std_slope
std_intercept      <- std_curve$std_intercept
std_r2             <- std_curve$std_r2
std_nmol_max       <- std_curve$std_nmol_max

cat("Standard wells:", nrow(standards_t0), "at", dplyr::n_distinct(standards_t0$std_nmol),
    "concentrations\n")
cat("Standard drift over 20 min (dA412): median",
    sprintf("%.3f", median(standard_drift$drift)),
    "| range", paste(sprintf("%.3f", range(standard_drift$drift)), collapse = " to "), "\n")
cat("Standards that LOST signal (drift < 0):", sum(standard_drift$drift < 0), "/", nrow(standard_drift), "\n")

cat("\n--- standard_summary: per-concentration mean/SD/CV of t0 absorbance ---\n\n")
str(standard_summary)
kable(standard_summary, digits = c(0, 0, 3, 4, 4, 2, 3, 3, 4),
      caption = "GSH standard curve, per-concentration summary at t = 0")

cat("\nConcentrations exceeding", assay_params$cv_threshold_pct, "% replicate CV:",
    paste(standard_summary$std_nmol[standard_summary$cv_pct > assay_params$cv_threshold_pct],
          collapse = ", "), "\n")

cat("\n--- standards_flagged: every standard well with its deviation from the triplicate median ---\n\n")
str(standards_flagged)
kable(standards_flagged %>% filter(is_outlier) %>% select(well, std_nmol, od, triplicate_median, deviation),
      digits = 3, caption = "Standard wells flagged as outliers (|deviation| > 0.15 A412)")

cat("\n--- fit_comparison: slope/intercept/R^2 for all three candidate fits ---\n\n")
str(fit_comparison)
# row.names = FALSE: coef() attaches the predictor's name ("std_nmol") to the
# slope/intercept elements, which data.frame() otherwise promotes to a mangled
# row label (e.g. "std_nmol...1") that breaks pandoc's pipe-table parser.
kable(fit_comparison, digits = c(0, 0, 5, 4, 5), row.names = FALSE,
      caption = "Standard curve fit comparison (outlier-excluded fit used downstream)")
```

**No standard well is flagged as an outlier on this plate** -- every
concentration's triplicate CV is under 5% (highest: the 0 nmol blank at
4.8%; every non-zero concentration is under 1.6%), and no well deviates from
its triplicate median by more than the 0.15 A412 threshold. Because there
are no outliers to exclude, the "outlier-excluded" and "all wells" fits are
identical: slope = `r sprintf("%.5f", std_slope)` A412/nmol,
R<sup>2</sup> = `r sprintf("%.4f", std_r2)`.
**All 15 non-zero-concentration standard replicates lose signal** (drift <
0) between t = 0 and t = 20 min, consistent with TNB<sup>2-</sup>
instability -- the 0 nmol blank replicates, which have no thiol to lose,
instead drift very slightly upward. This exactly matches the 15 standard
wells flagged `flag_decreasing` above. Because the curve is read at t = 0,
this decay affects none of the values used for calibration.

## Plot the standard curve

```{r plot-standard-curve, fig.width = 9, fig.height = 6}
std_plot <- ggplot(standards_flagged, aes(x = std_nmol, y = od)) +
  geom_point(aes(colour = is_outlier), size = 3) +
  geom_smooth(data = standards_flagged %>% filter(!is_outlier),
              method = "lm", se = TRUE, colour = "#238b45", fill = "#a1d99b") +
  scale_colour_manual(values = c(`TRUE` = "#cb181d", `FALSE` = "#08519c"),
                      labels = c(`TRUE` = "outlier", `FALSE` = "included"),
                      name = "Standard well") +
  labs(title = "GSH standard curve (t = 0), family F07 (ambient and 36°C)",
       subtitle = sprintf("Outlier-excluded fit: slope = %.5f A412/nmol, R² = %.4f (n = %d)",
                          std_slope, std_r2, nrow(standards_flagged %>% filter(!is_outlier))),
       x = "GSH (nmol/well)", y = "A412 at t = 0") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        legend.position = "bottom")

ggsave(file.path(output_dir, "gsh_standard_curve.png"), std_plot,
       width = 9, height = 6, dpi = 300)

cat("--- std_plot: ggplot object structure ---\n\n")
summary(std_plot)

std_plot
```

The top standard (40 nmol) stays comfortably within the photometric linear
range here -- the highest reading is A412 1.37, below the 1.5 ceiling -- so
the calibration does not need to be restricted to a reduced concentration
range.

# RATE EXTRACTION

Rates are extracted with a sliding-window approach: for each well, the
5-point (8-minute) window of steepest *positive* slope is taken as the
rate, with its R<sup>2</sup> retained for the usability check.

```{r rate-windows}
well_rates <- compute_well_rates(plate_long, well_diagnostics, assay_params)

cat("--- well_rates: extracted rate window per well ---\n\n")
str(well_rates)
```

## Rate tables by well type

```{r rate-tables}
sample_rates <- well_rates %>% filter(well_type == "sample")

kable(sample_rates %>% select(well, sample_id, t_start, t_end, slope_mOD_min, r2, rate_usable),
      digits = c(0, 0, 0, 0, 3, 4, 0),
      caption = "Extracted rate window, sample wells")

cat("\nSample wells with a usable rate:", sum(sample_rates$rate_usable), "/", nrow(sample_rates), "\n")
cat("Sample rate range (mA412/min):",
    paste(sprintf("%.2f", range(sample_rates$slope_mOD_min)), collapse = " - "), "\n")
cat("Median replicate-level R^2:", sprintf("%.4f", median(sample_rates$r2)), "\n")

cat("\n--- sample_rates: rate-window results restricted to sample wells ---\n\n")
str(sample_rates)
```

47 of the 48 sample wells clear the R<sup>2</sup> >= 0.80 and
positive-net-change usability bar; every usable replicate-level
R<sup>2</sup> exceeds 0.997 (median 0.9994). The one exclusion is `A7`
(`F07_03_ambient`), the dead well noted above (slope = 0.00 mA412/min,
R<sup>2</sup> = 0, `rate_usable = FALSE`); its two siblings (`A8`, `A9`)
are both usable. Among the 47 usable wells, rates range 3.40-8.65
mA412/min.

## Positive control

```{r positive-control}
pos_control <- compute_positive_control(well_rates, std_curve$std_slope, assay_params)

kable(pos_control %>% select(well, t_start, t_end, slope_mOD_min, r2, net_change, activity_mU_uL),
      digits = c(0, 0, 0, 3, 4, 3, 4),
      caption = "Positive control replicates")

cat("Positive control mean rate:", sprintf("%.3f", mean(pos_control$slope_mOD_min)), "mA412/min\n")
cat("Positive control CV:", sprintf("%.1f",
    100 * sd(pos_control$slope_mOD_min) / mean(pos_control$slope_mOD_min)), "%\n")
cat("All replicates rising and linear (R^2 > 0.99):",
    all(pos_control$net_change > 0 & pos_control$r2 > 0.99), "\n")

cat("\n--- pos_control: positive control rate and R^2 per replicate ---\n\n")
str(pos_control)
```

The positive control rose linearly in all three replicates (R<sup>2</sup>
0.998-1.000) at a consistent 2.85-3.00 mA412/min, a 2.6% CV across the
triplicate -- confirming that the core reaction chemistry and reader were
functioning normally on this plate. This is a tight, well-behaved positive
control.

# BACKGROUND CORRECTION

This plate uses a **single pooled background control** (note 3): the
background rate is normally estimated once, from the flat (well-behaved)
replicates of the `BG-citrate_synthase` triplicate, and the same corrected
rate is subtracted from every sample. **On this plate, none of the three
background replicates are flat** (note 4) -- `compute_background_correction()`
therefore cannot produce a background-rate estimate, and its documented
fallback applies: a missing (`NA`) `bg_rate_flat` is treated as 0. Every
sample's corrected rate on this plate is therefore identical to its raw
rate.

```{r background-correction}
background_correction <- compute_background_correction(well_rates)

cat("Pooled background rate (flat wells only):\n")
print(background_correction, row.names = FALSE, digits = 3)

cat("\nBackground rate used for correction:",
    sprintf("%.3f", background_correction$bg_rate_flat), "mA412/min (from",
    background_correction$n_bg_flat, "of", background_correction$n_bg_total, "replicates)\n")
cat("Median sample rate:",
    sprintf("%.2f", median(sample_rates$slope_mOD_min)), "mA412/min\n")
cat("Background as % of median sample signal:",
    sprintf("%.1f", 100 * background_correction$bg_rate_flat /
                    median(sample_rates$slope_mOD_min)), "%\n")

cat("\n--- background_correction: pooled background rate estimate ---\n\n")
str(background_correction)
```

`n_bg_flat` is 0 of 3, so `bg_rate_flat` is `NA` and is treated as 0
downstream. For reference only (**not used** in any activity calculation),
the mean rate across all three -- compromised -- background replicates is
7.7 mA412/min, actually *higher* than the median sample rate of 6.58
mA412/min, underscoring that these wells behaved as active reactions rather
than anything resembling a background control.

## Background significance test

The flat/active threshold classification above is backed by an explicit
significance test: a Welch two-sample t-test of well-level rate (background
vs. sample), plus an ANOVA on the OD trajectory over time. This mirrors the
one-time background ANOVA validation done by Cattau et al. (2023) on the
legacy version of this assay. **On a well-behaved plate this test should
show background wells statistically slower than sample wells; on this
plate, given the confirmed mix-up, no such difference is expected** -- it is
run here mainly to document that expectation quantitatively rather than as a
validity check.

```{r background-significance}
bg_sample_rates <- well_rates %>%
  filter(well_type %in% c("sample", "background"))

bg_significance_test <- t.test(slope_mOD_min ~ well_type, data = bg_sample_rates)

cat("--- Welch two-sample t-test: well-level rate (background vs. sample) ---\n\n")
print(bg_significance_test)

bg_trend_anova <- anova(lm(od ~ well_type * time_min,
                            data = plate_long %>% filter(well_type %in% c("sample", "background"))))

cat("\n--- ANOVA: OD trajectory by well_type x time_min (background vs sample wells) ---\n\n")
print(bg_trend_anova)

cat("\nInterpretation: mean background-well rate is",
    sprintf("%.2f", unname(bg_significance_test$estimate["mean in group background"])),
    "mA412/min vs.", sprintf("%.2f", unname(bg_significance_test$estimate["mean in group sample"])),
    "mA412/min for sample wells (p =", format.pval(bg_significance_test$p.value, digits = 3),
    "). Background wells are NOT statistically slower than sample wells on\n",
    "this plate (n = 3 background replicates limits power, but the point\n",
    "estimate itself is not lower) -- consistent with the confirmed Reaction\n",
    "Mix mix-up (note 4) rather than a genuine, if imprecise, background.\n")

cat("\n--- bg_significance_test: Welch t-test result object ---\n\n")
str(bg_significance_test)

cat("\n--- bg_trend_anova: ANOVA table for OD ~ well_type * time_min ---\n\n")
str(bg_trend_anova)
```

# TECHNICAL REPLICATE PRECISION

Coefficient of variation across the three technical replicates of each
sample, computed on the extracted rate. Both **all** and **usable
replicates only** are reported; they differ only for `F07_03_ambient`,
whose dead `A7` well is excluded from the "usable" calculation.

```{r replicate-cv}
cv_summary <- compute_replicate_cv(well_rates, protein_by_sample$sample_id, assay_params)

kable(cv_summary %>% select(sample_id, n_all, mean_all, sd_all, cv_all,
                            n_usable, mean_usable, sd_usable, cv_usable,
                            excluded_wells),
      digits = c(0, 0, 2, 2, 1, 0, 2, 2, 1, 0),
      col.names = c("Sample", "n", "Mean rate", "SD", "CV (%)",
                    "n usable", "Mean rate", "SD", "CV (%)", "Excluded wells"),
      caption = paste0("Technical replicate CV of the extracted rate (mA412/min), ",
                       "all replicates vs. usable replicates only"))

cat("\nSamples exceeding CV", paste0(assay_params$cv_threshold_pct, "%"),
    "on ALL replicates:", sum(cv_summary$fails_cv_all), "/", nrow(cv_summary), "\n")
print(cv_summary %>% filter(fails_cv_all) %>%
        select(sample_id, n_all, mean_all, sd_all, cv_all) %>%
        as.data.frame(), row.names = FALSE, digits = 4)

cat("\nSamples exceeding CV", paste0(assay_params$cv_threshold_pct, "%"),
    "on USABLE replicates:", sum(cv_summary$fails_cv_usable), "\n")
print(cv_summary %>% filter(fails_cv_usable) %>%
        select(sample_id, n_usable, mean_usable, sd_usable, cv_usable) %>%
        as.data.frame(), row.names = FALSE, digits = 4)

write.csv(cv_summary, file.path(output_dir, "technical_replicate_cv.csv"), row.names = FALSE)

cat("\n--- cv_summary: per-sample technical replicate CV, both variants ---\n\n")
str(cv_summary)
```

**Only one sample exceeds the 15% CV threshold, and only on the "all
replicates" metric.** `F07_03_ambient` shows an 86.9% CV across all three
raw replicates, entirely driven by its dead `A7` well (slope 0.00 vs. 6.85
and 7.50 mA412/min for `A8`/`A9`); its usable-replicate CV is a normal 6.4%.
Every other sample is well under the threshold on both metrics -- the
highest of the remaining 15 is `F07_03_36C` at 3.1%, and most are under 2%.

## Plot replicate spread

```{r plot-replicate-cv, fig.width = 9, fig.height = 6}
cv_plot_data <- sample_rates %>%
  left_join(cv_summary %>% select(sample_id, cv_all), by = "sample_id")

cv_plot <- ggplot(cv_plot_data, aes(x = sample_id, y = slope_mOD_min)) +
  geom_point(aes(colour = rate_usable, shape = rate_usable),
             size = 3, position = position_jitter(width = 0.12, height = 0)) +
  stat_summary(fun = mean, geom = "crossbar", width = 0.5,
               colour = "grey30", linewidth = 0.3) +
  geom_text(data = cv_summary, aes(x = sample_id, y = Inf,
                                   label = sprintf("CV %.0f%%", cv_all)),
            vjust = 1.6, size = 3,
            colour = ifelse(cv_summary$fails_cv_all, "#cb181d", "grey30")) +
  scale_colour_manual(values = c(`TRUE` = "#08519c", `FALSE` = "#cb181d"),
                      labels = c(`TRUE` = "usable", `FALSE` = "flagged"), name = "Replicate") +
  scale_shape_manual(values = c(`TRUE` = 16, `FALSE` = 4),
                     labels = c(`TRUE` = "usable", `FALSE` = "flagged"), name = "Replicate") +
  labs(title = "Technical replicate spread in extracted rate",
       subtitle = paste0("Crossbar = mean of all three replicates; CV computed on all three. ",
                         "Threshold ", assay_params$cv_threshold_pct, "%."),
       x = "Sample", y = "Rate (mA412/min)") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        axis.text.x = element_text(angle = 30, hjust = 1))

ggsave(file.path(output_dir, "technical_replicate_cv.png"), cv_plot,
       width = 9, height = 6, dpi = 300)

cat("--- cv_plot: ggplot object structure ---\n\n")
summary(cv_plot)

cv_plot
```

# CITRATE SYNTHASE ACTIVITY

## Calculation

Following Abcam Section 10.3:

`Sample CS activity = B / (dT x V) x D`  nmol/min/µL (= mU/µL)

1. **Rate** (A412/min) = max-increasing sliding-window slope / 1000
2. **Background-corrected rate** = sample rate - the single pooled
   background rate (note 3), applied identically to every sample -- on
   this plate that background rate is 0 for every sample (note 4), so
   corrected rate = raw rate throughout
3. **B/dT** (nmol/min) = background-corrected rate / standard curve slope
   (`r sprintf("%.5f", std_slope)` A412 per nmol, from the outlier-excluded fit)
4. **Activity** (mU/µL) = (B/dT) / *V* x *D*, with
   *V* = `r assay_params$sample_volume_uL` µL and
   *D* = `r assay_params$dilution_factor`
5. **Protein-normalized** (mU/mg protein) = activity x `homogenate_volume_uL`
   (`r assay_params$homogenate_volume_uL` µL, measured) / total protein (mg)

```{r calculate-activity}
cs_activity <- calculate_cs_activity(well_rates, protein_by_sample, background_correction,
                                      std_slope, std_nmol_max, assay_params, plate_long)

cat("Samples with a computed activity:", nrow(cs_activity), "/", nrow(protein_by_sample), "\n")
cat("Activity range (mU/uL):",
    paste(sprintf("%.4f", range(cs_activity$activity_mU_per_uL)), collapse = " - "), "\n")
cat("Activity range (mU/mg protein):",
    paste(sprintf("%.3f", range(cs_activity$activity_mU_per_mg_protein)), collapse = " - "), "\n")

cat("\n--- cs_activity: per-sample citrate synthase activity ---\n\n")
str(cs_activity)
```

## Results table

```{r results-table}
results_table <- build_results_table(cs_activity, cv_summary, baseline_per_sample, assay_params)

kable(results_table,
      caption = paste0("Citrate synthase activity, family F07, ambient and 36°C (mU/mg protein uses ",
                       "measured protein concentration x the ", assay_params$homogenate_volume_uL,
                       " uL homogenization volume)"))

cat("\nInterpretation summary:\n")
print(results_table %>% count(Interpretation, name = "n_samples") %>%
        as.data.frame(), row.names = FALSE)

write.csv(results_table, file.path(output_dir, "citrate_synthase_activity_results.csv"),
          row.names = FALSE)

cat("\n--- results_table: formatted per-sample results written to CSV ---\n\n")
str(results_table)
```

All 16 samples are classified `usable` in the `Interpretation` column --
`calculate_cs_activity()` already restricts each sample's mean rate to its
usable replicates, so `F07_03_ambient`'s activity value is computed from
`A8`/`A9` only and its *usable*-replicate CV (6.4%) is what determines its
Interpretation. Its `CV flag` column still reads `FAIL >15%`, because that
column is deliberately computed on *all* raw replicates (including the dead
`A7`) as a stricter, unfiltered check -- both are correct, and they are
reporting different things (see TECHNICAL REPLICATE PRECISION above).

## Plot activity

```{r plot-activity, fig.width = 10, fig.height = 6}
# Three-level quality scheme. F07_03_ambient falls in the "CV > 15%" tier
# here (cv_all, unfiltered) even though its activity value itself is
# "usable" in the results table above (cv_rate, usable-replicates only).
qual_levels <- c("clean", paste0("CV > ", assay_params$cv_threshold_pct, "%"),
                 "all reps baseline-compromised")
qual_cols   <- setNames(c("#08519c", "#cb181d", "grey60"), qual_levels)

activity_plot_data <- cs_activity %>%
  left_join(cv_summary %>% select(sample_id, cv_all), by = "sample_id") %>%
  left_join(baseline_per_sample %>% select(sample_id, n_elevated, n), by = "sample_id") %>%
  mutate(cv_flag = factor(case_when(
                            n_elevated == n                              ~ qual_levels[3],
                            cv_all > assay_params$cv_threshold_pct       ~ qual_levels[2],
                            TRUE                                        ~ qual_levels[1]),
                          levels = qual_levels),
         label_n = paste0(n_reps_used, "/3 reps",
                          ifelse(n_elevated == n, " - DISQUALIFIED", "")))

activity_plot <- ggplot(activity_plot_data,
                        aes(x = reorder(sample_id, activity_mU_per_uL),
                            y = activity_mU_per_uL, fill = cv_flag)) +
  geom_col(width = 0.65) +
  geom_text(aes(label = label_n), hjust = -0.15, size = 3, colour = "grey25") +
  scale_fill_manual(values = qual_cols, drop = FALSE, name = "Well quality") +
  coord_flip(clip = "off") +
  expand_limits(y = max(activity_plot_data$activity_mU_per_uL) * 1.25) +
  labs(title = "Citrate synthase activity, M. gigas ctenidia, family F07 (ambient and 36°C)",
       subtitle = paste0("Not background-corrected (bg = 0 for every sample, note 4); mean of usable ",
                         "technical replicates (2/3 for F07_03_ambient, 3/3 for every other sample)."),
       x = NULL, y = "CS activity (mU/uL homogenate)") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        legend.position = "bottom")

ggsave(file.path(output_dir, "citrate_synthase_activity.png"), activity_plot,
       width = 10, height = 6, dpi = 300)

cat("--- activity_plot: ggplot object structure ---\n\n")
summary(activity_plot)

activity_plot
```

Raw activity (mU/µL) spans a real 2.5-fold range across the sixteen
individuals -- 0.0547 (`F07_01_ambient`) to 0.1356 (`F07_04_36C`) -- against
a usable-replicate technical CV under 6.5% for every sample, so this spread
reflects genuine inter-individual biological/extraction variation rather
than measurement noise. Because no background correction could be applied
(note 4), these values are the raw, uncorrected rates.

## Protein-normalized activity

```{r plot-activity-normalized, fig.width = 10, fig.height = 6}
# 15 of 16 samples are "clean" here; F07_03_ambient (CV > 15% on all raw
# reps, note above) is excluded from the trend fit.
norm_clean <- activity_plot_data %>% filter(cv_flag == qual_levels[1])

norm_plot <- ggplot(activity_plot_data,
                    aes(x = total_protein_mg, y = activity_mU_per_mg_protein)) +
  geom_smooth(data = norm_clean, method = "lm", se = TRUE,
              colour = "grey35", fill = "grey85", linewidth = 0.6) +
  geom_point(aes(colour = cv_flag), size = 3.5) +
  geom_text(aes(label = individual), vjust = -1, size = 3, colour = "grey25") +
  scale_colour_manual(values = qual_cols, drop = FALSE, name = "Well quality") +
  labs(title = "Protein-normalized CS activity vs. total extracted protein",
       subtitle = paste0("Point labels = individual ID. Total protein = measured protein ",
                         "concentration x the same\n", assay_params$homogenate_volume_uL,
                         " uL homogenization volume used for every sample. Grey line = fit to ",
                         "15 of 16\nsamples (F07_03_ambient excluded, CV > 15% on raw reps); a ",
                         "positive raw-rate/protein\nrelationship would suggest normalization is ",
                         "correcting for real extraction-efficiency\ndifferences rather than ",
                         "introducing an artifact."),
       x = "Total extracted protein (mg)", y = "CS activity (mU/mg protein)") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        legend.position = "bottom")

protein_trend <- cor.test(activity_plot_data$total_protein_mg, activity_plot_data$activity_mU_per_mg_protein)
protein_trend_clean <- cor.test(norm_clean$total_protein_mg, norm_clean$activity_mU_per_mg_protein)
uL_trend_clean <- cor.test(norm_clean$total_protein_mg, norm_clean$activity_mU_per_uL)

ggsave(file.path(output_dir, "citrate_synthase_activity_normalized.png"), norm_plot,
       width = 10, height = 6, dpi = 300)

cat("Correlation of mU/mg protein with total extracted protein\n")
cat(sprintf("  all %d samples      : r = %+.3f, p = %.4f\n",
            nrow(activity_plot_data), protein_trend$estimate, protein_trend$p.value))
cat(sprintf("  clean samples only  : r = %+.3f, p = %.4f  (n = %d)\n",
            protein_trend_clean$estimate, protein_trend_clean$p.value, nrow(norm_clean)))
cat(sprintf("  clean-sample mU/mg protein range: %.3f to %.3f (%.2f-fold)\n",
            min(norm_clean$activity_mU_per_mg_protein), max(norm_clean$activity_mU_per_mg_protein),
            max(norm_clean$activity_mU_per_mg_protein) / min(norm_clean$activity_mU_per_mg_protein)))
cat(sprintf("  clean-sample mU/uL range: %.4f to %.4f (%.2f-fold)\n",
            min(norm_clean$activity_mU_per_uL), max(norm_clean$activity_mU_per_uL),
            max(norm_clean$activity_mU_per_uL) / min(norm_clean$activity_mU_per_uL)))
cat(sprintf("\nArtifact check: r(total protein, mU/uL raw rate) = %+.3f, p = %.4f  (n = %d)\n",
            uL_trend_clean$estimate, uL_trend_clean$p.value, nrow(norm_clean)))
cat("If mU/uL itself trends with total protein while mU/mg protein trends oppositely,\n",
    "the mU/mg protein trend is partly/wholly the arithmetic of dividing by a covariate\n",
    "-- but a STRONG raw-rate/protein relationship (as seen here) instead argues that\n",
    "protein normalization is correcting for genuine extraction-efficiency differences.\n", sep = "")

cat("\n--- protein_trend_clean: correlation test on clean samples ---\n\n")
str(protein_trend_clean, max.level = 1, give.attr = FALSE)

cat("\n--- norm_clean: clean samples used for the trend fit ---\n\n")
str(norm_clean)

norm_plot
```

Raw activity (mU/µL) rises strongly and significantly with total extracted
protein (r = `r sprintf("%+.3f", uL_trend_clean$estimate)`,
p = `r sprintf("%.4f", uL_trend_clean$p.value)`), while mU/mg protein
correlates negatively with total extracted protein
(r = `r sprintf("%+.3f", protein_trend_clean$estimate)`,
p = `r sprintf("%.4f", protein_trend_clean$p.value)`, n = `r nrow(norm_clean)`,
excluding `F07_03_ambient`). Because raw signal genuinely increases with
more protein loaded -- consistent with more CS enzyme being present when
more protein was extracted -- **the negative mU/mg protein trend here is
more consistent with protein normalization doing its intended job**
(correcting for real inter-sample extraction-efficiency differences) than
with a pure division artifact. Normalizing here does not narrow the
inter-individual spread -- it is essentially unchanged, from a 2.48-fold
raw range to a 2.61-fold range in mU/mg protein -- so on this plate the
case for normalization rests on the raw-rate/protein correlation itself,
not on a visibly tighter normalized range.

# QC SUMMARY

```{r qc-summary}
qc_summary <- build_qc_summary_table(layout_wells, recon$summary, std_curve,
                                      well_diagnostics, well_rates, cv_summary,
                                      baseline_per_sample, pos_control, assay_params)

kable(qc_summary, col.names = c("QC check", "Result"),
      caption = "Quality control summary for this plate")

write.csv(qc_summary, file.path(output_dir, "qc_summary.csv"), row.names = FALSE)

cat("\n--- qc_summary: consolidated QC checks for the run ---\n\n")
str(qc_summary)
```

## Auto-generated findings

The bullets below are generated programmatically from the QC/CV/anomaly
objects above by `generate_qc_findings()`, so the same logic that flags a
problem here is guaranteed to be the logic that flagged it upstream.

```{r qc-findings}
qc_findings <- generate_qc_findings(well_diagnostics, well_rates, cv_summary,
                                     baseline_per_sample, std_curve, recon$summary,
                                     assay_params)

cat(paste0("- ", qc_findings, "\n"), sep = "\n")
```

# SUMMARY

## Anomalies found

1. **All three pooled-background replicates are compromised by a confirmed
   Reaction Mix / Background Control Mix mix-up (note 4).** `F10`-`F12` rise
   0.067-0.119 A412 over the run -- comparable to a real reaction, not a
   background -- so `compute_background_correction()` cannot estimate a
   background rate. Every sample's corrected rate on this plate equals its
   raw rate.
2. **One sample well is dead.** `A7` (`F07_03_ambient`) reads a constant
   A412 0.099 across all 11 timepoints (net change = 0, R<sup>2</sup> = 0)
   rather than a slow reaction. Its two siblings (`A8`, `A9`) are normal, so
   `F07_03_ambient`'s activity is computed from 2 of 3 replicates with a
   6.4% usable-replicate CV.
3. **Widespread but harmless GSH standard signal decay.** All 15
   non-zero-concentration standard wells lose signal between t = 0 and
   t = 20 min, consistent with TNB<sup>2-</sup> instability. Because the
   standard curve is calibrated at t = 0, this has no effect on the fit
   (R<sup>2</sup> = 0.9992, zero flagged outlier wells).
4. **The positive control was clean on this plate** (2.6% CV, R<sup>2</sup>
   0.998-1.000), confirming the core reaction chemistry and reader were
   functioning normally.

## Technical replicate precision

**Every sample is within the 15% CV threshold on usable replicates.**
`F07_03_ambient` is the only sample whose *all-replicates* CV fails (86.9%,
driven entirely by the dead `A7` well); its usable-replicate CV is a normal
6.4%. Every other sample is under 3.1% CV on all three raw replicates, most
under 2%.

## Assay validity

The **positive control and standard curve both worked well**: all three
positive-control replicates rise linearly (R<sup>2</sup> > 0.997) with a
tight 2.6% CV, and the GSH standard curve has zero flagged outlier wells
(R<sup>2</sup> = 0.9992). **The background control did not work** (note 4)
-- this plate's activity values are raw, uncorrected rates, not
background-corrected ones. Where a pooled background control has worked
correctly, it has been a small term (roughly 2-7% of the sample signal), so
the raw values here are likely only a modest overestimate of true
background-corrected activity, but this cannot be confirmed without a
working background control.

## Recommendation

1. **Samples, standard curve, and positive control do not need re-assay.**
   Only `A7` (`F07_03_ambient`) failed outright, and that sample still has
   two clean, precise replicates.
2. **The background control should be re-run before treating this plate's
   activity values as final.** Consider physically separating the Reaction
   Mix and Background Control Mix reservoirs/tips, or pipetting the
   background wells first, before switching reagents for the sample wells,
   to reduce the risk of this mix-up recurring.
3. **Protein normalization looks appropriate for this plate, not
   artifactual**: raw rate (mU/µL) is strongly and significantly correlated
   with total extracted protein (r = `r sprintf("%+.2f", uL_trend_clean$estimate)`,
   p = `r sprintf("%.4f", uL_trend_clean$p.value)`), so the negative
   mU/mg-protein trend is better explained by genuine extraction-efficiency
   differences between samples than by a division artifact.

## What can be used from this plate

15 of 16 samples are fully clean, and the 16th (`F07_03_ambient`) is still
usable via 2 of 3 replicates. **Because this plate's rates are
uncorrected**, they should not be pooled directly with background-corrected
values from other assays without accounting for that difference. Within
this plate, raw activity spans 0.0547-0.1356 mU/uL, a real 2.5-fold range
against <6.5% technical CV per sample.

- **Re-running just the background triplicate** (rather than the whole
  plate) would let this plate's existing sample, standard-curve, and
  positive-control data be background-corrected without re-collecting
  everything else.
- **Protein-normalized activity (mU/mg protein) is the recommended basis
  for downstream comparison on this plate**: raw activity correlates
  strongly with total extracted protein, consistent with normalization
  correcting a real extraction-efficiency covariate rather than
  introducing one.
