---
title: "Gen5-20260813-mgig-sormi-BSA-F05-F07-reassay-protein"
author: "Sam White"
date: "2026-08-13"
output:
  github_document:
    toc: true
    number_sections: true
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
---

# BACKGROUND

Protein quantification (BCA/Bradford-style, 595&nbsp;nm) of *Magallana gigas*
(Pacific oyster) SoRMI re-assay plate, 9 samples plus an 8-point BSA standard
curve, each measured in triplicate wells. All 9 samples are re-assays of
samples previously flagged for re-assay in the `F05` and `F07` analyses
(5 from `F05` -- CV or standard-curve-range failures -- and 4 from `F07` --
CV failures), now run together on a single plate against a fresh standard
curve.

The raw Gen5 export
(`Gen5-20260813-mgig-BSA-F05-F07-reassay-absorbance.csv`) has the same
per-well layout as the earlier `F05`/`F07` exports (see those documents for
detail), but with one new wrinkle: sample names carry a dilution-factor
suffix, `<sample>-df.<N>`, e.g. `F05_01_ambient-df.0` or
`F05_03_ambient-df.2`. `df.0` means the well was read on **undiluted**
material; `df.N` (N > 0) means the material was diluted **N-fold** before
reading, so the well's back-calculated concentration must be multiplied by
`N` to recover the sample's true (pre-dilution) concentration -- `df.0` is
therefore treated as a multiplier of 1, not 0. This multiplication is kept
as a distinct, explicit step (see "Applying the standard curve to samples"
below) so the well-level (as-measured) and final (dilution-corrected)
concentrations both remain visible.

As before, this document ignores the instrument's pre-computed summary
columns and recomputes mean, SD, and CV independently in R via `dplyr`, for
both the raw absorbance (595&nbsp;nm) and the concentration.

**CV QC metric used for exclusion:** the >15% technical-replicate CV
exclusion rule (including for standard-curve points) is applied to **CV of
concentration** (matching what the instrument's own `CV (%)` column already
represents for this export -- verified against the raw file: e.g. `SPL1`'s
given Mean/SD/CV of 883.966/50.075/5.665 are exactly the mean/SD/CV of its
three `[Concentration]` values, not of its `595` values, and this holds
regardless of that sample's dilution factor since the raw `[Concentration]`
column reflects the as-measured, pre-correction well concentration).

Two edge cases in the raw `[Concentration]` column need explicit handling:

- **Censored reads.** Wells reading outside the instrument's own internal
  standard-curve range are reported as `<0.000` or `>2100.000` rather than a
  number. These are parsed to a `_censored` flag plus the boundary value,
  and excluded from concentration mean/SD/CV calculations (their true value
  is unknown, not equal to the boundary).
- **The zero (blank) BSA standard** (`STD8`) is *always* censored
  (`<0.000` on all three wells), by construction -- a true-zero standard's
  back-calculated concentration is not a meaningful quantity to compute a
  CV on, and this is not a data-quality problem. `STD8` is therefore exempt
  from the concentration-CV exclusion rule and is retained as the curve's
  origin point on the strength of its (excellent) 595 CV instead.

Any other group left with fewer than 2 valid (non-censored) concentration
replicates has an undefined (`NA`) concentration CV and is treated as
failing QC, consistent with the source file itself being unable to compute
a CV there either (`?????`). In this plate's data, every sample well yields
a numeric (non-censored) concentration, so this case only arises for the
zero BSA standard (which is exempted, above).

# SETUP

## Libraries
```{r setup, include=TRUE}
library(knitr)
library(ggplot2)
library(dplyr)
library(tidyr)

knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,         # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = "",        # Prevents appending '##' to beginning of lines in code output
  results = 'hold'     # Holds output so it's all printed together after code chunk
)
```

## Set output directory
```{r output-dir}
output_dir <- "../outputs/Gen5-20260813-mgig-BSA-F05-F07-reassay"
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
```

# DATA IMPORT

The raw file is a Gen5 per-well export with non-syntactic column names
(`[Concentration]`, `CV (%)`, etc.) and a UTF-8 byte-order mark on the first
header cell, so columns are renamed by position immediately after import
rather than referenced by their original names.

```{r read-raw-data}
data_path <- "../data/BSA/raw_absorbance/Gen5-20260813-mgig-BSA-F05-F07-reassay-absorbance.csv"

protein_raw <- read.csv(data_path, stringsAsFactors = FALSE, check.names = FALSE,
                         na.strings = c("", "NA"))
colnames(protein_raw) <- c("well_id", "name", "well", "std_conc_known", "abs_595",
                           "gen5_concentration", "gen5_count", "gen5_mean",
                           "gen5_sd", "gen5_cv")

cat("--- protein_raw: as imported from the Gen5 export ---\n")
str(protein_raw)
```

# RESHAPING

`well_id` and `name` are only populated on the first row of each
triplicate group in the raw export, so they are filled downward. Each row
is also classified as a `standard` (`well_id` starts with `STD`) or a
`sample` (`SPL`), and the instrument's censored concentration strings
(`<0.000`, `>2100.000`) are parsed into a numeric value plus a
`gen5_concentration_censored` flag. For sample rows, the `-df.<N>` suffix is
parsed off `name` into a clean `base_sample_name`, plus `dilution_code` (the
raw `N`) and `dilution_multiplier` (1 when `N == 0`, otherwise `N`).

```{r reshape-long}
protein_long <- protein_raw %>%
  select(well_id, name, well, std_conc_known, abs_595, gen5_concentration) %>%
  fill(well_id, name, .direction = "down") %>%
  mutate(
    type = if_else(grepl("^STD", well_id), "standard", "sample"),
    gen5_concentration_censored = grepl("^[<>]", gen5_concentration),
    gen5_concentration_numeric  = as.numeric(gsub("[<>]", "", gen5_concentration)),
    base_sample_name = if_else(type == "sample",
                                sub("-df\\.[0-9]+$", "", name),
                                NA_character_),
    dilution_code = if_else(type == "sample",
                             as.numeric(sub("^.*-df\\.([0-9]+)$", "\\1", name)),
                             NA_real_),
    dilution_multiplier = if_else(type == "sample",
                                   if_else(dilution_code == 0, 1, dilution_code),
                                   NA_real_)
  )

cat("--- protein_long: one row per well, group identifiers filled down, dilution parsed ---\n")
str(protein_long)
```

# MEAN VALUES

## Mean 595 nm absorbance
```{r mean-595}
mean_595_by_group <- protein_long %>%
  group_by(well_id, name, type, std_conc_known,
           base_sample_name, dilution_code, dilution_multiplier) %>%
  summarise(n_wells  = n(),
            mean_595 = mean(abs_595),
            .groups  = "drop")

cat("--- mean_595_by_group ---\n")
str(mean_595_by_group)
```

## Mean concentration
Censored wells (reading outside the instrument's own standard-curve range)
are dropped before averaging, since their true concentration is unknown.
```{r mean-concentration}
mean_conc_by_group <- protein_long %>%
  filter(!gen5_concentration_censored) %>%
  group_by(well_id, name, type, std_conc_known,
           base_sample_name, dilution_code, dilution_multiplier) %>%
  summarise(n_wells_valid_conc  = n(),
            mean_concentration  = mean(gen5_concentration_numeric),
            .groups = "drop")

cat("--- mean_conc_by_group ---\n")
str(mean_conc_by_group)
```

# STANDARD DEVIATION AND COEFFICIENT OF VARIATION

## SD and CV of 595 nm absorbance
```{r sd-cv-595}
sd_cv_595_by_group <- protein_long %>%
  group_by(well_id, name, type, std_conc_known,
           base_sample_name, dilution_code, dilution_multiplier) %>%
  summarise(
    n_wells  = n(),
    mean_595 = mean(abs_595),
    sd_595   = sd(abs_595),
    cv_595   = 100 * sd_595 / mean_595,
    .groups  = "drop"
  )

cat("--- sd_cv_595_by_group ---\n")
str(sd_cv_595_by_group)
```

## SD and CV of concentration
```{r sd-cv-concentration}
sd_cv_conc_by_group <- protein_long %>%
  filter(!gen5_concentration_censored) %>%
  group_by(well_id, name, type, std_conc_known,
           base_sample_name, dilution_code, dilution_multiplier) %>%
  summarise(
    n_wells_valid_conc  = n(),
    mean_concentration  = mean(gen5_concentration_numeric),
    sd_concentration    = sd(gen5_concentration_numeric),
    cv_concentration    = 100 * sd_concentration / mean_concentration,
    .groups = "drop"
  )

cat("--- sd_cv_conc_by_group ---\n")
str(sd_cv_conc_by_group)
```

## Combined per-group summary
```{r combine-group-stats}
group_stats <- sd_cv_595_by_group %>%
  left_join(
    sd_cv_conc_by_group %>%
      select(well_id, mean_concentration, sd_concentration, cv_concentration, n_wells_valid_conc),
    by = "well_id"
  ) %>%
  arrange(type, well_id)

cat("--- group_stats: one row per sample/standard, 595- and concentration-level stats ---\n")
str(group_stats)
kable(group_stats, digits = 3)
```

# CV-BASED EXCLUSION

Groups with CV(concentration) > 15% -- or an undefined concentration CV due
to insufficient valid (non-censored) replicates -- are flagged and excluded
from all downstream analysis, including standard-curve fitting. The zero
BSA standard (`STD8`) is exempt from this rule (see Background) and is
always retained.

```{r cv-exclusion}
group_stats <- group_stats %>%
  mutate(
    is_zero_standard = type == "standard" & !is.na(std_conc_known) & std_conc_known == 0,
    high_cv = !is_zero_standard & (is.na(cv_concentration) | cv_concentration > 15)
  )

excluded_groups <- group_stats %>% filter(high_cv)

cat("--- excluded_groups: CV(concentration) > 15% (or undefined), dropped from further analysis ---\n")
str(excluded_groups)
kable(excluded_groups %>%
        select(well_id, name, type, std_conc_known, n_wells_valid_conc,
               mean_concentration, sd_concentration, cv_concentration),
      digits = 3)

clean_group_stats <- group_stats %>% filter(!high_cv)

cat("\nRetained after CV exclusion:", nrow(clean_group_stats), "of", nrow(group_stats), "groups\n")
str(clean_group_stats)
```

# STANDARD CURVE

Fit using only the retained (CV &le; 15%) standard points: known BSA
concentration (`Conc/Dil`) on the x-axis, mean 595&nbsp;nm absorbance on the
y-axis. Dilution correction (below) has no bearing on the curve fit itself,
since standards are not diluted.

```{r fit-standard-curve}
standards_clean <- clean_group_stats %>% filter(type == "standard")

cat("--- standards_clean: standard points retained for curve fitting ---\n")
str(standards_clean)

standard_curve_model <- lm(mean_595 ~ std_conc_known, data = standards_clean)

curve_slope     <- unname(coef(standard_curve_model)[2])
curve_intercept <- unname(coef(standard_curve_model)[1])
curve_r2        <- summary(standard_curve_model)$r.squared

cat("Standard curve fit: 595 =", format(curve_slope, digits = 6), "* concentration +",
    format(curve_intercept, digits = 4), "\n")
cat("R2:", round(curve_r2, 5), "\n")
```

## Plot standard curve
```{r plot-standard-curve, fig.width=7, fig.height=5.5}
eq_label <- paste0("y = ", format(curve_slope, digits = 4, scientific = TRUE), "x + ",
                    sprintf("%.4f", curve_intercept),
                    "\nR² = ", sprintf("%.4f", curve_r2))

standard_curve_plot <- ggplot(standards_clean, aes(x = std_conc_known, y = mean_595)) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "steelblue", linewidth = 0.6) +
  geom_errorbar(aes(ymin = mean_595 - sd_595, ymax = mean_595 + sd_595),
                width = 30, color = "grey40") +
  geom_point(size = 2.6, color = "steelblue") +
  annotate("text", x = -Inf, y = Inf, label = eq_label,
           hjust = -0.08, vjust = 1.3, size = 3.6, fontface = "bold") +
  labs(x = "BSA standard concentration (µg/mL)",
       y = "Mean absorbance (595 nm)",
       title = "BSA standard curve",
       subtitle = "Points with technical-replicate CV(concentration) > 15% excluded (see below)") +
  theme_bw(base_size = 12)

print(standard_curve_plot)
ggsave(file.path(output_dir, "standard_curve_595.png"), standard_curve_plot,
       width = 7, height = 5.5, dpi = 300)
```

Standard points excluded from the curve:
```{r excluded-standards}
excluded_standards <- group_stats %>% filter(type == "standard", high_cv)

cat("--- excluded_standards ---\n")
str(excluded_standards)
kable(excluded_standards %>%
        select(well_id, std_conc_known, mean_concentration, sd_concentration, cv_concentration),
      digits = 3)
```

# APPLYING THE STANDARD CURVE TO SAMPLES

Samples that passed CV QC are back-calculated against the fitted curve to
give a **well-level** concentration (`calculated_concentration_well`) --
the concentration of whatever material was actually in the well, diluted or
not. Multiplying by each sample's `dilution_multiplier` then gives the
**final** concentration (`calculated_concentration`) -- the true
concentration of the original, pre-dilution sample, which is the value
reported in the results below. Samples whose mean 595 falls outside the
retained standards' absorbance range are flagged as extrapolated; this
range check is done on the as-measured 595 value and is unaffected by
dilution.

```{r back-calculate-samples}
samples_clean <- clean_group_stats %>%
  filter(type == "sample") %>%
  mutate(
    se_595                         = sd_595 / sqrt(n_wells),
    calculated_concentration_well  = (mean_595 - curve_intercept) / curve_slope,
    calculated_concentration       = calculated_concentration_well * dilution_multiplier,
    in_curve_range                 = mean_595 >= min(standards_clean$mean_595) &
                                       mean_595 <= max(standards_clean$mean_595)
  ) %>%
  arrange(base_sample_name)

cat("--- samples_clean: CV-QC-passing samples, quantified against the fitted standard curve ---\n")
cat("    (calculated_concentration_well = as-measured; calculated_concentration = dilution-corrected)\n")
str(samples_clean)
```

```{r out-of-range-samples}
cat("Samples with absorbance outside the retained standard curve's range (extrapolated):\n")
kable(samples_clean %>% filter(!in_curve_range) %>%
        select(well_id, base_sample_name, mean_595, calculated_concentration),
      digits = 3)
```

## Plot samples on standard curve

Every QC-passing sample is plotted at its own mean 595 and **well-level**
back-calculated concentration (i.e. before dilution correction, since that
is what is directly comparable to the curve's own x/y relationship), with
vertical error bars of &plusmn;1 standard error (of mean 595, across its
2-3 retained technical replicates), overlaid on the standard curve fit line
and retained standard points.

```{r plot-samples-on-curve, fig.width=8, fig.height=5.5}
samples_on_curve_plot <- ggplot() +
  geom_smooth(data = standards_clean, aes(x = std_conc_known, y = mean_595),
              method = "lm", formula = y ~ x, se = FALSE,
              color = "steelblue", linewidth = 0.6) +
  geom_point(data = standards_clean, aes(x = std_conc_known, y = mean_595),
             size = 2.6, color = "steelblue") +
  geom_errorbar(data = samples_clean,
                aes(x = calculated_concentration_well,
                    ymin = mean_595 - se_595, ymax = mean_595 + se_595),
                width = 30, color = "grey40") +
  geom_point(data = samples_clean,
             aes(x = calculated_concentration_well, y = mean_595, shape = in_curve_range),
             size = 2.4, color = "firebrick", alpha = 0.85) +
  scale_shape_manual(values = c(`TRUE` = 16, `FALSE` = 4),
                      labels = c(`TRUE` = "in range", `FALSE` = "extrapolated"),
                      name = "Sample status") +
  labs(x = "Well-level concentration (µg/mL, pre-dilution-correction)",
       y = "Mean absorbance (595 nm)",
       title = "BSA standard curve with QC-passing samples overlaid",
       subtitle = "Samples plotted at their well-level back-calculated concentration; error bars = ±1 SE of mean 595") +
  theme_bw(base_size = 12)

print(samples_on_curve_plot)
ggsave(file.path(output_dir, "standard_curve_with_samples.png"), samples_on_curve_plot,
       width = 8, height = 5.5, dpi = 300)
```

# RESULTS TABLE

```{r results-table}
results_table <- samples_clean %>%
  transmute(
    Sample                                   = base_sample_name,
    `Well ID`                                = well_id,
    `Dilution factor`                        = dilution_multiplier,
    `Mean 595`                               = round(mean_595, 3),
    `CV 595 (%)`                             = round(cv_595, 2),
    `CV concentration (%)`                   = round(cv_concentration, 2),
    `Well concentration (ug/mL)`             = round(calculated_concentration_well, 1),
    `Calculated concentration (ug/mL)`       = round(calculated_concentration, 1),
    `In standard curve range`                = ifelse(in_curve_range, "yes", "NO - extrapolated")
  ) %>%
  arrange(gsub("[0-9]+$", "", `Well ID`), as.numeric(gsub("^[A-Za-z]+", "", `Well ID`)))

kable(results_table)
```

Only samples passing *all* QC -- CV &le; 15% and within the retained
standard curve's range (i.e. not extrapolated) -- are written to the CSV
output.

```{r results-table-export}
results_table_export <- results_table %>% filter(`In standard curve range` == "yes")

cat("--- results_table_export: samples passing all QC (CV and in-range), written to CSV ---\n")
str(results_table_export)

write.csv(results_table_export, file.path(output_dir, "sample_protein_concentrations.csv"), row.names = FALSE)
```

# SUMMARY

- `r nrow(group_stats)` sample/standard triplicate groups were imported
  (`r sum(group_stats$type == "standard")` BSA standards,
  `r sum(group_stats$type == "sample")` samples).
- `r nrow(excluded_groups)` group(s) exceeded 15% technical-replicate
  CV(concentration), or had an undefined CV from insufficient valid
  replicates, and were excluded from all downstream analysis:
  `r paste(paste0(excluded_groups$well_id, " (", ifelse(excluded_groups$type == "sample", excluded_groups$base_sample_name, excluded_groups$name), ")"), collapse = ", ")`.
- The BSA standard curve was fit on the `r nrow(standards_clean)` remaining
  standard points: slope = `r format(curve_slope, digits = 4)`,
  intercept = `r format(curve_intercept, digits = 4)`,
  R&sup2; = `r round(curve_r2, 4)`.
- `r sum(!samples_clean$in_curve_range)` CV-QC-passing sample(s) fell
  outside the retained standard curve's absorbance range and are flagged as
  extrapolated in the results table above.

## Samples passing QC

Sample names are sorted naturally (ignoring stray punctuation) rather than
by plain lexicographic string order. Concentrations shown are the final,
dilution-corrected values.

```{r natural-sort-key}
natural_sort_key <- function(x) gsub("[^A-Za-z0-9]", "", x)
```

```{r summary-passing-samples}
passing_samples_table <- samples_clean %>%
  transmute(sample_name = base_sample_name,
            `concentration(ug/mL)` = round(calculated_concentration, 1)) %>%
  arrange(natural_sort_key(sample_name))

cat("--- passing_samples_table: samples passing CV QC, name + calculated concentration ---\n")
str(passing_samples_table)

kable(passing_samples_table)
```

## Samples recommended for re-assay

```{r summary-qc-objects}
failed_qc_samples    <- excluded_groups %>% filter(type == "sample") %>% arrange(base_sample_name)
extrapolated_samples <- samples_clean %>% filter(!in_curve_range) %>% arrange(base_sample_name)

cat("--- failed_qc_samples: samples excluded from analysis on CV grounds ---\n")
str(failed_qc_samples)
cat("\n--- extrapolated_samples: QC-passing samples outside the retained standard curve's range ---\n")
str(extrapolated_samples)
```

```{r reassay-table}
reassay_candidates <- bind_rows(
  failed_qc_samples %>%
    transmute(`Sample name` = base_sample_name,
              `Rationale for re-assay` = as.character(ifelse(
                is.na(cv_concentration),
                "CV undefined (< 2 valid replicates)",
                "CV > 15%"
              ))),
  extrapolated_samples %>%
    transmute(`Sample name` = base_sample_name,
              `Rationale for re-assay` = "outside standard curve range")
) %>%
  arrange(natural_sort_key(`Sample name`))

cat("--- reassay_candidates: all samples flagged for re-assay, with rationale ---\n")
str(reassay_candidates)

if (nrow(reassay_candidates) == 0) {
  cat("None -- every sample on this plate passed all QC.\n")
} else {
  kable(reassay_candidates)
}
```
