---
title: "Gen5-20260825-mgig-sormi-citrate_synthase-F05-F07-temperature-comparison"
author: "Sam White"
date: "2026-08-25"
output:
  github_document:
    toc: true
    number_sections: true
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
---

# BACKGROUND

Cross-plate comparison of citrate synthase (CS) activity in ctenidia of
*Magallana gigas* (Pacific oyster) from **two families** (`F05`, `F07`) at
**two temperature exposures** (**ambient** and **36 &deg;C**), assayed with
the
[Abcam Citrate Synthase Assay Kit (ab239712), v4a](https://github.com/RobertsLab/resources/blob/master/protocols/Commercial_Protocols/ABCAM-Citrate-Synthase-Assay-v4a-ab239712.pdf).

The question here is **not** how to compute activity -- that was done, plate
by plate, in the per-plate analyses listed below. This document takes those
finished per-plate results as its input and asks:

1. Does CS activity respond to 36 &deg;C exposure?
2. **Does that temperature response differ between families?** -- i.e. is
   there a family &times; temperature interaction?

All four family &times; temperature groups carry **n = 8** individuals, giving a
fully balanced 2 &times; 2 design (32 individuals total).

## Inputs: which plates are used

This analysis reads the `citrate_synthase_activity_results.csv` output of
**three** per-plate analyses:

| Family | Temperature | Source analysis | Plate date |
|---|---|---|---|
| `F05` | ambient | `Gen5-20260824-mgig-sormi-citrate_synthase-F05-ambient` | 2026-08-24 |
| `F05` | 36 &deg;C | `Gen5-20260814-mgig-citrate_synthase-F05-36C` | 2026-08-14 |
| `F07` | ambient + 36 &deg;C | `Gen5-20260824-mgig-sormi-citrate_synthase-F07-ambient_and_36C` | 2026-08-24 |

**The 2026-08-11 `F05` ambient plate is deliberately excluded** (note 1).

## Notes

1. **The 2026-08-11 `F05`-ambient plate is excluded from this analysis.**
   That plate required repeating because of poor data quality across both
   its samples and its standard curve (standard-curve R<sup>2</sup> of
   0.9678 after excluding 8 of 18 standard wells as outliers, 4 of 8 samples
   over the 15% technical-CV threshold, and one sample with all three
   replicates baseline-compromised). Those same `F05` ambient homogenates
   were re-assayed on 2026-08-24, and **the 2026-08-24 re-assay is the
   `F05` ambient data used here.**
2. **Background luminescence is NOT subtracted from any activity value in
   this document.** One retained plate
   (`Gen5-20260824-mgig-sormi-citrate_synthase-F07-ambient_and_36C`) has no
   usable background estimate -- Reaction Mix was added to all three of its
   pooled-background wells instead of Background Control Mix, so its
   background rate is unestimable and its per-plate analysis already fell
   back to a correction of 0. Rather than compare background-corrected
   values on two plates against uncorrected values on a third, **the
   background correction is removed from all plates** so every value is
   treated identically. This is a small change: where background *was*
   measurable it never exceeded ~10% of the raw rate (quantified in
   [BACKGROUND REMOVAL](#background-removal)).
3. **Activity is normalized to total extracted protein** (mU per mg
   protein), carried through from the per-plate analyses, which measured
   protein directly by BCA/Bradford-style assay on 2026-08-13. Tissue
   weight is not used.
4. **The `F05` temperature contrast spans two plates and two days**
   (36 &deg;C on 2026-08-14, ambient on 2026-08-24), whereas the `F07`
   temperature contrast sits **within a single plate** (2026-08-24). Plate
   and instrument state are compared across plates in
   [PLATE PROVENANCE](#plate-provenance).
5. **Individuals are treated as unpaired across temperature.** An animal is
   assayed at one exposure only; `F05_01_ambient` and `F05_01_36C` are
   different individuals from the same family, so no paired/repeated-measures
   structure is used.
6. **Statistics are computed on log<sub>10</sub>(activity).** Activity is a
   ratio-scale rate spanning roughly an order of magnitude across groups;
   the log scale makes the family &times; temperature interaction a test of
   *multiplicative* (fold-change) response, which is the biologically
   meaningful comparison for an enzyme rate, and it stabilizes the
   group-to-group spread in variance. Because the design is balanced,
   sequential (Type I) sums of squares from `aov()` are identical to Type II
   and Type III, so no special contrast handling is needed.

```{r setup, include = TRUE}
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)

knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,         # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = "",        # Prevents appending '##' to beginning of lines in code output
  results = 'hold'     # Holds output so it's all printed together after code chunk
)
```

## Analysis parameters

```{r analysis-params}
analysis_params <- list(
  subtract_background = FALSE,  # note 2: background is NOT subtracted anywhere
  alpha               = 0.05,   # significance level for ANOVA / Tukey
  cv_threshold_pct    = 15      # technical-CV threshold used by the per-plate analyses
)

cat("--- analysis_params: settings governing this comparison ---\n")
str(analysis_params)
```

## Input paths

```{r paths}
outputs_root <- "../outputs"

# Each retained per-plate analysis and the family/temperature it contributes.
plate_inputs <- tribble(
  ~plate_dir,                                                       ~contributes,
  "Gen5-20260824-mgig-sormi-citrate_synthase-F05-ambient",          "F05 ambient",
  "Gen5-20260814-mgig-citrate_synthase-F05-36C",                    "F05 36C",
  "Gen5-20260824-mgig-sormi-citrate_synthase-F07-ambient_and_36C",  "F07 ambient + 36C"
) %>%
  mutate(results_csv = file.path(outputs_root, plate_dir,
                                 "citrate_synthase_activity_results.csv"))

output_dir <- file.path(
  outputs_root,
  "Gen5-20260825-mgig-sormi-citrate_synthase-F05-F07-temperature-comparison")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

cat("--- plate_inputs: per-plate results files feeding this analysis ---\n")
print(as.data.frame(plate_inputs[, c("plate_dir", "contributes")]))
cat("\n--- all input files exist? ---\n")
print(setNames(file.exists(plate_inputs$results_csv), plate_inputs$plate_dir))
cat("\n--- output_dir ---\n")
str(output_dir)
```

## Plate provenance {#plate-provenance}

Per-plate calibration and instrument-state values, transcribed from the
three source analyses. These are the anchors that make cross-plate
comparison defensible: if the GSH standard curve and the CS positive control
agree across plates, then a rate measured on one plate means the same thing
as a rate measured on another.

```{r plate-provenance}
plate_provenance <- tribble(
  ~plate_dir,                                                      ~plate_date,  ~std_slope_A412_per_nmol, ~std_r2, ~pos_control_mA412_min, ~pos_control_cv_pct, ~background_usable,
  "Gen5-20260824-mgig-sormi-citrate_synthase-F05-ambient",         "2026-08-24", 0.03234,                  0.9983,  2.700,                  22.8,                TRUE,
  "Gen5-20260814-mgig-citrate_synthase-F05-36C",                   "2026-08-14", 0.03075,                  0.9985,  2.933,                  3.9,                 TRUE,
  "Gen5-20260824-mgig-sormi-citrate_synthase-F07-ambient_and_36C", "2026-08-24", 0.03140,                  0.9992,  2.933,                  2.6,                 FALSE
)

cat("--- plate_provenance: calibration + instrument state per plate ---\n")
str(as.data.frame(plate_provenance))
```

```{r plate-provenance-table}
plate_provenance %>%
  transmute(
    Plate            = plate_dir,
    Date             = plate_date,
    `Std slope (A412/nmol)` = std_slope_A412_per_nmol,
    `Std curve R^2`  = std_r2,
    `Pos ctrl (mA412/min)` = pos_control_mA412_min,
    `Pos ctrl CV (%)` = pos_control_cv_pct,
    `Background estimable` = ifelse(background_usable, "yes", "no (note 2)")
  ) %>%
  kable(caption = "Calibration and instrument state for each retained plate.")
```

```{r plate-provenance-spread}
prov_spread <- plate_provenance %>%
  summarise(
    slope_min   = min(std_slope_A412_per_nmol),
    slope_max   = max(std_slope_A412_per_nmol),
    slope_cv    = 100 * sd(std_slope_A412_per_nmol) / mean(std_slope_A412_per_nmol),
    pos_min     = min(pos_control_mA412_min),
    pos_max     = max(pos_control_mA412_min),
    pos_cv      = 100 * sd(pos_control_mA412_min) / mean(pos_control_mA412_min),
    min_std_r2  = min(std_r2)
  )

cat(sprintf(
  "Standard-curve slope across plates: %.5f-%.5f A412/nmol (CV %.1f%%)\n",
  prov_spread$slope_min, prov_spread$slope_max, prov_spread$slope_cv))
cat(sprintf(
  "Positive control across plates:     %.3f-%.3f mA412/min (CV %.1f%%)\n",
  prov_spread$pos_min, prov_spread$pos_max, prov_spread$pos_cv))
cat(sprintf("Worst standard-curve R^2:           %.4f\n", prov_spread$min_std_r2))
```

# LOAD PER-PLATE RESULTS

## Read the results CSVs

```{r read-results}
read_plate_results <- function(results_csv, plate_dir) {
  read.csv(results_csv, check.names = FALSE, stringsAsFactors = FALSE) %>%
    mutate(plate_dir = plate_dir)
}

plate_results_raw <- Map(read_plate_results,
                         plate_inputs$results_csv,
                         plate_inputs$plate_dir) %>%
  bind_rows()

cat("--- plate_results_raw: all per-plate result rows, stacked ---\n")
str(as.data.frame(plate_results_raw))
```

```{r read-results-check}
cat("Rows read per plate:\n")
print(table(plate_results_raw$plate_dir))
cat("\nColumns present in every input file:\n")
print(names(plate_results_raw))
```

## Keep only the columns this analysis needs

The per-plate results tables carry QC columns alongside the activity
values. This analysis needs the identity of each individual, its protein
normalization, both the raw and background-corrected rate (to undo the
background subtraction -- note 2), the published activity, and the QC verdict.

```{r select-columns}
cs_all <- plate_results_raw %>%
  transmute(
    sample_id        = Sample,
    family           = Family,
    temperature      = Temperature,
    individual       = sprintf("%02d", as.integer(Individual)),
    plate_dir        = plate_dir,
    protein_mg       = `Total protein (mg)`,
    reps_used        = `Reps used`,
    cv_used_pct      = `CV used reps (%)`,
    rate_raw         = `Rate (mA412/min)`,
    rate_bg          = `BG rate (mA412/min)`,
    rate_corrected   = `Corrected rate (mA412/min)`,
    activity_published = `Activity (mU/mg protein)`,
    cv_flag          = `CV flag`,
    interpretation   = Interpretation
  ) %>%
  left_join(plate_provenance %>% select(plate_dir, plate_date, std_slope_A412_per_nmol),
            by = "plate_dir") %>%
  arrange(family, temperature, individual)

cat("--- cs_all: one row per individual, columns needed downstream ---\n")
str(as.data.frame(cs_all))
```

## Verify the design is complete and balanced

```{r check-design}
design_counts <- cs_all %>%
  count(family, temperature, name = "n_individuals")

cat("--- design_counts: individuals per family x temperature cell ---\n")
print(as.data.frame(design_counts))

cat(sprintf("\nTotal individuals: %d\n", nrow(cs_all)))
cat(sprintf("Duplicated sample_ids: %d\n", sum(duplicated(cs_all$sample_id))))
cat(sprintf("Rows with missing protein_mg: %d\n", sum(is.na(cs_all$protein_mg))))
cat(sprintf("Rows with missing raw rate: %d\n", sum(is.na(cs_all$rate_raw))))
cat(sprintf("Design balanced (all cells equal n): %s\n",
            length(unique(design_counts$n_individuals)) == 1))
```

# BACKGROUND REMOVAL {#background-removal}

## Why the correction is removed

Per note 2, the `F07` plate has no estimable background, so its published
"corrected" rate already equals its raw rate. The two `F05` plates *do*
carry a background subtraction. Comparing corrected `F05` values against
uncorrected `F07` values would put a systematic, plate-dependent offset
directly into the family contrast -- so instead the correction is stripped
from every plate.

## How much does the correction actually matter?

```{r background-magnitude}
bg_impact <- cs_all %>%
  mutate(bg_pct_of_raw = 100 * rate_bg / rate_raw) %>%
  group_by(plate_dir) %>%
  summarise(
    n              = n(),
    bg_rate_min    = min(rate_bg),
    bg_rate_max    = max(rate_bg),
    bg_pct_min     = min(bg_pct_of_raw),
    bg_pct_max     = max(bg_pct_of_raw),
    .groups = "drop"
  )

cat("--- bg_impact: size of the background correction, per plate ---\n")
print(as.data.frame(bg_impact))
```

```{r background-magnitude-table}
bg_impact %>%
  transmute(
    Plate = plate_dir,
    n     = n,
    `BG rate (mA412/min)` = sprintf("%.3f - %.3f", bg_rate_min, bg_rate_max),
    `BG as % of raw rate` = sprintf("%.1f - %.1f", bg_pct_min, bg_pct_max)
  ) %>%
  kable(caption = paste("Magnitude of the background correction being removed.",
                        "The F07 plate has no estimable background (note 2),",
                        "so its correction is already zero."))
```

## Recompute activity without background subtraction

Activity is **linear** in the background-corrected rate: every other term in
the Abcam calculation (standard-curve slope, sample volume, dilution factor,
homogenate volume, total protein) is a per-sample constant. So the
uncorrected activity is recovered exactly by rescaling the published
activity by `rate_raw / rate_corrected`, with no need to re-derive anything
from the raw kinetic traces.

```{r remove-background}
stopifnot(!analysis_params$subtract_background)

cs_activity <- cs_all %>%
  mutate(
    # Scale factor undoing the background subtraction; exactly 1 where the
    # per-plate analysis already applied no correction (F07 plate).
    bg_undo_factor  = rate_raw / rate_corrected,
    activity_nobg   = activity_published * bg_undo_factor,
    log10_activity  = log10(activity_nobg)
  )

cat("--- cs_activity: activity with background subtraction removed ---\n")
str(as.data.frame(cs_activity))
```

## Independently verify the rescaling

The rescaling above is checked against a from-scratch recomputation of
activity out of the raw rate and each plate's own standard-curve slope,
using the Abcam formula
(`nmol/min = (rate/1000) / std_slope`; `mU/uL = nmol/min / sample_volume * D`;
`mU/mg = mU/uL * homogenate_volume / total_protein_mg`). The two paths must
agree to within the rounding of the input CSVs.

```{r verify-rescaling}
# Assay constants shared by all three plates (identical in all source analyses).
sample_volume_uL     <- 2
dilution_factor      <- 1
homogenate_volume_uL <- 350

verify <- cs_activity %>%
  mutate(
    activity_recomputed =
      ((rate_raw / 1000) / std_slope_A412_per_nmol / sample_volume_uL *
         dilution_factor) * homogenate_volume_uL / protein_mg,
    pct_diff = 100 * abs(activity_recomputed / activity_nobg - 1)
  )

cat(sprintf("Max %% difference, rescaled vs recomputed-from-raw: %.3f%%\n",
            max(verify$pct_diff)))
cat(sprintf("All within 1%% (input CSV rounding): %s\n", max(verify$pct_diff) < 1))
```

```{r background-effect-summary}
cat(sprintf(
  "Removing the background correction changes activity by %.1f%%-%.1f%% (median %.1f%%).\n",
  100 * (min(cs_activity$bg_undo_factor) - 1),
  100 * (max(cs_activity$bg_undo_factor) - 1),
  100 * (median(cs_activity$bg_undo_factor) - 1)))
```

# ANALYSIS DATA SET

## Set factor levels

```{r factor-levels}
# Group levels are ordered family-major so that, on a single x axis, each
# family's ambient and 36C boxes sit side by side.
group_levels <- c("F05 ambient", "F05 36C", "F07 ambient", "F07 36C")

cs_analysis <- cs_activity %>%
  mutate(
    family      = factor(family, levels = c("F05", "F07")),
    temperature = factor(temperature, levels = c("ambient", "36C")),
    group       = factor(paste(family, temperature), levels = group_levels),
    plate_dir   = factor(plate_dir)
  )

cat("--- cs_analysis: analysis-ready data with ordered factors ---\n")
cat("family levels:     ", paste(levels(cs_analysis$family), collapse = ", "), "\n")
cat("temperature levels:", paste(levels(cs_analysis$temperature), collapse = ", "), "\n")
cat("group levels:      ", paste(levels(cs_analysis$group), collapse = " | "), "\n")
cat("n rows:            ", nrow(cs_analysis), "\n")
```

## Colour palette

Colours are drawn from the **Okabe-Ito** qualitative palette, which is
designed to remain distinguishable under deuteranopia, protanopia and
tritanopia (Okabe & Ito 2008, *Color Universal Design*). One hue is bound to
each family and reused in every figure in this document, so colour alone
identifies the family throughout. Temperature is additionally encoded by
point shape and by a light/dark tint of the family hue, so **no figure relies
on colour as its only channel** -- each remains readable in greyscale and to
a colourblind reader.

```{r plot-settings}
# Okabe-Ito colourblind-safe qualitative palette.
okabe_ito <- c(
  blue          = "#0072B2",
  vermillion    = "#D55E00",
  bluishgreen   = "#009E73",
  orange        = "#E69F00",
  skyblue       = "#56B4E9",
  reddishpurple = "#CC79A7",
  yellow        = "#F0E442",
  black         = "#000000"
)

# Family identity: one hue per family.
family_colours <- c(F05 = okabe_ito[["blue"]], F07 = okabe_ito[["vermillion"]])

# Group identity: each family keeps its hue, with the 36C member of the pair
# as the lighter tint, so family and temperature are both readable from fill.
group_colours <- c(
  "F05 ambient" = okabe_ito[["blue"]],
  "F05 36C"     = okabe_ito[["skyblue"]],
  "F07 ambient" = okabe_ito[["vermillion"]],
  "F07 36C"     = okabe_ito[["orange"]]
)

stopifnot(all(levels(cs_analysis$group) %in% names(group_colours)))

cat("--- family_colours: hue bound to each family in every figure ---\n")
print(family_colours)
cat("\n--- group_colours: fill bound to each family x temperature group ---\n")
print(group_colours)
```

## QC status of the analysis set

Every individual is retained. The per-plate analyses already excluded
unusable technical replicates before averaging; the flags below describe
which individuals had a replicate dropped or a wide replicate CV, so any
group difference can be checked against QC status rather than silently
inheriting it.

```{r qc-status}
qc_status <- cs_analysis %>%
  group_by(family, temperature) %>%
  summarise(
    n                   = n(),
    n_full_triplicate   = sum(reps_used == "3/3"),
    n_cv_flagged        = sum(cv_flag != "pass"),
    n_not_usable        = sum(interpretation != "usable"),
    max_cv_used_pct     = max(cv_used_pct),
    .groups = "drop"
  )

cat("--- qc_status: QC composition of each group ---\n")
print(as.data.frame(qc_status))
```

```{r qc-status-table}
qc_status %>%
  transmute(
    Family = family, Temperature = temperature, n = n,
    `Full 3/3 reps` = n_full_triplicate,
    `CV-flagged`    = n_cv_flagged,
    `Not usable`    = n_not_usable,
    `Worst CV of used reps (%)` = max_cv_used_pct
  ) %>%
  kable(caption = paste0("QC composition per group. CV-flagged counts",
                         " individuals whose CV across ALL replicates exceeded ",
                         analysis_params$cv_threshold_pct,
                         "%; such individuals still contribute a mean over",
                         " their retained replicates."))
```

# DESCRIPTIVE STATISTICS

## Per-group summary

```{r group-summary}
group_summary <- cs_analysis %>%
  group_by(family, temperature) %>%
  summarise(
    n        = n(),
    mean_act = mean(activity_nobg),
    sd_act   = sd(activity_nobg),
    cv_act   = 100 * sd(activity_nobg) / mean(activity_nobg),
    median_act = median(activity_nobg),
    min_act  = min(activity_nobg),
    max_act  = max(activity_nobg),
    mean_rate_raw = mean(rate_raw),
    sd_rate_raw   = sd(rate_raw),
    .groups  = "drop"
  )

cat("--- group_summary: activity (mU/mg protein), background NOT subtracted ---\n")
print(as.data.frame(group_summary))
```

```{r group-summary-table}
group_summary %>%
  transmute(
    Family = family, Temperature = temperature, n = n,
    `Mean (mU/mg)`   = round(mean_act, 1),
    `SD`             = round(sd_act, 1),
    `CV (%)`         = round(cv_act, 1),
    `Median`         = round(median_act, 1),
    `Range`          = sprintf("%.1f - %.1f", min_act, max_act),
    `Raw rate (mA412/min)` = sprintf("%.2f +/- %.2f", mean_rate_raw, sd_rate_raw)
  ) %>%
  kable(caption = paste("Citrate synthase activity per family x temperature",
                        "group, with background subtraction removed (note 2)."))
```

## Temperature effect within each family

```{r within-family-effect}
temp_effect <- group_summary %>%
  select(family, temperature, mean_act) %>%
  pivot_wider(names_from = temperature, values_from = mean_act) %>%
  mutate(
    abs_change    = `36C` - ambient,
    fold_change   = `36C` / ambient,
    pct_change    = 100 * (fold_change - 1)
  )

cat("--- temp_effect: ambient -> 36C change in mean activity, per family ---\n")
print(as.data.frame(temp_effect))
```

```{r within-family-effect-table}
temp_effect %>%
  transmute(
    Family = family,
    `Ambient mean (mU/mg)` = round(ambient, 1),
    `36C mean (mU/mg)`     = round(`36C`, 1),
    `Change (mU/mg)`       = round(abs_change, 1),
    `Fold change`          = round(fold_change, 2),
    `% change`             = round(pct_change, 1)
  ) %>%
  kable(caption = "Direction and size of the 36C response within each family.")
```

## Individual-level table

```{r individual-table}
individual_table <- cs_analysis %>%
  arrange(family, temperature, individual) %>%
  transmute(
    Sample = sample_id, Family = family, Individual = individual,
    Temperature = temperature,
    `Protein (mg)` = protein_mg,
    `Raw rate (mA412/min)` = rate_raw,
    `Activity, no BG (mU/mg)` = round(activity_nobg, 1),
    `Activity, published (mU/mg)` = activity_published,
    `Reps used` = reps_used,
    `CV used (%)` = cv_used_pct,
    `QC` = interpretation
  )

write.csv(individual_table,
          file.path(output_dir, "cs_activity_all_families_no_background.csv"),
          row.names = FALSE)

kable(individual_table,
      caption = paste("All 32 individuals with background subtraction",
                      "removed, alongside the originally published",
                      "background-corrected value for reference."))
```

# STATISTICS

## Two-way ANOVA on log10 activity

The model is `log10(activity) ~ family * temperature`. The **interaction
term is the formal test of the question** in this document: does the
temperature response differ between families? Because the design is balanced
(8 per cell), sequential sums of squares are unambiguous.

```{r anova-fit}
cs_aov <- aov(log10_activity ~ family * temperature, data = cs_analysis)

cat("--- cs_aov: two-way factorial ANOVA on log10 activity ---\n")
print(summary(cs_aov))
```

```{r anova-table}
aov_tidy <- as.data.frame(summary(cs_aov)[[1]])
aov_tidy$Term <- trimws(rownames(aov_tidy))

anova_table <- aov_tidy %>%
  transmute(
    Term    = Term,
    Df      = Df,
    `Sum Sq`  = round(`Sum Sq`, 4),
    `Mean Sq` = round(`Mean Sq`, 4),
    `F value` = ifelse(is.na(`F value`), NA, round(`F value`, 3)),
    `p value` = ifelse(is.na(`Pr(>F)`), NA, signif(`Pr(>F)`, 4)),
    Significant = ifelse(is.na(`Pr(>F)`), "",
                         ifelse(`Pr(>F)` < analysis_params$alpha, "yes", "no"))
  )

kable(anova_table, row.names = FALSE,
      caption = sprintf(paste("Two-way ANOVA of log10 CS activity.",
                              "The family:temperature row is the test of",
                              "whether families differ in temperature",
                              "response (alpha = %.2f)."),
                        analysis_params$alpha))
```

```{r anova-effects}
p_family      <- aov_tidy$`Pr(>F)`[aov_tidy$Term == "family"]
p_temperature <- aov_tidy$`Pr(>F)`[aov_tidy$Term == "temperature"]
p_interaction <- aov_tidy$`Pr(>F)`[aov_tidy$Term == "family:temperature"]

cat(sprintf("family effect:              p = %.5g\n", p_family))
cat(sprintf("temperature effect:         p = %.5g\n", p_temperature))
cat(sprintf("family x temperature:       p = %.5g\n", p_interaction))
cat(sprintf("\nInteraction significant at alpha = %.2f: %s\n",
            analysis_params$alpha, p_interaction < analysis_params$alpha))
```

## Variance explained

```{r effect-sizes}
ss <- aov_tidy$`Sum Sq`
names(ss) <- aov_tidy$Term
ss_total <- sum(ss)

eta_sq <- tibble(
  Term          = names(ss),
  `Sum Sq`      = round(ss, 4),
  `% of total variance` = round(100 * ss / ss_total, 1)
)

cat("--- eta_sq: partition of variance in log10 activity ---\n")
print(as.data.frame(eta_sq))
```

```{r effect-sizes-table}
kable(eta_sq, caption = paste("Partition of total variance in log10 activity",
                              "(eta-squared). Residuals represent",
                              "between-individual variation within groups."))
```

## Residual diagnostics

```{r residual-diagnostics}
resid_shapiro <- shapiro.test(residuals(cs_aov))
resid_bartlett <- bartlett.test(log10_activity ~ group, data = cs_analysis)

cat(sprintf("Shapiro-Wilk on residuals:  W = %.4f, p = %.4g\n",
            resid_shapiro$statistic, resid_shapiro$p.value))
cat(sprintf("Bartlett homogeneity:       K^2 = %.4f, p = %.4g\n",
            resid_bartlett$statistic, resid_bartlett$p.value))
cat(sprintf("\nResiduals consistent with normality (p > %.2f): %s\n",
            analysis_params$alpha, resid_shapiro$p.value > analysis_params$alpha))
cat(sprintf("Group variances homogeneous (p > %.2f): %s\n",
            analysis_params$alpha, resid_bartlett$p.value > analysis_params$alpha))
```

```{r plot-diagnostics, fig.width = 9, fig.height = 4.5}
diag_df <- tibble(
  fitted   = fitted(cs_aov),
  residual = residuals(cs_aov),
  group    = cs_analysis$group
)

resid_fitted_plot <- ggplot(diag_df,
                            aes(x = fitted, y = residual,
                                colour = group, shape = group)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(size = 2.5, alpha = 0.9) +
  scale_colour_manual(values = group_colours, name = "Group") +
  scale_shape_manual(values = c(16, 17, 15, 18), name = "Group") +
  labs(title = "Residuals vs fitted values",
       subtitle = "Two-way ANOVA on log10 CS activity",
       x = "Fitted log10 activity", y = "Residual") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"))

qq_df <- diag_df %>%
  arrange(residual) %>%
  mutate(theoretical = qnorm(ppoints(n())))

qq_plot <- ggplot(qq_df, aes(x = theoretical, y = residual)) +
  geom_abline(slope = sd(qq_df$residual), intercept = mean(qq_df$residual),
              linetype = "dashed", colour = "grey40") +
  geom_point(aes(colour = group, shape = group), size = 2.5, alpha = 0.9) +
  scale_colour_manual(values = group_colours, name = "Group") +
  scale_shape_manual(values = c(16, 17, 15, 18), name = "Group") +
  labs(title = "Normal Q-Q plot of residuals",
       subtitle = sprintf("Shapiro-Wilk p = %.3g", resid_shapiro$p.value),
       x = "Theoretical quantile", y = "Residual") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"))

ggsave(file.path(output_dir, "anova_residuals_vs_fitted.png"), resid_fitted_plot,
       width = 9, height = 4.5, dpi = 300)
ggsave(file.path(output_dir, "anova_qq_residuals.png"), qq_plot,
       width = 9, height = 4.5, dpi = 300)

print(resid_fitted_plot)
print(qq_plot)
```

## Tukey HSD pairwise contrasts

```{r tukey}
tukey_res <- TukeyHSD(cs_aov, which = "family:temperature")

cat("--- tukey_res: all pairwise group contrasts on log10 activity ---\n")
print(tukey_res)
```

```{r tukey-table}
tukey_df <- as.data.frame(tukey_res$`family:temperature`)
tukey_df$Contrast <- rownames(tukey_df)

tukey_table <- tukey_df %>%
  transmute(
    Contrast = Contrast,
    `Diff (log10)`  = round(diff, 4),
    `Fold change`   = round(10^diff, 2),
    `95% CI (fold)` = sprintf("%.2f - %.2f", 10^lwr, 10^upr),
    `p adj`         = signif(`p adj`, 4),
    Significant     = ifelse(`p adj` < analysis_params$alpha, "yes", "no")
  ) %>%
  arrange(`p adj`)

write.csv(tukey_table, file.path(output_dir, "tukey_pairwise_contrasts.csv"),
          row.names = FALSE)

kable(tukey_table, row.names = FALSE,
      caption = paste("Tukey HSD contrasts between all four groups.",
                      "Differences on the log10 scale are back-transformed to",
                      "fold changes; a fold change of 1 means no difference."))
```

## The two within-family temperature contrasts

Pulled out of the Tukey table because these two rows, read together, are the
interaction: how each family responds to 36 &deg;C.

```{r within-family-contrasts}
# TukeyHSD labels interaction levels with a ":" separator, e.g. "F05:36C".
within_family <- tukey_table %>%
  filter(Contrast %in% c("F05:36C-F05:ambient", "F07:36C-F07:ambient"))

stopifnot(nrow(within_family) == 2)

cat("--- within_family: the ambient vs 36C contrast inside each family ---\n")
print(as.data.frame(within_family))
```

```{r within-family-contrasts-table}
kable(within_family, row.names = FALSE,
      caption = "Within-family ambient vs 36C contrasts (Tukey-adjusted).")
```

# FIGURES

## Activity by family and temperature

Box plots of all four groups on a single axis, with every individual
overlaid (n = 8 per group, so the underlying points are worth showing rather
than summarising away). Boxes are the standard Tukey construction: the middle
line is the median, the box spans the interquartile range, and the whiskers
reach the most extreme point within 1.5 &times; IQR of the box.

### Significance notation

The brackets are built **from the Tukey HSD table computed above**, not
entered by hand, so the annotation cannot drift from the statistics.

Four of the six pairwise contrasts are annotated: the two **within-family**
temperature contrasts and the two **within-temperature** family contrasts --
i.e. every pair differing in exactly one factor. The two remaining
"diagonal" contrasts (`F05 ambient` vs `F07 36C`, and `F05 36C` vs
`F07 ambient`) change family *and* temperature at once and so cannot be
attributed to either factor; they are not annotated. This is not selection
by outcome -- one omitted diagonal is significant and the other is not -- and
**all six contrasts remain in the Tukey table and in
`tukey_pairwise_contrasts.csv`**. Non-significant retained contrasts are
labelled `ns` rather than left blank.

```{r significance-brackets}
# Star notation used on the plot.
sig_stars <- function(p) {
  ifelse(p < 0.001, "***",
  ifelse(p < 0.01,  "**",
  ifelse(p < 0.05,  "*", "ns")))
}

# x positions of each group on the plot's discrete axis.
group_x <- setNames(seq_along(group_levels), group_levels)

# TukeyHSD row names are "<groupB>-<groupA>" with ":" between factor levels.
sig_brackets <- tukey_df %>%
  transmute(
    contrast = Contrast,
    p_adj    = `p adj`,
    g2       = gsub(":", " ", sub("-.*$", "", Contrast)),
    g1       = gsub(":", " ", sub("^.*-", "", Contrast))
  ) %>%
  # Keep only pairs differing in exactly one factor (see prose above).
  mutate(
    fam1 = sub(" .*$", "", g1), fam2 = sub(" .*$", "", g2),
    tmp1 = sub("^\\S+ ", "", g1), tmp2 = sub("^\\S+ ", "", g2)
  ) %>%
  filter((fam1 == fam2) | (tmp1 == tmp2)) %>%
  mutate(
    xlo   = pmin(group_x[g1], group_x[g2]),
    xhi   = pmax(group_x[g1], group_x[g2]),
    span  = xhi - xlo,
    label = sig_stars(p_adj)
  ) %>%
  arrange(span, xlo)

stopifnot(nrow(sig_brackets) == 4)

# Pack brackets onto as few tiers as possible: shortest spans lowest, and two
# brackets share a tier when their x ranges do not overlap. This keeps the
# annotation band shallow so it does not crowd the data.
tier_of <- integer(nrow(sig_brackets))
tier_end <- numeric(0)  # right-most x reached on each tier so far
for (i in seq_len(nrow(sig_brackets))) {
  placed <- FALSE
  for (t in seq_along(tier_end)) {
    if (sig_brackets$xlo[i] > tier_end[t] + 0.25) {   # 0.25 = visual gap
      tier_of[i] <- t; tier_end[t] <- sig_brackets$xhi[i]; placed <- TRUE; break
    }
  }
  if (!placed) {
    tier_end <- c(tier_end, sig_brackets$xhi[i])
    tier_of[i] <- length(tier_end)
  }
}

y_data_max <- max(cs_analysis$activity_nobg)

sig_brackets <- sig_brackets %>%
  mutate(
    tier   = tier_of,
    y_top  = y_data_max * (1.05 + 0.075 * (tier - 1)),
    y_tick = y_data_max * 0.018
  )

cat("--- sig_brackets: significance annotations derived from tukey_df ---\n")
print(as.data.frame(sig_brackets[, c("contrast", "p_adj", "label",
                                     "xlo", "xhi", "tier")]))
cat(sprintf("\nBrackets packed onto %d tiers (from %d contrasts).\n",
            max(sig_brackets$tier), nrow(sig_brackets)))
```

```{r legend-caption}
sig_caption <- paste0(
  "Significance: Tukey HSD on log10 activity. ",
  "*** p < 0.001,  ** p < 0.01,  * p < 0.05,  ns = not significant.\n",
  "Brackets show the four contrasts differing in exactly one factor; ",
  "all six pairwise contrasts are in tukey_pairwise_contrasts.csv.")

cat(sig_caption, "\n")
```

```{r plot-activity-by-group, fig.width = 10, fig.height = 7}
set.seed(42)  # reproducible jitter

activity_plot <- ggplot(cs_analysis, aes(x = group, y = activity_nobg)) +
  geom_boxplot(aes(fill = group), outlier.shape = NA,
               width = 0.6, alpha = 0.55, linewidth = 0.5, colour = "grey25") +
  geom_jitter(aes(shape = temperature), width = 0.13, height = 0,
              size = 2.4, alpha = 0.9, colour = "grey15") +
  # Significance brackets: horizontal span + downward ticks + star label.
  geom_segment(data = sig_brackets,
               aes(x = xlo, xend = xhi, y = y_top, yend = y_top),
               inherit.aes = FALSE, linewidth = 0.4, colour = "grey30") +
  geom_segment(data = sig_brackets,
               aes(x = xlo, xend = xlo, y = y_top, yend = y_top - y_tick),
               inherit.aes = FALSE, linewidth = 0.4, colour = "grey30") +
  geom_segment(data = sig_brackets,
               aes(x = xhi, xend = xhi, y = y_top, yend = y_top - y_tick),
               inherit.aes = FALSE, linewidth = 0.4, colour = "grey30") +
  geom_text(data = sig_brackets,
            aes(x = (xlo + xhi) / 2, y = y_top, label = label),
            inherit.aes = FALSE, vjust = -0.3, size = 3.8) +
  scale_fill_manual(values = group_colours, guide = "none") +
  scale_shape_manual(values = c(ambient = 16, `36C` = 17),
                     name = "Temperature") +
  scale_y_continuous(expand = expansion(mult = c(0.04, 0.06))) +
  labs(
    title = "Family F05 loses citrate synthase activity at 36 C; family F07 does not",
    subtitle = sprintf(
      paste("Box = median and IQR, whiskers to 1.5 x IQR; each point is one",
            "individual (n = 8 per group). Background not subtracted.",
            "\nfamily x temperature p = %.2g"),
      p_interaction),
    x = NULL,
    y = "CS activity (mU / mg protein)",
    caption = sig_caption
  ) +
  theme_bw() +
  theme(plot.title    = element_text(size = 13, face = "bold"),
        plot.subtitle = element_text(size = 9.5),
        plot.caption  = element_text(size = 8.5, hjust = 0, colour = "grey30"),
        axis.text.x   = element_text(size = 11),
        legend.position = "right")

ggsave(file.path(output_dir, "cs_activity_by_family_temperature.png"), activity_plot,
       width = 10, height = 7, dpi = 300)

print(activity_plot)
```

## Interaction plot

The two lines are the interaction: parallel lines would mean both families
respond to temperature the same way. Family labels sit at the right end of
each line, so no legend is needed.

```{r plot-interaction, fig.width = 9, fig.height = 6}
interaction_df <- group_summary %>%
  mutate(
    family      = factor(family, levels = c("F05", "F07")),
    temperature = factor(temperature, levels = c("ambient", "36C")),
    se_act      = sd_act / sqrt(n),
    ci_half     = qt(0.975, df = n - 1) * se_act
  )

label_df <- interaction_df %>% filter(temperature == "36C")

interaction_plot <- ggplot(interaction_df,
                           aes(x = temperature, y = mean_act,
                               colour = family, group = family)) +
  geom_line(linewidth = 1.1) +
  geom_errorbar(aes(ymin = mean_act - ci_half, ymax = mean_act + ci_half),
                width = 0.07, linewidth = 0.7) +
  geom_point(size = 3.6) +
  geom_text(data = label_df, aes(label = family),
            hjust = -0.35, size = 4.2, fontface = "bold", show.legend = FALSE) +
  scale_colour_manual(values = family_colours, guide = "none") +
  scale_x_discrete(expand = expansion(mult = c(0.12, 0.28))) +
  labs(
    title = "The two families respond to 36 C differently (non-parallel lines)",
    subtitle = sprintf(
      "Group mean +/- 95%% CI, n = 8 per group. Background not subtracted. family x temperature p = %.2g",
      p_interaction),
    x = "Temperature exposure",
    y = "Mean CS activity (mU / mg protein)"
  ) +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"))

ggsave(file.path(output_dir, "cs_activity_interaction_plot.png"), interaction_plot,
       width = 9, height = 6, dpi = 300)

print(interaction_plot)
```

## Per-individual activity

```{r plot-individuals, fig.width = 10, fig.height = 6}
individual_plot <- ggplot(
  cs_analysis %>% mutate(label = paste0(family, "_", individual)),
  aes(x = reorder(label, activity_nobg), y = activity_nobg,
      colour = family, shape = temperature)) +
  geom_segment(aes(xend = reorder(label, activity_nobg), y = 0, yend = activity_nobg),
               linewidth = 0.4, colour = "grey75") +
  geom_point(size = 3) +
  scale_colour_manual(values = family_colours, name = "Family") +
  scale_shape_manual(values = c(ambient = 16, `36C` = 17), name = "Temperature") +
  coord_flip() +
  labs(
    title = "Per-individual CS activity across both families and exposures",
    subtitle = "Background not subtracted. Each row is one individual.",
    x = NULL,
    y = "CS activity (mU / mg protein)"
  ) +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        axis.text.y = element_text(size = 7),
        legend.position = "right")

ggsave(file.path(output_dir, "cs_activity_per_individual.png"), individual_plot,
       width = 10, height = 6, dpi = 300)

print(individual_plot)
```

## Effect of removing the background correction

A check that the decision in note 2 does not drive any group difference: if
removing the correction mattered, points would sit far off the 1:1 line.

```{r plot-background-effect, fig.width = 9, fig.height = 6}
bg_plot <- ggplot(cs_analysis,
                  aes(x = activity_published, y = activity_nobg, colour = family)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
  geom_point(aes(shape = temperature), size = 2.8, alpha = 0.9) +
  scale_colour_manual(values = family_colours, name = "Family") +
  scale_shape_manual(values = c(ambient = 16, `36C` = 17), name = "Temperature") +
  labs(
    title = "Removing the background correction shifts activity by at most ~11%",
    subtitle = paste("Dashed line is 1:1 (no change). Points above the line are",
                     "the two F05 plates, where a background rate was estimable;",
                     "F07 points sit on the line."),
    x = "Activity as published, background subtracted (mU / mg protein)",
    y = "Activity with background NOT subtracted (mU / mg protein)"
  ) +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        plot.subtitle = element_text(size = 9))

ggsave(file.path(output_dir, "background_correction_effect.png"), bg_plot,
       width = 9, height = 6, dpi = 300)

print(bg_plot)
```

# SUMMARY

```{r summary-numbers}
f05_fold <- temp_effect$fold_change[temp_effect$family == "F05"]
f07_fold <- temp_effect$fold_change[temp_effect$family == "F07"]

f05_p <- within_family$`p adj`[within_family$Contrast == "F05:36C-F05:ambient"]
f07_p <- within_family$`p adj`[within_family$Contrast == "F07:36C-F07:ambient"]

amb_contrast <- tukey_table %>% filter(Contrast == "F07:ambient-F05:ambient")

cat(sprintf("F05: 36C / ambient = %.2f-fold (Tukey p = %.3g)\n", f05_fold, f05_p))
cat(sprintf("F07: 36C / ambient = %.2f-fold (Tukey p = %.3g)\n", f07_fold, f07_p))
cat(sprintf("Ambient F07 vs F05: %.2f-fold (Tukey p = %.3g)\n",
            amb_contrast$`Fold change`, amb_contrast$`p adj`))
cat(sprintf("Interaction p = %.3g | %% variance explained by interaction = %.1f%%\n",
            p_interaction,
            eta_sq$`% of total variance`[eta_sq$Term == "family:temperature"]))
```

## Findings

```{r summary-findings, results = 'asis'}
cat(sprintf("
1. **The temperature response differs between the two families.** The
   family &times; temperature interaction is significant (p = %.2g), accounting
   for %.1f%% of the total variance in log10 activity. This is the central
   result: a single \"heat effect\" on CS activity does not describe both
   families.

2. **Family `F05` loses CS activity at 36 &deg;C.** Mean activity falls from
   %.1f to %.1f mU/mg protein, a **%.2f-fold change (%.0f%%)**, and the
   contrast survives Tukey adjustment (p = %.2g).

3. **Family `F07` shows essentially no temperature response.** Mean activity
   moves from %.1f to %.1f mU/mg protein -- a %.2f-fold change, entirely
   non-significant (p = %.2g). The `F07` 95%% CI at 36 &deg;C overlaps its
   ambient CI substantially.

4. **The two families are indistinguishable at ambient temperature**
   (%.2f-fold, p = %.2g). The families differ only *after* heat exposure,
   which is what makes this an interaction rather than a baseline family
   difference.

5. **Removing the background correction does not drive any of this.** Across
   all 32 individuals the correction was worth %.1f%%-%.1f%% of activity
   (median %.1f%%), and both families' ambient groups shift in the same
   direction, so no group contrast depends on the choice.
",
p_interaction,
eta_sq$`% of total variance`[eta_sq$Term == "family:temperature"],
temp_effect$ambient[temp_effect$family == "F05"],
temp_effect$`36C`[temp_effect$family == "F05"],
f05_fold, temp_effect$pct_change[temp_effect$family == "F05"], f05_p,
temp_effect$ambient[temp_effect$family == "F07"],
temp_effect$`36C`[temp_effect$family == "F07"],
f07_fold, f07_p,
amb_contrast$`Fold change`, amb_contrast$`p adj`,
100 * (min(cs_analysis$bg_undo_factor) - 1),
100 * (max(cs_analysis$bg_undo_factor) - 1),
100 * (median(cs_analysis$bg_undo_factor) - 1)))
```

## Interpretation and caveats

```{r summary-caveats, results = 'asis'}
cat(sprintf("
- **Statistical model.** Two-way ANOVA on log10 activity, balanced at n = 8
  per cell. Residuals are consistent with normality (Shapiro-Wilk
  p = %.3g) and group variances are homogeneous (Bartlett p = %.3g), so the
  parametric model is appropriate here. Working on the log scale makes the
  interaction a test of *fold-change* response, which is the meaningful
  comparison for an enzyme rate.

- **Assay design.** The `F05` ambient-vs-36 &deg;C comparison is made across
  two plates run ten days apart (2026-08-14 and 2026-08-24), whereas the
  `F07` comparison is made within a single plate (note 4). Cross-plate
  comparability is supported by the calibration and instrument-state values
  in [PLATE PROVENANCE](#plate-provenance): standard-curve slope varies by
  only %.1f%% CV across plates (all R<sup>2</sup> >= %.4f) and the CS
  positive control by %.1f%% CV.

- **Background correction.** No activity value in this document has
  background subtracted (note 2), so all four groups are treated
  identically. The `F07` plate's background wells received the wrong reagent
  and cannot be recovered; a future plate with correct Background Control Mix
  would let this choice be checked rather than assumed.

- **Scope.** Two families, eight individuals per family per exposure. These
  results describe these two families and do not establish how broadly
  family-dependent heat sensitivity extends across *M. gigas*.
",
resid_shapiro$p.value, resid_bartlett$p.value,
prov_spread$slope_cv, prov_spread$min_std_r2, prov_spread$pos_cv))
```

## Output files

```{r output-manifest}
manifest <- tibble(
  file = list.files(output_dir),
) %>%
  mutate(
    size_kb = round(file.size(file.path(output_dir, file)) / 1024, 1),
    type    = case_when(
      grepl("\\.csv$", file) ~ "table",
      grepl("\\.png$", file) ~ "figure",
      TRUE                   ~ "documentation"
    )
  ) %>%
  arrange(type, file)

kable(manifest, caption = sprintf("Files written to `%s`.", output_dir))
```

# SESSION INFO

```{r session-info}
sessionInfo()
```
