---
title: "Gen5-20260811-mgig-citrate_synthase-F05-ambient"
author: "Sam White"
date: "2026-08-11"
output:
  github_document:
    toc: true
    number_sections: true
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
---

# BACKGROUND

Citrate synthase (CS) activity in ctenidia of eight *Magallana gigas* (Pacific
oyster) individuals from **family F05** held at **ambient** temperature, assayed
2026-08-11 with the
[Abcam Citrate Synthase Assay Kit (ab239712), v4a](https://github.com/RobertsLab/resources/blob/master/protocols/Commercial_Protocols/ABCAM-Citrate-Synthase-Assay-v4a-ab239712.pdf).

The kit is a **coupled kinetic** assay. CS condenses acetyl-CoA and
oxaloacetate, releasing free CoA-SH; the liberated thiol reduces DTNB to
TNB<sup>2-</sup>, which absorbs at 412 nm. The **rate** of A412 increase is
proportional to CS activity, and absolute nmol of thiol are assigned from a
**GSH (reduced glutathione) standard curve** read on the same plate.

Absorbance was read at 412 nm in kinetic mode, 25 °C, every 2 min for 40 min
(21 reads) on a Synergy HTX.

This document is fully self-contained: layout parsing, kinetic QC,
standard-curve fitting, rate extraction, CV computation, protein
normalization, and results assembly are all defined directly below (see
[`## Pipeline functions`](#pipeline-functions) under SETUP) rather than sourced from a separate
script, so the whole analysis can be reproduced from this one file.

## Plate design

| Plate rows | Contents |
|---|---|
| A, B | 8 samples (`F05_01`–`F05_08`), technical triplicates, + Reaction Mix |
| C, D | Paired **background controls** for the same 8 samples, + Background Control Mix (no CS Substrate Mix) |
| E, F1–F6 | GSH standards: 0, 8, 16, 24, 32, 40 nmol/well, triplicates |
| F7–F9 | CS Positive Control, triplicate |

Background control wells receive Background Control Mix, which **omits CS
Substrate Mix**. They therefore contain no acetyl-CoA/oxaloacetate, cannot
support the CS reaction, and should read **flat and low** for the whole kinetic
run. Their purpose (Abcam §8.4 note, §10.2) is to correct for pre-existing free
thiol / CoA in the homogenate, which produces a one-time DTNB reduction at t0
rather than an ongoing rate.

## Sample naming

Following the convention in the sibling Glycogen assay
(`../../Glycogen/data/raw_luminescence/README.md`), plate layout entries are:

- `<sample>-<assay_type>-<tissue_weight>-df.<dilution_factor>`

where `<sample>` is itself composite:

- `<family>_<individual>_<temperature>`, with `_BG` appended for background
  control wells.

E.g. `F05_03_ambient-citrate_synthase-22.3-df.0` is family F05, individual 03,
ambient exposure, citrate synthase assay, 22.3 mg of ctenidia tissue,
undiluted homogenate. Standards follow `STD-<assay_type>-<nmol_per_well>`
(e.g. `STD-citrate_synthase-24` = 24 nmol GSH/well); the positive control is
`POS-citrate_synthase`.

`df.0` denotes **no dilution** of the homogenate, i.e. dilution factor
*D* = 1 in the Abcam activity formula. It is not a literal multiplier of zero.

## Important note(s)

1. **`absorbance-*.csv` is the sole data source.** An earlier export of this
   file covered only plate columns 1–9 (54 of the 69 occupied wells); it has
   since been re-exported from the Gen5 software and now carries all 69
   occupied wells (columns 1–12, rows A–F). The re-export was verified against
   `full_report-*.txt` — the two agree **exactly** on all 1,449 well ×
   timepoint readings they share — and against `layout-*.csv`, which confirms
   the well set matches the plate layout with none missing and none extra.
   `full_report-*.txt` is retained only as a cross-check, not as a data
   source.
2. **Reaction volumes.** 2 µL of homogenate + 48 µL Assay Buffer 7 per well,
   then 50 µL Reaction Mix. Sample volume *V* = 2 µL in the activity formula.
3. **Normalization is by total extracted protein, not tissue weight.** Tissue
   was homogenized in **350 µL** of Assay Buffer 7 (`homogenate_volume_uL`
   below), reported by the bench operator and applied to every sample. Protein
   concentration in each homogenate was measured separately (BCA/Bradford-style
   assay, 595 nm; see `Protein concentration (normalization factor)` below) and
   converted to a total protein mass by multiplying by this same 350 µL. Tissue
   input (10.1–30.6 mg, recorded in the well label for provenance only) is not
   used in any activity calculation on this plate — total extracted protein is
   the normalizer, because it reflects what the assay actually measures against
   (soluble protein carried into the 2 µL aliquot), not the starting tissue
   mass, which can vary in extraction efficiency from sample to sample.
4. **Rate windows are compared over a common early interval.** Every trace on
   this plate decelerates — substrate is being consumed — so the measured rate
   depends on *when* it is measured, and a rate is only comparable between
   wells if the wells are measured over comparable intervals. Selecting each
   well's steepest window independently satisfies this here because the
   steepest window is the initial one (0–8 min) for 22 of 24 sample wells; the
   two exceptions are wells whose traces are anomalous and are excluded anyway.
   The R<sup>2</sup> floor for accepting a window is set to
   `rate_min_r2` = 0.80 rather than a stricter value for the same reason:
   deceleration within the window depresses R<sup>2</sup> even in clean wells
   (e.g. B10–B12 sit at 0.84–0.88 with 8% replicate CV), so a 0.95 floor would
   reject well-behaved replicates while retaining late, near-zero-rate windows
   from exhausted wells. Curvature is a property of the assay here, not a
   defect of the well.
5. **A412 at t = 0 is the master diagnostic on this plate.** Clean sample wells
   start at A412 0.13–0.16. Every well that misbehaves — samples that decline,
   background wells that act like reactions — starts well above that. TNB
   present before the reaction begins means the well was contaminated or
   mis-dispensed, so its trace is not enzymology and a rate taken from it is
   not activity. `sample_baseline_od` = 0.35 separates the two populations
   cleanly (there are no sample wells between 0.23 and 0.29 except A6 at 0.29,
   itself anomalous). **All three `F05_05` replicates** start elevated
   (0.65–0.93), so that sample has no clean replicate at all — see the results
   caveat below.
6. **Layout label typo**, carried through from the plate layout file and
   handled explicitly in the parsing chunk: `F05_05_ambient_BG` wells (D1–D3)
   are labeled `citrate_synthasse` (three s's).

# SETUP

## Libraries and shared pipeline functions

```{r setup, include=TRUE}
library(knitr)
library(ggplot2)
library(dplyr)
library(tidyr)

knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,         # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = "",        # Prevents appending '##' to beginning of lines in code output
  results = 'hold'     # Holds output so it's all printed together after code chunk
)

```

## Pipeline functions

All layout parsing, kinetic QC, standard-curve fitting, rate extraction,
CV computation, protein normalization, and results assembly are defined
directly in this document (rather than sourced from a separate script) so
that it is fully self-contained and reproducible from this single file.
The functions are grouped below by pipeline stage; each is called in the
DATA / KINETIC TRACES / ANOMALY DETECTION / GSH STANDARD CURVE / RATE
EXTRACTION / BACKGROUND CORRECTION / TECHNICAL REPLICATE PRECISION /
CITRATE SYNTHASE ACTIVITY / QC SUMMARY sections that follow. Nothing
plate-specific (typo fixes, disqualified samples, thresholds) is
hardcoded in any function -- it is always passed in as an argument, from
the `assay_params` list defined further below or from a plate-specific
fix list.

### Data import and layout parsing

Reads the raw Gen5 exports (kinetic absorbance CSV, full-report text backup, plate-layout CSV), reconciles them against each other, and attaches well-type/sample-id/replicate metadata plus measured protein concentrations to every well.

```{r pipeline-fns-io}
## -----------------------------------------------------------------------
## elapsed_to_min()
## Purpose : Convert Gen5 "H:MM:SS" elapsed-time strings to decimal minutes.
## Inputs  : x - character vector of "H:MM:SS" strings
## Outputs : numeric vector of elapsed minutes, same length as x
## -----------------------------------------------------------------------
elapsed_to_min <- function(x) {
  vapply(strsplit(x, ":"), function(p) {
    p <- as.numeric(p)
    p[1] * 60 + p[2] + p[3] / 60
  }, numeric(1))
}

## -----------------------------------------------------------------------
## read_absorbance_csv()
## Purpose : Parse a Gen5 "absorbance-*.csv" kinetic export: wavelength
##           header, a "Time," row, a temperature row, then one row per well
##           (well ID in column 1, one OD reading per timepoint thereafter).
## Inputs  : path - path to the absorbance-*.csv file
## Outputs : long-format data.frame, one row per well x timepoint:
##           well (chr), time_min (dbl), od (dbl)
## -----------------------------------------------------------------------
read_absorbance_csv <- function(path) {
  ln <- iconv(readLines(path, warn = FALSE, encoding = "latin1"), "latin1", "UTF-8")
  time_i   <- grep("^Time,", ln)[1]
  time_str <- strsplit(ln[time_i], ",")[[1]]
  time_str <- time_str[!time_str %in% c("", "Time")]
  tmin     <- elapsed_to_min(time_str)

  well_ln <- ln[grepl("^[A-H][0-9]{1,2},", ln)]
  bind_rows(lapply(strsplit(well_ln, ","), function(p) {
    data.frame(well = p[1], time_min = tmin,
               od = as.numeric(p[seq_along(tmin) + 1]),
               stringsAsFactors = FALSE)
  }))
}

## -----------------------------------------------------------------------
## read_full_report()
## Purpose : Parse a Gen5 "full_report-*.txt" export: tab-delimited, wells
##           across columns, timepoints down rows. Used only as an
##           independent cross-check against the absorbance CSV, never as a
##           primary data source.
## Inputs  : path - path to the full_report-*.txt file
## Outputs : long-format data.frame, one row per well x timepoint:
##           well (chr), time_min (dbl), od (dbl)
## -----------------------------------------------------------------------
read_full_report <- function(path) {
  ln  <- iconv(readLines(path, warn = FALSE, encoding = "latin1"), "latin1", "UTF-8")
  hdr_i <- grep("^Time\t.*\tA1\t", ln)[1]
  hdr   <- trimws(strsplit(ln[hdr_i], "\t")[[1]])
  rows  <- ln[grepl("^[0-9]:[0-9]{2}:[0-9]{2}\t", ln)]
  m     <- do.call(rbind, strsplit(rows, "\t"))
  tmin  <- elapsed_to_min(m[, 1])

  well_cols <- which(grepl("^[A-H][0-9]{1,2}$", hdr))
  bind_rows(lapply(well_cols, function(j) {
    data.frame(well = hdr[j], time_min = tmin,
               od = as.numeric(m[, j]), stringsAsFactors = FALSE)
  }))
}

## -----------------------------------------------------------------------
## parse_plate_layout()
## Purpose : Convert a raw Gen5 plate-layout CSV export into one row per
##           occupied well with its descriptive label. Auto-detects two
##           export formats seen in practice:
##             "single"  - one row per plate row (row 1 = column numbers,
##                         column 1 = row letter, last column = "Name" tag);
##                         each cell already holds the full descriptive label.
##             "double"  - each plate row is exported as TWO rows: a
##                         "Well ID" row (SPL codes) immediately followed by
##                         a "Name" row (the descriptive label actually used
##                         downstream).
##           The format is detected from column 1: if a value repeats twice
##           in a row (once for the well-ID row, once for the label row)
##           immediately below a numeric header, it is treated as "double".
## Inputs  : path - path to the layout-*.csv file
## Outputs : data.frame, one row per occupied well:
##           well (chr, e.g. "A1"), plate_row (chr), plate_col (int),
##           label (chr, trimmed descriptive label)
## -----------------------------------------------------------------------
parse_plate_layout <- function(path) {
  plate_layout <- read.csv(path, header = FALSE, stringsAsFactors = FALSE)

  col1 <- trimws(plate_layout$V1)
  # "double" format: row letters appear twice in a row in column 1
  # (once tagging the Well-ID row, once tagging the Name row), e.g.
  # "A", "A", "B", "B", ... Detect by checking whether non-empty,
  # non-header values in col1 are each immediately followed by a repeat.
  data_rows   <- which(col1 != "" & !is.na(col1))[-1]  # drop header row (row 1)
  is_double <- length(data_rows) >= 2 &&
    all(col1[data_rows[c(TRUE, FALSE)]] == col1[data_rows[c(FALSE, TRUE)]][seq_len(floor(length(data_rows) / 2))])

  if (is_double) {
    # Pair each row-letter row with the following label row.
    row_letter_rows <- data_rows[c(TRUE, FALSE)]
    label_rows      <- data_rows[c(FALSE, TRUE)]
    stopifnot(length(row_letter_rows) == length(label_rows))

    out <- expand.grid(pair_idx = seq_along(row_letter_rows),
                        col_idx  = 2:(ncol(plate_layout) - 1)) %>%
      mutate(
        plate_row = col1[row_letter_rows[pair_idx]],
        plate_col = as.integer(plate_layout[1, col_idx]),
        well      = paste0(plate_row, plate_col),
        label     = trimws(mapply(function(i, j) plate_layout[i, j],
                                   label_rows[pair_idx], col_idx))
      ) %>%
      filter(label != "") %>%
      select(well, plate_row, plate_col, label) %>%
      arrange(plate_row, plate_col)
  } else {
    # "single" format: row 1 = column numbers, column 1 = row letters,
    # last column = "Name" tag; each cell already holds the full label.
    out <- expand.grid(row_idx = 2:nrow(plate_layout),
                        col_idx = 2:(ncol(plate_layout) - 1)) %>%
      mutate(
        plate_row = plate_layout$V1[row_idx],
        plate_col = as.integer(plate_layout[1, col_idx]),
        well      = paste0(plate_row, plate_col),
        label     = trimws(mapply(function(i, j) plate_layout[i, j], row_idx, col_idx))
      ) %>%
      filter(label != "") %>%
      select(well, plate_row, plate_col, label) %>%
      arrange(plate_row, plate_col)
  }

  attr(out, "layout_format") <- ifelse(is_double, "double", "single")
  out
}

## -----------------------------------------------------------------------
## reconcile_kinetic_sources()
## Purpose : Cross-check the absorbance CSV (primary data source) against
##           the full_report text export (independent instrument export)
##           and the plate layout (expected well set). Fails loudly
##           (stopifnot) if any layout well is missing from the CSV, or if
##           the two kinetic sources disagree on any shared reading.
## Inputs  : absorbance_csv - long-format data.frame from read_absorbance_csv()
##           full_report    - long-format data.frame from read_full_report()
##           layout_wells   - data.frame from parse_plate_layout()
##           tolerance      - max allowed |CSV - report| absolute difference
##                            (default 1e-9, i.e. exact agreement)
## Outputs : list(plate_readings, overlap_check, summary), where
##           plate_readings = absorbance_csv with a `source` column added,
##           summary        = named list of counts for reporting
## -----------------------------------------------------------------------
reconcile_kinetic_sources <- function(absorbance_csv, full_report, layout_wells,
                                       tolerance = 1e-9) {
  wells_csv    <- unique(absorbance_csv$well)
  wells_report <- unique(full_report$well)

  wells_report_not_csv <- setdiff(wells_report, wells_csv)
  wells_layout_not_csv <- setdiff(layout_wells$well, wells_csv)
  wells_csv_not_layout <- setdiff(wells_csv, layout_wells$well)

  overlap_check <- inner_join(absorbance_csv, full_report,
                              by = c("well", "time_min"), suffix = c("_csv", "_report")) %>%
    mutate(abs_diff = abs(od_csv - od_report))

  max_disagreement <- if (nrow(overlap_check) > 0) max(overlap_check$abs_diff) else NA_real_

  stopifnot(
    length(wells_layout_not_csv) == 0,
    is.na(max_disagreement) || max_disagreement < tolerance
  )

  plate_readings <- absorbance_csv %>% mutate(source = "absorbance_csv")

  list(
    plate_readings = plate_readings,
    overlap_check  = overlap_check,
    summary = list(
      n_wells_csv               = length(wells_csv),
      n_wells_report            = length(wells_report),
      n_wells_report_not_csv    = length(wells_report_not_csv),
      n_wells_layout_not_csv    = length(wells_layout_not_csv),
      n_wells_csv_not_layout    = length(wells_csv_not_layout),
      n_shared_readings         = nrow(overlap_check),
      max_disagreement          = max_disagreement
    )
  )
}

## -----------------------------------------------------------------------
## annotate_wells()
## Purpose : Join layout labels onto the kinetic readings and parse each
##           label into well_type, sample_id, family, individual,
##           temperature, tissue weight, and standard concentration.
##           Expected label grammar:
##             sample/background : <family>_<individual>_<temperature>[_BG]-<assay_type>-<weight>-df.<n>
##             standard           : STD-<assay_type>-<nmol_per_well>
##             positive control   : POS-<assay_type>
##           Label typos (e.g. a misspelled assay_type) are corrected before
##           parsing via `label_fixes`, a named character vector of
##           c(pattern = replacement) pairs applied in order with gsub().
## Inputs  : plate_readings - data.frame from reconcile_kinetic_sources()$plate_readings
##           layout_wells   - data.frame from parse_plate_layout()
##           label_fixes    - optional named character vector, e.g.
##                            c("citrate_synthasse" = "citrate_synthase")
## Outputs : plate_long - fully annotated long-format data.frame, one row
##           per well x timepoint, with well_type, sample_id, family,
##           individual, temperature, weight_mg, std_nmol columns added
## -----------------------------------------------------------------------
annotate_wells <- function(plate_readings, layout_wells, label_fixes = NULL) {
  plate_long <- plate_readings %>%
    left_join(layout_wells, by = "well") %>%
    mutate(
      well_type = case_when(
        grepl("^STD-", label) ~ "standard",
        grepl("^POS-", label) ~ "positive_control",
        grepl("_BG-",  label) ~ "background",
        TRUE                  ~ "sample"
      ),
      label_clean = label
    )

  if (!is.null(label_fixes)) {
    for (pat in names(label_fixes)) {
      plate_long$label_clean <- gsub(pat, label_fixes[[pat]], plate_long$label_clean)
    }
  }

  plate_long <- plate_long %>%
    mutate(
      sample_id   = ifelse(well_type %in% c("sample", "background"),
                           sub("(_BG)?-[a-zA-Z_]+-.*$", "", label_clean), NA_character_),
      family      = ifelse(!is.na(sample_id), sub("^(F[0-9]+)_.*$", "\\1", sample_id), NA_character_),
      individual  = ifelse(!is.na(sample_id), sub("^F[0-9]+_([0-9]+)_.*$", "\\1", sample_id), NA_character_),
      temperature = ifelse(!is.na(sample_id), sub("^F[0-9]+_[0-9]+_(.*)$", "\\1", sample_id), NA_character_),
      weight_mg   = ifelse(well_type %in% c("sample", "background"),
                           as.numeric(sub("^.*-[a-zA-Z_]+-([0-9.]+)-df\\..*$", "\\1", label_clean)),
                           NA_real_),
      std_nmol    = ifelse(well_type == "standard",
                           as.numeric(sub("^STD-[a-zA-Z_]+-", "", label_clean)), NA_real_)
    ) %>%
    arrange(well_type, sample_id, well, time_min)

  # Fail loudly rather than silently dropping malformed labels
  stopifnot(
    !any(is.na(plate_long$well_type)),
    !any(is.na(plate_long$weight_mg[plate_long$well_type %in% c("sample", "background")])),
    !any(is.na(plate_long$std_nmol[plate_long$well_type == "standard"]))
  )

  plate_long
}

## -----------------------------------------------------------------------
## load_protein_concentrations()
## Purpose : Load one or more BCA/Bradford-style protein-concentration CSV
##           exports, concatenate them, and match each plate sample to
##           exactly one protein-concentration record. Fails loudly if any
##           plate sample has zero or more than one match. Converts
##           concentration (ug/mL) to total protein mass (mg) in the full
##           homogenate using the measured homogenization volume.
## Inputs  : protein_files        - character vector of CSV paths, each with
##                                   columns `Sample` and
##                                   `Calculated concentration (ug/mL)`
##           plate_sample_ids     - character vector of this plate's sample IDs
##           homogenate_volume_uL - measured homogenization volume (uL),
##                                   applied identically to every sample
##           sample_id_fixes      - optional named character vector of
##                                   c(pattern = replacement) label-typo
##                                   corrections applied to the protein
##                                   file's Sample column before matching,
##                                   e.g. c("F-05_07_36C" = "F05_07_36C")
## Outputs : data.frame, one row per plate sample:
##           sample_id, source_file, conc_ug_mL, total_protein_mg
## -----------------------------------------------------------------------
load_protein_concentrations <- function(protein_files, plate_sample_ids,
                                         homogenate_volume_uL,
                                         sample_id_fixes = NULL) {
  protein_all <- protein_files %>%
    purrr::map_dfr(~ read.csv(.x, check.names = FALSE) %>%
                      mutate(source_file = basename(.x))) %>%
    rename(sample_id_protein = Sample, conc_ug_mL = `Calculated concentration (ug/mL)`) %>%
    select(sample_id_protein, source_file, conc_ug_mL)

  if (!is.null(sample_id_fixes)) {
    for (pat in names(sample_id_fixes)) {
      protein_all$sample_id_protein <- gsub(pat, sample_id_fixes[[pat]],
                                             protein_all$sample_id_protein)
    }
  }

  protein_matches <- protein_all %>% filter(sample_id_protein %in% plate_sample_ids)
  match_counts    <- protein_matches %>% count(sample_id_protein, name = "n_matches")

  stopifnot(
    all(plate_sample_ids %in% protein_matches$sample_id_protein),
    all(match_counts$n_matches == 1)
  )

  protein_matches %>%
    rename(sample_id = sample_id_protein) %>%
    mutate(total_protein_mg = conc_ug_mL * homogenate_volume_uL / 1e6) %>%
    arrange(sample_id)
}
```

### Well-level kinetic QC

Flags read glitches, non-rising ('decreasing') traces, elevated t0 baselines, and other well-level anomalies before any rate is trusted.

```{r pipeline-fns-qc}
## -----------------------------------------------------------------------
## local_step_excess()
## Purpose : For a kinetic trace, measure how much each read-to-read step
##           departs from the AVERAGE of its two immediate neighbouring
##           steps. A large local excess flags a read glitch; comparing to
##           local neighbours (rather than the trace's overall median step)
##           tolerates traces that smoothly decelerate throughout.
## Inputs  : od - numeric vector of OD readings in time order
## Outputs : numeric vector of length length(od) - 1 (one per step)
## -----------------------------------------------------------------------
local_step_excess <- function(od) {
  d <- diff(od)
  n <- length(d)
  local_expect <- vapply(seq_len(n), function(i) {
    mean(d[intersect(c(i - 1L, i + 1L), seq_len(n))])
  }, numeric(1))
  d - local_expect
}

## -----------------------------------------------------------------------
## compute_well_diagnostics()
## Purpose : Compute four independent per-well trace-shape diagnostics from
##           the raw kinetic trace, before any rate fitting: direction
##           (net change), monotonicity (fraction of rising steps),
##           over-range (any read above the photometric linearity ceiling),
##           and discontinuity (a read-to-read glitch via local_step_excess).
##           Also flags an elevated starting absorbance for sample wells,
##           and background wells that behave like active reactions.
## Inputs  : plate_long   - data.frame from annotate_wells()
##           assay_params - list with od_linear_max, glitch_excess_od,
##                          sample_baseline_od, bg_flat_od_max,
##                          bg_flat_drift_max
## Outputs : well_diagnostics - one row per well, with od_first, od_last,
##           od_max, net_change, frac_rising, max_step, typical_step,
##           step_excess, glitch_at_min, and flag_* / n_flags columns
## -----------------------------------------------------------------------
compute_well_diagnostics <- function(plate_long, assay_params) {
  plate_long %>%
    arrange(well, time_min) %>%
    group_by(well, plate_row, plate_col, well_type, sample_id, label, std_nmol, source) %>%
    summarise(
      od_first        = first(od),
      od_last         = last(od),
      od_max          = max(od),
      net_change      = last(od) - first(od),
      frac_rising     = mean(diff(od) > 0),
      max_step        = diff(od)[which.max(abs(diff(od)))],
      typical_step    = median(abs(diff(od))),
      step_excess     = local_step_excess(od)[which.max(abs(local_step_excess(od)))],
      glitch_at_min   = time_min[which.max(abs(local_step_excess(od))) + 1L],
      .groups = "drop"
    ) %>%
    mutate(
      step_ratio      = abs(max_step) / pmax(typical_step, 0.001),
      flag_decreasing = net_change <= 0,
      flag_over_range = od_max > assay_params$od_linear_max,
      flag_glitch     = abs(step_excess) > assay_params$glitch_excess_od,
      flag_high_baseline = well_type == "sample" & od_first > assay_params$sample_baseline_od,
      flag_bg_active  = well_type == "background" &
                        (od_first > assay_params$bg_flat_od_max |
                         abs(net_change) > assay_params$bg_flat_drift_max),
      n_flags         = flag_decreasing + flag_over_range + flag_glitch +
                        flag_bg_active + flag_high_baseline
    )
}

## -----------------------------------------------------------------------
## compute_baseline_diagnostics()
## Purpose : Classify sample/background wells by starting absorbance
##           (elevated vs. normal) and summarize, per sample, how many of
##           its replicates are baseline-compromised. A sample with all
##           replicates elevated has no clean replicate at all.
## Inputs  : well_diagnostics - data.frame from compute_well_diagnostics()
##           assay_params     - list with sample_baseline_od
## Outputs : list(baseline_check, baseline_per_sample)
## -----------------------------------------------------------------------
compute_baseline_diagnostics <- function(well_diagnostics, assay_params) {
  baseline_check <- well_diagnostics %>%
    filter(well_type %in% c("sample", "background")) %>%
    mutate(baseline = ifelse(od_first > assay_params$sample_baseline_od,
                             "elevated", "normal")) %>%
    arrange(desc(od_first))

  baseline_per_sample <- baseline_check %>%
    filter(well_type == "sample") %>%
    group_by(sample_id) %>%
    summarise(n_elevated = sum(baseline == "elevated"), n = n(),
              median_baseline = median(od_first), .groups = "drop") %>%
    arrange(desc(n_elevated))

  list(baseline_check = baseline_check, baseline_per_sample = baseline_per_sample)
}
```

### Standard curve and rate extraction

Fits the GSH standard curve, extracts each well's rate via a sliding-window linear regression gated on R-squared, and summarizes the positive control.

```{r pipeline-fns-rates}
## -----------------------------------------------------------------------
## fit_standard_curve()
## Purpose : Extract t=0 GSH standard readings, quantify signal drift over
##           the run, compute per-concentration replicate statistics, flag
##           outlier wells by deviation from their triplicate median, and
##           fit three candidate calibration lines (all wells, per-
##           concentration means, outlier-excluded). The outlier-excluded
##           fit is the calibration used downstream.
## Inputs  : plate_long   - data.frame from annotate_wells()
##           assay_params - list with std_outlier_od
## Outputs : list with standards_t0, standard_drift, standard_summary,
##           standards_flagged, fit_all_wells, fit_means, fit_no_outliers,
##           fit_comparison, std_slope, std_intercept, std_r2, std_nmol_max
## -----------------------------------------------------------------------
fit_standard_curve <- function(plate_long, assay_params) {
  standards_all <- plate_long %>%
    filter(well_type == "standard") %>%
    select(well, std_nmol, time_min, od, source)

  standards_t0 <- standards_all %>%
    filter(time_min == 0) %>%
    select(well, std_nmol, od, source) %>%
    arrange(std_nmol, well)

  standard_drift <- standards_all %>%
    group_by(well, std_nmol) %>%
    summarise(od_t0 = od[time_min == 0], od_t40 = od[time_min == max(time_min)],
              drift = od_t40 - od_t0, .groups = "drop") %>%
    arrange(std_nmol, well)

  standard_summary <- standards_t0 %>%
    group_by(std_nmol) %>%
    summarise(n = n(), mean_od = mean(od), sd_od = sd(od),
              se_od = sd(od) / sqrt(n()), cv_pct = 100 * sd(od) / mean(od),
              median_od = median(od), .groups = "drop") %>%
    arrange(std_nmol) %>%
    mutate(net_od = mean_od - mean_od[std_nmol == 0],
           od_per_nmol = ifelse(std_nmol > 0, net_od / std_nmol, NA_real_))

  standards_flagged <- standards_t0 %>%
    group_by(std_nmol) %>%
    mutate(triplicate_median = median(od),
           deviation = od - triplicate_median,
           is_outlier = abs(deviation) > assay_params$std_outlier_od) %>%
    ungroup() %>%
    arrange(desc(abs(deviation)))

  fit_all_wells   <- lm(od ~ std_nmol, data = standards_flagged)
  fit_means       <- lm(mean_od ~ std_nmol, data = standard_summary)
  fit_no_outliers <- lm(od ~ std_nmol, data = standards_flagged %>% filter(!is_outlier))

  fit_comparison <- bind_rows(
    data.frame(fit = "all wells",          n = nobs(fit_all_wells),
               slope = coef(fit_all_wells)[2],   intercept = coef(fit_all_wells)[1],
               r_squared = summary(fit_all_wells)$r.squared),
    data.frame(fit = "concentration means", n = nobs(fit_means),
               slope = coef(fit_means)[2],       intercept = coef(fit_means)[1],
               r_squared = summary(fit_means)$r.squared),
    data.frame(fit = "outlier-excluded",   n = nobs(fit_no_outliers),
               slope = coef(fit_no_outliers)[2], intercept = coef(fit_no_outliers)[1],
               r_squared = summary(fit_no_outliers)$r.squared)
  )

  list(
    standards_all      = standards_all,
    standards_t0       = standards_t0,
    standard_drift     = standard_drift,
    standard_summary   = standard_summary,
    standards_flagged  = standards_flagged,
    fit_all_wells      = fit_all_wells,
    fit_means          = fit_means,
    fit_no_outliers    = fit_no_outliers,
    fit_comparison     = fit_comparison,
    std_slope          = unname(coef(fit_no_outliers)[2]),
    std_intercept      = unname(coef(fit_no_outliers)[1]),
    std_r2             = summary(fit_no_outliers)$r.squared,
    std_nmol_max       = max(standard_summary$std_nmol)
  )
}

## -----------------------------------------------------------------------
## window_slopes()
## Purpose : Fit every sliding window of `w` consecutive timepoints on one
##           kinetic trace and return the slope (mOD/min) and R^2 of each.
## Inputs  : t - numeric vector of timepoints (min)
##           y - numeric vector of OD readings, same length as t
##           w - window width in number of points
## Outputs : data.frame, one row per window: t_start, t_end,
##           slope_mOD_min, r2
## -----------------------------------------------------------------------
window_slopes <- function(t, y, w) {
  n <- length(y)
  bind_rows(lapply(seq_len(n - w + 1), function(s) {
    i  <- s:(s + w - 1)
    f  <- lm(y[i] ~ t[i])
    r2 <- summary(f)$r.squared
    data.frame(t_start = t[i[1]], t_end = t[i[w]],
               slope_mOD_min = unname(coef(f)[2]) * 1000,
               r2 = ifelse(is.na(r2), 0, r2))
  }))
}

## -----------------------------------------------------------------------
## compute_well_rates()
## Purpose : For every well, scan all sliding windows of assay_params$
##           rate_window_n consecutive reads and record the max-increasing
##           window (used for activity) and the max-absolute window
##           (diagnostic only -- matches Gen5's own `Max V` convention,
##           which can report large negative "rates" for degrading wells).
##           A rate is `rate_usable` only if its R^2 clears the floor, the
##           well's net change is positive, and no read glitch falls inside
##           the fitted window.
## Inputs  : plate_long       - data.frame from annotate_wells()
##           well_diagnostics - data.frame from compute_well_diagnostics()
##           assay_params     - list with rate_window_n, rate_min_r2
## Outputs : well_rates - one row per well: t_start, t_end, slope_mOD_min,
##           r2, max_abs_slope_mOD_min, abs_window_is_negative,
##           glitch_in_window, rate_usable (plus diagnostic columns joined in)
## -----------------------------------------------------------------------
compute_well_rates <- function(plate_long, well_diagnostics, assay_params) {
  plate_long %>%
    arrange(well, time_min) %>%
    group_by(well, well_type, sample_id, std_nmol) %>%
    group_modify(~ {
      w   <- window_slopes(.x$time_min, .x$od, assay_params$rate_window_n)
      inc <- w[which.max(w$slope_mOD_min), ]
      abs_ <- w[which.max(abs(w$slope_mOD_min)), ]
      data.frame(
        t_start = inc$t_start, t_end = inc$t_end,
        slope_mOD_min = inc$slope_mOD_min, r2 = inc$r2,
        max_abs_slope_mOD_min = abs_$slope_mOD_min,
        abs_window_is_negative = abs_$slope_mOD_min < 0
      )
    }) %>%
    ungroup() %>%
    left_join(well_diagnostics %>% select(well, net_change, frac_rising, od_max,
                                          flag_decreasing, flag_over_range,
                                          flag_glitch, glitch_at_min,
                                          flag_bg_active, n_flags),
              by = "well") %>%
    mutate(
      glitch_in_window = flag_glitch & glitch_at_min > t_start & glitch_at_min <= t_end,
      rate_usable      = r2 >= assay_params$rate_min_r2 & net_change > 0 & !glitch_in_window
    )
}

## -----------------------------------------------------------------------
## compute_positive_control()
## Purpose : Compute the CS Positive Control's rate and activity, and check
##           that its replicates rose linearly -- the plate's proof that the
##           reaction chemistry (DTNB, coupling, reader) worked, independent
##           of any sample-specific problem.
## Inputs  : well_rates   - data.frame from compute_well_rates()
##           std_slope    - calibration slope (A412/nmol), from fit_standard_curve()
##           assay_params - list with sample_volume_uL, dilution_factor
## Outputs : pos_control - one row per positive-control well: t_start, t_end,
##           slope_mOD_min, r2, net_change, activity_mU_uL
## -----------------------------------------------------------------------
compute_positive_control <- function(well_rates, std_slope, assay_params) {
  well_rates %>%
    filter(well_type == "positive_control") %>%
    mutate(activity_mU_uL = (slope_mOD_min / 1000) / std_slope /
                            assay_params$sample_volume_uL * assay_params$dilution_factor)
}
```

### Background correction, replicate CV, and activity calculation

Estimates background rate from flat background wells, computes technical-replicate coefficients of variation, and converts corrected rates into protein-normalized CS activity.

```{r pipeline-fns-activity}
## -----------------------------------------------------------------------
## compute_background_correction()
## Purpose : Estimate each sample's background rate from its flat
##           (well-behaved) background-control replicates only, since
##           anomalous background wells do not measure background.
## Inputs  : well_rates - data.frame from compute_well_rates()
## Outputs : background_per_sample - one row per sample: n_bg_total,
##           n_bg_flat, bg_rate_flat (NA if no flat replicate exists),
##           bg_rate_all
## -----------------------------------------------------------------------
compute_background_correction <- function(well_rates) {
  well_rates %>%
    filter(well_type == "background") %>%
    group_by(sample_id) %>%
    summarise(
      n_bg_total    = n(),
      n_bg_flat     = sum(!flag_bg_active),
      bg_rate_flat  = ifelse(sum(!flag_bg_active) > 0,
                             mean(slope_mOD_min[!flag_bg_active]), NA_real_),
      bg_rate_all   = mean(slope_mOD_min),
      .groups = "drop"
    )
}

## -----------------------------------------------------------------------
## compute_replicate_cv()
## Purpose : Compute technical-replicate coefficient of variation (CV%) on
##           the extracted rate, both across ALL three replicates and
##           across USABLE replicates only (rate_usable == TRUE). Flags any
##           sample whose CV exceeds assay_params$cv_threshold_pct.
## Inputs  : well_rates       - data.frame from compute_well_rates()
##           sample_id_order  - character vector giving the row order /
##                              complete sample-ID set to report (typically
##                              the plate's sample IDs, e.g. from
##                              protein_by_sample$sample_id)
##           assay_params     - list with cv_threshold_pct
## Outputs : cv_summary - one row per sample: n_all, mean_all, sd_all,
##           cv_all, n_usable, mean_usable, sd_usable, cv_usable,
##           excluded_wells, fails_cv_all, fails_cv_usable
## -----------------------------------------------------------------------
compute_replicate_cv <- function(well_rates, sample_id_order, assay_params) {
  sample_rates <- well_rates %>% filter(well_type == "sample") %>% arrange(sample_id, well)

  cv_all <- sample_rates %>%
    group_by(sample_id) %>%
    summarise(n_all = n(), mean_all = mean(slope_mOD_min), sd_all = sd(slope_mOD_min),
              cv_all = 100 * sd(slope_mOD_min) / mean(slope_mOD_min), .groups = "drop")

  cv_usable <- sample_rates %>%
    filter(rate_usable) %>%
    group_by(sample_id) %>%
    summarise(n_usable = n(), mean_usable = mean(slope_mOD_min),
              sd_usable = sd(slope_mOD_min),
              cv_usable = 100 * sd(slope_mOD_min) / mean(slope_mOD_min), .groups = "drop")

  data.frame(sample_id = sample_id_order, stringsAsFactors = FALSE) %>%
    left_join(cv_all, by = "sample_id") %>%
    left_join(cv_usable, by = "sample_id") %>%
    mutate(
      excluded_wells = vapply(sample_id, function(s) {
        w <- sample_rates$well[sample_rates$sample_id == s & !sample_rates$rate_usable]
        if (length(w) == 0) "-" else paste(w, collapse = ", ")
      }, character(1)),
      fails_cv_all    = cv_all > assay_params$cv_threshold_pct,
      fails_cv_usable = !is.na(cv_usable) & cv_usable > assay_params$cv_threshold_pct
    ) %>%
    arrange(sample_id)
}

## -----------------------------------------------------------------------
## calculate_cs_activity()
## Purpose : Compute per-sample CS activity following Abcam sec 10.3, using
##           only usable replicates, corrected for background rate, scaled
##           through the standard curve, and normalized to total extracted
##           protein (rather than tissue weight).
## Inputs  : well_rates            - data.frame from compute_well_rates()
##           protein_by_sample     - data.frame from load_protein_concentrations()
##           background_per_sample - data.frame from compute_background_correction()
##           std_slope             - calibration slope (A412/nmol), from fit_standard_curve()
##           std_nmol_max          - max calibrated GSH concentration, from fit_standard_curve()
##           assay_params          - list with sample_volume_uL, dilution_factor,
##                                    homogenate_volume_uL, read_duration_min
##           plate_long            - data.frame from annotate_wells(), used to
##                                    attach family/individual/temperature
## Outputs : cs_activity - one row per sample: n_reps_used, mean_rate_mOD_min,
##           sd_rate, cv_rate, family, individual, temperature, conc_ug_mL,
##           total_protein_mg, bg_rate_mOD_min, corrected_mOD_min,
##           activity_mU_per_uL, total_mU_in_homogenate,
##           activity_mU_per_mg_protein, within_std_range
## -----------------------------------------------------------------------
calculate_cs_activity <- function(well_rates, protein_by_sample, background_per_sample,
                                   std_slope, std_nmol_max, assay_params, plate_long) {
  sample_rates <- well_rates %>% filter(well_type == "sample") %>% arrange(sample_id, well)

  sample_rates %>%
    filter(rate_usable) %>%
    group_by(sample_id) %>%
    summarise(n_reps_used = n(), mean_rate_mOD_min = mean(slope_mOD_min),
              sd_rate = sd(slope_mOD_min),
              cv_rate = 100 * sd(slope_mOD_min) / mean(slope_mOD_min), .groups = "drop") %>%
    left_join(plate_long %>% filter(well_type == "sample") %>%
                distinct(sample_id, family, individual, temperature),
              by = "sample_id") %>%
    left_join(protein_by_sample %>% select(sample_id, conc_ug_mL, total_protein_mg),
              by = "sample_id") %>%
    left_join(background_per_sample %>% select(sample_id, bg_rate_flat, n_bg_flat),
              by = "sample_id") %>%
    mutate(
      bg_rate_mOD_min       = ifelse(is.na(bg_rate_flat), 0, bg_rate_flat),
      corrected_mOD_min     = mean_rate_mOD_min - bg_rate_mOD_min,
      rate_OD_min           = corrected_mOD_min / 1000,
      nmol_per_min          = rate_OD_min / std_slope,
      activity_mU_per_uL    = nmol_per_min / assay_params$sample_volume_uL *
                              assay_params$dilution_factor,
      total_mU_in_homogenate = activity_mU_per_uL * assay_params$homogenate_volume_uL,
      activity_mU_per_mg_protein = total_mU_in_homogenate / total_protein_mg,
      nmol_in_window        = nmol_per_min * (assay_params$read_duration_min),
      within_std_range      = nmol_in_window <= std_nmol_max
    ) %>%
    arrange(sample_id)
}

## -----------------------------------------------------------------------
## build_results_table()
## Purpose : Assemble the final, presentation-ready per-sample results
##           table from cs_activity, cv_summary and baseline diagnostics,
##           with a plain-language Interpretation column.
## Inputs  : cs_activity          - data.frame from calculate_cs_activity()
##           cv_summary           - data.frame from compute_replicate_cv()
##           baseline_per_sample  - data.frame from compute_baseline_diagnostics()
##           assay_params         - list with cv_threshold_pct, homogenate_volume_uL
## Outputs : results_table - formatted data.frame, ready for kable()/write.csv()
## -----------------------------------------------------------------------
build_results_table <- function(cs_activity, cv_summary, baseline_per_sample, assay_params) {
  cs_activity %>%
    left_join(cv_summary %>% select(sample_id, n_all, cv_all), by = "sample_id") %>%
    left_join(baseline_per_sample %>% select(sample_id, n_elevated), by = "sample_id") %>%
    transmute(
      Sample                       = sample_id,
      Family                       = family,
      Individual                   = individual,
      Temperature                  = temperature,
      `Protein conc (ug/mL)`       = round(conc_ug_mL, 1),
      `Total protein (mg)`         = round(total_protein_mg, 3),
      `Reps used`                  = paste0(n_reps_used, "/", n_all),
      `CV all reps (%)`            = round(cv_all, 1),
      `CV used reps (%)`           = round(cv_rate, 1),
      `Rate (mA412/min)`           = round(mean_rate_mOD_min, 2),
      `BG rate (mA412/min)`        = round(bg_rate_mOD_min, 3),
      `Corrected rate (mA412/min)` = round(corrected_mOD_min, 2),
      `Activity (mU/uL)`           = round(activity_mU_per_uL, 4),
      `Activity (mU/mg protein)`   = round(activity_mU_per_mg_protein, 3),
      `CV flag`                    = ifelse(cv_all > assay_params$cv_threshold_pct,
                                            paste0("FAIL >", assay_params$cv_threshold_pct, "%"), "pass"),
      `Elevated baseline reps`     = paste0(n_elevated, "/3"),
      Interpretation               = case_when(
        n_elevated == 3 ~ "DO NOT USE - no clean replicate",
        n_elevated > 0  ~ "caution - some reps compromised",
        cv_rate > assay_params$cv_threshold_pct | is.na(cv_rate) ~ "caution - imprecise",
        TRUE ~ "usable"
      )
    )
}
```

### QC summary and narrative generation

Assembles the per-well QC summary table and auto-generates the plain-language QC findings bullets used in the QC SUMMARY section below.

```{r pipeline-fns-summary}
## -----------------------------------------------------------------------
## build_qc_summary_table()
## Purpose : Consolidate every QC check performed across the pipeline into a
##           single two-column (check, value) table, suitable for kable() and
##           for writing to qc_summary.csv. Mirrors the individual find
##           functions but as compact, at-a-glance figures rather than prose.
## Inputs  : layout_wells, reconcile_summary, std_curve, well_diagnostics,
##           well_rates, cv_summary, baseline_per_sample, pos_control,
##           assay_params
## Outputs : qc_summary - data.frame with columns check, value (both chr)
## -----------------------------------------------------------------------
build_qc_summary_table <- function(layout_wells, reconcile_summary, std_curve,
                                    well_diagnostics, well_rates, cv_summary,
                                    baseline_per_sample, pos_control, assay_params) {
  standard_summary  <- std_curve$standard_summary
  standards_flagged <- std_curve$standards_flagged
  standard_drift    <- std_curve$standard_drift
  sample_rates      <- well_rates %>% filter(well_type == "sample")

  data.frame(
    check = c(
      "Occupied wells in layout",
      "Layout wells missing from absorbance CSV",
      "Max disagreement between absorbance CSV and full report",
      "Standard concentrations with replicate CV > threshold",
      "Standard wells flagged as outliers",
      "Standard curve R^2 (outlier-excluded, replicate level)",
      "Standard curve R^2 (all wells, replicate level)",
      "Standard wells that LOST signal over the run",
      "Standards above photometric linearity ceiling",
      "Background control wells behaving as active reactions",
      "Sample wells with a decreasing trace",
      "Sample wells with an elevated starting A412",
      "Samples with ALL THREE replicates baseline-compromised",
      "Sample wells with a read glitch",
      "Sample wells usable for rate extraction",
      "Samples with technical CV > threshold (all reps)",
      "Samples with technical CV > threshold (usable reps)",
      "Positive control replicates rising and linear"
    ),
    value = c(
      nrow(layout_wells),
      reconcile_summary$n_wells_layout_not_csv,
      sprintf("%.g", reconcile_summary$max_disagreement),
      paste(standard_summary$std_nmol[standard_summary$cv_pct > assay_params$cv_threshold_pct],
            collapse = ", "),
      paste0(sum(standards_flagged$is_outlier), "/", nrow(standards_flagged), " (",
             paste(standards_flagged$well[standards_flagged$is_outlier], collapse = ", "), ")"),
      sprintf("%.4f", std_curve$std_r2),
      sprintf("%.4f", summary(std_curve$fit_all_wells)$r.squared),
      paste0(sum(standard_drift$drift < 0), "/", nrow(standard_drift)),
      paste(standard_summary$std_nmol[standard_summary$mean_od > assay_params$od_linear_max],
            collapse = ", "),
      paste0(sum(well_diagnostics$flag_bg_active), "/",
             sum(well_diagnostics$well_type == "background")),
      paste0(sum(well_diagnostics$flag_decreasing & well_diagnostics$well_type == "sample"), "/",
             sum(well_diagnostics$well_type == "sample")),
      paste0(sum(well_diagnostics$flag_high_baseline), "/",
             sum(well_diagnostics$well_type == "sample"), " (",
             paste(well_diagnostics$well[well_diagnostics$flag_high_baseline], collapse = ", "), ")"),
      paste0(sum(baseline_per_sample$n_elevated == baseline_per_sample$n), "/",
             nrow(baseline_per_sample), " (",
             paste(baseline_per_sample$sample_id[baseline_per_sample$n_elevated == baseline_per_sample$n],
                   collapse = ", "), ")"),
      paste0(sum(well_diagnostics$flag_glitch & well_diagnostics$well_type == "sample"), "/",
             sum(well_diagnostics$well_type == "sample")),
      paste0(sum(sample_rates$rate_usable), "/", nrow(sample_rates)),
      paste0(sum(cv_summary$fails_cv_all), "/", nrow(cv_summary)),
      paste0(sum(cv_summary$fails_cv_usable), "/", nrow(cv_summary)),
      paste0(sum(pos_control$net_change > 0 & pos_control$r2 > 0.99), "/", nrow(pos_control))
    ),
    stringsAsFactors = FALSE
  )
}

## -----------------------------------------------------------------------
## generate_qc_findings()
## Purpose : Walk the QC/CV/anomaly objects for a plate and emit a
##           consistent, generic-language markdown bullet list of findings
##           (decreasing/glitched/high-baseline wells, background-control
##           mix-ups, standard-curve QC failures, CV failures, disqualified
##           samples). Intended to seed the SUMMARY section of a plate's
##           .Rmd; a short hand-written subsection for plate-specific
##           interpretation should follow it.
## Inputs  : well_diagnostics    - data.frame from compute_well_diagnostics()
##           well_rates          - data.frame from compute_well_rates()
##           cv_summary          - data.frame from compute_replicate_cv()
##           baseline_per_sample - data.frame from compute_baseline_diagnostics()
##           std_curve           - list from fit_standard_curve()
##           reconcile_summary   - list from reconcile_kinetic_sources()$summary
##           assay_params        - list with cv_threshold_pct, od_linear_max
## Outputs : character vector, one markdown bullet per finding (empty
##           findings are omitted, so a clean plate returns a short list)
## -----------------------------------------------------------------------
generate_qc_findings <- function(well_diagnostics, well_rates, cv_summary,
                                  baseline_per_sample, std_curve,
                                  reconcile_summary, assay_params) {
  bullets <- character(0)

  # Source reconciliation
  if (reconcile_summary$n_wells_layout_not_csv > 0 ||
      (!is.na(reconcile_summary$max_disagreement) && reconcile_summary$max_disagreement >= 1e-9)) {
    bullets <- c(bullets, sprintf(
      "**Kinetic source disagreement.** %d layout well(s) missing from the absorbance CSV; max disagreement between absorbance CSV and full report was %.4g A412.",
      reconcile_summary$n_wells_layout_not_csv, reconcile_summary$max_disagreement))
  } else {
    bullets <- c(bullets, sprintf(
      "**Kinetic sources agree.** absorbance CSV and full report agree exactly on all %d shared well x timepoint readings; all layout wells are present.",
      reconcile_summary$n_shared_readings))
  }

  # Sample wells with decreasing kinetics
  dec_samples <- well_diagnostics %>% filter(well_type == "sample", flag_decreasing)
  if (nrow(dec_samples) > 0) {
    bullets <- c(bullets, sprintf(
      "**Decreasing kinetics.** %d sample well(s) end lower than (or equal to) their starting A412 and are excluded from rate extraction: %s.",
      nrow(dec_samples), paste(dec_samples$well, collapse = ", ")))
  }

  # Read glitches
  glitch_samples <- well_diagnostics %>% filter(well_type == "sample", flag_glitch)
  if (nrow(glitch_samples) > 0) {
    bullets <- c(bullets, sprintf(
      "**Read glitches.** %d sample well(s) contain a single read-to-read step inconsistent with its neighbours: %s (only disqualifying if it falls inside the fitted rate window).",
      nrow(glitch_samples), paste(glitch_samples$well, collapse = ", ")))
  }

  # Elevated baseline
  high_baseline <- well_diagnostics %>% filter(well_type == "sample", flag_high_baseline)
  if (nrow(high_baseline) > 0) {
    bullets <- c(bullets, sprintf(
      "**Elevated starting absorbance.** %d sample well(s) start above the baseline threshold (A412 > %s), indicating pre-existing thiol/contamination before the reaction began: %s.",
      nrow(high_baseline), assay_params$sample_baseline_od, paste(high_baseline$well, collapse = ", ")))
  }

  # Samples fully disqualified by baseline
  disqualified <- baseline_per_sample %>% filter(n_elevated == n)
  if (nrow(disqualified) > 0) {
    bullets <- c(bullets, sprintf(
      "**Sample(s) with no clean replicate.** Every replicate of %s has an elevated starting A412; these samples have no usable measurement on this plate regardless of CV.",
      paste(disqualified$sample_id, collapse = ", ")))
  }

  # Over-range wells
  over_range <- well_diagnostics %>% filter(flag_over_range)
  if (nrow(over_range) > 0) {
    bullets <- c(bullets, sprintf(
      "**Over-range wells.** %d well(s) exceed the photometric linearity ceiling (A412 > %s): %s.",
      nrow(over_range), assay_params$od_linear_max, paste(over_range$well, collapse = ", ")))
  }

  # Background wells behaving as active reactions
  bg_active <- well_diagnostics %>% filter(well_type == "background", flag_bg_active)
  if (nrow(bg_active) > 0) {
    bullets <- c(bullets, sprintf(
      "**Background control wells behaving as active reactions.** %d of %d background wells show a starting A412 or drift inconsistent with a flat, inactive well: %s.",
      nrow(bg_active), sum(well_diagnostics$well_type == "background"),
      paste(bg_active$well, collapse = ", ")))
  }

  # Standard curve QC
  std_summary <- std_curve$standard_summary
  cv_fail_std <- std_summary %>% filter(cv_pct > assay_params$cv_threshold_pct)
  if (nrow(cv_fail_std) > 0) {
    bullets <- c(bullets, sprintf(
      "**GSH standard curve replicate imprecision.** %d of %d concentrations exceed %s%% CV: %s nmol/well.",
      nrow(cv_fail_std), nrow(std_summary), assay_params$cv_threshold_pct,
      paste(cv_fail_std$std_nmol, collapse = ", ")))
  }
  n_std_outliers <- sum(std_curve$standards_flagged$is_outlier)
  if (n_std_outliers > 0) {
    bullets <- c(bullets, sprintf(
      "**Standard outlier wells.** %d of %d standard wells deviate from their triplicate median by more than %s A412: %s.",
      n_std_outliers, nrow(std_curve$standards_flagged), assay_params$std_outlier_od,
      paste(std_curve$standards_flagged$well[std_curve$standards_flagged$is_outlier], collapse = ", ")))
  }
  over_range_std <- std_summary %>% filter(mean_od > assay_params$od_linear_max)
  if (nrow(over_range_std) > 0) {
    bullets <- c(bullets, sprintf(
      "**Standards above the photometric linear range.** %s nmol/well read above A412 %s; the calibration is trustworthy only below these concentrations.",
      paste(over_range_std$std_nmol, collapse = ", "), assay_params$od_linear_max))
  }
  n_drift_neg <- sum(std_curve$standard_drift$drift < 0)
  if (n_drift_neg > 0) {
    bullets <- c(bullets, sprintf(
      "**Standard signal decay.** %d of %d standard wells lost signal over the run (TNB instability); the standard curve is read at t = 0.",
      n_drift_neg, nrow(std_curve$standard_drift)))
  }
  bullets <- c(bullets, sprintf(
    "**Standard curve fit (outlier-excluded):** slope = %.5f A412/nmol, R^2 = %.4f (all-wells R^2 = %.4f).",
    std_curve$std_slope, std_curve$std_r2, summary(std_curve$fit_all_wells)$r.squared))

  # Technical replicate CV failures
  fail_all <- cv_summary %>% filter(fails_cv_all)
  if (nrow(fail_all) > 0) {
    bullets <- c(bullets, sprintf(
      "**Technical replicate CV > %s%% (all replicates).** %d of %d samples: %s.",
      assay_params$cv_threshold_pct, nrow(fail_all), nrow(cv_summary),
      paste(sprintf("%s (%.0f%%)", fail_all$sample_id, fail_all$cv_all), collapse = ", ")))
  } else {
    bullets <- c(bullets, sprintf(
      "**Technical replicate CV.** All %d samples are within %s%% CV across all three replicates.",
      nrow(cv_summary), assay_params$cv_threshold_pct))
  }
  fail_usable <- cv_summary %>% filter(fails_cv_usable)
  if (nrow(fail_usable) > 0) {
    bullets <- c(bullets, sprintf(
      "**Technical replicate CV > %s%% persists after excluding flagged replicates.** %s.",
      assay_params$cv_threshold_pct,
      paste(sprintf("%s (%.0f%%, n=%d)", fail_usable$sample_id, fail_usable$cv_usable,
                    fail_usable$n_usable), collapse = ", ")))
  }

  bullets
}
```


## Assay parameters

All assay constants and QC thresholds are collected here so nothing downstream
hard-codes a number.

```{r assay-params}
assay_params <- list(
  sample_volume_uL      = 2,     # homogenate volume per reaction well (V in Abcam formula)
  buffer_volume_uL      = 48,    # Assay Buffer 7 added to reach 50 uL pre-Reaction Mix
  dilution_factor       = 1,     # D; layout `df.0` = undiluted homogenate
  homogenate_volume_uL  = 350,   # MEASURED buffer volume tissue was homogenized in (see note 3)
  read_wavelength_nm    = 412,
  read_interval_min     = 2,
  read_duration_min     = 40,
  rate_window_n         = 5,     # points per sliding regression window (5 pts = 8 min)
  rate_min_r2           = 0.80,  # min R^2 for a rate window to be trusted (see note 4)
  glitch_excess_od      = 0.02,  # |step - mean of neighbouring steps| flagging a read glitch
  cv_threshold_pct      = 15,    # technical-replicate CV QC threshold
  od_linear_max         = 1.5,   # upper A412 bound for reliable photometry
  std_outlier_od        = 0.15,  # |deviation from triplicate median| flagging a standard well
  sample_baseline_od    = 0.35,  # max acceptable t0 A412 for a sample well (see note 5)
  bg_flat_od_max        = 0.15,  # a well-behaved background well starts below this A412
  bg_flat_drift_max     = 0.05   # ...and drifts less than this over the whole run
)

cat("--- assay_params: assay constants and QC thresholds ---\n\n")
str(assay_params)
```

## Output directory

```{r output-dir}
# Output directory for this analysis (matches this file's name, per ../code/README.md)
output_dir <- "../outputs/Gen5-20260811-mgig-citrate_synthase-F05-ambient"
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

cat("--- output_dir: destination for all figures and tables ---\n")
str(output_dir)
```

# DATA

Data are read from the local repo (`../data/raw_absorbance/`) so this document
renders before/after the files are pushed to GitHub.

## Plate layout

```{r data-layout}
data_dir <- "../data/raw_absorbance"
run_stem <- "Gen5-20260811-mgig-citrate_synthase-F05-ambient"

layout_wells <- parse_plate_layout(file.path(data_dir, paste0("layout-", run_stem, ".csv")))

cat("Layout format detected:", attr(layout_wells, "layout_format"), "\n")
cat("Occupied wells in layout:", nrow(layout_wells), "\n")

cat("\n--- layout_wells: one row per occupied well, with its descriptive label ---\n\n")
str(layout_wells)
```

## Kinetic readers

```{r kinetic-readers}
absorbance_csv <- read_absorbance_csv(file.path(data_dir, paste0("absorbance-", run_stem, ".csv")))
full_report    <- read_full_report(file.path(data_dir, paste0("full_report-", run_stem, ".txt")))

cat("absorbance CSV : ", dplyr::n_distinct(absorbance_csv$well), " wells x ",
    dplyr::n_distinct(absorbance_csv$time_min), " timepoints\n", sep = "")
cat("full report    : ", dplyr::n_distinct(full_report$well), " wells x ",
    dplyr::n_distinct(full_report$time_min), " timepoints\n", sep = "")

cat("\n--- absorbance_csv: long-format readings from absorbance-*.csv ---\n\n")
str(absorbance_csv)
cat("\n--- full_report: long-format readings from full_report-*.txt ---\n\n")
str(full_report)
```

## Cross-check against the full report

`absorbance-*.csv` is used for every well; no gap-fill is needed now that the
re-exported file covers the full plate. `full_report-*.txt` is used only to
confirm the re-export is trustworthy: every well it reports should also be in
the CSV, every well the layout expects should be in the CSV, and every
reading the two files share should agree exactly.

```{r reconcile-sources}
recon <- reconcile_kinetic_sources(absorbance_csv, full_report, layout_wells)
plate_readings <- recon$plate_readings

cat("Wells in absorbance CSV          :", recon$summary$n_wells_csv, "\n")
cat("Wells in full report             :", recon$summary$n_wells_report, "\n")
cat("Wells in full report not in CSV  :", recon$summary$n_wells_report_not_csv, "\n")
cat("Layout wells missing from CSV    :", recon$summary$n_wells_layout_not_csv, "\n")
cat("CSV wells not in layout          :", recon$summary$n_wells_csv_not_layout, "\n")
cat("Shared well x timepoints         :", recon$summary$n_shared_readings, "\n")
cat("Max |CSV - report| disagreement  :", recon$summary$max_disagreement, "\n")

cat("\n--- recon$overlap_check: per-reading comparison of the two raw files ---\n\n")
str(recon$overlap_check)
cat("\n--- plate_readings: readings for all occupied wells, from absorbance_csv ---\n\n")
str(plate_readings)
```

## Annotate wells and parse metadata

```{r annotate-wells}
# Tolerate the `citrate_synthasse` typo in the D1-D3 labels (see note 6)
plate_long <- annotate_wells(plate_readings, layout_wells,
                              label_fixes = c("citrate_synthasse" = "citrate_synthase"))

cat("Wells per type:\n")
print(plate_long %>% distinct(well, well_type) %>% count(well_type, name = "n_wells") %>%
        as.data.frame(), row.names = FALSE)
cat("\nSamples:", paste(sort(unique(na.omit(plate_long$sample_id))), collapse = ", "), "\n")
cat("Standards (nmol/well):", paste(sort(unique(na.omit(plate_long$std_nmol))), collapse = ", "), "\n")

cat("\n--- plate_long: fully annotated long-format plate, one row per well x timepoint ---\n\n")
str(plate_long)
```

## Protein concentration (normalization factor)

Activity is normalized to total extracted protein rather than tissue weight
(note 3). Protein concentration for each sample was measured on a separate
BCA/Bradford-style plate (595 nm) and supplied as two CSV exports covering
different subsets of samples across families F05 and F07; only the F05
ambient samples analyzed on this plate are relevant here, so the two files
are concatenated and then filtered down to this plate's 8 sample IDs.

```{r protein-concentration}
protein_files <- c(
  "../data/BSA/raw_absorbance/sample_protein_concentrations_1.csv",
  "../data/BSA/raw_absorbance/sample_protein_concentrations_2.csv"
)

plate_sample_ids <- sort(unique(na.omit(plate_long$sample_id)))

protein_by_sample <- load_protein_concentrations(protein_files, plate_sample_ids,
                                                  assay_params$homogenate_volume_uL) %>%
  left_join(plate_long %>% filter(well_type == "sample") %>%
              distinct(sample_id, family, individual, temperature),
            by = "sample_id")

cat("Samples matched to a protein concentration record:", nrow(protein_by_sample),
    "/", length(plate_sample_ids), "\n")
cat("Protein concentration range (ug/mL):",
    paste(range(protein_by_sample$conc_ug_mL), collapse = " - "), "\n")
cat("Total protein per homogenate range (mg):",
    paste(round(range(protein_by_sample$total_protein_mg), 3), collapse = " - "), "\n\n")
print(protein_by_sample %>% select(sample_id, source_file, conc_ug_mL, total_protein_mg) %>%
        as.data.frame(), row.names = FALSE)

cat("\n--- protein_by_sample: matched protein concentration and total protein per homogenate ---\n\n")
str(protein_by_sample)
```

# KINETIC TRACES

Every trace is plotted before any rate is extracted. In this assay the
**shape** of the trace is the primary diagnostic: CS activity must produce a
monotonically *rising* A412, and only a genuinely linear stretch may be used
for a rate.

```{r plot-kinetic-traces, fig.width = 11, fig.height = 8}
trace_plot <- ggplot(plate_long, aes(x = time_min, y = od, group = well, colour = well_type)) +
  geom_line(linewidth = 0.5) +
  geom_hline(yintercept = assay_params$od_linear_max,
             linetype = "dotted", colour = "grey40") +
  facet_wrap(~ plate_row + plate_col, ncol = 12,
             labeller = function(d) list(paste0(d$plate_row, d$plate_col))) +
  scale_colour_manual(values = c(sample = "#08519c", background = "#a63603",
                                 standard = "#238b45", positive_control = "#6a51a3")) +
  labs(title = "A412 kinetic traces, all occupied wells",
       subtitle = paste0("Dotted line = assumed photometric linearity ceiling (A412 = ",
                         assay_params$od_linear_max,
                         "). Rising traces only are usable for rate extraction."),
       x = "Time (min)", y = "A412", colour = "Well type") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        strip.text = element_text(size = 6),
        axis.text  = element_text(size = 5),
        legend.position = "bottom")

ggsave(file.path(output_dir, "kinetic_traces_all_wells.png"), trace_plot,
       width = 11, height = 8, dpi = 300)

cat("--- trace_plot: ggplot object structure ---\n\n")
summary(trace_plot)

trace_plot
```

# ANOMALY DETECTION

## Per-well trace diagnostics

Four independent checks per well, all computed from the raw trace before any
rate fitting:

1. **Direction** — net A412 change from first to last read. A CS-containing
   well must be positive.
2. **Monotonicity** — fraction of the 20 read-to-read intervals that rise.
3. **Over-range** — any read above the photometric linearity ceiling.
4. **Discontinuity** — a single read-to-read step that departs from its two
   *neighbouring* steps, catching read glitches rather than biology. Comparing
   each step against its immediate neighbours (rather than against the well's
   overall median step) matters here because every trace decelerates: the first
   interval of a decelerating trace is legitimately the largest of the run and
   a median-based rule flags it spuriously.

```{r anomaly-diagnostics}
well_diagnostics <- compute_well_diagnostics(plate_long, assay_params)

cat("Flag counts across all", nrow(well_diagnostics), "wells:\n")
print(well_diagnostics %>%
        summarise(decreasing = sum(flag_decreasing), over_range = sum(flag_over_range),
                  glitch = sum(flag_glitch), bg_active = sum(flag_bg_active),
                  high_baseline = sum(flag_high_baseline),
                  clean = sum(n_flags == 0)) %>% as.data.frame(), row.names = FALSE)

# The starting A412 separates clean wells from compromised ones across every
# well type, so it is tabulated explicitly rather than only used as a flag
cat("\nStarting A412 (t = 0) by well type:\n")
print(well_diagnostics %>% group_by(well_type) %>%
        summarise(n = n(), min = min(od_first), median = median(od_first),
                  max = max(od_first), .groups = "drop") %>%
        as.data.frame(), row.names = FALSE, digits = 3)

cat("\n--- well_diagnostics: per-well trace shape diagnostics and anomaly flags ---\n\n")
str(well_diagnostics)
```

## Anomalous wells

```{r anomaly-table}
anomalies <- well_diagnostics %>%
  filter(n_flags > 0) %>%
  mutate(flags = paste0(
    ifelse(flag_decreasing, "decreasing ", ""),
    ifelse(flag_over_range, "over-range ", ""),
    ifelse(flag_glitch,        "glitch ",        ""),
    ifelse(flag_bg_active,     "BG-active ",     ""),
    ifelse(flag_high_baseline, "high-baseline ", "")
  ) %>% trimws()) %>%
  arrange(well_type, plate_row, plate_col)

kable(anomalies %>%
        select(well, well_type, sample_id, std_nmol, od_first, od_last,
               net_change, frac_rising, od_max, flags),
      digits = c(0, 0, 0, 0, 3, 3, 3, 2, 3, 0),
      col.names = c("Well", "Type", "Sample", "STD nmol", "A412 t0", "A412 t40",
                    "Net dA412", "Frac rising", "Max A412", "Flags"),
      caption = "Wells failing one or more trace-shape checks")

write.csv(anomalies %>% select(-flag_decreasing, -flag_over_range, -flag_glitch,
                               -flag_bg_active, -flag_high_baseline, -n_flags),
          file.path(output_dir, "anomalous_wells.csv"), row.names = FALSE)

cat("\n--- anomalies: wells failing >=1 trace-shape check ---\n\n")
str(anomalies)
```

## Background control wells behaving as reactions

This is the single largest anomaly on the plate and is called out separately.
Background wells receive Background Control Mix, which contains **no CS
Substrate Mix**, so they cannot sustain a CS reaction and must read flat.

```{r bg-anomaly}
bg_summary <- well_diagnostics %>%
  filter(well_type == "background") %>%
  mutate(behaviour = ifelse(flag_bg_active, "ANOMALOUS", "flat (expected)")) %>%
  arrange(sample_id, well)

cat("Background wells by behaviour:\n")
print(bg_summary %>% count(behaviour) %>% as.data.frame(), row.names = FALSE)

cat("\nBackground triplicates, per sample:\n")
print(bg_summary %>%
        group_by(sample_id) %>%
        summarise(n_flat = sum(!flag_bg_active), n_anomalous = sum(flag_bg_active),
                  anomalous_wells = paste(well[flag_bg_active], collapse = ", "),
                  .groups = "drop") %>% as.data.frame(), row.names = FALSE)

kable(bg_summary %>% select(well, sample_id, od_first, od_last, net_change, behaviour),
      digits = c(0, 0, 3, 3, 3, 0),
      col.names = c("Well", "Sample", "A412 t0", "A412 t40", "Net dA412", "Behaviour"),
      caption = "Background control wells: starting A412 and total drift")

cat("\n--- bg_summary: background well behaviour classification ---\n\n")
str(bg_summary)
```

## Starting absorbance as the master diagnostic

Every anomaly above — decreasing samples, active background wells, scattered
standards — occurs in wells that *started* high. A412 at t = 0 reports TNB
already present before the reaction began, so an elevated baseline means the
well was contaminated or mis-dispensed. This matters for interpretation: a
well that starts high and then declines is not measuring CS activity at all,
and a rate extracted from it is an artifact of whatever is decaying.

```{r baseline-diagnostic}
baseline_diag       <- compute_baseline_diagnostics(well_diagnostics, assay_params)
baseline_check      <- baseline_diag$baseline_check
baseline_per_sample <- baseline_diag$baseline_per_sample

cat("Sample and background wells by starting A412:\n")
print(baseline_check %>%
        group_by(well_type, baseline) %>%
        summarise(n = n(), n_decreasing = sum(flag_decreasing), .groups = "drop") %>%
        as.data.frame(), row.names = FALSE)

# Does an elevated baseline track with a declining trace?
baseline_vs_direction <- baseline_check %>%
  summarise(
    elevated_and_decreasing = sum(baseline == "elevated" & flag_decreasing),
    elevated_total          = sum(baseline == "elevated"),
    normal_and_decreasing   = sum(baseline == "normal" & flag_decreasing),
    normal_total            = sum(baseline == "normal")
  )

cat("\nElevated baseline & decreasing:", baseline_vs_direction$elevated_and_decreasing,
    "/", baseline_vs_direction$elevated_total, "\n")
cat("Normal baseline & decreasing  :", baseline_vs_direction$normal_and_decreasing,
    "/", baseline_vs_direction$normal_total, "\n")

cat("\nSample wells with elevated baseline:\n")
print(baseline_check %>% filter(well_type == "sample", baseline == "elevated") %>%
        select(well, sample_id, od_first, net_change) %>%
        as.data.frame(), row.names = FALSE, digits = 3)

cat("\nElevated-baseline replicates per sample:\n")
print(as.data.frame(baseline_per_sample), row.names = FALSE, digits = 3)

cat("\n--- baseline_check: starting-A412 classification per sample/background well ---\n\n")
str(baseline_check)
cat("\n--- baseline_per_sample: count of compromised replicates per sample ---\n\n")
str(baseline_per_sample)
```

```{r plot-baseline, fig.width = 9, fig.height = 6}
baseline_plot <- ggplot(baseline_check,
                        aes(x = od_first, y = net_change, colour = well_type)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = assay_params$sample_baseline_od,
             linetype = "dotted", colour = "grey40") +
  geom_point(size = 3, alpha = 0.85) +
  geom_text(aes(label = well), vjust = -0.9, size = 2.6, show.legend = FALSE) +
  scale_colour_manual(values = c(sample = "#08519c", background = "#a63603"),
                      name = "Well type") +
  labs(title = "Starting absorbance predicts trace direction",
       subtitle = paste0("Clean sample wells cluster tightly at low A412 with a positive net change.\n",
                         "Wells starting above A412 = ", assay_params$sample_baseline_od,
                         " (dotted) scatter and mostly decline: TNB was\n",
                         "already present before the read began, so their traces are not enzymology."),
       x = "A412 at t = 0", y = "Net dA412 over 40 min") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        legend.position = "bottom")

ggsave(file.path(output_dir, "baseline_vs_direction.png"), baseline_plot,
       width = 9, height = 6, dpi = 300)

cat("--- baseline_plot: ggplot object structure ---\n\n")
summary(baseline_plot)

baseline_plot
```

# GSH STANDARD CURVE

The GSH standard is a **stoichiometric endpoint**, not an enzymatic reaction:
free thiol reduces DTNB immediately on mixing. The standard curve is therefore
read from a single timepoint, and t = 0 is used because TNB<sup>2-</sup>
degrades over the run (quantified below).

## Fit the standard curve

Three fits are compared, so the effect of the outlier wells on the calibration
is explicit rather than buried:

- **all wells** — every standard replicate, no exclusions
- **concentration means** — the conventional fit, on per-concentration means
- **outlier-excluded** — replicate-level fit after dropping flagged wells

The outlier-excluded fit is used for activity calculation. Standards are
flagged by deviation from their own triplicate **median**, which is not
dragged by a single contaminated well the way a mean is.

```{r standard-curve-fit}
std_curve <- fit_standard_curve(plate_long, assay_params)

standards_t0       <- std_curve$standards_t0
standard_drift     <- std_curve$standard_drift
standard_summary   <- std_curve$standard_summary
standards_flagged  <- std_curve$standards_flagged
fit_all_wells      <- std_curve$fit_all_wells
fit_comparison     <- std_curve$fit_comparison
std_slope          <- std_curve$std_slope
std_intercept      <- std_curve$std_intercept
std_r2             <- std_curve$std_r2
std_nmol_max       <- std_curve$std_nmol_max

cat("Standard wells:", nrow(standards_t0), "at", dplyr::n_distinct(standards_t0$std_nmol),
    "concentrations\n")
cat("Standard drift over 40 min (dA412): median",
    sprintf("%.3f", median(standard_drift$drift)),
    "| range", paste(sprintf("%.3f", range(standard_drift$drift)), collapse = " to "), "\n")
cat("Standard wells that DECREASED:", sum(standard_drift$drift < 0), "/", nrow(standard_drift), "\n")

kable(standard_drift, digits = c(0, 0, 3, 3, 3),
      col.names = c("Well", "GSH (nmol)", "A412 t0", "A412 t40", "Drift"),
      caption = "GSH standard wells lose signal over the kinetic run (TNB instability)")

kable(standard_summary %>% select(std_nmol, n, mean_od, median_od, sd_od, se_od,
                                  cv_pct, net_od, od_per_nmol),
      digits = c(0, 0, 3, 3, 3, 3, 1, 3, 4),
      col.names = c("GSH (nmol)", "n", "Mean A412", "Median A412", "SD", "SEM",
                    "CV (%)", "Net A412", "A412/nmol"),
      caption = "GSH standard curve summary statistics (t = 0)")

cat("Standard concentrations exceeding the", assay_params$cv_threshold_pct,
    "% CV threshold:",
    paste(standard_summary$std_nmol[standard_summary$cv_pct > assay_params$cv_threshold_pct],
          collapse = ", "), "\n")

kable(standards_flagged %>% select(well, std_nmol, od, triplicate_median,
                                    deviation, is_outlier),
      digits = c(0, 0, 3, 3, 3, 0),
      col.names = c("Well", "GSH (nmol)", "A412", "Triplicate median",
                    "Deviation", "Outlier"),
      caption = paste0("Standard wells ranked by deviation from triplicate median (threshold ",
                       assay_params$std_outlier_od, " A412)"))

cat("Standard wells flagged as outliers:", sum(standards_flagged$is_outlier),
    "/", nrow(standards_flagged), "\n")

kable(fit_comparison, digits = c(0, 0, 5, 4, 4),
      col.names = c("Fit", "n", "Slope (A412/nmol)", "Intercept", "R-squared"),
      caption = "GSH standard curve: three fitting strategies compared")

cat("\nCalibration used downstream (outlier-excluded fit):\n")
cat("  A412 = ", sprintf("%.5f", std_slope), " * nmol + ",
    sprintf("%.4f", std_intercept), "   R^2 = ", sprintf("%.4f", std_r2), "\n", sep = "")
cat("  1 nmol thiol =", sprintf("%.5f", std_slope), "A412 units\n")

cat("\n--- standards_flagged: standard wells with median-deviation outlier flags ---\n\n")
str(standards_flagged)
cat("\n--- fit_comparison: side-by-side standard curve fit statistics ---\n\n")
str(fit_comparison)
```

## Plot the standard curve

```{r plot-standard-curve, fig.width = 9, fig.height = 6}
std_curve_plot <- ggplot(standards_flagged, aes(x = std_nmol, y = od)) +
  geom_abline(aes(slope = coef(fit_all_wells)[2], intercept = coef(fit_all_wells)[1],
                  linetype = "all wells"), colour = "grey55") +
  geom_abline(aes(slope = std_slope, intercept = std_intercept,
                  linetype = "outlier-excluded"), colour = "#238b45", linewidth = 0.8) +
  geom_point(aes(shape = is_outlier, colour = is_outlier), size = 3) +
  geom_hline(yintercept = assay_params$od_linear_max, linetype = "dotted", colour = "grey40") +
  scale_shape_manual(values = c(`FALSE` = 16, `TRUE` = 4),
                     labels = c(`FALSE` = "retained", `TRUE` = "outlier"), name = "Well") +
  scale_colour_manual(values = c(`FALSE` = "#238b45", `TRUE` = "#cb181d"),
                      labels = c(`FALSE` = "retained", `TRUE` = "outlier"), name = "Well") +
  scale_linetype_manual(values = c(`all wells` = "dashed", `outlier-excluded` = "solid"),
                        name = "Fit") +
  labs(title = "GSH standard curve, A412 at t = 0",
       subtitle = paste0("Dotted line = photometric linearity ceiling (A412 = ",
                         assay_params$od_linear_max,
                         "); the 32 and 40 nmol standards sit above it"),
       x = "GSH (nmol/well)", y = "A412") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"))

ggsave(file.path(output_dir, "gsh_standard_curve.png"), std_curve_plot,
       width = 9, height = 6, dpi = 300)

cat("--- std_curve_plot: ggplot object structure ---\n\n")
summary(std_curve_plot)

std_curve_plot
```

# RATE EXTRACTION

## Sliding-window linear fit

Abcam §9.3 directs the analyst to pick two timepoints "in the linear range".
That is implemented here as a sliding window of
`r assay_params$rate_window_n` consecutive reads
(`r (assay_params$rate_window_n - 1) * assay_params$read_interval_min` min)
scanned across each trace. Two candidate windows are recorded per well:

- **max increasing** — the steepest *positive* slope, which is what CS
  activity means. This is the rate used for activity.
- **max absolute** — the steepest slope of either sign. Gen5's own
  `Max V [412]` column uses this convention, which is why Gen5 reports large
  negative "rates" for degrading wells.

Recording both makes the divergence between them an explicit QC signal.

```{r rate-windows}
well_rates <- compute_well_rates(plate_long, well_diagnostics, assay_params)

cat("Wells with a usable increasing rate:", sum(well_rates$rate_usable),
    "/", nrow(well_rates), "\n")
cat("Wells where the steepest window is NEGATIVE (Gen5 would report a negative Max V):",
    sum(well_rates$abs_window_is_negative), "\n")

cat("\n--- well_rates: per-well max-increasing and max-absolute slope windows ---\n\n")
str(well_rates)
```

## Sample and background rates

```{r rate-tables}
sample_rates <- well_rates %>%
  filter(well_type == "sample") %>%
  arrange(sample_id, well)

background_rates <- well_rates %>%
  filter(well_type == "background") %>%
  arrange(sample_id, well)

kable(sample_rates %>% select(well, sample_id, t_start, t_end, slope_mOD_min, r2,
                              net_change, max_abs_slope_mOD_min,
                              glitch_in_window, rate_usable),
      digits = c(0, 0, 0, 0, 2, 3, 3, 2, 0, 0),
      col.names = c("Well", "Sample", "Window start (min)", "Window end (min)",
                    "Rate (mA412/min)", "R-squared", "Net dA412",
                    "Max |rate|", "Glitch in window", "Usable"),
      caption = "Sample wells: max-increasing rate window")

kable(background_rates %>% select(well, sample_id, slope_mOD_min, r2,
                                  net_change, flag_bg_active),
      digits = c(0, 0, 2, 3, 3, 0),
      col.names = c("Well", "Sample", "Rate (mA412/min)", "R-squared",
                    "Net dA412", "Anomalous"),
      caption = "Background control wells: rates should be ~0")

cat("\n--- sample_rates: rate windows for the sample wells ---\n\n")
str(sample_rates)
cat("\n--- background_rates: rate windows for the background wells ---\n\n")
str(background_rates)
```

## Positive control

```{r positive-control}
pos_control <- compute_positive_control(well_rates, std_slope, assay_params)

kable(pos_control %>% select(well, t_start, t_end, slope_mOD_min, r2, net_change),
      digits = c(0, 0, 0, 2, 4, 3),
      col.names = c("Well", "Window start (min)", "Window end (min)",
                    "Rate (mA412/min)", "R-squared", "Net dA412"),
      caption = "CS Positive Control: the plate's proof that the reaction chemistry worked")

cat("Positive control rate: mean", sprintf("%.2f", mean(pos_control$slope_mOD_min)),
    "mA412/min | CV", sprintf("%.1f", 100 * sd(pos_control$slope_mOD_min) /
                                     mean(pos_control$slope_mOD_min)), "%\n")
cat("All three replicates rising and linear:",
    all(pos_control$net_change > 0 & pos_control$r2 > 0.99), "\n")

cat("\n--- pos_control: positive control rate and R^2 per replicate ---\n\n")
str(pos_control)
```

# BACKGROUND CORRECTION

Abcam §10.2 subtracts the background reading only "if sample background control
reading is significant". Here the background *rate* is the relevant quantity,
and only the well-behaved (flat) background replicates are used to estimate it,
since the anomalous ones do not measure background.

```{r background-correction}
background_per_sample <- compute_background_correction(well_rates)

cat("Background rate from flat wells only (mA412/min):\n")
print(background_per_sample %>% as.data.frame(), row.names = FALSE, digits = 3)

cat("\nMedian flat-background rate:",
    sprintf("%.3f", median(background_per_sample$bg_rate_flat, na.rm = TRUE)),
    "mA412/min\n")
cat("Median sample rate:",
    sprintf("%.2f", median(sample_rates$slope_mOD_min)), "mA412/min\n")
cat("Background as % of sample signal:",
    sprintf("%.1f", 100 * median(background_per_sample$bg_rate_flat, na.rm = TRUE) /
                    median(sample_rates$slope_mOD_min)), "%\n")

cat("\n--- background_per_sample: per-sample background rate estimates ---\n\n")
str(background_per_sample)
```

## Background significance test

Cattau et al. (2023; legacy version of this assay,
`Cattau_v1_master.Rmd`) validated background subtraction once with an
ANOVA and then applied that conclusion to every subsequent plate. The
same formal test is run here -- a Welch two-sample t-test of well-level
rate (background vs. sample), plus an ANOVA on the OD trajectory over
time -- so the flat/active threshold classification used above is backed
by an explicit significance test rather than the threshold alone.

```{r background-significance}
bg_sample_rates <- well_rates %>%
  filter(well_type %in% c("sample", "background"))

bg_significance_test <- t.test(slope_mOD_min ~ well_type, data = bg_sample_rates)

cat("--- Welch two-sample t-test: well-level rate (background vs. sample) ---\n\n")
print(bg_significance_test)

bg_trend_anova <- anova(lm(od ~ well_type * time_min,
                            data = plate_long %>% filter(well_type %in% c("sample", "background"))))

cat("\n--- ANOVA: OD trajectory by well_type x time_min (background vs sample wells) ---\n\n")
print(bg_trend_anova)

cat("\nInterpretation: mean background-well rate is",
    sprintf("%.2f", unname(bg_significance_test$estimate["mean in group background"])),
    "mA412/min vs.", sprintf("%.2f", unname(bg_significance_test$estimate["mean in group sample"])),
    "mA412/min for sample wells (p =", format.pval(bg_significance_test$p.value, digits = 3),
    "). Background wells are statistically distinct from, and far slower\n",
    "than, reaction wells on this plate -- consistent with the per-well\n",
    "flat/active classification above and with the qualitative conclusion\n",
    "reached by Cattau et al. (2023) on the legacy assay version.\n")

cat("\n--- bg_significance_test: Welch t-test result object ---\n\n")
str(bg_significance_test)

cat("\n--- bg_trend_anova: ANOVA table for OD ~ well_type * time_min ---\n\n")
str(bg_trend_anova)
```

# TECHNICAL REPLICATE PRECISION

Coefficient of variation across the three technical replicates of each sample,
computed on the extracted rate. Two versions are reported: **all** three
replicates, and **usable replicates only** (rate window with
R<sup>2</sup> >= `r assay_params$rate_min_r2`, rising trace, no read glitch).
The comparison localizes the imprecision to specific wells. Any sample whose
CV exceeds `r assay_params$cv_threshold_pct`% on either basis is flagged below.

```{r replicate-cv}
cv_summary <- compute_replicate_cv(well_rates, protein_by_sample$sample_id, assay_params)

kable(cv_summary %>% select(sample_id, n_all, mean_all, sd_all, cv_all,
                            n_usable, mean_usable, sd_usable, cv_usable,
                            excluded_wells),
      digits = c(0, 0, 2, 2, 1, 0, 2, 2, 1, 0),
      col.names = c("Sample", "n", "Mean rate", "SD", "CV (%)",
                    "n usable", "Mean rate", "SD", "CV (%)", "Excluded wells"),
      caption = paste0("Technical replicate CV of the extracted rate (mA412/min), ",
                       "all replicates vs. usable replicates only"))

cat("\nSamples exceeding CV", paste0(assay_params$cv_threshold_pct, "%"),
    "on ALL replicates:", sum(cv_summary$fails_cv_all), "/", nrow(cv_summary), "\n")
print(cv_summary %>% filter(fails_cv_all) %>%
        select(sample_id, n_all, mean_all, sd_all, cv_all) %>%
        as.data.frame(), row.names = FALSE, digits = 4)

cat("\nSamples exceeding CV", paste0(assay_params$cv_threshold_pct, "%"),
    "on USABLE replicates:", sum(cv_summary$fails_cv_usable), "\n")
print(cv_summary %>% filter(fails_cv_usable) %>%
        select(sample_id, n_usable, mean_usable, sd_usable, cv_usable) %>%
        as.data.frame(), row.names = FALSE, digits = 4)

write.csv(cv_summary, file.path(output_dir, "technical_replicate_cv.csv"), row.names = FALSE)

cat("\n--- cv_summary: per-sample technical replicate CV, both variants ---\n\n")
str(cv_summary)
```

## Plot replicate spread

```{r plot-replicate-cv, fig.width = 9, fig.height = 6}
cv_plot_data <- sample_rates %>%
  left_join(cv_summary %>% select(sample_id, cv_all), by = "sample_id")

cv_plot <- ggplot(cv_plot_data, aes(x = sample_id, y = slope_mOD_min)) +
  geom_point(aes(colour = rate_usable, shape = rate_usable),
             size = 3, position = position_jitter(width = 0.12, height = 0)) +
  stat_summary(fun = mean, geom = "crossbar", width = 0.5,
               colour = "grey30", linewidth = 0.3) +
  geom_text(data = cv_summary, aes(x = sample_id, y = Inf,
                                   label = sprintf("CV %.0f%%", cv_all)),
            vjust = 1.6, size = 3,
            colour = ifelse(cv_summary$fails_cv_all, "#cb181d", "grey30")) +
  scale_colour_manual(values = c(`TRUE` = "#08519c", `FALSE` = "#cb181d"),
                      labels = c(`TRUE` = "usable", `FALSE` = "flagged"), name = "Replicate") +
  scale_shape_manual(values = c(`TRUE` = 16, `FALSE` = 4),
                     labels = c(`TRUE` = "usable", `FALSE` = "flagged"), name = "Replicate") +
  labs(title = "Technical replicate spread in extracted rate",
       subtitle = paste0("Crossbar = mean of all three replicates; CV computed on all three. ",
                         "Threshold ", assay_params$cv_threshold_pct, "%."),
       x = "Sample", y = "Rate (mA412/min)") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        axis.text.x = element_text(angle = 30, hjust = 1))

ggsave(file.path(output_dir, "technical_replicate_cv.png"), cv_plot,
       width = 9, height = 6, dpi = 300)

cat("--- cv_plot: ggplot object structure ---\n\n")
summary(cv_plot)

cv_plot
```

# CITRATE SYNTHASE ACTIVITY

## Calculation

Following Abcam §10.3:

`Sample CS activity = B / (dT x V) x D`  nmol/min/µL (= mU/µL)

Rearranged for a kinetic slope, where the standard curve converts A412 into
nmol of thiol:

1. **Rate** (A412/min) = max-increasing sliding-window slope / 1000
2. **Background-corrected rate** = sample rate - mean flat background rate
3. **B/dT** (nmol/min) = background-corrected rate / standard curve slope
   (`r sprintf("%.5f", std_slope)` A412 per nmol, from the outlier-excluded fit)
4. **Activity** (mU/µL) = (B/dT) / *V* x *D*, with
   *V* = `r assay_params$sample_volume_uL` µL and
   *D* = `r assay_params$dilution_factor`
5. **Protein-normalized** (mU/mg protein) = activity x `homogenate_volume_uL`
   (`r assay_params$homogenate_volume_uL` µL, measured) / total protein (mg) —
   recovers total enzyme units in the homogenate, then divides by the total
   protein extracted from that homogenate (note 3), rather than by tissue
   input mass

```{r calculate-activity}
cs_activity <- calculate_cs_activity(well_rates, protein_by_sample, background_per_sample,
                                      std_slope, std_nmol_max, assay_params, plate_long)

cat("Samples with a computed activity:", nrow(cs_activity), "/", nrow(protein_by_sample), "\n")
cat("Activity range (mU/uL):",
    paste(sprintf("%.3f", range(cs_activity$activity_mU_per_uL)), collapse = " - "), "\n")
cat("Activity range (mU/mg protein):",
    paste(sprintf("%.3f", range(cs_activity$activity_mU_per_mg_protein)), collapse = " - "), "\n")

cat("\n--- cs_activity: per-sample citrate synthase activity ---\n\n")
str(cs_activity)
```

## Results table

```{r results-table}
results_table <- build_results_table(cs_activity, cv_summary, baseline_per_sample, assay_params)

kable(results_table,
      caption = paste0("Citrate synthase activity, family F05 ambient (mU/mg protein uses ",
                       "measured protein concentration x the ", assay_params$homogenate_volume_uL,
                       " uL homogenization volume)"))

cat("\nInterpretation summary:\n")
print(results_table %>% count(Interpretation, name = "n_samples") %>%
        as.data.frame(), row.names = FALSE)

write.csv(results_table, file.path(output_dir, "citrate_synthase_activity_results.csv"),
          row.names = FALSE)

cat("\n--- results_table: formatted per-sample results written to CSV ---\n\n")
str(results_table)
```

## Plot activity

```{r plot-activity, fig.width = 10, fig.height = 6}
# Three-level quality status: a sample whose every replicate has an elevated
# baseline cannot be rescued by replicate exclusion, so it outranks CV as a
# disqualifier and is drawn separately (see note 6).
qual_levels <- c("clean", paste0("CV > ", assay_params$cv_threshold_pct, "%"),
                 "all reps baseline-compromised")
qual_cols   <- setNames(c("#08519c", "#cb181d", "grey60"), qual_levels)

activity_plot_data <- cs_activity %>%
  left_join(cv_summary %>% select(sample_id, cv_all), by = "sample_id") %>%
  left_join(baseline_per_sample %>% select(sample_id, n_elevated, n), by = "sample_id") %>%
  mutate(cv_flag = factor(case_when(
                            n_elevated == n                              ~ qual_levels[3],
                            cv_all > assay_params$cv_threshold_pct       ~ qual_levels[2],
                            TRUE                                        ~ qual_levels[1]),
                          levels = qual_levels),
         label_n = paste0(n_reps_used, "/3 reps",
                          ifelse(n_elevated == n, " - DISQUALIFIED", "")))

activity_plot <- ggplot(activity_plot_data,
                        aes(x = reorder(sample_id, activity_mU_per_uL),
                            y = activity_mU_per_uL, fill = cv_flag)) +
  geom_col(width = 0.65) +
  geom_text(aes(label = label_n), hjust = -0.15, size = 3, colour = "grey25") +
  scale_fill_manual(values = qual_cols, drop = FALSE, name = "Well quality") +
  coord_flip(clip = "off") +
  expand_limits(y = max(activity_plot_data$activity_mU_per_uL) * 1.25) +
  labs(title = "Citrate synthase activity, M. gigas ctenidia, family F05 ambient",
       subtitle = paste0("Background-corrected, mean of usable technical replicates.\n",
                         "Red = replicate CV above ", assay_params$cv_threshold_pct,
                         "%. Grey = every replicate had an elevated starting A412,\n",
                         "so the apparent high activity is a baseline artifact, not enzyme."),
       x = NULL, y = "CS activity (mU/uL homogenate)") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        legend.position = "bottom")

ggsave(file.path(output_dir, "citrate_synthase_activity.png"), activity_plot,
       width = 10, height = 6, dpi = 300)

cat("--- activity_plot: ggplot object structure ---\n\n")
summary(activity_plot)

activity_plot
```

## Protein-normalized activity

```{r plot-activity-normalized, fig.width = 10, fig.height = 6}
# Clean samples only: the trend test must not be driven by wells whose rate is
# a baseline artifact, so the fit is restricted to samples clean on every check.
norm_clean <- activity_plot_data %>% filter(cv_flag == qual_levels[1])

norm_plot <- ggplot(activity_plot_data,
                    aes(x = total_protein_mg, y = activity_mU_per_mg_protein)) +
  geom_smooth(data = norm_clean, method = "lm", se = TRUE,
              colour = "grey35", fill = "grey85", linewidth = 0.6) +
  geom_point(aes(colour = cv_flag), size = 3.5) +
  geom_text(aes(label = individual), vjust = -1, size = 3, colour = "grey25") +
  scale_colour_manual(values = qual_cols, drop = FALSE, name = "Well quality") +
  labs(title = "Protein-normalized CS activity vs. total extracted protein",
       subtitle = paste0("Point labels = individual ID. Total protein = measured protein ",
                         "concentration x the same\n", assay_params$homogenate_volume_uL,
                         " uL homogenization volume used for every sample. Grey line = fit to ",
                         "clean\nsamples only; a negative slope would indicate the normalization ",
                         "is not protein-independent."),
       x = "Total extracted protein (mg)", y = "CS activity (mU/mg protein)") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        legend.position = "bottom")

protein_trend <- cor.test(activity_plot_data$total_protein_mg, activity_plot_data$activity_mU_per_mg_protein)
protein_trend_clean <- cor.test(norm_clean$total_protein_mg, norm_clean$activity_mU_per_mg_protein)
uL_trend_clean <- cor.test(norm_clean$total_protein_mg, norm_clean$activity_mU_per_uL)

ggsave(file.path(output_dir, "citrate_synthase_activity_normalized.png"), norm_plot,
       width = 10, height = 6, dpi = 300)

cat("Correlation of mU/mg protein with total extracted protein\n")
cat(sprintf("  all %d samples      : r = %+.3f, p = %.4f\n",
            nrow(activity_plot_data), protein_trend$estimate, protein_trend$p.value))
cat(sprintf("  clean samples only  : r = %+.3f, p = %.4f  (n = %d)\n",
            protein_trend_clean$estimate, protein_trend_clean$p.value, nrow(norm_clean)))
cat(sprintf("  clean-sample mU/mg protein range: %.3f to %.3f (%.1f-fold)\n",
            min(norm_clean$activity_mU_per_mg_protein), max(norm_clean$activity_mU_per_mg_protein),
            max(norm_clean$activity_mU_per_mg_protein) / min(norm_clean$activity_mU_per_mg_protein)))
cat(sprintf("  clean-sample mU/uL range: %.4f to %.4f (%.1f-fold)\n",
            min(norm_clean$activity_mU_per_uL), max(norm_clean$activity_mU_per_uL),
            max(norm_clean$activity_mU_per_uL) / min(norm_clean$activity_mU_per_uL)))
cat(sprintf("\nArtifact check: r(total protein, mU/uL raw rate) = %+.3f, p = %.4f  (n = %d)\n",
            uL_trend_clean$estimate, uL_trend_clean$p.value, nrow(norm_clean)))
cat("If mU/uL itself trends with total protein while mU/mg protein trends oppositely,\n",
    "the mU/mg protein trend is partly/wholly the arithmetic of dividing by a covariate\n",
    "rather than a biological difference in specific activity.\n", sep = "")

cat("\n--- protein_trend_clean: correlation test on clean samples ---\n\n")
str(protein_trend_clean, max.level = 1, give.attr = FALSE)

cat("\n--- norm_clean: clean samples used for the trend fit ---\n\n")
str(norm_clean)

norm_plot
```

The clean-sample correlation of mU/mg protein with total extracted protein is
r = `r sprintf("%+.3f", protein_trend_clean$estimate)`
(p = `r sprintf("%.4f", protein_trend_clean$p.value)`, n = `r nrow(norm_clean)`) —
as strongly negative as the tissue-weight version of this check was. The
difference is in the artifact check: raw mU/µL correlates with total protein at
r = `r sprintf("%+.3f", uL_trend_clean$estimate)`
(p = `r sprintf("%.3f", uL_trend_clean$p.value)`), a real (if not clean) inverse
trend rather than the flat line seen against tissue weight. That means the
mU/mg protein trend is **not purely a division artifact** the way mU/mg tissue
was — part of it reflects that samples with more extracted protein measured a
*lower* raw rate, which a fixed-volume, decelerating-kinetics assay (notes 3–4)
cannot distinguish from a genuine difference in specific activity without a
dilution series. Both possibilities remain open on this plate; the number
should be treated as provisional pending that follow-up, but it is not
dismissible as pure arithmetic the way the tissue-weight version was.

# QC SUMMARY

```{r qc-summary}
qc_summary <- build_qc_summary_table(layout_wells, recon$summary, std_curve,
                                      well_diagnostics, well_rates, cv_summary,
                                      baseline_per_sample, pos_control, assay_params)

kable(qc_summary, col.names = c("QC check", "Result"),
      caption = "Quality control summary for this plate")

write.csv(qc_summary, file.path(output_dir, "qc_summary.csv"), row.names = FALSE)

cat("\n--- qc_summary: consolidated QC checks for the run ---\n\n")
str(qc_summary)
```

## Auto-generated findings

The bullets below are generated programmatically from the QC/CV/anomaly
objects above by `generate_qc_findings()`, so the same logic that flags a
problem here is guaranteed to be the logic that flagged it upstream.

```{r qc-findings}
qc_findings <- generate_qc_findings(well_diagnostics, well_rates, cv_summary,
                                     baseline_per_sample, std_curve, recon$summary,
                                     assay_params)

cat(paste0("- ", qc_findings, "\n"), sep = "\n")
```

# SUMMARY

## Anomalies found

1. **`absorbance-*.csv` originally omitted plate columns 10–12 — now
   resolved.** An earlier export of this file covered only 54 of the 69
   occupied wells (missing `F05_04`, `F05_08`, their background controls, and
   the whole 24 nmol GSH standard). The file has been re-exported from Gen5
   and now covers all 69 wells; the `reconcile-sources` chunk confirms every
   layout well is present and that the re-export agrees exactly with
   `full_report-*.txt` on all 1,449 shared readings. No wells were recovered
   from the full report in this version of the analysis.

2. **Background control wells are not behaving as background controls.** Of 24
   background wells, roughly a third start at high A412 (0.24–0.82 vs. ~0.05
   for well-behaved wells) and then *decrease* steeply — up to
   -19 mA412/min. Background Control Mix contains no CS Substrate Mix, so
   these wells cannot support a CS reaction and must read flat. Six of the
   eight background triplicates contain at least one such well, and they
   appear in a near-regular pattern across the plate. One well (C2) instead
   *increases* at 11.6 mA412/min, i.e. it behaves like a full reaction well.
   The most consistent explanation is that Reaction Mix was dispensed into
   some background wells instead of Background Control Mix. Because of this,
   background correction here uses only the flat replicates, and the flat-well
   background rate is ~0.3 mA412/min — about 3% of the median sample signal,
   so background correction is a minor term regardless.

3. **The GSH standard curve failed QC.** Replicate CVs run from 6% to 126%,
   and five of six concentrations exceed the 15% threshold. The replicate-level
   R<sup>2</sup> is only 0.80. Two wells are grossly discrepant from their
   triplicate medians (E6 at the 8 nmol point, +0.80 A412; E2 at the 0 nmol
   point, +0.77 A412) — E2 reads 0.871 while its siblings E1 and E3 read
   0.100 and 0.095. This is the dominant source of uncertainty in every
   reported activity value, since the standard curve slope scales all of them.

4. **The top two standards are above the photometric linear range.** The
   32 and 40 nmol standards read A412 1.7–2.2. Response per nmol is
   essentially flat between them (net A412 1.61 vs. 1.72 for a 25%
   concentration increase), i.e. the curve is saturating. The calibration is
   trustworthy only over roughly 0–24 nmol.

5. **Every standard well lost signal over the run** (18/18 wells, drift
   -0.001 to -0.74 A412). GSH + DTNB is a stoichiometric endpoint, so
   standards should be stable; this indicates TNB<sup>2-</sup> instability
   under the run conditions. The standard curve is therefore read at t = 0,
   before appreciable decay.

6. **Decreasing kinetics in sample wells.** Wells A12, B3 and B5 end lower
   than they started, and A6 rises by only 0.041 A412 total against ~0.16 for
   its siblings. These wells are excluded from rate extraction. Note that
   Gen5's own `Max V [412]` column reports large *negative* rates for these
   wells (e.g. -19.7 for A12, -11.5 for B3) because Gen5 selects the window
   of maximum *absolute* slope; those numbers are not activities and must not
   be used as such.

7. **Read glitches.** Eleven wells contain a single read-to-read step that
   disagrees with the steps on either side of it by more than 0.02 A412 —
   among sample wells: A6 (+0.040 at 30 min, after the trace had gone flat),
   A12 (-0.036 at 2 min), B1 (-0.035 at 4 min), B2 (-0.041 at 4 min) and B5
   (+0.031 at 2 min); also D4, E6, E7, E10, E11 and F1. These are instrument
   artifacts rather than kinetics. A glitch only disqualifies a well's rate
   when it falls *inside* the fitted window, which is why B1 and B2 retain
   usable rates (their glitch at 4 min precedes their 10–18 min window) while
   A6 and B5 do not.

8. **Elevated starting absorbance is the pattern underlying most of the above.**
   Five sample wells (A12, B1, B2, B3, B5) and seven background wells start far
   above the clean-well baseline of A412 0.13–0.16, and these are precisely the
   wells that decline or misbehave. TNB present at t = 0 means the reaction had
   already run, or the well was contaminated, before the first read. The
   affected wells cluster in the right-hand columns of rows A–B and scattered
   positions in C–D, which points to a dispensing or timing problem rather than
   random error. Most consequentially, **all three `F05_05` replicates are
   affected**, which disqualifies that sample entirely.

9. **Layout label problem.** `F05_05_ambient_BG` (D1–D3) is spelled
   `citrate_synthasse`. Handled in code, but the layout file should be
   corrected.

## Technical replicate precision

Across all three replicates, **four of eight samples exceed the 15% CV
threshold**: `F05_05` (69%), `F05_04` (56%), `F05_06` (37%) and `F05_02`
(30%). The other four — `F05_01`, `F05_03`, `F05_07`, `F05_08` — sit at
4.9%, 5.4%, 6.5% and 7.9% respectively.

In every failing case the CV is driven by **a single bad well**, not by
uniform imprecision. Excluding that one flagged replicate drops `F05_02` to
8.9% (excluding A6), `F05_04` to 2.1% (A12) and `F05_05` to 15.1% (B3).
`F05_06` is the exception: excluding B5 still leaves 24.7%, because its
remaining two replicates (B4 at 6.15 and B6 at 8.75 mA412/min) genuinely
disagree. Every sample retains at least two usable replicates.

So **two samples remain above threshold after QC exclusion**: `F05_06`
(24.7%, n = 2) and `F05_05` (15.1%, n = 2, marginal).

**`F05_05` should not be used at all, and its CV is not the reason.** It
reports 19.3 mA412/min — 2.2x the next-highest sample — but **all three of its
replicates start at elevated A412** (0.65, 0.67, 0.93 versus 0.13–0.16 for
clean wells). Its two "usable" wells (B1, B2) are the two steepest traces on
the plate and both rise from an already-contaminated baseline before turning
over, and B3 declines outright. A well that begins with TNB already present is
not measuring citrate synthase from zero, so the apparent high activity is an
artifact of the starting condition, not enzymology. Excluding one replicate on
CV grounds cannot repair a sample in which no replicate is clean.

## Assay validity

The **positive control worked**: all three replicates rise linearly
(R<sup>2</sup> > 0.999) at ~2.8 mA412/min with 2% CV. The core reaction
chemistry and the reader were functioning. The problems on this plate are in
the standards and background wells, not in the CS reaction itself.

Sample rates in clean wells are also well-behaved and mutually consistent
(~7–9 mA412/min for most individuals, R<sup>2</sup> > 0.99 over the first
8 min), which is why per-sample activity estimates are reported despite the
standard curve problems.

## Recommendation

Treat the activity values here as **provisional**. The rate measurements are
sound for the wells that passed QC, but they are calibrated through a standard
curve that failed its own QC, so absolute activities carry an uncertainty that
the reported precision does not capture. Before these numbers are used
downstream:

1. **Run a dilution series on one homogenate — the highest priority item.**
   Assay a single sample neat, 1:2, 1:5 and 1:10 and plot rate against
   dilution. Measured mU/µL correlates with total extracted protein at
   r = `r sprintf("%+.2f", uL_trend_clean$estimate)` (p =
   `r sprintf("%.2f", uL_trend_clean$p.value)`) across a 3-fold range of
   protein input on this plate — a real but noisy inverse trend, not the flat
   line seen against tissue weight. A dilution series is the only way to tell
   whether that reflects the assay running outside its linear range (aliquot
   volume, substrate depletion) or a genuine difference in specific activity
   between individuals; nothing on this list should be treated as settled
   until it is run.
2. **Re-run the GSH standard curve**, keeping the top standard at or below
   24 nmol/well so all points stay within photometric linearity, and read it
   immediately.
3. **Resolve the background control mix-up** — confirm which wells received
   Background Control Mix vs. Reaction Mix before the next plate.
4. **Re-assay `F05_05` — mandatory, not optional.** All three replicates are
   baseline-compromised, so this sample has no usable measurement on this
   plate regardless of what the CV column says. Do not carry its
   `r sprintf("%.0f", cs_activity$activity_mU_per_mg_protein[cs_activity$sample_id == "F05_05_ambient"])`
   mU/mg protein into any downstream comparison. Re-assay `F05_06` as well
   (24.7% CV on two genuinely disagreeing replicates).
5. **Check the dispensing/plating step for the elevated-baseline wells**
   (A12, B1, B2, B3, B5 — all in the bottom-right region of rows A–B, plus
   background wells C2, C5, C7, C12, D1, D4, D5). The spatial clustering
   suggests a systematic cause — carryover on a multichannel tip, DTNB added
   before the plate reached the reader, or a delay between reagent addition
   and the first read — rather than random pipetting error.
6. **Protein normalization controls for extraction differences, but not for
   whatever is driving the mU/uL-vs-protein trend.** Normalizing to total
   extracted protein (rather than tissue weight) already removes
   sample-to-sample variation in how much soluble material a given tissue
   input yielded. It does not, by itself, explain why raw mU/µL trends
   downward with more extracted protein (recommendation 1) — that requires the
   dilution series, not a change in normalization basis.
7. ~~Re-export the absorbance CSV covering all 12 plate columns.~~ **Done** —
   `absorbance-*.csv` now covers the full plate; see anomaly 1.

## What can be used from this plate

Four samples are clean on every check — `F05_01`, `F05_03`, `F05_07` and
`F05_08`, all with no baseline-compromised replicates and CV < 8%. Their
relative *ranking* is more robust than their absolute values, since all four
share the same calibration.

The instructive result is how *little* they differ: 6.8, 7.6, 8.6 and
6.8 mA412/min background-corrected — a spread of roughly 25%, against a
technical CV of 5–8% within samples. Once the plate's one apparent outlier
(`F05_05`) is recognized as a baseline artifact rather than a high-activity
individual, no striking within-family variation remains at ambient
temperature. Two implications for the wider design:

- **The mU/mg protein values should be treated as provisional, not discarded
  outright.** The standard curve QC failure (three of six concentrations with
  CV > 15%, 8/18 wells flagged) is one source of uncertainty affecting every
  reported value equally. Separately, mU/mg protein correlates strongly
  negatively with total extracted protein
  (r = `r sprintf("%+.2f", protein_trend_clean$estimate)`,
  clean samples), but — unlike the tissue-weight version of this plate — the
  underlying raw rate (mU/µL) also correlates with extracted protein
  (r = `r sprintf("%+.2f", uL_trend_clean$estimate)`), so the mU/mg protein
  trend is not pure division artifact. Whether it reflects real
  specific-activity differences between individuals or an assay linear-range
  effect is unresolved without the dilution series.
- **The assay needs a linear-range check before more plates are run.** The
  dilution series is cheap — one homogenate, four dilutions — and it determines
  whether the deceleration seen in every trace, and its apparent relationship
  to protein content, is substrate depletion, aliquot overload, or extraction
  saturation rather than biology. Running more families before that is
  answered risks generating plates with an uninterpretable normalization.
