---
title: "02-NaK-ATPase-vs-family-survival"
author: "Steven Roberts"
date: "2026-10-02"
output:
  github_document:
    toc: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
```

# Overview

Compare family-level gill Na<sup>+</sup>/K<sup>+</sup>-ATPase activity (June
2026 sampling; see `01-NaK-ATPase-202606-sampling.Rmd`) with family-level
survival from the heat-survivorship cross-experiment summary
(`heat-survivorship/code/09-mgig-survivorship-cross-experiment-summary.qmd`).

Only 9 families are shared, so correlations are descriptive.

# Data

## Na/K-ATPase

```{r read-atpase}
atpase <- read_tsv("../data/202606-sampling.tsv") %>%
  rename(ATPase = `ATPase (umol ADP/mg protein/hr)`)

sampling_log <- read_csv("../../sampling-event-metadata/june-2026/sampling-log.csv") %>%
  select(`Tube ID`, Condition, `Oyster Family ID`)

nak <- atpase %>%
  left_join(sampling_log, by = "Tube ID") %>%
  mutate(family = str_remove(`Oyster Family ID`, "Family "))

stopifnot(!any(is.na(nak$Condition)))
```

Family summaries: median activity under each condition (median is robust to
the failed M21 assay), and the heat response as the 36C - Ambient difference.

```{r family-atpase}
fam_atpase <- nak %>%
  group_by(family, Condition) %>%
  summarise(median_ATPase = median(ATPase), .groups = "drop") %>%
  pivot_wider(names_from = Condition, values_from = median_ATPase) %>%
  mutate(delta_36C_minus_Ambient = `36C` - Ambient)

knitr::kable(fam_atpase, digits = 2)
```

## Survival

```{r read-survival}
surv_dir <- "../../heat-survivorship/outputs/09-mgig-survivorship-cross-experiment-summary"

fam_rank <- read_csv(file.path(surv_dir, "family_ranking.csv"),
                     col_types = cols(family = col_character()))
fam_exp  <- read_csv(file.path(surv_dir, "family_by_experiment.csv"),
                     col_types = cols(family = col_character()))

fam <- fam_atpase %>% inner_join(fam_rank, by = "family")
nrow(fam)
```

# ATPase vs overall survival ranking

`composite_score` is the mean within-experiment survival percentile (0-100,
higher = hardier) across all seven survival experiments.

```{r overall-plot, fig.width=10, fig.height=4}
fam_long <- fam %>%
  pivot_longer(c(Ambient, `36C`, delta_36C_minus_Ambient),
               names_to = "atpase_metric", values_to = "atpase_value") %>%
  mutate(atpase_metric = factor(atpase_metric,
                                levels = c("Ambient", "36C", "delta_36C_minus_Ambient"),
                                labels = c("Ambient", "36C", "36C - Ambient")))

overall_cor <- fam_long %>%
  group_by(atpase_metric) %>%
  summarise(rho = cor(atpase_value, composite_score, method = "spearman"),
            p   = cor.test(atpase_value, composite_score, method = "spearman", exact = TRUE)$p.value,
            .groups = "drop")

knitr::kable(overall_cor, digits = 3)

ggplot(fam_long, aes(x = atpase_value, y = composite_score)) +
  geom_smooth(method = "lm", se = FALSE, colour = "grey60", linewidth = 0.5) +
  geom_point(size = 2) +
  geom_text(aes(label = family), nudge_y = 3, size = 3) +
  geom_text(data = overall_cor,
            aes(label = sprintf("rho = %.2f, p = %.2f", rho, p)),
            x = Inf, y = -Inf, hjust = 1.1, vjust = -0.8, size = 3, inherit.aes = FALSE) +
  facet_wrap(~ atpase_metric, scales = "free_x") +
  labs(x = "Median Na+/K+-ATPase activity (umol ADP / mg protein / hr)",
       y = "Composite survival score\n(higher = hardier)") +
  theme_bw()
```

# ATPase vs survival in each experiment

Spearman correlation between family ATPase metrics and within-experiment
survival proportion. Experiments with few families (or near-total mortality)
carry little information.

```{r per-experiment, fig.width=8, fig.height=4}
exp_cor <- fam_exp %>%
  inner_join(fam_atpase, by = "family") %>%
  pivot_longer(c(Ambient, `36C`, delta_36C_minus_Ambient),
               names_to = "atpase_metric", values_to = "atpase_value") %>%
  group_by(exp_id, stressor, atpase_metric) %>%
  summarise(n_fam = n(),
            rho   = suppressWarnings(cor(atpase_value, surv_prop, method = "spearman")),
            .groups = "drop") %>%
  mutate(atpase_metric = factor(atpase_metric,
                                levels = c("Ambient", "36C", "delta_36C_minus_Ambient"),
                                labels = c("Ambient", "36C", "36C - Ambient")),
         exp_label = paste0(exp_id, " (n=", n_fam, ")"))

ggplot(exp_cor, aes(x = atpase_metric, y = exp_label, fill = rho)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = ifelse(is.na(rho), "NA", sprintf("%.2f", rho))), size = 3) +
  scale_fill_gradient2(low = "#B2182B", mid = "white", high = "#2166AC",
                       limits = c(-1, 1), na.value = "grey85") +
  labs(x = "Family ATPase metric", y = NULL, fill = "Spearman rho",
       title = "ATPase vs family survival proportion, by experiment") +
  theme_minimal()
```

# Within-family heat response vs survival

Does the change in ATPase between 36C and Ambient *within* a family predict
survival? The heat response is expressed three ways:

- `diff`: difference in medians (36C - Ambient), as above
- `log2ratio`: log2(median 36C / median Ambient), a relative change
- `cohens_d`: standardized mean difference, scaling the change by within-family
  variability

`wilcox_p` tests whether 36C and Ambient differ within each family.

```{r heat-response}
heat_resp <- nak %>%
  group_by(family) %>%
  summarise(
    diff      = median(ATPase[Condition == "36C"]) - median(ATPase[Condition == "Ambient"]),
    log2ratio = log2(median(ATPase[Condition == "36C"]) / median(ATPase[Condition == "Ambient"])),
    cohens_d  = (mean(ATPase[Condition == "36C"]) - mean(ATPase[Condition == "Ambient"])) /
      sqrt((var(ATPase[Condition == "36C"]) + var(ATPase[Condition == "Ambient"])) / 2),
    wilcox_p  = suppressWarnings(
      wilcox.test(ATPase[Condition == "36C"], ATPase[Condition == "Ambient"])$p.value),
    .groups = "drop"
  ) %>%
  inner_join(fam_rank, by = "family") %>%
  arrange(desc(composite_score))

heat_resp %>%
  select(family, composite_score, mean_surv_prop, diff, log2ratio, cohens_d, wilcox_p) %>%
  knitr::kable(digits = 3)
```

```{r heat-response-cor}
heat_long <- heat_resp %>%
  pivot_longer(c(diff, log2ratio, cohens_d),
               names_to = "response_metric", values_to = "response_value") %>%
  mutate(response_metric = factor(response_metric, levels = c("diff", "log2ratio", "cohens_d")))

heat_cor <- heat_long %>%
  pivot_longer(c(composite_score, mean_surv_prop),
               names_to = "survival_metric", values_to = "survival_value") %>%
  group_by(response_metric, survival_metric) %>%
  summarise(rho = cor(response_value, survival_value, method = "spearman"),
            p   = cor.test(response_value, survival_value, method = "spearman", exact = TRUE)$p.value,
            .groups = "drop")

knitr::kable(heat_cor, digits = 3)
```

```{r heat-response-plot, fig.width=10, fig.height=4}
heat_lab <- filter(heat_cor, survival_metric == "composite_score")

ggplot(heat_long, aes(x = response_value, y = composite_score)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey70") +
  geom_smooth(method = "lm", se = FALSE, colour = "grey60", linewidth = 0.5) +
  geom_point(aes(shape = wilcox_p < 0.05), size = 2.5) +
  geom_text(aes(label = family), nudge_y = 3, size = 3) +
  geom_text(data = heat_lab,
            aes(label = sprintf("rho = %.2f, p = %.2f", rho, p)),
            x = -Inf, y = Inf, hjust = -0.1, vjust = 1.5, size = 3, inherit.aes = FALSE) +
  scale_shape_manual(values = c(`FALSE` = 1, `TRUE` = 16),
                     labels = c(`FALSE` = "no", `TRUE` = "yes"),
                     name = "36C vs Ambient\np < 0.05") +
  facet_wrap(~ response_metric, scales = "free_x",
             labeller = as_labeller(c(diff = "36C - Ambient (median)",
                                      log2ratio = "log2(36C / Ambient)",
                                      cohens_d = "Cohen's d"))) +
  labs(x = "Within-family ATPase heat response",
       y = "Composite survival score\n(higher = hardier)") +
  theme_bw()
```

The heat response does not track survival under any of the three metrics
(Spearman rho 0.12-0.18, all p > 0.6). The two hardiest families (5, 2) held
or slightly raised ATPase at 36C, but family 9 (second hardiest) showed the
clearest decline and family 6 (near the bottom) the largest increase. Only
families 9 and 10 changed significantly within family, and they sit at
opposite ends of the survival ranking.
