---
title: "02-NaK-ATPase-vs-family-survival"
author: "Steven Roberts"
date: "2026-10-02"
output:
  github_document:
    toc: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
```

# Overview

Compare family-level gill Na<sup>+</sup>/K<sup>+</sup>-ATPase activity (June
2026 sampling; see `01-NaK-ATPase-202606-sampling.Rmd`) with family-level
survival from the heat-survivorship cross-experiment summary
(`heat-survivorship/code/09-mgig-survivorship-cross-experiment-summary.qmd`).

Only 9 families are shared, so correlations are descriptive.

# Data

## Na/K-ATPase

```{r read-atpase}
atpase <- read_tsv("../data/202606-sampling.tsv") %>%
  rename(ATPase = `ATPase (umol ADP/mg protein/hr)`)

sampling_log <- read_csv("../../sampling-event-metadata/june-2026/sampling-log.csv") %>%
  select(`Tube ID`, Condition, `Oyster Family ID`)

nak <- atpase %>%
  left_join(sampling_log, by = "Tube ID") %>%
  mutate(family = str_remove(`Oyster Family ID`, "Family "))

stopifnot(!any(is.na(nak$Condition)))
```

Family summaries: median activity under each condition (median is robust to
the failed M21 assay), and the heat response as the 36C - Ambient difference.

```{r family-atpase}
fam_atpase <- nak %>%
  group_by(family, Condition) %>%
  summarise(median_ATPase = median(ATPase), .groups = "drop") %>%
  pivot_wider(names_from = Condition, values_from = median_ATPase) %>%
  mutate(delta_36C_minus_Ambient = `36C` - Ambient)

knitr::kable(fam_atpase, digits = 2)
```

## Survival

```{r read-survival}
surv_dir <- "../../heat-survivorship/outputs/09-mgig-survivorship-cross-experiment-summary"

fam_rank <- read_csv(file.path(surv_dir, "family_ranking.csv"),
                     col_types = cols(family = col_character()))
fam_exp  <- read_csv(file.path(surv_dir, "family_by_experiment.csv"),
                     col_types = cols(family = col_character()))

fam <- fam_atpase %>% inner_join(fam_rank, by = "family")
nrow(fam)
```

# ATPase vs overall survival ranking

`composite_score` is the mean within-experiment survival percentile (0-100,
higher = hardier) across all seven survival experiments.

```{r overall-plot, fig.width=10, fig.height=4}
fam_long <- fam %>%
  pivot_longer(c(Ambient, `36C`, delta_36C_minus_Ambient),
               names_to = "atpase_metric", values_to = "atpase_value") %>%
  mutate(atpase_metric = factor(atpase_metric,
                                levels = c("Ambient", "36C", "delta_36C_minus_Ambient"),
                                labels = c("Ambient", "36C", "36C - Ambient")))

overall_cor <- fam_long %>%
  group_by(atpase_metric) %>%
  summarise(rho = cor(atpase_value, composite_score, method = "spearman"),
            p   = cor.test(atpase_value, composite_score, method = "spearman", exact = TRUE)$p.value,
            .groups = "drop")

knitr::kable(overall_cor, digits = 3)

ggplot(fam_long, aes(x = atpase_value, y = composite_score)) +
  geom_smooth(method = "lm", se = FALSE, colour = "grey60", linewidth = 0.5) +
  geom_point(size = 2) +
  geom_text(aes(label = family), nudge_y = 3, size = 3) +
  geom_text(data = overall_cor,
            aes(label = sprintf("rho = %.2f, p = %.2f", rho, p)),
            x = Inf, y = -Inf, hjust = 1.1, vjust = -0.8, size = 3, inherit.aes = FALSE) +
  facet_wrap(~ atpase_metric, scales = "free_x") +
  labs(x = "Median Na+/K+-ATPase activity (umol ADP / mg protein / hr)",
       y = "Composite survival score\n(higher = hardier)") +
  theme_bw()
```

# ATPase vs survival in each experiment

Spearman correlation between family ATPase metrics and within-experiment
survival proportion. Experiments with few families (or near-total mortality)
carry little information.

```{r per-experiment, fig.width=8, fig.height=4}
exp_cor <- fam_exp %>%
  inner_join(fam_atpase, by = "family") %>%
  pivot_longer(c(Ambient, `36C`, delta_36C_minus_Ambient),
               names_to = "atpase_metric", values_to = "atpase_value") %>%
  group_by(exp_id, stressor, atpase_metric) %>%
  summarise(n_fam = n(),
            rho   = suppressWarnings(cor(atpase_value, surv_prop, method = "spearman")),
            .groups = "drop") %>%
  mutate(atpase_metric = factor(atpase_metric,
                                levels = c("Ambient", "36C", "delta_36C_minus_Ambient"),
                                labels = c("Ambient", "36C", "36C - Ambient")),
         exp_label = paste0(exp_id, " (n=", n_fam, ")"))

ggplot(exp_cor, aes(x = atpase_metric, y = exp_label, fill = rho)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = ifelse(is.na(rho), "NA", sprintf("%.2f", rho))), size = 3) +
  scale_fill_gradient2(low = "#B2182B", mid = "white", high = "#2166AC",
                       limits = c(-1, 1), na.value = "grey85") +
  labs(x = "Family ATPase metric", y = NULL, fill = "Spearman rho",
       title = "ATPase vs family survival proportion, by experiment") +
  theme_minimal()
```

# Within-family heat response vs survival

Does the change in ATPase between 36C and Ambient *within* a family predict
survival? The heat response is expressed three ways:

- `diff`: difference in medians (36C - Ambient), as above
- `log2ratio`: log2(median 36C / median Ambient), a relative change
- `cohens_d`: standardized mean difference, scaling the change by within-family
  variability

`wilcox_p` tests whether 36C and Ambient differ within each family.

```{r heat-response}
heat_resp <- nak %>%
  group_by(family) %>%
  summarise(
    diff      = median(ATPase[Condition == "36C"]) - median(ATPase[Condition == "Ambient"]),
    log2ratio = log2(median(ATPase[Condition == "36C"]) / median(ATPase[Condition == "Ambient"])),
    cohens_d  = (mean(ATPase[Condition == "36C"]) - mean(ATPase[Condition == "Ambient"])) /
      sqrt((var(ATPase[Condition == "36C"]) + var(ATPase[Condition == "Ambient"])) / 2),
    wilcox_p  = suppressWarnings(
      wilcox.test(ATPase[Condition == "36C"], ATPase[Condition == "Ambient"])$p.value),
    .groups = "drop"
  ) %>%
  inner_join(fam_rank, by = "family") %>%
  arrange(desc(composite_score))

heat_resp %>%
  select(family, composite_score, mean_surv_prop, diff, log2ratio, cohens_d, wilcox_p) %>%
  knitr::kable(digits = 3)
```

```{r heat-response-cor}
heat_long <- heat_resp %>%
  pivot_longer(c(diff, log2ratio, cohens_d),
               names_to = "response_metric", values_to = "response_value") %>%
  mutate(response_metric = factor(response_metric, levels = c("diff", "log2ratio", "cohens_d")))

heat_cor <- heat_long %>%
  pivot_longer(c(composite_score, mean_surv_prop),
               names_to = "survival_metric", values_to = "survival_value") %>%
  group_by(response_metric, survival_metric) %>%
  summarise(rho = cor(response_value, survival_value, method = "spearman"),
            p   = cor.test(response_value, survival_value, method = "spearman", exact = TRUE)$p.value,
            .groups = "drop")

knitr::kable(heat_cor, digits = 3)
```

```{r heat-response-plot, fig.width=10, fig.height=4}
heat_lab <- filter(heat_cor, survival_metric == "composite_score")

ggplot(heat_long, aes(x = response_value, y = composite_score)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey70") +
  geom_smooth(method = "lm", se = FALSE, colour = "grey60", linewidth = 0.5) +
  geom_point(aes(shape = wilcox_p < 0.05), size = 2.5) +
  geom_text(aes(label = family), nudge_y = 3, size = 3) +
  geom_text(data = heat_lab,
            aes(label = sprintf("rho = %.2f, p = %.2f", rho, p)),
            x = -Inf, y = Inf, hjust = -0.1, vjust = 1.5, size = 3, inherit.aes = FALSE) +
  scale_shape_manual(values = c(`FALSE` = 1, `TRUE` = 16),
                     labels = c(`FALSE` = "no", `TRUE` = "yes"),
                     name = "36C vs Ambient\np < 0.05") +
  facet_wrap(~ response_metric, scales = "free_x",
             labeller = as_labeller(c(diff = "36C - Ambient (median)",
                                      log2ratio = "log2(36C / Ambient)",
                                      cohens_d = "Cohen's d"))) +
  labs(x = "Within-family ATPase heat response",
       y = "Composite survival score\n(higher = hardier)") +
  theme_bw()
```

The heat response does not track survival under any of the three metrics
(Spearman rho 0.12-0.18, all p > 0.6). The two hardiest families (5, 2) held
or slightly raised ATPase at 36C, but family 9 (second hardiest) showed the
clearest decline and family 6 (near the bottom) the largest increase. Only
families 9 and 10 changed significantly within family, and they sit at
opposite ends of the survival ranking.

# Family rankings: ATPase vs survival

Rank families (1 = highest) by survival `composite_score` and by three ATPase
measures: baseline (Ambient median), 36C median, and the relative change at 36C
(`log2ratio`, where rank 1 = largest gain). Spearman rho is the correlation of
these ranks; Kendall tau is a more conservative rank-agreement measure for
small samples.

```{r rank-table}
fam_ranks <- heat_resp %>%
  select(family, composite_score, log2ratio) %>%
  left_join(fam_atpase %>% select(family, Ambient, `36C`), by = "family") %>%
  mutate(
    survival_rank = rank(-composite_score),
    ambient_rank  = rank(-Ambient),
    heat_rank     = rank(-`36C`),
    relative_rank = rank(-log2ratio)
  ) %>%
  arrange(survival_rank)

fam_ranks %>%
  select(family, survival_rank, composite_score, ambient_rank, Ambient,
         heat_rank, `36C`, relative_rank, log2ratio) %>%
  knitr::kable(digits = 2)
```

```{r rank-cor}
rank_long <- fam_ranks %>%
  pivot_longer(c(ambient_rank, heat_rank, relative_rank),
               names_to = "atpase_rank", values_to = "atpase_rank_value") %>%
  mutate(atpase_rank = factor(atpase_rank,
                              levels = c("ambient_rank", "heat_rank", "relative_rank"),
                              labels = c("Ambient (baseline)", "36C", "Relative change at 36C")))

rank_cor <- rank_long %>%
  group_by(atpase_rank) %>%
  summarise(
    spearman_rho = cor(atpase_rank_value, survival_rank, method = "spearman"),
    spearman_p   = cor.test(atpase_rank_value, survival_rank, method = "spearman", exact = TRUE)$p.value,
    kendall_tau  = cor(atpase_rank_value, survival_rank, method = "kendall"),
    kendall_p    = cor.test(atpase_rank_value, survival_rank, method = "kendall", exact = TRUE)$p.value,
    .groups = "drop"
  )

knitr::kable(rank_cor, digits = 3)
```

```{r rank-plot, fig.width=10, fig.height=4}
rank_labels <- rank_cor %>%
  mutate(label = sprintf("%s\nrho = %.2f, tau = %.2f", atpase_rank, spearman_rho, kendall_tau)) %>%
  select(atpase_rank, label) %>%
  deframe()

ggplot(rank_long, aes(x = atpase_rank_value, y = survival_rank)) +
  geom_line(data = tibble(atpase_rank_value = 1:9, survival_rank = 1:9),
            linetype = "dashed", colour = "grey70") +
  geom_point(size = 2.5) +
  geom_text(aes(label = family), nudge_x = 0.3, nudge_y = -0.3, size = 3) +
  scale_x_continuous(breaks = 1:9) +
  scale_y_reverse(breaks = 1:9) +
  facet_wrap(~ atpase_rank, labeller = as_labeller(rank_labels)) +
  labs(x = "ATPase rank (1 = highest)",
       y = "Survival rank (1 = hardiest)",
       caption = "Dashed line: identical ranks") +
  theme_bw()
```

Baseline (Ambient) ATPase gives the strongest, and inverse, ranking agreement:
the three families with the highest baseline activity (10, 3, 1) rank 8th, 5th
and 6th for survival, and the hardiest family (5) has the lowest baseline. With
9 families this is not significant (Spearman p = 0.15; Kendall p = 0.26). The
36C ranking is weaker in the same direction, and the relative change at 36C
does not align with survival.

# Mixed-model estimates of family ATPase

The family summaries above are medians of 15 oysters per condition, and most of
the variation in ATPase is among oysters within a family. Here a mixed model
fit to the individual oysters estimates each family's baseline and heat
response, with an uncertainty for each, and that uncertainty is carried into
the comparison with survival.

The failed M21 assay is dropped here because, unlike the medians, model means
are sensitive to it. Ambient and 36C are different oysters, so the heat
response is a between-group contrast within each family.

```{r mm-data}
mm <- nak %>%
  filter(`Tube ID` != "M21") %>%
  mutate(Condition = factor(Condition, levels = c("Ambient", "36C")))

mm %>% count(Condition)
```

## How much of the variation is between families?

Separate random-intercept models for each condition partition the variance into
between-family and within-family (among-oyster) components. The ICC is the
between-family share. Reliability is how well a family mean from `n` oysters
tracks the family's true mean, and `n_for_0.8` is the number of oysters per
family needed for a reliability of 0.8.

```{r mm-variance-partition}
var_part <- mm %>%
  group_by(Condition) %>%
  group_modify(function(df, key) {
    fit <- lme4::lmer(ATPase ~ 1 + (1 | family), data = df)
    vc  <- as.data.frame(lme4::VarCorr(fit))
    tibble(
      n_per_family = mean(table(df$family)),
      var_family   = vc$vcov[vc$grp == "family"],
      var_oyster   = vc$vcov[vc$grp == "Residual"]
    )
  }) %>%
  ungroup() %>%
  mutate(
    ICC         = var_family / (var_family + var_oyster),
    reliability = var_family / (var_family + var_oyster / n_per_family),
    n_for_0.8   = ceiling(4 * var_oyster / var_family)
  )

knitr::kable(var_part, digits = 3)
```

## Joint model

Model: ATPase ~ Condition, with a random intercept (family baseline) and random
Condition slope (family heat response) for each family. The residual variance
is allowed to differ between conditions, since Ambient is more variable among
oysters than 36C.

```{r mm-fit}
fit_hom <- nlme::lme(ATPase ~ Condition, random = ~ 1 + Condition | family,
                     data = mm, method = "REML")
fit_het <- nlme::lme(ATPase ~ Condition, random = ~ 1 + Condition | family,
                     weights = nlme::varIdent(form = ~ 1 | Condition),
                     data = mm, method = "REML")
fit_int <- nlme::lme(ATPase ~ Condition, random = ~ 1 | family,
                     weights = nlme::varIdent(form = ~ 1 | Condition),
                     data = mm, method = "REML")

# Residual variance: one value vs one per condition
anova(fit_hom, fit_het)

# Do families differ in heat response? (random slope vs intercept only;
# the p value is conservative because the variance is tested at its boundary)
anova(fit_int, fit_het)

summary(fit_het)$tTable %>% knitr::kable(digits = 3)
nlme::VarCorr(fit_het)

resid_sd <- fit_het$sigma *
  coef(fit_het$modelStruct$varStruct, unconstrained = FALSE, allCoef = TRUE)
resid_sd
```

## Family estimates with uncertainty

The family estimates (BLUPs) are pulled toward the overall mean in proportion
to how noisy each family's data are. Their conditional covariance is
(Z'R^-1^Z + G^-1^)^-1^ for each family, given the fitted variance components.

```{r mm-blups}
G     <- as.matrix(nlme::getVarCov(fit_het))
beta  <- nlme::fixef(fit_het)
re    <- nlme::ranef(fit_het)

blup_cov <- map(set_names(rownames(re)), function(f) {
  d  <- filter(mm, family == f)
  Z  <- model.matrix(~ Condition, d)
  Ri <- diag(1 / resid_sd[as.character(d$Condition)]^2)
  solve(t(Z) %*% Ri %*% Z + solve(G))
})

fam_mm <- tibble(
  family   = rownames(re),
  b0       = re[, 1],
  b1       = re[, 2],
  se_b0    = map_dbl(blup_cov, ~ sqrt(.x[1, 1])),
  se_b1    = map_dbl(blup_cov, ~ sqrt(.x[2, 2])),
  se_36C   = map_dbl(blup_cov, ~ sqrt(sum(.x)))
) %>%
  mutate(
    mm_Ambient = beta[1] + b0,
    mm_36C     = beta[1] + beta[2] + b0 + b1,
    mm_heat    = beta[2] + b1
  ) %>%
  inner_join(fam_rank %>% select(family, composite_score), by = "family") %>%
  left_join(fam_atpase, by = "family") %>%
  arrange(desc(composite_score))

fam_mm %>%
  select(family, composite_score,
         Ambient, mm_Ambient, se_b0,
         `36C`, mm_36C, se_36C,
         delta_36C_minus_Ambient, mm_heat, se_b1) %>%
  knitr::kable(digits = 2)
```

```{r mm-shrinkage-plot, fig.width=10, fig.height=4}
shrink <- bind_rows(
  fam_mm %>% transmute(family, metric = "Ambient", raw = Ambient,
                       est = mm_Ambient, se = se_b0),
  fam_mm %>% transmute(family, metric = "36C", raw = `36C`,
                       est = mm_36C, se = se_36C),
  fam_mm %>% transmute(family, metric = "36C - Ambient", raw = delta_36C_minus_Ambient,
                       est = mm_heat, se = se_b1)
) %>%
  mutate(metric = factor(metric, levels = c("Ambient", "36C", "36C - Ambient")),
         family = factor(family, levels = rev(fam_mm$family)))

ggplot(shrink, aes(y = family)) +
  geom_segment(aes(x = raw, xend = est, yend = family), colour = "grey70") +
  geom_point(aes(x = raw), shape = 1, size = 2) +
  geom_pointrange(aes(x = est, xmin = est - 1.96 * se, xmax = est + 1.96 * se),
                  size = 0.3) +
  facet_wrap(~ metric, scales = "free_x") +
  labs(x = "ATPase (umol ADP / mg protein / hr)",
       y = "Family (ordered by survival, hardiest at top)",
       caption = "Open circles: raw family medians. Filled: mixed-model estimate with 95% interval.") +
  theme_bw()
```

## Survival correlation with uncertainty carried through

Point correlations use the model estimates. To carry the uncertainty in each
family's estimate into the correlation, family effects are drawn 4000 times
from their conditional distributions and the Spearman rho with
`composite_score` is recomputed each time. Fixed effects shift every family
equally, so they do not affect the ranks.

```{r mm-survival-cor}
mm_point <- fam_mm %>%
  summarise(
    Ambient         = cor(mm_Ambient, composite_score, method = "spearman"),
    `36C`           = cor(mm_36C,     composite_score, method = "spearman"),
    `36C - Ambient` = cor(mm_heat,    composite_score, method = "spearman")
  ) %>%
  pivot_longer(everything(), names_to = "metric", values_to = "rho_point")

set.seed(20261003)
n_draw <- 4000
draws <- map_dfr(seq_len(nrow(fam_mm)), function(i) {
  f <- fam_mm$family[i]
  b <- MASS::mvrnorm(n_draw, mu = c(fam_mm$b0[i], fam_mm$b1[i]), Sigma = blup_cov[[f]])
  tibble(draw = seq_len(n_draw), family = f, b0 = b[, 1], b1 = b[, 2])
}) %>%
  left_join(fam_mm %>% select(family, composite_score), by = "family")

mm_draw_cor <- draws %>%
  group_by(draw) %>%
  summarise(
    Ambient         = cor(b0,      composite_score, method = "spearman"),
    `36C`           = cor(b0 + b1, composite_score, method = "spearman"),
    `36C - Ambient` = cor(b1,      composite_score, method = "spearman"),
    .groups = "drop"
  ) %>%
  pivot_longer(-draw, names_to = "metric", values_to = "rho")

mm_cor <- mm_draw_cor %>%
  group_by(metric) %>%
  summarise(rho_median = median(rho),
            lower_95   = quantile(rho, 0.025),
            upper_95   = quantile(rho, 0.975),
            prob_neg   = mean(rho < 0),
            .groups = "drop") %>%
  left_join(mm_point, by = "metric") %>%
  mutate(metric = factor(metric, levels = c("Ambient", "36C", "36C - Ambient"))) %>%
  arrange(metric) %>%
  select(metric, rho_point, rho_median, lower_95, upper_95, prob_neg)

knitr::kable(mm_cor, digits = 2)
```

```{r mm-survival-cor-plot, fig.width=8, fig.height=3.5}
ggplot(mm_draw_cor %>%
         mutate(metric = factor(metric, levels = c("Ambient", "36C", "36C - Ambient"))),
       aes(x = rho)) +
  geom_histogram(binwidth = 1 / 30, fill = "grey60") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_wrap(~ metric) +
  labs(x = "Spearman rho with composite survival score",
       y = "Draws",
       caption = "Distribution reflects uncertainty in family ATPase estimates only, not sampling of families.") +
  theme_bw()
```

These intervals reflect only measurement uncertainty within families. They do
not include the larger uncertainty from having only 9 families, so they are
narrower than a full test against survival would be.

Summary of the mixed-model results:

- **Most variation is among oysters, not families.** The between-family share
  (ICC) is 0.08 at Ambient and 0.13 at 36C, so a 15-oyster family mean has a
  reliability of only 0.56 (Ambient) and 0.70 (36C). Reaching 0.8 would take
  about 47 and 27 oysters per family. Ambient is also more variable among
  oysters than 36C (residual SD 1.47 vs 0.96; likelihood ratio p < 0.001).
- **ATPase drops at 36C overall, but there is no evidence families differ in
  how much.** The population heat effect is -0.43 (p = 0.04), but adding a
  family-specific heat response does not improve the model (p = 0.40). Once
  shrunk, every family's heat response is negative (-0.05 to -0.88, versus raw
  differences of -1.74 to +0.79). The apparent gains in families 6 and 2 are
  mostly noise. That explains why the heat response does not track survival
  under any metric above.
- **Baseline activity carries the survival signal.** The model-estimated
  Ambient activity correlates with survival at rho = -0.67 (median across draws
  -0.55, 95% interval -0.83 to -0.03; negative in 98% of draws), slightly
  stronger than the median-based -0.53. 36C activity is weaker in the same
  direction (-0.37), and the heat response does not align with survival (0.23,
  interval spanning zero).
- **This remains a hypothesis to test.** The intervals cover measurement
  uncertainty within families but not the uncertainty from sampling only 9
  families, which dominates. The negative baseline association is the one
  ATPase measure worth carrying forward, prespecified, to new families or the
  multi-assay model.
