---
title: "00.00-mgig-A2-dilutions_tests.Rmd"
author: "Sam White"
date: "2026-01-12"
output: 
  github_document:
    toc: true
    number_sections: true
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
bibliography: references.bib
---
# BACKGROUND

This was the initial testing of the [Glycogen-Glo Assay (Promega)](https://github.com/RobertsLab/resources/blob/master/protocols/Commercial_Protocols/Promega_Glycogen_Glo_Assay.pdf) (GitHub; PDF) using _M.gigas_ (Pacific oyster) ctenidia homegenate from [20251210](https://robertslab.github.io/sams-notebook/posts/2025/2025-12-10-Tissue-Collection-and-Homogenization---M.gigas-Ctenidia-for-Glycogen-Glo-Assay/) (notebook entry).

The idea was to identify a range (i.e. a very rough min/max) of starting tissue amounts that will fall within the range of the Glycogen-Glo standard curves. Additionally, this was an initial walk-through of using this kit to help figure out a general protocol that identified any tips/tricks/pitfalls to aid in developing a Standard Operating Protocol (SOP) for the [Roberts Lab Handbook](https://robertslab.github.io/resources/).

## Important note(s)

Insoluble material was _not_ pelleted prior to beginning assay, which led to pipetting issues. Dilution factor `2` was skipped in this analysis due to some of these issues.




```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = FALSE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = ""         # Prevents appending '##' to beginning of lines in code output
)
```

# DATA

Data contains two CSV files:
1. Plate layout: Specifies the contents of each well in the 96-well plate.
2. Raw luminescence readings: Contains the luminescence values measured from the plate.

There are two standard curves on the plate. Both curves have three replicates each:
- Glycogen Standard Curve: Rows E, F, G (5, 6, 7), Columns 1-5
- Glucose Standard Curve: Rows E, F, G (5, 6, 7), Columns 8-12

There is a single sample (`A2`) with six dilution factors for both glycogen and glucose measurements. Dilutions were run in duplicate:
- Glycogen Samples: Rows A and B (1, 2), Columns 1-6
- Glucose Samples: Rows C and D (3, 4), Columns 1-6

Sample name layouts are as follows:

- `<sample>-<analyte>-<tissue amount in mg>-df.<dilution factor>`

E.g. `A2-glyc-50-df.10`

Where:
- `<sample>`: Sample identifier (e.g., A2)
- `<analyte>`: Either "glyc" for glycogen or "gluc" for glucose
- `<tissue amount in mg>`: Amount of tissue _originally collected_ in mg.
- `<dilution factor>`: Dilution factor applied during sample preparation. E.g a dilution factor of `1` means no dilution, `10` means a 1:10 dilution, etc.



```{r data, include=TRUE}

plate_layout <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/layout-Gen5-20260112-mgig-glycogen_glo-dilutions_test.csv", header = FALSE)
raw_luminescence <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/raw_lum-Gen5-20260112-mgig-glycogen_glo-dilutions_test.csv", header = FALSE)

print("Plate layout:")
str(plate_layout)

cat("\n\n")

print("Raw luminescence:")
str(raw_luminescence)
```

# STANDARD CURVES

## Glycogen Standard Curve

### Extract luminescence data

```{r extract-glycogen-data, include=TRUE}
# Extract glycogen standard curve data from plate layout and raw luminescence.
# Rows E, F, G (rows 5, 6, 7) correspond to glycogen standards
# Columns 1-5 contain the standard curve concentrations

# Extract concentration values from plate layout (row 5, columns 1-5)
# Parse concentration from labels like "STD-glyc-20" -> 20
concentrations <- as.numeric(gsub("STD-glyc-", "", plate_layout[5, 1:5]))

# Extract standard curve luminescence values for each replicate from raw_luminescence
row_E <- as.numeric(raw_luminescence[5, 1:5])  # Row 5 (E)
row_F <- as.numeric(raw_luminescence[6, 1:5])  # Row 6 (F)
row_G <- as.numeric(raw_luminescence[7, 1:5])  # Row 7 (G)

# Extract sample data from rows A and B (columns 1, 3, 4, 5, 6)
# Parse dilution factors from labels like "A2-glyc-50-df.10" -> 10
# Skip column 2 (df.2)
sample_cols <- c(1, 3, 4, 5, 6)
dilution_factors_row_A <- as.numeric(gsub(".*-df\\.", "", plate_layout[1, sample_cols]))
dilution_factors_row_B <- as.numeric(gsub(".*-df\\.", "", plate_layout[2, sample_cols]))
sample_luminescence_row_A <- as.numeric(raw_luminescence[1, sample_cols])
sample_luminescence_row_B <- as.numeric(raw_luminescence[2, sample_cols])

# Combine the two rows of samples
dilution_factors <- c(dilution_factors_row_A, dilution_factors_row_B)
sample_luminescence <- c(sample_luminescence_row_A, sample_luminescence_row_B)
```


### Calculate summary statistics and linear regression
```{r glycogen-summary-data, include=TRUE}
#| results: hold


# Calculate mean and standard error for each concentration
means <- numeric(length(concentrations))
std_errors <- numeric(length(concentrations))
std_devs <- numeric(length(concentrations))

for (i in 1:length(concentrations)) {
  values <- c(row_E[i], row_F[i], row_G[i])
  means[i] <- mean(values)
  std_devs[i] <- sd(values)
  std_errors[i] <- sd(values) / sqrt(length(values))
}

# Create data frame for plotting
glycogen_summary_data <- data.frame(
  concentration = concentrations,
  mean_luminescence = means,
  se = std_errors,
  sd = std_devs,
  cv = (std_devs / means) * 100
)

# Calculate linear regression and R-squared
lm_model <- lm(mean_luminescence ~ concentration, data = glycogen_summary_data)
r_squared <- summary(lm_model)$r.squared
slope <- coef(lm_model)[2]
intercept <- coef(lm_model)[1]

# Create sample data frame
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Get unique dilution factors
unique_dilution_factors <- unique(dilution_factors_row_A)

# Calculate mean and SE for each dilution factor across replicates
sample_mean_lum <- numeric(length(unique_dilution_factors))
sample_se_lum <- numeric(length(unique_dilution_factors))
sample_mean_glyc_conc <- numeric(length(unique_dilution_factors))

for (i in 1:length(unique_dilution_factors)) {
  df_val <- unique_dilution_factors[i]
  # Get luminescence values for this dilution factor from both rows
  lum_values <- c(sample_luminescence_row_A[dilution_factors_row_A == df_val],
                  sample_luminescence_row_B[dilution_factors_row_B == df_val])
  sample_mean_lum[i] <- mean(lum_values)
  sample_se_lum[i] <- sd(lum_values) / sqrt(length(lum_values))
  # Calculate concentration from mean luminescence
  sample_mean_glyc_conc[i] <- (sample_mean_lum[i] - intercept) / slope
}

sample_data <- data.frame(
  dilution_factor = unique_dilution_factors,
  mean_luminescence = sample_mean_lum,
  se = sample_se_lum,
  glycogen_conc = sample_mean_glyc_conc,
  label = paste0("df.", unique_dilution_factors)
)
```

### Plot Glycogen Standard Curve with Sample Points
```{r plot-glycogen-data, fig.width=10, fig.height=6, include=TRUE}
# Create the plot
plot <- ggplot(glycogen_summary_data, aes(x = concentration, y = mean_luminescence)) +
  geom_smooth(aes(linetype = "Std Curve Best Fit Line"), method = "lm", se = FALSE, 
              color = "coral", linewidth = 1) +
  geom_errorbar(aes(ymin = mean_luminescence - se, ymax = mean_luminescence + se),
                width = 0.1, linewidth = 1, color = "darkblue") +
  geom_point(aes(color = "Standard Curve", shape = "Standard Curve"), size = 4) +
  geom_errorbar(data = sample_data, aes(x = glycogen_conc, y = mean_luminescence,
                ymin = mean_luminescence - se, ymax = mean_luminescence + se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = sample_data, aes(x = glycogen_conc, y = mean_luminescence,
             color = label, shape = label), size = 4) +
  scale_color_manual(name = "",
                     values = c("Standard Curve" = "steelblue",
                                "df.1" = "darkred",
                                "df.2" = "darkorange",
                                "df.20" = "darkgreen",
                                "df.200" = "purple",
                                "df.2000" = "brown",
                                "df.0" = "pink")) +
  scale_shape_manual(name = "",
                     values = c("Standard Curve" = 16,
                                "df.1" = 17,
                                "df.2" = 15,
                                "df.20" = 18,
                                "df.200" = 8,
                                "df.2000" = 4,
                                "df.0" = 3)) +
  scale_linetype_manual(name = "",
                        values = c("Std Curve Best Fit Line" = "dashed")) +
  annotate("label", x = max(glycogen_summary_data$concentration) * 0.75, 
           y = max(glycogen_summary_data$mean_luminescence) * 0.15,
           label = sprintf("y = %.2fx + %.2f\nR² = %.4f", slope, intercept, r_squared),
           size = 3.5, fontface = "bold", fill = "white", 
           color = "coral", label.padding = unit(0.3, "lines")) +
  labs(
    title = "Glycogen Standard Curve",
    x = "Glycogen Concentration (µg/µL)",
    y = "Luminescence",
    caption = "Error bars represent standard error of the mean (SEM)"
  ) +
  scale_x_continuous(breaks = concentrations) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_line(linetype = "dashed", color = "grey70")
  )

# Save the plot
ggsave("../output/00.00-mgig-A2-dilutions_tests.Rmd/glycogen_plot.png", plot, width = 10, height = 6, dpi = 300)

# Display the plot
plot

# Print summary statistics
cat("Glycogen Standard Curve Summary:\n")
cat(rep("=", 50), "\n", sep = "")
for (i in 1:nrow(glycogen_summary_data)) {
  cat(sprintf("Concentration: %g µg/µL\n", glycogen_summary_data$concentration[i]))
  cat(sprintf("  Mean Luminescence: %.2f\n", glycogen_summary_data$mean_luminescence[i]))
  cat(sprintf("  Standard Error: %.2f\n", glycogen_summary_data$se[i]))
  cat(sprintf("  CV%%: %.2f%%\n\n", glycogen_summary_data$cv[i]))
}
```

## Glucose Standard Curve

### Extract luminescence data

```{r extract-glucose-data, include=TRUE}
# Extract glucose standard curve data from plate layout and raw luminescence.
# Rows E, F, G (rows 5, 6, 7) correspond to glucose standards
# Columns 8-12 contain the standard curve concentrations

# Extract concentration values from plate layout (row 5, columns 8-12)
# Parse concentration from labels like "STD-gluc-100" -> 100
gluc_concentrations <- as.numeric(gsub("STD-gluc-", "", plate_layout[5, 8:12]))

# Extract standard curve luminescence values for each replicate from raw_luminescence
gluc_row_E <- as.numeric(raw_luminescence[5, 8:12])  # Row 5 (E)
gluc_row_F <- as.numeric(raw_luminescence[6, 8:12])  # Row 6 (F)
gluc_row_G <- as.numeric(raw_luminescence[7, 8:12])  # Row 7 (G)

# Extract sample data from rows C and D (rows 3 and 4, columns 1, 3, 4, 5, 6)
# Parse dilution factors from labels like "A2-gluc-50-df.10" -> 10
# Skip column 2 (df.2)
gluc_sample_cols <- c(1, 3, 4, 5, 6)
gluc_dilution_factors_row_C <- as.numeric(gsub(".*-df\\.", "", plate_layout[3, gluc_sample_cols]))
gluc_dilution_factors_row_D <- as.numeric(gsub(".*-df\\.", "", plate_layout[4, gluc_sample_cols]))
gluc_sample_luminescence_row_C <- as.numeric(raw_luminescence[3, gluc_sample_cols])
gluc_sample_luminescence_row_D <- as.numeric(raw_luminescence[4, gluc_sample_cols])

# Combine the two rows of samples
gluc_dilution_factors <- c(gluc_dilution_factors_row_C, gluc_dilution_factors_row_D)
gluc_sample_luminescence <- c(gluc_sample_luminescence_row_C, gluc_sample_luminescence_row_D)
```


### Calculate summary statistics and linear regression
```{r glucose-summary-data, include=TRUE}
#| results: hold


# Calculate mean and standard error for each concentration
gluc_means <- numeric(length(gluc_concentrations))
gluc_std_errors <- numeric(length(gluc_concentrations))
gluc_std_devs <- numeric(length(gluc_concentrations))

for (i in 1:length(gluc_concentrations)) {
  values <- c(gluc_row_E[i], gluc_row_F[i], gluc_row_G[i])
  gluc_means[i] <- mean(values)
  gluc_std_devs[i] <- sd(values)
  gluc_std_errors[i] <- sd(values) / sqrt(length(values))
}

# Create data frame for plotting
glucose_summary_data <- data.frame(
  concentration = gluc_concentrations,
  mean_luminescence = gluc_means,
  se = gluc_std_errors,
  sd = gluc_std_devs,
  cv = (gluc_std_devs / gluc_means) * 100
)

# Calculate linear regression and R-squared
gluc_lm_model <- lm(mean_luminescence ~ concentration, data = glucose_summary_data)
gluc_r_squared <- summary(gluc_lm_model)$r.squared
gluc_slope <- coef(gluc_lm_model)[2]
gluc_intercept <- coef(gluc_lm_model)[1]

# Create sample data frame
# Calculate glucose concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Get unique dilution factors
gluc_unique_dilution_factors <- unique(gluc_dilution_factors_row_C)

# Calculate mean and SE for each dilution factor across replicates
gluc_sample_mean_lum <- numeric(length(gluc_unique_dilution_factors))
gluc_sample_se_lum <- numeric(length(gluc_unique_dilution_factors))
gluc_sample_mean_conc <- numeric(length(gluc_unique_dilution_factors))

for (i in 1:length(gluc_unique_dilution_factors)) {
  df_val <- gluc_unique_dilution_factors[i]
  # Get luminescence values for this dilution factor from both rows
  lum_values <- c(gluc_sample_luminescence_row_C[gluc_dilution_factors_row_C == df_val],
                  gluc_sample_luminescence_row_D[gluc_dilution_factors_row_D == df_val])
  gluc_sample_mean_lum[i] <- mean(lum_values)
  gluc_sample_se_lum[i] <- sd(lum_values) / sqrt(length(lum_values))
  # Calculate concentration from mean luminescence
  gluc_sample_mean_conc[i] <- (gluc_sample_mean_lum[i] - gluc_intercept) / gluc_slope
}

gluc_sample_data <- data.frame(
  dilution_factor = gluc_unique_dilution_factors,
  mean_luminescence = gluc_sample_mean_lum,
  se = gluc_sample_se_lum,
  glucose_conc = gluc_sample_mean_conc,
  label = paste0("df.", gluc_unique_dilution_factors)
)
```

### Plot Glucose Standard Curve with Sample Points
```{r plot-glucose-data, fig.width=10, fig.height=6, include=TRUE}
# Create the plot
gluc_plot <- ggplot(glucose_summary_data, aes(x = concentration, y = mean_luminescence)) +
  geom_smooth(aes(linetype = "Std Curve Best Fit Line"), method = "lm", se = FALSE, 
              color = "coral", linewidth = 1) +
  geom_errorbar(aes(ymin = mean_luminescence - se, ymax = mean_luminescence + se),
                width = 0.1, linewidth = 1, color = "darkblue") +
  geom_point(aes(color = "Standard Curve", shape = "Standard Curve"), size = 4) +
  geom_errorbar(data = gluc_sample_data, aes(x = glucose_conc, y = mean_luminescence,
                ymin = mean_luminescence - se, ymax = mean_luminescence + se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = gluc_sample_data, aes(x = glucose_conc, y = mean_luminescence,
             color = label, shape = label), size = 4) +
  scale_color_manual(name = "",
                     values = c("Standard Curve" = "steelblue",
                                "df.1" = "darkred",
                                "df.20" = "darkgreen",
                                "df.200" = "purple",
                                "df.2000" = "brown",
                                "df.0" = "pink")) +
  scale_shape_manual(name = "",
                     values = c("Standard Curve" = 16,
                                "df.1" = 17,
                                "df.20" = 18,
                                "df.200" = 8,
                                "df.2000" = 4,
                                "df.0" = 3)) +
  scale_linetype_manual(name = "",
                        values = c("Std Curve Best Fit Line" = "dashed")) +
  annotate("label", x = max(glucose_summary_data$concentration) * 0.75, 
           y = max(glucose_summary_data$mean_luminescence) * 0.15,
           label = sprintf("y = %.2fx + %.2f\nR² = %.4f", gluc_slope, gluc_intercept, gluc_r_squared),
           size = 3.5, fontface = "bold", fill = "white", 
           color = "coral", label.padding = unit(0.3, "lines")) +
  labs(
    title = "Glucose Standard Curve",
    x = "Glucose Concentration (µg/µL)",
    y = "Luminescence",
    caption = "Error bars represent standard error of the mean (SEM)"
  ) +
  scale_x_continuous(breaks = gluc_concentrations) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_line(linetype = "dashed", color = "grey70")
  )

# Save the plot
ggsave("../output/00.00-mgig-A2-dilutions_tests.Rmd/glucose_plot.png", gluc_plot, width = 10, height = 6, dpi = 300)

# Display the plot
gluc_plot

# Print summary statistics
cat("Glucose Standard Curve Summary:\n")
cat(rep("=", 50), "\n", sep = "")
for (i in 1:nrow(glucose_summary_data)) {
  cat(sprintf("Concentration: %g µg/µL\n", glucose_summary_data$concentration[i]))
  cat(sprintf("  Mean Luminescence: %.2f\n", glucose_summary_data$mean_luminescence[i]))
  cat(sprintf("  Standard Error: %.2f\n", glucose_summary_data$se[i]))
  cat(sprintf("  CV%%: %.2f%%\n\n", glucose_summary_data$cv[i]))
}
```

# SAMPLE A2 QUANTIFICATION

```{r sample-quantification, include=TRUE}
#| results: hold

# Function to calculate concentration from luminescence using linear model
# Returns concentration or "OUT OF RANGE" message
calculate_concentration <- function(luminescence, slope, intercept, min_conc, max_conc, units) {
  # Calculate concentration from y = mx + b -> x = (y - b) / m
  concentration <- (luminescence - intercept) / slope
  
  # Check if within range
  if (concentration < min_conc) {
    return("OUT OF RANGE - TOO LOW")
  } else if (concentration > max_conc) {
    return("OUT OF RANGE - TOO HIGH")
  } else {
    return(sprintf("%.4f %s", concentration, units))
  }
}

# Glycogen calculations
cat("GLYCOGEN ASSAY RESULTS (Sample A2):\n")
cat(rep("=", 100), "\n", sep = "")
cat(sprintf("%-15s %-20s %-25s %s\n", "Dilution Factor", "Luminescence", "Calculated Glycogen", "Total Glycogen"))
cat(rep("-", 100), "\n", sep = "")

glycogen_min <- min(concentrations)
glycogen_max <- max(concentrations)

# Sort by dilution factor
sample_data_sorted <- sample_data[order(sample_data$dilution_factor), ]

for (i in 1:nrow(sample_data_sorted)) {
  result <- calculate_concentration(
    sample_data_sorted$mean_luminescence[i],
    slope,
    intercept,
    glycogen_min,
    glycogen_max,
    "µg/µL"
  )
  
  # Calculate total glycogen (concentration in original sample before dilution)
  conc <- (sample_data_sorted$mean_luminescence[i] - intercept) / slope
  df <- sample_data_sorted$dilution_factor[i]
  # Handle division by zero for df=0
  if (df == 0) {
    total_result <- "N/A"
  } else {
    total_glyc <- conc * df
    # Check if within range after accounting for dilution
    if (total_glyc < glycogen_min * df) {
      total_result <- "OUT OF RANGE - TOO LOW"
    } else if (total_glyc > glycogen_max * df) {
      total_result <- "OUT OF RANGE - TOO HIGH"
    } else {
      total_result <- sprintf("%.4f µg/µL", total_glyc)
    }
  }
  
  cat(sprintf("%-15s %-20.0f %-25s %s\n", 
              sample_data_sorted$label[i], 
              sample_data_sorted$mean_luminescence[i], 
              result,
              total_result))
}

cat("\n\n")

# Glucose calculations
cat("GLUCOSE ASSAY RESULTS (Sample A2):\n")
cat(rep("=", 100), "\n", sep = "")
cat(sprintf("%-15s %-20s %-25s %s\n", "Dilution Factor", "Luminescence", "Calculated Glucose", "Total Glucose"))
cat(rep("-", 100), "\n", sep = "")

glucose_min <- min(gluc_concentrations)
glucose_max <- max(gluc_concentrations)

# Sort by dilution factor
gluc_sample_data_sorted <- gluc_sample_data[order(gluc_sample_data$dilution_factor), ]

for (i in 1:nrow(gluc_sample_data_sorted)) {
  result <- calculate_concentration(
    gluc_sample_data_sorted$mean_luminescence[i],
    gluc_slope,
    gluc_intercept,
    glucose_min,
    glucose_max,
    "µM"
  )
  
  # Calculate total glucose (concentration in original sample before dilution)
  conc <- (gluc_sample_data_sorted$mean_luminescence[i] - gluc_intercept) / gluc_slope
  df <- gluc_sample_data_sorted$dilution_factor[i]
  # Handle division by zero for df=0
  if (df == 0) {
    total_result <- "N/A"
  } else {
    total_gluc <- conc * df
    # Check if within range after accounting for dilution
    if (total_gluc < glucose_min * df) {
      total_result <- "OUT OF RANGE - TOO LOW"
    } else if (total_gluc > glucose_max * df) {
      total_result <- "OUT OF RANGE - TOO HIGH"
    } else {
      total_result <- sprintf("%.4f µM", total_gluc)
    }
  }
  
  cat(sprintf("%-15s %-20.0f %-25s %s\n", 
              gluc_sample_data_sorted$label[i], 
              gluc_sample_data_sorted$mean_luminescence[i], 
              result,
              total_result))
}
```



# DISCUSSION

