---
title: "Gen5-20260218-mgig-glycogenglo"
author: "Hazel Abrahamson-Amerine"
date: "2026-03-06"
output:  
    github_document:
      toc: true
      number_sections: true
---

```{r setup, include=TRUE}
library(knitr)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = "",         # Prevents appending '##' to beginning of lines in code output
  results = 'hold'     # Holds output so it's all printed together after code chunk
)
```

```{r data, include=TRUE}
#Samples C7.df50, C7.df100, D2, E4-E7
plate_layout <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/layout-Gen5-20260218-mgig-glycogenglo.csv", header = FALSE)
raw_luminescence <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/raw_lum-Gen5-20260218-mgig-glycogenglo.csv", header = FALSE)
weights <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/20251210-mgig-ctenidia_weights-glycogen_assay.csv")

cat("Plate layout:\n")
str(plate_layout)

cat("\n\n")

cat("Raw luminescence:\n")
str(raw_luminescence)

#standard curve for sample E5.df50
plate_layout2 <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/layout-Gen5-20260330A-tempprelim-mgig-glycogengglo.csv", header = FALSE)
raw_luminescence2 <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/raw_lum-Gen5-20260330A-tempprelim-mgig-glycogengglo.csv", header = FALSE)

cat("Plate layout2:\n")
str(plate_layout2)

cat("\n\n")

cat("Raw luminescence2:\n")
str(raw_luminescence2)

#sample E5.df50 - on same plate as temp stress prelim samples
plate_layout3 <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/layout-Gen5-20260330B-tempprelim-mgig-glycogengglo.csv", header = FALSE)
raw_luminescence3 <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/raw_lum-Gen5-20260330B-tempprelim-mgig-glycogengglo.csv", header = FALSE)

cat("Plate layout3:\n")
str(plate_layout3)

cat("\n\n")

cat("Raw luminescence3:\n")
str(raw_luminescence3)
```

# Standard curves for 20260218

## Glycogen Standard Curve

### Extract glycogen luminescence data

```{r glycogen standard data, include=TRUE}
# Extract glycogen standard curve data from plate layout and raw luminescence.
# Rows F, G, H (rows 6, 7, 8) correspond to glycogen standards
# Columns 1-5 contain the standard curve concentrations

# Extract concentration values from plate layout (row 6, columns 1-5)
# Parse concentration from labels like "STD-glyc-20" -> 20
glyc_concentrations <- as.numeric(gsub("STD-glyc-", "", plate_layout[6, 1:5]))
glyc_concentrations

# Extract standard curve luminescence values for each replicate from raw_luminescence
glyc_row_F <- as.numeric(raw_luminescence[6, 1:5])  # Row 6 (F)
glyc_row_G <- as.numeric(raw_luminescence[7, 1:5])  # Row 7 (G)
glyc_row_H <- as.numeric(raw_luminescence[8, 1:5])  # Row 8 (H)
```

```{r glycogen samples, include=TRUE}
#Extract glycogen sample data
glyc_sample_cols1 <- c(1,2,3)
glyc_C7_dilution50 <- as.numeric(gsub(".*-df\\.", "", plate_layout[1, 1]))
glyc_C7_luminescence50 <- as.numeric(raw_luminescence[1, glyc_sample_cols1])

glyc_C7_dilution100 <- as.numeric(gsub(".*-df\\.", "", plate_layout[2, 1]))
glyc_C7_luminescence100 <- as.numeric(raw_luminescence[2, glyc_sample_cols1])

glyc_D2_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout[3, 1]))
glyc_D2_luminescence <- as.numeric(raw_luminescence[3, glyc_sample_cols1])

glyc_E4_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout[4, 1]))
glyc_E4_luminescence <- as.numeric(raw_luminescence[4, glyc_sample_cols1])

glyc_E5_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout[5, 1]))
glyc_E5_luminescence <- as.numeric(raw_luminescence[5, glyc_sample_cols1])


glyc_sample_cols2 <- c(4,5,6)
glyc_E6_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout[1, 4]))
glyc_E6_luminescence <- as.numeric(raw_luminescence[1, glyc_sample_cols2])

glyc_E7_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout[2, 4]))
glyc_E7_luminescence <- as.numeric(raw_luminescence[2, glyc_sample_cols2])
```

### Glycogen standard curve summary statistics and linear regression

```{r glycogen summary statistics and linear regression,include=TRUE}
# Calculate mean and standard error for each concentration
glyc_means <- numeric(length(glyc_concentrations))
glyc_std_errors <- numeric(length(glyc_concentrations))
glyc_std_devs <- numeric(length(glyc_concentrations))

for (i in 1:length(glyc_concentrations)) {
  glyc_values <- c(glyc_row_F[i], glyc_row_G[i], glyc_row_H[i])
  glyc_means[i] <- mean(glyc_values)
  glyc_std_devs[i] <- sd(glyc_values)
  glyc_std_errors[i] <- sd(glyc_values) / sqrt(length(glyc_values))
}

# Create data frame for plotting
glycogen_summary_data <- data.frame(
  glyc_concentration = glyc_concentrations,
  glyc_mean_luminescence = glyc_means,
  glyc_se = glyc_std_errors,
  glyc_sd = glyc_std_devs,
  glyc_cv = (glyc_std_devs / glyc_means) * 100
)

# Calculate linear regression and R-squared
lm_model <- lm(glyc_mean_luminescence ~ glyc_concentration, data = glycogen_summary_data)
glyc_r_squared <- summary(lm_model)$r.squared
glyc_slope <- coef(lm_model)[2]
glyc_intercept <- coef(lm_model)[1]
```

### Calculate sample glycogen levels

```{r add C7 df 50 glycogen to standard curve, include=TRUE}
# Create sample data frame
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_C7_mean_lum50 <- numeric(length(glyc_C7_dilution50))
glyc_C7_se_lum50 <- numeric(length(glyc_C7_dilution50))
glyc_C7_mean_conc50 <- numeric(length(glyc_C7_dilution50))

for (i in 1:length(glyc_C7_dilution50)) {
  df_val <- glyc_C7_dilution50[i]
  # Get luminescence values for this dilution factor from both rows
   glyc_C7_lum_values50 <- c(glyc_C7_luminescence50[glyc_C7_dilution50 == df_val])
  glyc_C7_mean_lum50[i] <- mean( glyc_C7_lum_values50)
  glyc_C7_se_lum50[i] <- sd( glyc_C7_lum_values50) / sqrt(length( glyc_C7_lum_values50))
  # Calculate concentration from mean luminescence
  glyc_C7_mean_conc50[i] <- (glyc_C7_mean_lum50[i] - glyc_intercept) / glyc_slope
}

glyc_C7_data50 <- data.frame(
  glyc_C7_dilution50_factor = glyc_C7_dilution50,
  glyc_C7_mean_luminescence50 = glyc_C7_mean_lum50,
  glyc_C7_se50 = glyc_C7_se_lum50,
  glyc_C7_conc50 =  glyc_C7_mean_conc50,
  label = paste0("C7df.", glyc_C7_dilution50)
)
glyc_C7_mean_lum50
glyc_C7_mean_conc50
```

```{r add C7 df 100 glycogen to standard curve, include=TRUE}
# Create sample data frame
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_C7_mean_lum100 <- numeric(length(glyc_C7_dilution100))
glyc_C7_se_lum100 <- numeric(length(glyc_C7_dilution100))
glyc_C7_mean_conc100 <- numeric(length(glyc_C7_dilution100))

for (i in 1:length(glyc_C7_dilution100)) {
  df_val <- glyc_C7_dilution100[i]
  # Get luminescence values for this dilution factor from both rows
   glyc_C7_lum_values100 <- c(glyc_C7_luminescence100[glyc_C7_dilution100 == df_val])
  glyc_C7_mean_lum100[i] <- mean( glyc_C7_lum_values100)
  glyc_C7_se_lum100[i] <- sd( glyc_C7_lum_values100) / sqrt(length( glyc_C7_lum_values100))
  # Calculate concentration from mean luminescence
  glyc_C7_mean_conc100[i] <- (glyc_C7_mean_lum100[i] - glyc_intercept) / glyc_slope
}

glyc_C7_data100 <- data.frame(
  glyc_C7_dilution100_factor = glyc_C7_dilution100,
  glyc_C7_mean_luminescence100 = glyc_C7_mean_lum100,
  glyc_C7_se100 = glyc_C7_se_lum100,
  glyc_C7_conc100 =  glyc_C7_mean_conc100,
  label = paste0("C7df.", glyc_C7_dilution100)
)
glyc_C7_mean_lum100
glyc_C7_mean_conc100
```

```{r add D2 glycogen to standard curve, include=TRUE}
# Create sample data frame
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_D2_mean_lum <- numeric(length(glyc_D2_dilution))
glyc_D2_se_lum <- numeric(length(glyc_D2_dilution))
glyc_D2_mean_conc <- numeric(length(glyc_D2_dilution))

for (i in 1:length(glyc_D2_dilution)) {
  df_val <- glyc_D2_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_D2_lum_values <- c(glyc_D2_luminescence[glyc_D2_dilution == df_val])
  glyc_D2_mean_lum[i] <- mean(glyc_D2_lum_values)
  glyc_D2_se_lum[i] <- sd(glyc_D2_lum_values) / sqrt(length(glyc_D2_lum_values))
  # Calculate concentration from mean luminescence
  glyc_D2_mean_conc[i] <- (glyc_D2_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_D2_data <- data.frame(
  glyc_D2_dilution_factor = glyc_D2_dilution,
  glyc_D2_mean_luminescence = glyc_D2_mean_lum,
  glyc_D2_se = glyc_D2_se_lum,
  glyc_D2_conc =  glyc_D2_mean_conc,
  label = paste0("D2df.", glyc_D2_dilution)
)
glyc_D2_mean_lum
glyc_D2_mean_conc
```


```{r add E4 glycogen to standard curve, include=TRUE}
# Create sample data frame
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_E4_mean_lum <- numeric(length(glyc_E4_dilution))
glyc_E4_se_lum <- numeric(length(glyc_E4_dilution))
glyc_E4_mean_conc <- numeric(length(glyc_E4_dilution))

for (i in 1:length(glyc_E4_dilution)) {
  df_val <- glyc_E4_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_E4_lum_values <- c(glyc_E4_luminescence[glyc_E4_dilution == df_val])
  glyc_E4_mean_lum[i] <- mean(glyc_E4_lum_values)
  glyc_E4_se_lum[i] <- sd(glyc_E4_lum_values) / sqrt(length(glyc_E4_lum_values))
  # Calculate concentration from mean luminescence
  glyc_E4_mean_conc[i] <- (glyc_E4_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_E4_data <- data.frame(
  glyc_E4_dilution_factor = glyc_E4_dilution,
  glyc_E4_mean_luminescence = glyc_E4_mean_lum,
  glyc_E4_se = glyc_E4_se_lum,
  glyc_E4_conc =  glyc_E4_mean_conc,
  label = paste0("E4df.", glyc_E4_dilution)
)
glyc_E4_mean_lum
glyc_E4_mean_conc
```

```{r add E5 glycogen to standard curve, include=TRUE}
# Create sample data frame
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_E5_mean_lum <- numeric(length(glyc_E5_dilution))
glyc_E5_se_lum <- numeric(length(glyc_E5_dilution))
glyc_E5_mean_conc <- numeric(length(glyc_E5_dilution))

for (i in 1:length(glyc_E5_dilution)) {
  df_val <- glyc_E5_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_E5_lum_values <- c(glyc_E5_luminescence[glyc_E5_dilution == df_val])
  glyc_E5_mean_lum[i] <- mean(glyc_E5_lum_values)
  glyc_E5_se_lum[i] <- sd(glyc_E5_lum_values) / sqrt(length(glyc_E5_lum_values))
  # Calculate concentration from mean luminescence
  glyc_E5_mean_conc[i] <- (glyc_E5_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_E5_data <- data.frame(
  glyc_E5_dilution_factor = glyc_E5_dilution,
  glyc_E5_mean_luminescence = glyc_E5_mean_lum,
  glyc_E5_se = glyc_E5_se_lum,
  glyc_E5_conc =  glyc_E5_mean_conc,
  label = paste0("E5df.", glyc_E5_dilution)
)
glyc_E5_mean_lum
glyc_E5_mean_conc
```

```{r add E6 glycogen to standard curve, include=TRUE}
# Create sample data frame
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_E6_mean_lum <- numeric(length(glyc_E6_dilution))
glyc_E6_se_lum <- numeric(length(glyc_E6_dilution))
glyc_E6_mean_conc <- numeric(length(glyc_E6_dilution))

for (i in 1:length(glyc_E6_dilution)) {
  df_val <- glyc_E6_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_E6_lum_values <- c(glyc_E6_luminescence[glyc_E6_dilution == df_val])
  glyc_E6_mean_lum[i] <- mean(glyc_E6_lum_values)
  glyc_E6_se_lum[i] <- sd(glyc_E6_lum_values) / sqrt(length(glyc_E6_lum_values))
  # Calculate concentration from mean luminescence
  glyc_E6_mean_conc[i] <- (glyc_E6_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_E6_data <- data.frame(
  glyc_E6_dilution_factor = glyc_E6_dilution,
  glyc_E6_mean_luminescence = glyc_E6_mean_lum,
  glyc_E6_se = glyc_E6_se_lum,
  glyc_E6_conc =  glyc_E6_mean_conc,
  label = paste0("E6df.", glyc_E6_dilution)
)
glyc_E6_mean_lum
glyc_E6_mean_conc
```

```{r add E7 glycogen to standard curve, include=TRUE}
# Create sample data frame
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_E7_mean_lum <- numeric(length(glyc_E7_dilution))
glyc_E7_se_lum <- numeric(length(glyc_E7_dilution))
glyc_E7_mean_conc <- numeric(length(glyc_E7_dilution))

for (i in 1:length(glyc_E7_dilution)) {
  df_val <- glyc_E7_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_E7_lum_values <- c(glyc_E7_luminescence[glyc_E7_dilution == df_val])
  glyc_E7_mean_lum[i] <- mean(glyc_E7_lum_values)
  glyc_E7_se_lum[i] <- sd(glyc_E7_lum_values) / sqrt(length(glyc_E7_lum_values))
  # Calculate concentration from mean luminescence
  glyc_E7_mean_conc[i] <- (glyc_E7_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_E7_data <- data.frame(
  glyc_E7_dilution_factor = glyc_E7_dilution,
  glyc_E7_mean_luminescence = glyc_E7_mean_lum,
  glyc_E7_se = glyc_E7_se_lum,
  glyc_E7_conc =  glyc_E7_mean_conc,
  label = paste0("E7df.", glyc_E7_dilution)
)
glyc_E7_mean_lum
glyc_E7_mean_conc
```

### Plot glycogen standard curve, sample points

```{r plot glycogen standard curve with samples, include=TRUE}
# Create the plot
glyc_plot <- ggplot(glycogen_summary_data, aes(x = glyc_concentration, y = glyc_mean_luminescence)) +
  geom_smooth(aes(linetype = "Std Curve Best Fit Line"), method = "lm", se = FALSE, 
              color = "coral", linewidth = 1) +
  geom_errorbar(aes(ymin = glyc_mean_luminescence - glyc_se, ymax = glyc_mean_luminescence + glyc_se),
                width = 0.1, linewidth = 1, color = "darkblue") +
  geom_point(aes(color = "Standard Curve", shape = "Standard Curve"), size = 4) +

  geom_errorbar(data = glyc_C7_data50, aes(x = glyc_C7_conc50, y = glyc_C7_mean_luminescence50,
                ymin = glyc_C7_mean_luminescence50 - glyc_C7_se50, ymax = glyc_C7_mean_luminescence50 + glyc_C7_se50,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_C7_data50, aes(x = glyc_C7_conc50, y = glyc_C7_mean_luminescence50,
             color = label, shape = label), size = 4) +

  geom_errorbar(data = glyc_C7_data100, aes(x = glyc_C7_conc100, y = glyc_C7_mean_luminescence100,
                ymin = glyc_C7_mean_luminescence100 - glyc_C7_se100, ymax = glyc_C7_mean_luminescence100 + glyc_C7_se100,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_C7_data100, aes(x = glyc_C7_conc100, y = glyc_C7_mean_luminescence100,
             color = label, shape = label), size = 4) +
  
     geom_errorbar(data = glyc_D2_data, aes(x = glyc_D2_conc, y = glyc_D2_mean_luminescence,
                ymin = glyc_D2_mean_luminescence - glyc_D2_se, ymax = glyc_D2_mean_luminescence + glyc_D2_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_D2_data, aes(x = glyc_D2_conc, y = glyc_D2_mean_luminescence,
             color = label, shape = label), size = 4) +

   geom_errorbar(data = glyc_E4_data, aes(x = glyc_E4_conc, y = glyc_E4_mean_luminescence,
                ymin = glyc_E4_mean_luminescence - glyc_E4_se, ymax = glyc_E4_mean_luminescence + glyc_E4_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_E4_data, aes(x = glyc_E4_conc, y = glyc_E4_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_E5_data, aes(x = glyc_E5_conc, y = glyc_E5_mean_luminescence,
                ymin = glyc_E5_mean_luminescence - glyc_E5_se, ymax = glyc_E5_mean_luminescence + glyc_E5_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_E5_data, aes(x = glyc_E5_conc, y = glyc_E5_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_E6_data, aes(x = glyc_E6_conc, y = glyc_E6_mean_luminescence,
                ymin = glyc_E6_mean_luminescence - glyc_E6_se, ymax = glyc_E6_mean_luminescence + glyc_E6_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_E6_data, aes(x = glyc_E6_conc, y = glyc_E6_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_E7_data, aes(x = glyc_E7_conc, y = glyc_E7_mean_luminescence,
                ymin = glyc_E7_mean_luminescence - glyc_E7_se, ymax = glyc_E7_mean_luminescence + glyc_E7_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_E7_data, aes(x = glyc_E7_conc, y = glyc_E7_mean_luminescence,
             color = label, shape = label), size = 4) +

  
  
  scale_color_manual(name = "",
                     values = c("Standard Curve" = "steelblue",
                                "C7df.50" = "darkred",
                                "C7df.100" = "darkorange",
                                "D2df.20" = "cyan",
                                "E4df.20" = "darkgreen",
                                "E5df.20" = "purple",
                                "E6df.20" = "brown",
                                "E7df.20" = "green3"
                                )) +
  scale_shape_manual(name = "",
                     values = c("Standard Curve" = 16,
                                 "C7df.50" = 17,
                                 "C7df.100" = 15,
                                 "D2df.20" = 2,
                                 "E4df.20" = 18,
                                 "E5df.20" = 8,
                                 "E6df.20" = 4,
                                 "E7df.20" = 5
                                )) +
 scale_linetype_manual(name = "",
                        values = c("Std Curve Best Fit Line" = "dashed")) +
  annotate("label", x = (max(glycogen_summary_data$glyc_concentration) * 1.05), 
           y = (max(glycogen_summary_data$glyc_mean_luminescence) * 0.15),
           label = sprintf("y = %.2fx + %.2f\nR² = %.4f", glyc_slope, glyc_intercept, glyc_r_squared),
           size = 3.5, fontface = "bold", fill = "white", 
           color = "coral", label.padding = unit(0.3, "lines")) +
  labs(
    title = "Glycogen Standard Curve",
    x = "Glycogen Concentration (µg/µL)",
    y = "Luminescence",
    caption = "Error bars represent standard error of the mean (SEM)"
  ) +
  scale_x_continuous(breaks = glyc_concentrations) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_line(linetype = "dashed", color = "grey70")
  )


# Display the plot
glyc_plot
glyc_C7_data50
glyc_C7_data100
glyc_D2_data
glyc_E4_data
glyc_E5_data
glyc_E6_data
glyc_E7_data

# Print summary statistics
cat("glycogen Standard Curve Summary:\n")
cat(rep("=", 50), "\n", sep = "")
for (i in 1:nrow(glycogen_summary_data)) {
  cat(sprintf("Concentration: %g µg/µL\n", glycogen_summary_data$glyc_concentration[i]))
  cat(sprintf("  Mean Luminescence: %.2f\n",glycogen_summary_data$glyc_mean_luminescence[i]))
  cat(sprintf("  Standard Error: %.2f\n", glycogen_summary_data$glyc_se[i]))
  cat(sprintf("  CV%%: %.2f%%\n\n", glycogen_summary_data$glyc_cv[i]))
}
```

### Glycogen summary table

```{r, sample glycogen calculations}
tab <- matrix(c(glyc_C7_dilution50, glyc_C7_mean_lum50, glyc_C7_mean_conc50,  (glyc_C7_dilution50*glyc_C7_mean_conc50), 
                glyc_C7_dilution100, glyc_C7_mean_lum100, glyc_C7_mean_conc100,  (glyc_C7_dilution100*glyc_C7_mean_conc100), 
                glyc_D2_dilution, glyc_D2_mean_lum, glyc_D2_mean_conc,  (glyc_D2_dilution*glyc_D2_mean_conc), 
                glyc_E4_dilution, glyc_E4_mean_lum, glyc_E4_mean_conc,  (glyc_E4_dilution*glyc_E4_mean_conc), 
                glyc_E5_dilution, glyc_E5_mean_lum, glyc_E5_mean_conc,  (glyc_E5_dilution*glyc_E5_mean_conc), 
                glyc_E6_dilution, glyc_E6_mean_lum, glyc_E6_mean_conc,  (glyc_E6_dilution*glyc_E6_mean_conc), 
                glyc_E7_dilution, glyc_E7_mean_lum, glyc_E7_mean_conc,  (glyc_E7_dilution*glyc_E7_mean_conc)), ncol=4, byrow=TRUE)
colnames(tab) <- c('Dilution factor','Luminescence','Calculated Glycogen (ug/uL)', 'Total glycogen (ug/uL)')
rownames(tab) <- c('C7_df50','C7_df100','D2','E4','E5','E6','E7')
tab <- as.table(tab)
tab
```

## Glucose Standard Curve

### Extract glucose luminescence data

```{r glucose standard data, include=TRUE}
# Extract glucose standard curve data from plate layout and raw luminescence.
# Rows F, G, H (rows 6, 7, 8) correspond to glucose standards
# Columns 7-11 contain the standard curve concentrations

# Extract concentration values from plate layout (row 6, columns 7-11)
# Parse concentration from labels like "STD-glu-20" -> 20
glu_concentrations <- as.numeric(gsub("STD-glu-", "", plate_layout[6, 7:11]))
glu_concentrations

# Extract standard curve luminescence values for each replicate from raw_luminescence
glu_row_F <- as.numeric(raw_luminescence[6, 7:11])  # Row 6 (F)
glu_row_G <- as.numeric(raw_luminescence[7, 7:11])  # Row 7 (G)
glu_row_H <- as.numeric(raw_luminescence[8, 7:11])  # Row 8 (H)
```

```{r glucose samples, include=TRUE}
glu_sample_cols1 <- c(7,8,9)
glu_C7_dilution50 <- as.numeric(gsub(".*-df\\.", "", plate_layout[1, 1]))
glu_C7_luminescence50 <- as.numeric(raw_luminescence[1, glu_sample_cols1])

glu_C7_dilution100 <- as.numeric(gsub(".*-df\\.", "", plate_layout[2, 1]))
glu_C7_luminescence100 <- as.numeric(raw_luminescence[2, glu_sample_cols1])

glu_D2_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout[3, 1]))
glu_D2_luminescence <- as.numeric(raw_luminescence[3, glu_sample_cols1])

glu_E4_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout[4, 1]))
glu_E4_luminescence <- as.numeric(raw_luminescence[4, glu_sample_cols1])

glu_E5_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout[5, 1]))
glu_E5_luminescence <- as.numeric(raw_luminescence[5, glu_sample_cols1])


glu_sample_cols2 <- c(10,11,12)
glu_E6_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout[1, 4]))
glu_E6_luminescence <- as.numeric(raw_luminescence[1, glu_sample_cols2])

glu_E7_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout[2, 4]))
glu_E7_luminescence <- as.numeric(raw_luminescence[2, glu_sample_cols2])
```

### Glucose standard curve summary statistics and linear regression

```{r glucose summary statistics and linear regression,include=TRUE}
# Calculate mean and standard error for each concentration
glu_means <- numeric(length(glu_concentrations))
glu_std_errors <- numeric(length(glu_concentrations))
glu_std_devs <- numeric(length(glu_concentrations))

for (i in 1:length(glu_concentrations)) {
  glu_values <- c(glu_row_F[i], glu_row_G[i], glu_row_H[i])
  glu_means[i] <- mean(glu_values)
  glu_std_devs[i] <- sd(glu_values)
  glu_std_errors[i] <- sd(glu_values) / sqrt(length(glu_values))
}

# Create data frame for plotting
glucose_summary_data <- data.frame(
  glu_concentration = glu_concentrations,
  glu_mean_luminescence = glu_means,
  glu_se = glu_std_errors,
  glu_sd = glu_std_devs,
  glu_cv = (glu_std_devs / glu_means) * 100
)

# Calculate linear regression and R-squared
lm_model <- lm(glu_mean_luminescence ~ glu_concentration, data = glucose_summary_data)
glu_r_squared <- summary(lm_model)$r.squared
glu_slope <- coef(lm_model)[2]
glu_intercept <- coef(lm_model)[1]
```

### Calculate sample glucose levels

```{r add C7 df 50 glucose to standard curve, include=TRUE}
# Create sample data frame
# Calculate glucose concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glu_C7_mean_lum50 <- numeric(length(glu_C7_dilution50))
glu_C7_se_lum50 <- numeric(length(glu_C7_dilution50))
glu_C7_mean_conc50 <- numeric(length(glu_C7_dilution50))

for (i in 1:length(glu_C7_dilution50)) {
  df_val <- glu_C7_dilution50[i]
  # Get luminescence values for this dilution factor from both rows
   glu_C7_lum_values50 <- c(glu_C7_luminescence50[glu_C7_dilution50 == df_val])
  glu_C7_mean_lum50[i] <- mean( glu_C7_lum_values50)
  glu_C7_se_lum50[i] <- sd( glu_C7_lum_values50) / sqrt(length( glu_C7_lum_values50))
  # Calculate concentration from mean luminescence
  glu_C7_mean_conc50[i] <- (glu_C7_mean_lum50[i] - glu_intercept) / glu_slope
}

glu_C7_data50 <- data.frame(
  glu_C7_dilution50_factor = glu_C7_dilution50,
  glu_C7_mean_luminescence50 = glu_C7_mean_lum50,
  glu_C7_se50 = glu_C7_se_lum50,
  glu_C7_conc50 =  glu_C7_mean_conc50,
  label = paste0("C7df.", glu_C7_dilution50)
)
glu_C7_mean_lum50
glu_C7_mean_conc50
```

```{r add C7 df 100 glucose to standard curve, include=TRUE}
# Create sample data frame
# Calculate glucose concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glu_C7_mean_lum100 <- numeric(length(glu_C7_dilution100))
glu_C7_se_lum100 <- numeric(length(glu_C7_dilution100))
glu_C7_mean_conc100 <- numeric(length(glu_C7_dilution100))

for (i in 1:length(glu_C7_dilution100)) {
  df_val <- glu_C7_dilution100[i]
  # Get luminescence values for this dilution factor from both rows
   glu_C7_lum_values100 <- c(glu_C7_luminescence100[glu_C7_dilution100 == df_val])
  glu_C7_mean_lum100[i] <- mean( glu_C7_lum_values100)
  glu_C7_se_lum100[i] <- sd( glu_C7_lum_values100) / sqrt(length( glu_C7_lum_values100))
  # Calculate concentration from mean luminescence
  glu_C7_mean_conc100[i] <- (glu_C7_mean_lum100[i] - glu_intercept) / glu_slope
}

glu_C7_data100 <- data.frame(
  glu_C7_dilution100_factor = glu_C7_dilution100,
  glu_C7_mean_luminescence100 = glu_C7_mean_lum100,
  glu_C7_se100 = glu_C7_se_lum100,
  glu_C7_conc100 =  glu_C7_mean_conc100,
  label = paste0("C7df.", glu_C7_dilution100)
)
glu_C7_mean_lum100
glu_C7_mean_conc100
```

```{r add D2 glucose to standard curve, include=TRUE}
# Create sample data frame
# Calculate glucose concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glu_D2_mean_lum <- numeric(length(glu_D2_dilution))
glu_D2_se_lum <- numeric(length(glu_D2_dilution))
glu_D2_mean_conc <- numeric(length(glu_D2_dilution))

for (i in 1:length(glu_D2_dilution)) {
  df_val <- glu_D2_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glu_D2_lum_values <- c(glu_D2_luminescence[glu_D2_dilution == df_val])
  glu_D2_mean_lum[i] <- mean(glu_D2_lum_values)
  glu_D2_se_lum[i] <- sd(glu_D2_lum_values) / sqrt(length(glu_D2_lum_values))
  # Calculate concentration from mean luminescence
  glu_D2_mean_conc[i] <- (glu_D2_mean_lum[i] - glu_intercept) / glu_slope
}

glu_D2_data <- data.frame(
  glu_D2_dilution_factor = glu_D2_dilution,
  glu_D2_mean_luminescence = glu_D2_mean_lum,
  glu_D2_se = glu_D2_se_lum,
  glu_D2_conc =  glu_D2_mean_conc,
  label = paste0("D2df.", glu_D2_dilution)
)
glu_D2_mean_lum
glu_D2_mean_conc
```


```{r add E4 glucose to standard curve, include=TRUE}
# Create sample data frame
# Calculate glucose concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glu_E4_mean_lum <- numeric(length(glu_E4_dilution))
glu_E4_se_lum <- numeric(length(glu_E4_dilution))
glu_E4_mean_conc <- numeric(length(glu_E4_dilution))

for (i in 1:length(glu_E4_dilution)) {
  df_val <- glu_E4_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glu_E4_lum_values <- c(glu_E4_luminescence[glu_E4_dilution == df_val])
  glu_E4_mean_lum[i] <- mean(glu_E4_lum_values)
  glu_E4_se_lum[i] <- sd(glu_E4_lum_values) / sqrt(length(glu_E4_lum_values))
  # Calculate concentration from mean luminescence
  glu_E4_mean_conc[i] <- (glu_E4_mean_lum[i] - glu_intercept) / glu_slope
}

glu_E4_data <- data.frame(
  glu_E4_dilution_factor = glu_E4_dilution,
  glu_E4_mean_luminescence = glu_E4_mean_lum,
  glu_E4_se = glu_E4_se_lum,
  glu_E4_conc =  glu_E4_mean_conc,
  label = paste0("E4df.", glu_E4_dilution)
)
glu_E4_mean_lum
glu_E4_mean_conc
```

```{r add E5 glucose to standard curve, include=TRUE}
# Create sample data frame
# Calculate glucose concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glu_E5_mean_lum <- numeric(length(glu_E5_dilution))
glu_E5_se_lum <- numeric(length(glu_E5_dilution))
glu_E5_mean_conc <- numeric(length(glu_E5_dilution))

for (i in 1:length(glu_E5_dilution)) {
  df_val <- glu_E5_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glu_E5_lum_values <- c(glu_E5_luminescence[glu_E5_dilution == df_val])
  glu_E5_mean_lum[i] <- mean(glu_E5_lum_values)
  glu_E5_se_lum[i] <- sd(glu_E5_lum_values) / sqrt(length(glu_E5_lum_values))
  # Calculate concentration from mean luminescence
  glu_E5_mean_conc[i] <- (glu_E5_mean_lum[i] - glu_intercept) / glu_slope
}

glu_E5_data <- data.frame(
  glu_E5_dilution_factor = glu_E5_dilution,
  glu_E5_mean_luminescence = glu_E5_mean_lum,
  glu_E5_se = glu_E5_se_lum,
  glu_E5_conc =  glu_E5_mean_conc,
  label = paste0("E5df.", glu_E5_dilution)
)
glu_E5_mean_lum
glu_E5_mean_conc
```

```{r add E6 glucose to standard curve, include=TRUE}
# Create sample data frame
# Calculate glucose concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glu_E6_mean_lum <- numeric(length(glu_E6_dilution))
glu_E6_se_lum <- numeric(length(glu_E6_dilution))
glu_E6_mean_conc <- numeric(length(glu_E6_dilution))

for (i in 1:length(glu_E6_dilution)) {
  df_val <- glu_E6_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glu_E6_lum_values <- c(glu_E6_luminescence[glu_E6_dilution == df_val])
  glu_E6_mean_lum[i] <- mean(glu_E6_lum_values)
  glu_E6_se_lum[i] <- sd(glu_E6_lum_values) / sqrt(length(glu_E6_lum_values))
  # Calculate concentration from mean luminescence
  glu_E6_mean_conc[i] <- (glu_E6_mean_lum[i] - glu_intercept) / glu_slope
}

glu_E6_data <- data.frame(
  glu_E6_dilution_factor = glu_E6_dilution,
  glu_E6_mean_luminescence = glu_E6_mean_lum,
  glu_E6_se = glu_E6_se_lum,
  glu_E6_conc =  glu_E6_mean_conc,
  label = paste0("E6df.", glu_E6_dilution)
)
glu_E6_mean_lum
glu_E6_mean_conc
```

```{r add E7 glucose to standard curve, include=TRUE}
# Create sample data frame
# Calculate glucose concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glu_E7_mean_lum <- numeric(length(glu_E7_dilution))
glu_E7_se_lum <- numeric(length(glu_E7_dilution))
glu_E7_mean_conc <- numeric(length(glu_E7_dilution))

for (i in 1:length(glu_E7_dilution)) {
  df_val <- glu_E7_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glu_E7_lum_values <- c(glu_E7_luminescence[glu_E7_dilution == df_val])
  glu_E7_mean_lum[i] <- mean(glu_E7_lum_values)
  glu_E7_se_lum[i] <- sd(glu_E7_lum_values) / sqrt(length(glu_E7_lum_values))
  # Calculate concentration from mean luminescence
  glu_E7_mean_conc[i] <- (glu_E7_mean_lum[i] - glu_intercept) / glu_slope
}

glu_E7_data <- data.frame(
  glu_E7_dilution_factor = glu_E7_dilution,
  glu_E7_mean_luminescence = glu_E7_mean_lum,
  glu_E7_se = glu_E7_se_lum,
  glu_E7_conc =  glu_E7_mean_conc,
  label = paste0("E7df.", glu_E7_dilution)
)
glu_E7_mean_lum
glu_E7_mean_conc
```

### Plot glucose standard curve, sample points

```{r plot glucose standard curve with samples, include=TRUE}
# Create the plot
glu_plot <- ggplot(glucose_summary_data, aes(x = glu_concentration, y = glu_mean_luminescence)) +
  geom_smooth(aes(linetype = "Std Curve Best Fit Line"), method = "lm", se = FALSE, 
              color = "coral", linewidth = 1) +
  geom_errorbar(aes(ymin = glu_mean_luminescence - glu_se, ymax = glu_mean_luminescence + glu_se),
                width = 0.1, linewidth = 1, color = "darkblue") +
  geom_point(aes(color = "Standard Curve", shape = "Standard Curve"), size = 4) +

  geom_errorbar(data = glu_C7_data50, aes(x = glu_C7_conc50, y = glu_C7_mean_luminescence50,
                ymin = glu_C7_mean_luminescence50 - glu_C7_se50, ymax = glu_C7_mean_luminescence50 + glu_C7_se50,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glu_C7_data50, aes(x = glu_C7_conc50, y = glu_C7_mean_luminescence50,
             color = label, shape = label), size = 4) +

  geom_errorbar(data = glu_C7_data100, aes(x = glu_C7_conc100, y = glu_C7_mean_luminescence100,
                ymin = glu_C7_mean_luminescence100 - glu_C7_se100, ymax = glu_C7_mean_luminescence100 + glu_C7_se100,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glu_C7_data100, aes(x = glu_C7_conc100, y = glu_C7_mean_luminescence100,
             color = label, shape = label), size = 4) +
  
     geom_errorbar(data = glu_D2_data, aes(x = glu_D2_conc, y = glu_D2_mean_luminescence,
                ymin = glu_D2_mean_luminescence - glu_D2_se, ymax = glu_D2_mean_luminescence + glu_D2_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glu_D2_data, aes(x = glu_D2_conc, y = glu_D2_mean_luminescence,
             color = label, shape = label), size = 4) +

   geom_errorbar(data = glu_E4_data, aes(x = glu_E4_conc, y = glu_E4_mean_luminescence,
                ymin = glu_E4_mean_luminescence - glu_E4_se, ymax = glu_E4_mean_luminescence + glu_E4_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glu_E4_data, aes(x = glu_E4_conc, y = glu_E4_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glu_E5_data, aes(x = glu_E5_conc, y = glu_E5_mean_luminescence,
                ymin = glu_E5_mean_luminescence - glu_E5_se, ymax = glu_E5_mean_luminescence + glu_E5_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glu_E5_data, aes(x = glu_E5_conc, y = glu_E5_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glu_E6_data, aes(x = glu_E6_conc, y = glu_E6_mean_luminescence,
                ymin = glu_E6_mean_luminescence - glu_E6_se, ymax = glu_E6_mean_luminescence + glu_E6_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glu_E6_data, aes(x = glu_E6_conc, y = glu_E6_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glu_E7_data, aes(x = glu_E7_conc, y = glu_E7_mean_luminescence,
                ymin = glu_E7_mean_luminescence - glu_E7_se, ymax = glu_E7_mean_luminescence + glu_E7_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glu_E7_data, aes(x = glu_E7_conc, y = glu_E7_mean_luminescence,
             color = label, shape = label), size = 4) +

  
  
  scale_color_manual(name = "",
                     values = c("Standard Curve" = "steelblue",
                                "C7df.50" = "darkred",
                                "C7df.100" = "darkorange",
                                "D2df.20" = "cyan",
                                "E4df.20" = "darkgreen",
                                "E5df.20" = "purple",
                                "E6df.20" = "brown",
                                "E7df.20" = "green3"
                                )) +
  scale_shape_manual(name = "",
                     values = c("Standard Curve" = 16,
                                 "C7df.50" = 17,
                                 "C7df.100" = 15,
                                 "D2df.20" = 2,
                                 "E4df.20" = 18,
                                 "E5df.20" = 8,
                                 "E6df.20" = 4,
                                 "E7df.20" = 5
                                )) +
 scale_linetype_manual(name = "",
                        values = c("Std Curve Best Fit Line" = "dashed")) +
  annotate("label", x = (max(glucose_summary_data$glu_concentration) * 0.75), 
           y = (max(glucose_summary_data$glu_mean_luminescence) * 0.15),
           label = sprintf("y = %.2fx + %.2f\nR² = %.4f", glu_slope, glu_intercept, glu_r_squared),
           size = 3.5, fontface = "bold", fill = "white", 
           color = "coral", label.padding = unit(0.3, "lines")) +
  labs(
    title = "glucose Standard Curve",
    x = "glucose Concentration (µg/µL)",
    y = "Luminescence",
    caption = "Error bars represent standard error of the mean (SEM)"
  ) +
  scale_x_continuous(breaks = glu_concentrations) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_line(linetype = "dashed", color = "grey70")
  )


# Display the plot
glu_plot
glu_C7_data50
glu_C7_data100
glu_D2_data
glu_E4_data
glu_E5_data
glu_E6_data
glu_E7_data

# Print summary statistics
cat("glucose Standard Curve Summary:\n")
cat(rep("=", 50), "\n", sep = "")
for (i in 1:nrow(glucose_summary_data)) {
  cat(sprintf("Concentration: %g µg/µL\n", glucose_summary_data$glu_concentration[i]))
  cat(sprintf("  Mean Luminescence: %.2f\n",glucose_summary_data$glu_mean_luminescence[i]))
  cat(sprintf("  Standard Error: %.2f\n", glucose_summary_data$glu_se[i]))
  cat(sprintf("  CV%%: %.2f%%\n\n", glucose_summary_data$glu_cv[i]))
}
```

### Glucose summary table

```{r, sample glucose calculations}
tab <- matrix(c(glu_C7_dilution50, glu_C7_mean_lum50, glu_C7_mean_conc50,  (glu_C7_dilution50*glu_C7_mean_conc50), 
                glu_C7_dilution100, glu_C7_mean_lum100, glu_C7_mean_conc100,  (glu_C7_dilution100*glu_C7_mean_conc100), 
                glu_D2_dilution, glu_D2_mean_lum, glu_D2_mean_conc,  (glu_D2_dilution*glu_D2_mean_conc), 
                glu_E4_dilution, glu_E4_mean_lum, glu_E4_mean_conc,  (glu_E4_dilution*glu_E4_mean_conc), 
                glu_E5_dilution, glu_E5_mean_lum, glu_E5_mean_conc,  (glu_E5_dilution*glu_E5_mean_conc), 
                glu_E6_dilution, glu_E6_mean_lum, glu_E6_mean_conc,  (glu_E6_dilution*glu_E6_mean_conc), 
                glu_E7_dilution, glu_E7_mean_lum, glu_E7_mean_conc,  (glu_E7_dilution*glu_E7_mean_conc)), ncol=4, byrow=TRUE)
colnames(tab) <- c('Dilution factor','Luminescence','Calculated glucose (ug/uL)', 'Total glucose (ug/uL)')
rownames(tab) <- c('C7_df50','C7_df100','D2','E4','E5','E6','E7')
tab <- as.table(tab)
tab
```
# Standard curves for 20260330B - sample E5 re-run

## Glycogen Standard Curve

### Extract glycogen luminescence data

```{r glycogen standard data E5 re-run, include=TRUE}
# Extract glycogen standard curve data from plate layout and raw luminescence.
# Rows F, G, H (rows 6, 7, 8) correspond to glycogen standards
# Columns 1-5 contain the standard curve concentrations

# Extract concentration values from plate layout (row 6, columns 1-5)
# Parse concentration from labels like "STD-glyc-20" -> 20
glyc_concentrations2 <- as.numeric(gsub("STD-glyc-", "", plate_layout2[6, 1:5]))
glyc_concentrations2

# Extract standard curve luminescence values for each replicate from raw_luminescence
glyc_row_F2 <- as.numeric(raw_luminescence2[6, 1:5])  # Row 6 (F)
glyc_row_G2 <- as.numeric(raw_luminescence2[7, 1:5])  # Row 7 (G)
glyc_row_H2 <- as.numeric(raw_luminescence2[8, 1:5])  # Row 8 (H)
```
```{r glycogen sample E5 re-run, include=TRUE}
#Extract glycogen sample data
glyc_sample_cols1 <- c(1,2,3)

glyc_E5_dilution2 <- as.numeric(gsub(".*-df\\.", "", plate_layout3[6, 1]))
glyc_E5_luminescence2 <- as.numeric(raw_luminescence3[6, glyc_sample_cols1])
glyc_E5_dilution2
glyc_E5_luminescence2
```

### Glycogen standard curve summary statistics and linear regression

```{r glycogen summary statistics and linear regression E5 re-run,include=TRUE}
# Calculate mean and standard error for each concentration
glyc_means2 <- numeric(length(glyc_concentrations2))
glyc_std_errors2 <- numeric(length(glyc_concentrations2))
glyc_std_devs2 <- numeric(length(glyc_concentrations2))

for (i in 1:length(glyc_concentrations2)) {
  glyc_values2 <- c(glyc_row_F2[i], glyc_row_G2[i], glyc_row_H2[i])
  glyc_means2[i] <- mean(glyc_values2)
  glyc_std_devs2[i] <- sd(glyc_values2)
  glyc_std_errors2[i] <- sd(glyc_values2) / sqrt(length(glyc_values2))
}

# Create data frame for plotting
glycogen_summary_data2 <- data.frame(
  glyc_concentration2 = glyc_concentrations2,
  glyc_mean_luminescence2 = glyc_means2,
  glyc_se2 = glyc_std_errors2,
  glyc_sd2 = glyc_std_devs2,
  glyc_cv2 = (glyc_std_devs2 / glyc_means2) * 100
)

# Calculate linear regression and R-squared
lm_model2 <- lm(glyc_mean_luminescence2 ~ glyc_concentration2, data = glycogen_summary_data2)
glyc_r_squared2 <- summary(lm_model2)$r.squared
glyc_slope2 <- coef(lm_model2)[2]
glyc_intercept2 <- coef(lm_model2)[1]

glyc_r_squared2
glyc_slope2
glyc_intercept2
```

### Calculate sample glycogen levels

```{r add E5 re-run glycogen to standard curve, include=TRUE}
# Create sample data frame
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_E5_mean_lum2 <- numeric(length(glyc_E5_dilution2))
glyc_E5_se_lum2 <- numeric(length(glyc_E5_dilution2))
glyc_E5_mean_conc2 <- numeric(length(glyc_E5_dilution2))

for (i in 1:length(glyc_E5_dilution2)) {
  df_val <- glyc_E5_dilution2[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_E5_lum2_values <- c(glyc_E5_luminescence2[glyc_E5_dilution2 == df_val])
  glyc_E5_mean_lum2[i] <- mean(glyc_E5_lum2_values)
  glyc_E5_se_lum2[i] <- sd(glyc_E5_lum2_values) / sqrt(length(glyc_E5_lum2_values))
  # Calculate concentration from mean luminescence
  glyc_E5_mean_conc2[i] <- (glyc_E5_mean_lum2[i] - glyc_intercept2) / glyc_slope2
}

glyc_E5_data2 <- data.frame(
  glyc_E5_dilution2_factor = glyc_E5_dilution2,
  glyc_E5_mean_luminescence2 = glyc_E5_mean_lum2,
  glyc_E5_se2 = glyc_E5_se_lum2,
  glyc_E5_conc2 =  glyc_E5_mean_conc2,
  label = paste0("E5df.", glyc_E5_dilution2)
)
glyc_E5_mean_lum2
glyc_E5_mean_conc2
```
### Plot glycogen standard curve, sample points

```{r plot glycogen standard curve from 20260330 with E5 re-run, include=TRUE}
# Create the plot2
glyc_plot2 <- ggplot(glycogen_summary_data2, aes(x = glyc_concentration2, y = glyc_mean_luminescence2)) +
  geom_smooth(aes(linetype = "Std Curve Best Fit Line"), method = "lm", se = FALSE, 
              color = "coral", linewidth = 1) +
  geom_errorbar(aes(ymin = glyc_mean_luminescence2 - glyc_se2, ymax = glyc_mean_luminescence2 + glyc_se2),
                width = 0.1, linewidth = 1, color = "darkblue") +
  geom_point(aes(color = "Standard Curve", shape = "Standard Curve"), size = 4) +

  geom_errorbar(data = glyc_E5_data2, aes(x = glyc_E5_conc2, y = glyc_E5_mean_luminescence2,
                ymin = glyc_E5_mean_luminescence2 - glyc_E5_se2, ymax = glyc_E5_mean_luminescence2 + glyc_E5_se2,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_E5_data2, aes(x = glyc_E5_conc2, y = glyc_E5_mean_luminescence2,
             color = label, shape = label), size = 4) +
  
  scale_color_manual(name = "",
                     values = c("Standard Curve" = "steelblue",
                                "E5df.50" = "darkred"
                                )) +
  scale_shape_manual(name = "",
                     values = c("Standard Curve" = 16,
                                 "E5df.50" = 17
                                )) +
 scale_linetype_manual(name = "",
                        values = c("Std Curve Best Fit Line" = "dashed")) +
  annotate("label", x = (max(glycogen_summary_data2$glyc_concentration2) * 0.85), 
           y = (max(glycogen_summary_data2$glyc_mean_luminescence2) * 0.15),
           label = sprintf("y = %.2fx + %.2f\nR² = %.4f", glyc_slope2, glyc_intercept2, glyc_r_squared2),
           size = 3.5, fontface = "bold", fill = "white", 
           color = "coral", label.padding = unit(0.3, "lines")) +
  labs(
    title = "Glycogen Standard Curve E5 re-run",
    x = "Glycogen Concentration (µg/µL)",
    y = "Luminescence",
    caption = "Error bars represent standard error of the mean (SEM)"
  ) +
  scale_x_continuous(breaks = glyc_concentrations2) +
  theme_bw() +
  theme(
    plot2.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_line(linetype = "dashed", color = "grey70")
  )


# Display the plot2
glyc_plot2
glyc_E5_data2


# Print summary statistics
cat("Glycogen Standard Curve Summary:\n")
cat(rep("=", 50), "\n", sep = "")
for (i in 1:nrow(glycogen_summary_data2)) {
  cat(sprintf("Concentration: %g µg/µL\n", glycogen_summary_data2$glyc_concentration2[i]))
  cat(sprintf("  Mean Luminescence: %.2f\n",glycogen_summary_data2$glyc_mean_luminescence2[i]))
  cat(sprintf("  Standard Error: %.2f\n", glycogen_summary_data2$glyc_se2[i]))
  cat(sprintf("  CV%%: %.2f%%\n\n", glycogen_summary_data2$glyc_cv2[i]))
}
```

## Glucose Standard Curve

### Extract glucose luminescence data

```{r glucose standard data E5 re-run, include=TRUE}
# Extract glucose standard curve data from plate layout and raw luminescence.
# Rows F, G, H (rows 6, 7, 8) correspond to glucose standards
# Columns 7-11 contain the standard curve concentrations

# Extract concentration values from plate layout (row 6, columns 7-11)
# Parse concentration from labels like "STD-glyc-20" -> 20
glu_concentrations2 <- as.numeric(gsub("STD-glu-", "", plate_layout2[6, 7:11]))
glu_concentrations2

# Extract standard curve luminescence values for each replicate from raw_luminescence
glu_row_F2 <- as.numeric(raw_luminescence2[6, 7:11])  # Row 6 (F)
glu_row_G2 <- as.numeric(raw_luminescence2[7, 7:11])  # Row 7 (G)
glu_row_H2 <- as.numeric(raw_luminescence2[8, 7:11])  # Row 8 (H)
```

```{r glucose sample E5 re-run, include=TRUE}
glu_sample_cols1 <- c(7,8,9)

glu_E5_dilution2 <- as.numeric(gsub(".*-df\\.", "", plate_layout3[6, 7]))
glu_E5_luminescence2 <- as.numeric(raw_luminescence3[6, glu_sample_cols1])
glu_E5_dilution2
glu_E5_luminescence2
```

### Glucose standard curve summary statistics and linear regression

```{r glucose summary statistics and linear regression E5 re-run,include=TRUE}
# Calculate mean and standard error for each concentration
glu_means2 <- numeric(length(glu_concentrations2))
glu_std_errors2 <- numeric(length(glu_concentrations2))
glu_std_devs2 <- numeric(length(glu_concentrations2))

for (i in 1:length(glu_concentrations2)) {
  glu_values2 <- c(glu_row_F2[i], glu_row_G2[i], glu_row_H2[i])
  glu_means2[i] <- mean(glu_values2)
  glu_std_devs2[i] <- sd(glu_values2)
  glu_std_errors2[i] <- sd(glu_values2) / sqrt(length(glu_values2))
}

# Create data frame for plotting
glucose_summary_data2 <- data.frame(
  glu_concentration2 = glu_concentrations2,
  glu_mean_luminescence2 = glu_means2,
  glu_se2 = glu_std_errors2,
  glu_sd2 = glu_std_devs2,
  glu_cv2 = (glu_std_devs2 / glu_means2) * 100
)

# Calculate linear regression and R-squared
lm_model2 <- lm(glu_mean_luminescence2 ~ glu_concentration2, data = glucose_summary_data2)
glu_r_squared2 <- summary(lm_model2)$r.squared
glu_slope2 <- coef(lm_model2)[2]
glu_intercept2 <- coef(lm_model2)[1]

glu_r_squared2
glu_slope2
glu_intercept2
```

### Calculate sample glucose levels

```{r add E5 re-run glucose to standard curve, include=TRUE}
# Create sample data frame
# Calculate glucose concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glu_E5_mean_lum2 <- numeric(length(glu_E5_dilution2))
glu_E5_se_lum2 <- numeric(length(glu_E5_dilution2))
glu_E5_mean_conc2 <- numeric(length(glu_E5_dilution2))

for (i in 1:length(glu_E5_dilution2)) {
  df_val <- glu_E5_dilution2[i]
  # Get luminescence values for this dilution factor from both rows
  glu_E5_lum2_values <- c(glu_E5_luminescence2[glu_E5_dilution2 == df_val])
  glu_E5_mean_lum2[i] <- mean(glu_E5_lum2_values)
  glu_E5_se_lum2[i] <- sd(glu_E5_lum2_values) / sqrt(length(glu_E5_lum2_values))
  # Calculate concentration from mean luminescence
  glu_E5_mean_conc2[i] <- (glu_E5_mean_lum2[i] - glu_intercept2) / glu_slope2
}

glu_E5_data2 <- data.frame(
  glu_E5_dilution2_factor = glu_E5_dilution2,
  glu_E5_mean_luminescence2 = glu_E5_mean_lum2,
  glu_E5_se2 = glu_E5_se_lum2,
  glu_E5_conc2 =  glu_E5_mean_conc2,
  label = paste0("E5df.", glu_E5_dilution2)
)
glu_E5_mean_lum2
glu_E5_mean_conc2
```

### Plot glucose standard curve, sample points

```{r plot glucose standard curve from 20260330 with E5 re-run, include=TRUE}
# Create the plot
glu_plot2 <- ggplot(glucose_summary_data2, aes(x = glu_concentration2, y = glu_mean_luminescence2)) +
  geom_smooth(aes(linetype = "Std Curve Best Fit Line"), method = "lm", se = FALSE, 
              color = "coral", linewidth = 1) +
  geom_errorbar(aes(ymin = glu_mean_luminescence2 - glu_se2, ymax = glu_mean_luminescence2 + glu_se2),
                width = 0.1, linewidth = 1, color = "darkblue") +
  geom_point(aes(color = "Standard Curve", shape = "Standard Curve"), size = 4) +

  geom_errorbar(data = glu_E5_data2, aes(x = glu_E5_conc2, y = glu_E5_mean_luminescence2,
                ymin = glu_E5_mean_luminescence2 - glu_E5_se2, ymax = glu_E5_mean_luminescence2 + glu_E5_se2,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glu_E5_data2, aes(x = glu_E5_conc2, y = glu_E5_mean_luminescence2,
             color = label, shape = label), size = 4) +
  
  scale_color_manual(name = "",
                     values = c("Standard Curve" = "steelblue",
                                "E5df.50" = "darkred"
                                )) +
  scale_shape_manual(name = "",
                     values = c("Standard Curve" = 16,
                                 "E5df.50" = 17
                                )) +
 scale_linetype_manual(name = "",
                        values = c("Std Curve Best Fit Line" = "dashed")) +
  annotate("label", x = (max(glucose_summary_data2$glu_concentration2) * 0.85), 
           y = (max(glucose_summary_data2$glu_mean_luminescence2) * 0.15),
           label = sprintf("y = %.2fx + %.2f\nR² = %.4f", glu_slope2, glu_intercept2, glu_r_squared2),
           size = 3.5, fontface = "bold", fill = "white", 
           color = "coral", label.padding = unit(0.3, "lines")) +
  labs(
    title = "Glucose Standard Curve E5 re-run",
    x = "Glucose Concentration (µg/µL)",
    y = "Luminescence",
    caption = "Error bars represent standard error of the mean (SEM)"
  ) +
  scale_x_continuous(breaks = glu_concentrations2) +
  theme_bw() +
  theme(
    plot2.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_line(linetype = "dashed", color = "grey70")
  )


# Display the plot2
glu_plot2
glu_E5_data2


# Print summary statistics
cat("Glucose Standard Curve Summary:\n")
cat(rep("=", 50), "\n", sep = "")
for (i in 1:nrow(glucose_summary_data2)) {
  cat(sprintf("Concentration: %g µg/µL\n", glucose_summary_data2$glu_concentration2[i]))
  cat(sprintf("  Mean Luminescence: %.2f\n",glucose_summary_data2$glu_mean_luminescence2[i]))
  cat(sprintf("  Standard Error: %.2f\n", glucose_summary_data2$glu_se2[i]))
  cat(sprintf("  CV%%: %.2f%%\n\n", glucose_summary_data2$glu_cv2[i]))
}
```


# Overall summary table

Notes: the ‘glycogen’ readout includes background glucose - the ‘Glycogen-Glucose’ number is a more accurate measurement of glycogen levels, as this removes the contribution of background glucose

Sample E5 was out of the standard curve range on 20260218, so was re-run with further dilutions on 20260330, (alongside samples from a separate temp stress experiment) - this data is included as 'E5 re-run'.

```{r, Overall summary table}
tab <- matrix(c(as.numeric(weights[23,2]), glyc_C7_dilution50, glyc_C7_mean_conc50, glu_C7_mean_conc50, (glyc_C7_mean_conc50-glu_C7_mean_conc50), ((glyc_C7_mean_conc50-glu_C7_mean_conc50)*(glyc_C7_dilution50/as.numeric(weights[23,2]))), 
                as.numeric(weights[23,2]), glyc_C7_dilution100, glyc_C7_mean_conc100, glu_C7_mean_conc100, (glyc_C7_mean_conc100-glu_C7_mean_conc100), ((glyc_C7_mean_conc100-glu_C7_mean_conc100)*(glyc_C7_dilution100/as.numeric(weights[23,2]))),
                as.numeric(weights[26,2]), glyc_D2_dilution, glyc_D2_mean_conc, glu_D2_mean_conc, (glyc_D2_mean_conc-glu_D2_mean_conc), ((glyc_D2_mean_conc-glu_D2_mean_conc)*(glyc_D2_dilution/as.numeric(weights[26,2]))), 
                as.numeric(weights[36,2]), glyc_E4_dilution, glyc_E4_mean_conc, glu_E4_mean_conc, (glyc_E4_mean_conc-glu_E4_mean_conc), ((glyc_E4_mean_conc-glu_E4_mean_conc)*(glyc_E4_dilution/as.numeric(weights[36,2]))),
                as.numeric(weights[37,2]), glyc_E5_dilution, glyc_E5_mean_conc, glu_E5_mean_conc, (glyc_E5_mean_conc-glu_E5_mean_conc), ((glyc_E5_mean_conc-glu_E5_mean_conc)*(glyc_E5_dilution/as.numeric(weights[37,2]))),
                as.numeric(weights[37,2]), glyc_E5_dilution2, glyc_E5_mean_conc2, glu_E5_mean_conc2, (glyc_E5_mean_conc2-glu_E5_mean_conc2), ((glyc_E5_mean_conc2-glu_E5_mean_conc2)*(glyc_E5_dilution2/as.numeric(weights[37,2]))),
                as.numeric(weights[38,2]), glyc_E6_dilution, glyc_E6_mean_conc, glu_E6_mean_conc, (glyc_E6_mean_conc-glu_E6_mean_conc), ((glyc_E6_mean_conc-glu_E6_mean_conc)*(glyc_E6_dilution/as.numeric(weights[38,2]))), 
                as.numeric(weights[39,2]), glyc_E7_dilution, glyc_E7_mean_conc, glu_E7_mean_conc, (glyc_E7_mean_conc-glu_E7_mean_conc), ((glyc_E7_mean_conc-glu_E7_mean_conc)*(glyc_E7_dilution/as.numeric(weights[39,2])))
                           ), ncol=6, byrow=TRUE)

colnames(tab) <- c('Weight (mg)', 'df', 'Glycogen (ug/uL)', 'Glucose (uM)', 'Glycogen-Glucose (ug/uL)', '(Glycogen-Glucose)*df/weight (ug/uL/mg)')
rownames(tab) <- c('C7.df50','C7.df100','D2','E4','E5','E5 re-run', 'E6','E7')
tab <- as.table(tab)
tab
```

