---
title: "Gen5-20260629AB-mgig-glycogenglo"
author: "Hazel Abrahamson-Amerine"
date: "2026-07-07"
output:  
    github_document:
      toc: true
      number_sections: true
---

```{r setup, include=TRUE}
library(knitr)
library(ggplot2)
library(dplyr)
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = "",         # Prevents appending '##' to beginning of lines in code output
  results = 'hold'     # Holds output so it's all printed together after code chunk
)
```

#Load plate reader, luminescence, and sample weight data
```{r data, include=TRUE}
#load data
#plate 1 - samples 48.0A-48.2F
plate_layout1 <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/layout-Gen5-20260629A-mgig-glycogenglo.csv", header = FALSE)
raw_luminescence1 <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/raw_lum-Gen5-20260629A-mgig-glycogenglo.csv", header = FALSE)

cat("Plate layout1:\n")
str(plate_layout1)

cat("\n\n")

cat("Raw luminescence1:\n")
str(raw_luminescence1)

#plate 2 - samples 18.0A-24.0C, 0.0A, 0.0B
plate_layout2 <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/layout-Gen5-20260629B-mgig-glycogenglo.csv", header = FALSE)
raw_luminescence2 <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/raw_luminescence/raw_lum-Gen5-20260629B-mgig-glycogenglo.csv", header = FALSE)

cat("Plate layout2:\n")
str(plate_layout2)

cat("\n\n")

cat("Raw luminescence2:\n")
str(raw_luminescence2)

#sample weights
sample_weights <- read.csv("https://raw.githubusercontent.com/RobertsLab/sormi-assay-development/refs/heads/main/Glycogen/data/20260625B_mgig_wholetissue_weights_GlycogenGlo.csv", header = FALSE)

cat("Sample Weights:\n")
str(sample_weights)
```

# STANDARD CURVES

## Glycogen Standard Curve

### Extract luminescence data

```{r glycogen standard data, include=TRUE}
# Extract glycogen standard curve data from plate layout and raw luminescence.
# Rows F, G, H (rows 6, 7, 8) correspond to glycogen standards
# Columns 1-5 contain the standard curve concentrations

# Extract concentration values from plate layout (row 6, columns 1-5)
# Parse concentration from labels like "STD-glyc-20" -> 20
glyc_concentrations <- as.numeric(gsub("STD-glyc-", "", plate_layout1[6, 1:5]))
glyc_concentrations

# Extract standard curve luminescence values for each replicate from raw_luminescence
glyc_row_F <- as.numeric(raw_luminescence1[6, 1:5])  # Row 6 (F)
glyc_row_G <- as.numeric(raw_luminescence1[7, 1:5])  # Row 7 (G)
glyc_row_H <- as.numeric(raw_luminescence1[8, 1:5])  # Row 8 (H)
```

```{r glycogen samples, include=TRUE}
#Extract glycogen sample data
#plate 1
glyc_sample_cols1 <- c(1,2,3)
glyc_0.0A_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[1, 1]))
glyc_0.0A_luminescence <- as.numeric(raw_luminescence1[1, glyc_sample_cols1])

glyc_0.0B_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[2, 1]))
glyc_0.0B_luminescence <- as.numeric(raw_luminescence1[2, glyc_sample_cols1])

glyc_0.0C_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[3, 1]))
glyc_0.0C_luminescence <- as.numeric(raw_luminescence1[3, glyc_sample_cols1])

glyc_0.0D_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[4, 1]))
glyc_0.0D_luminescence <- as.numeric(raw_luminescence1[4, glyc_sample_cols1])

glyc_0.0E_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[5, 1]))
glyc_0.0E_luminescence <- as.numeric(raw_luminescence1[5, glyc_sample_cols1])

glyc_sample_cols2 <- c(4,5,6)
glyc_0.0F_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[1, 4]))
glyc_0.0F_luminescence <- as.numeric(raw_luminescence1[1, glyc_sample_cols2])

glyc_0.0G_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[2, 4]))
glyc_0.0G_luminescence <- as.numeric(raw_luminescence1[2, glyc_sample_cols2])

glyc_0.0H_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[3, 4]))
glyc_0.0H_luminescence <- as.numeric(raw_luminescence1[3, glyc_sample_cols2])

glyc_0.0I_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[4, 4]))
glyc_0.0I_luminescence <- as.numeric(raw_luminescence1[4, glyc_sample_cols2])

glyc_0.0J_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[5, 4]))
glyc_0.0J_luminescence <- as.numeric(raw_luminescence1[5, glyc_sample_cols2])

glyc_sample_cols3 <- c(7,8,9)
glyc_48.0A_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[1, 7]))
glyc_48.0A_luminescence <- as.numeric(raw_luminescence1[1, glyc_sample_cols3]) 

glyc_48.0B_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[2, 7]))
glyc_48.0B_luminescence <- as.numeric(raw_luminescence1[2, glyc_sample_cols3]) 

glyc_48.0C_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[3, 7]))
glyc_48.0C_luminescence <- as.numeric(raw_luminescence1[3, glyc_sample_cols3]) 

glyc_48.0D_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[4, 7]))
glyc_48.0D_luminescence <- as.numeric(raw_luminescence1[4, glyc_sample_cols3]) 

glyc_48.0E_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[5, 7]))
glyc_48.0E_luminescence <- as.numeric(raw_luminescence1[5, glyc_sample_cols3])

glyc_48.0F_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[6, 7]))
glyc_48.0F_luminescence <- as.numeric(raw_luminescence1[6, glyc_sample_cols3]) 

glyc_48.0G_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[7, 7]))
glyc_48.0G_luminescence <- as.numeric(raw_luminescence1[7, glyc_sample_cols3])

glyc_48.0H_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[8, 7]))
glyc_48.0H_luminescence <- as.numeric(raw_luminescence1[8, glyc_sample_cols3]) 

glyc_sample_cols4 <- c(10,11,12)
glyc_48.0I_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[1, 10]))
glyc_48.0I_luminescence <- as.numeric(raw_luminescence1[1, glyc_sample_cols4]) 

glyc_48.0J_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[2, 10]))
glyc_48.0J_luminescence <- as.numeric(raw_luminescence1[2, glyc_sample_cols4]) 

glyc_48.2A_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[3, 10]))
glyc_48.2A_luminescence <- as.numeric(raw_luminescence1[3, glyc_sample_cols4])

glyc_48.2B_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[4, 10]))
glyc_48.2B_luminescence <- as.numeric(raw_luminescence1[4, glyc_sample_cols4]) 

glyc_48.2C_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[5, 10]))
glyc_48.2C_luminescence <- as.numeric(raw_luminescence1[5, glyc_sample_cols4]) 

glyc_48.2D_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[6, 10]))
glyc_48.2D_luminescence <- as.numeric(raw_luminescence1[6, glyc_sample_cols4]) 

glyc_48.2E_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[7, 10]))
glyc_48.2E_luminescence <- as.numeric(raw_luminescence1[7, glyc_sample_cols4]) 

glyc_48.2F_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout1[8, 10]))
glyc_48.2F_luminescence <- as.numeric(raw_luminescence1[8, glyc_sample_cols4]) 

#plate2
glyc_sample_cols1 <- c(1,2,3)
glyc_48.2G_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[1, 1]))
glyc_48.2G_luminescence <- as.numeric(raw_luminescence2[1, glyc_sample_cols1]) 

glyc_48.2H_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[2, 1]))
glyc_48.2H_luminescence <- as.numeric(raw_luminescence2[2, glyc_sample_cols1]) 

glyc_48.2I_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[3, 1]))
glyc_48.2I_luminescence <- as.numeric(raw_luminescence2[3, glyc_sample_cols1])

glyc_48.2J_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[4, 1]))
glyc_48.2J_luminescence <- as.numeric(raw_luminescence2[4, glyc_sample_cols1])

glyc_48.7A_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[5, 1]))
glyc_48.7A_luminescence <- as.numeric(raw_luminescence2[5, glyc_sample_cols1]) 

glyc_48.7B_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[6, 1]))
glyc_48.7B_luminescence <- as.numeric(raw_luminescence2[6, glyc_sample_cols1]) 

glyc_48.7C_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[7, 1]))
glyc_48.7C_luminescence <- as.numeric(raw_luminescence2[7, glyc_sample_cols1]) 

glyc_48.7D_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[8, 1]))
glyc_48.7D_luminescence <- as.numeric(raw_luminescence2[8, glyc_sample_cols1]) 

glyc_sample_cols2 <- c(4,5,6)
glyc_48.7E_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[1, 4]))
glyc_48.7E_luminescence <- as.numeric(raw_luminescence2[1, glyc_sample_cols2]) 

glyc_48.7F_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[2, 4]))
glyc_48.7F_luminescence <- as.numeric(raw_luminescence2[2, glyc_sample_cols2]) 

glyc_48.7G_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[3, 4]))
glyc_48.7G_luminescence <- as.numeric(raw_luminescence2[3, glyc_sample_cols2]) 

glyc_48.7H_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[4, 4]))
glyc_48.7H_luminescence <- as.numeric(raw_luminescence2[4, glyc_sample_cols2]) 

glyc_48.7I_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[5, 4]))
glyc_48.7I_luminescence <- as.numeric(raw_luminescence2[5, glyc_sample_cols2]) 

glyc_48.7J_dilution <- as.numeric(gsub(".*-df\\.", "", plate_layout2[6, 4]))
glyc_48.7J_luminescence <- as.numeric(raw_luminescence2[6, glyc_sample_cols2]) 

```

### Glycogen standard curve summary statistics and linear regression

```{r glycogen summary statistics and linear regression,include=TRUE}
# Calculate mean and standard error for each concentration
glyc_means <- numeric(length(glyc_concentrations))
glyc_std_errors <- numeric(length(glyc_concentrations))
glyc_std_devs <- numeric(length(glyc_concentrations))

for (i in 1:length(glyc_concentrations)) {
  glyc_values <- c(glyc_row_F[i], glyc_row_G[i], glyc_row_H[i])
  glyc_means[i] <- mean(glyc_values)
  glyc_std_devs[i] <- sd(glyc_values)
  glyc_std_errors[i] <- sd(glyc_values) / sqrt(length(glyc_values))
}

# Create data frame for plotting
glycogen_summary_data <- data.frame(
  glyc_concentration = glyc_concentrations,
  glyc_mean_luminescence = glyc_means,
  glyc_se = glyc_std_errors,
  glyc_sd = glyc_std_devs,
  glyc_cv = (glyc_std_devs / glyc_means) * 100
)

# Calculate linear regression and R-squared
lm_model <- lm(glyc_mean_luminescence ~ glyc_concentration, data = glycogen_summary_data)
glyc_r_squared <- summary(lm_model)$r.squared
glyc_slope <- coef(lm_model)[2]
glyc_intercept <- coef(lm_model)[1]
```

### Calculate sample glycogen levels

```{r add 0.0A** glycogen to standard curve, include=TRUE}
# Create sample data frame for 0.0A**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_0.0A_mean_lum <- numeric(length(glyc_0.0A_dilution))
glyc_0.0A_se_lum <- numeric(length(glyc_0.0A_dilution))
glyc_0.0A_mean_conc <- numeric(length(glyc_0.0A_dilution))

for (i in 1:length(glyc_0.0A_dilution)) {
  df_val <- glyc_0.0A_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_0.0A_lum_values <- c(glyc_0.0A_luminescence[glyc_0.0A_dilution == df_val])
  glyc_0.0A_mean_lum[i] <- mean(glyc_0.0A_lum_values)
  glyc_0.0A_se_lum[i] <- sd(glyc_0.0A_lum_values) / sqrt(length(glyc_0.0A_lum_values))
  # Calculate concentration from mean luminescence
  glyc_0.0A_mean_conc[i] <- (glyc_0.0A_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_0.0A_data <- data.frame(
  glyc_0.0A_dilution_factor = glyc_0.0A_dilution,
  glyc_0.0A_mean_luminescence = glyc_0.0A_mean_lum,
  glyc_0.0A_se = glyc_0.0A_se_lum,
  glyc_0.0A_conc =  glyc_0.0A_mean_conc,
  label = paste0("0.0Adf.", glyc_0.0A_dilution)
)
glyc_0.0A_mean_lum
glyc_0.0A_mean_conc
```

```{r add 0.0B** glycogen to standard curve, include=TRUE}
# Create sample data frame for 0.0B**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_0.0B_mean_lum <- numeric(length(glyc_0.0B_dilution))
glyc_0.0B_se_lum <- numeric(length(glyc_0.0B_dilution))
glyc_0.0B_mean_conc <- numeric(length(glyc_0.0B_dilution))

for (i in 1:length(glyc_0.0B_dilution)) {
  df_val <- glyc_0.0B_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_0.0B_lum_values <- c(glyc_0.0B_luminescence[glyc_0.0B_dilution == df_val])
  glyc_0.0B_mean_lum[i] <- mean(glyc_0.0B_lum_values)
  glyc_0.0B_se_lum[i] <- sd(glyc_0.0B_lum_values) / sqrt(length(glyc_0.0B_lum_values))
  # Calculate concentration from mean luminescence
  glyc_0.0B_mean_conc[i] <- (glyc_0.0B_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_0.0B_data <- data.frame(
  glyc_0.0B_dilution_factor = glyc_0.0B_dilution,
  glyc_0.0B_mean_luminescence = glyc_0.0B_mean_lum,
  glyc_0.0B_se = glyc_0.0B_se_lum,
  glyc_0.0B_conc =  glyc_0.0B_mean_conc,
  label = paste0("0.0Bdf.", glyc_0.0B_dilution)
)
glyc_0.0B_mean_lum
glyc_0.0B_mean_conc
```

```{r add 0.0C** glycogen to standard curve, include=TRUE}
# Create sample data frame for 0.0C**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_0.0C_mean_lum <- numeric(length(glyc_0.0C_dilution))
glyc_0.0C_se_lum <- numeric(length(glyc_0.0C_dilution))
glyc_0.0C_mean_conc <- numeric(length(glyc_0.0C_dilution))

for (i in 1:length(glyc_0.0C_dilution)) {
  df_val <- glyc_0.0C_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_0.0C_lum_values <- c(glyc_0.0C_luminescence[glyc_0.0C_dilution == df_val])
  glyc_0.0C_mean_lum[i] <- mean(glyc_0.0C_lum_values)
  glyc_0.0C_se_lum[i] <- sd(glyc_0.0C_lum_values) / sqrt(length(glyc_0.0C_lum_values))
  # Calculate concentration from mean luminescence
  glyc_0.0C_mean_conc[i] <- (glyc_0.0C_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_0.0C_data <- data.frame(
  glyc_0.0C_dilution_factor = glyc_0.0C_dilution,
  glyc_0.0C_mean_luminescence = glyc_0.0C_mean_lum,
  glyc_0.0C_se = glyc_0.0C_se_lum,
  glyc_0.0C_conc =  glyc_0.0C_mean_conc,
  label = paste0("0.0Cdf.", glyc_0.0C_dilution)
)
glyc_0.0C_mean_lum
glyc_0.0C_mean_conc
```

```{r add 0.0D** glycogen to standard curve, include=TRUE}
# Create sample data frame for 0.0D**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_0.0D_mean_lum <- numeric(length(glyc_0.0D_dilution))
glyc_0.0D_se_lum <- numeric(length(glyc_0.0D_dilution))
glyc_0.0D_mean_conc <- numeric(length(glyc_0.0D_dilution))

for (i in 1:length(glyc_0.0D_dilution)) {
  df_val <- glyc_0.0D_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_0.0D_lum_values <- c(glyc_0.0D_luminescence[glyc_0.0D_dilution == df_val])
  glyc_0.0D_mean_lum[i] <- mean(glyc_0.0D_lum_values)
  glyc_0.0D_se_lum[i] <- sd(glyc_0.0D_lum_values) / sqrt(length(glyc_0.0D_lum_values))
  # Calculate concentration from mean luminescence
  glyc_0.0D_mean_conc[i] <- (glyc_0.0D_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_0.0D_data <- data.frame(
  glyc_0.0D_dilution_factor = glyc_0.0D_dilution,
  glyc_0.0D_mean_luminescence = glyc_0.0D_mean_lum,
  glyc_0.0D_se = glyc_0.0D_se_lum,
  glyc_0.0D_conc =  glyc_0.0D_mean_conc,
  label = paste0("0.0Ddf.", glyc_0.0D_dilution)
)
glyc_0.0D_mean_lum
glyc_0.0D_mean_conc
```

```{r add 0.0E** glycogen to standard curve, include=TRUE}
# Create sample data frame for 0.0E**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_0.0E_mean_lum <- numeric(length(glyc_0.0E_dilution))
glyc_0.0E_se_lum <- numeric(length(glyc_0.0E_dilution))
glyc_0.0E_mean_conc <- numeric(length(glyc_0.0E_dilution))

for (i in 1:length(glyc_0.0E_dilution)) {
  df_val <- glyc_0.0E_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_0.0E_lum_values <- c(glyc_0.0E_luminescence[glyc_0.0E_dilution == df_val])
  glyc_0.0E_mean_lum[i] <- mean(glyc_0.0E_lum_values)
  glyc_0.0E_se_lum[i] <- sd(glyc_0.0E_lum_values) / sqrt(length(glyc_0.0E_lum_values))
  # Calculate concentration from mean luminescence
  glyc_0.0E_mean_conc[i] <- (glyc_0.0E_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_0.0E_data <- data.frame(
  glyc_0.0E_dilution_factor = glyc_0.0E_dilution,
  glyc_0.0E_mean_luminescence = glyc_0.0E_mean_lum,
  glyc_0.0E_se = glyc_0.0E_se_lum,
  glyc_0.0E_conc =  glyc_0.0E_mean_conc,
  label = paste0("0.0Edf.", glyc_0.0E_dilution)
)
glyc_0.0E_mean_lum
glyc_0.0E_mean_conc
```

```{r add 0.0F** glycogen to standard curve, include=TRUE}
# Create sample data frame for 0.0F**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_0.0F_mean_lum <- numeric(length(glyc_0.0F_dilution))
glyc_0.0F_se_lum <- numeric(length(glyc_0.0F_dilution))
glyc_0.0F_mean_conc <- numeric(length(glyc_0.0F_dilution))

for (i in 1:length(glyc_0.0F_dilution)) {
  df_val <- glyc_0.0F_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_0.0F_lum_values <- c(glyc_0.0F_luminescence[glyc_0.0F_dilution == df_val])
  glyc_0.0F_mean_lum[i] <- mean(glyc_0.0F_lum_values)
  glyc_0.0F_se_lum[i] <- sd(glyc_0.0F_lum_values) / sqrt(length(glyc_0.0F_lum_values))
  # Calculate concentration from mean luminescence
  glyc_0.0F_mean_conc[i] <- (glyc_0.0F_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_0.0F_data <- data.frame(
  glyc_0.0F_dilution_factor = glyc_0.0F_dilution,
  glyc_0.0F_mean_luminescence = glyc_0.0F_mean_lum,
  glyc_0.0F_se = glyc_0.0F_se_lum,
  glyc_0.0F_conc =  glyc_0.0F_mean_conc,
  label = paste0("0.0Fdf.", glyc_0.0F_dilution)
)
glyc_0.0F_mean_lum
glyc_0.0F_mean_conc
```

```{r add 0.0G** glycogen to standard curve, include=TRUE}
# Create sample data frame for 0.0G**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_0.0G_mean_lum <- numeric(length(glyc_0.0G_dilution))
glyc_0.0G_se_lum <- numeric(length(glyc_0.0G_dilution))
glyc_0.0G_mean_conc <- numeric(length(glyc_0.0G_dilution))

for (i in 1:length(glyc_0.0G_dilution)) {
  df_val <- glyc_0.0G_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_0.0G_lum_values <- c(glyc_0.0G_luminescence[glyc_0.0G_dilution == df_val])
  glyc_0.0G_mean_lum[i] <- mean(glyc_0.0G_lum_values)
  glyc_0.0G_se_lum[i] <- sd(glyc_0.0G_lum_values) / sqrt(length(glyc_0.0G_lum_values))
  # Calculate concentration from mean luminescence
  glyc_0.0G_mean_conc[i] <- (glyc_0.0G_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_0.0G_data <- data.frame(
  glyc_0.0G_dilution_factor = glyc_0.0G_dilution,
  glyc_0.0G_mean_luminescence = glyc_0.0G_mean_lum,
  glyc_0.0G_se = glyc_0.0G_se_lum,
  glyc_0.0G_conc =  glyc_0.0G_mean_conc,
  label = paste0("0.0Gdf.", glyc_0.0G_dilution)
)
glyc_0.0G_mean_lum
glyc_0.0G_mean_conc
```

```{r add 0.0H** glycogen to standard curve, include=TRUE}
# Create sample data frame for 0.0H**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_0.0H_mean_lum <- numeric(length(glyc_0.0H_dilution))
glyc_0.0H_se_lum <- numeric(length(glyc_0.0H_dilution))
glyc_0.0H_mean_conc <- numeric(length(glyc_0.0H_dilution))

for (i in 1:length(glyc_0.0H_dilution)) {
  df_val <- glyc_0.0H_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_0.0H_lum_values <- c(glyc_0.0H_luminescence[glyc_0.0H_dilution == df_val])
  glyc_0.0H_mean_lum[i] <- mean(glyc_0.0H_lum_values)
  glyc_0.0H_se_lum[i] <- sd(glyc_0.0H_lum_values) / sqrt(length(glyc_0.0H_lum_values))
  # Calculate concentration from mean luminescence
  glyc_0.0H_mean_conc[i] <- (glyc_0.0H_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_0.0H_data <- data.frame(
  glyc_0.0H_dilution_factor = glyc_0.0H_dilution,
  glyc_0.0H_mean_luminescence = glyc_0.0H_mean_lum,
  glyc_0.0H_se = glyc_0.0H_se_lum,
  glyc_0.0H_conc =  glyc_0.0H_mean_conc,
  label = paste0("0.0Hdf.", glyc_0.0H_dilution)
)
glyc_0.0H_mean_lum
glyc_0.0H_mean_conc
```

```{r add 0.0I** glycogen to standard curve, include=TRUE}
# Create sample data frame for 0.0I**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_0.0I_mean_lum <- numeric(length(glyc_0.0I_dilution))
glyc_0.0I_se_lum <- numeric(length(glyc_0.0I_dilution))
glyc_0.0I_mean_conc <- numeric(length(glyc_0.0I_dilution))

for (i in 1:length(glyc_0.0I_dilution)) {
  df_val <- glyc_0.0I_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_0.0I_lum_values <- c(glyc_0.0I_luminescence[glyc_0.0I_dilution == df_val])
  glyc_0.0I_mean_lum[i] <- mean(glyc_0.0I_lum_values)
  glyc_0.0I_se_lum[i] <- sd(glyc_0.0I_lum_values) / sqrt(length(glyc_0.0I_lum_values))
  # Calculate concentration from mean luminescence
  glyc_0.0I_mean_conc[i] <- (glyc_0.0I_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_0.0I_data <- data.frame(
  glyc_0.0I_dilution_factor = glyc_0.0I_dilution,
  glyc_0.0I_mean_luminescence = glyc_0.0I_mean_lum,
  glyc_0.0I_se = glyc_0.0I_se_lum,
  glyc_0.0I_conc =  glyc_0.0I_mean_conc,
  label = paste0("0.0Idf.", glyc_0.0I_dilution)
)
glyc_0.0I_mean_lum
glyc_0.0I_mean_conc
```

```{r add 0.0J** glycogen to standard curve, include=TRUE}
# Create sample data frame for 0.0J**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_0.0J_mean_lum <- numeric(length(glyc_0.0J_dilution))
glyc_0.0J_se_lum <- numeric(length(glyc_0.0J_dilution))
glyc_0.0J_mean_conc <- numeric(length(glyc_0.0J_dilution))

for (i in 1:length(glyc_0.0J_dilution)) {
  df_val <- glyc_0.0J_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_0.0J_lum_values <- c(glyc_0.0J_luminescence[glyc_0.0J_dilution == df_val])
  glyc_0.0J_mean_lum[i] <- mean(glyc_0.0J_lum_values)
  glyc_0.0J_se_lum[i] <- sd(glyc_0.0J_lum_values) / sqrt(length(glyc_0.0J_lum_values))
  # Calculate concentration from mean luminescence
  glyc_0.0J_mean_conc[i] <- (glyc_0.0J_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_0.0J_data <- data.frame(
  glyc_0.0J_dilution_factor = glyc_0.0J_dilution,
  glyc_0.0J_mean_luminescence = glyc_0.0J_mean_lum,
  glyc_0.0J_se = glyc_0.0J_se_lum,
  glyc_0.0J_conc =  glyc_0.0J_mean_conc,
  label = paste0("0.0Jdf.", glyc_0.0J_dilution)
)
glyc_0.0J_mean_lum
glyc_0.0J_mean_conc
```

```{r add 48.0A** glycogen to standard curve, include=TRUE}
# Create sample data frame for for 48.0A**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.0A_mean_lum <- numeric(length(glyc_48.0A_dilution))
glyc_48.0A_se_lum <- numeric(length(glyc_48.0A_dilution))
glyc_48.0A_mean_conc <- numeric(length(glyc_48.0A_dilution))

for (i in 1:length(glyc_48.0A_dilution)) {
  df_val <- glyc_48.0A_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.0A_lum_values <- c(glyc_48.0A_luminescence[glyc_48.0A_dilution == df_val])
  glyc_48.0A_mean_lum[i] <- mean(glyc_48.0A_lum_values)
  glyc_48.0A_se_lum[i] <- sd(glyc_48.0A_lum_values) / sqrt(length(glyc_48.0A_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.0A_mean_conc[i] <- (glyc_48.0A_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.0A_data <- data.frame(
  glyc_48.0A_dilution_factor = glyc_48.0A_dilution,
  glyc_48.0A_mean_luminescence = glyc_48.0A_mean_lum,
  glyc_48.0A_se = glyc_48.0A_se_lum,
  glyc_48.0A_conc =  glyc_48.0A_mean_conc,
  label = paste0("48.0Adf.", glyc_48.0A_dilution)
)
glyc_48.0A_mean_lum
glyc_48.0A_mean_conc
```

```{r add 48.0B** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.0B**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.0B_mean_lum <- numeric(length(glyc_48.0B_dilution))
glyc_48.0B_se_lum <- numeric(length(glyc_48.0B_dilution))
glyc_48.0B_mean_conc <- numeric(length(glyc_48.0B_dilution))

for (i in 1:length(glyc_48.0B_dilution)) {
  df_val <- glyc_48.0B_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.0B_lum_values <- c(glyc_48.0B_luminescence[glyc_48.0B_dilution == df_val])
  glyc_48.0B_mean_lum[i] <- mean(glyc_48.0B_lum_values)
  glyc_48.0B_se_lum[i] <- sd(glyc_48.0B_lum_values) / sqrt(length(glyc_48.0B_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.0B_mean_conc[i] <- (glyc_48.0B_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.0B_data <- data.frame(
  glyc_48.0B_dilution_factor = glyc_48.0B_dilution,
  glyc_48.0B_mean_luminescence = glyc_48.0B_mean_lum,
  glyc_48.0B_se = glyc_48.0B_se_lum,
  glyc_48.0B_conc =  glyc_48.0B_mean_conc,
  label = paste0("48.0Bdf.", glyc_48.0B_dilution)
)
glyc_48.0B_mean_lum
glyc_48.0B_mean_conc
```

```{r add 48.0C** glycogen to standard curve, include=TRUE}
# Create sample data frame 48.0C**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.0C_mean_lum <- numeric(length(glyc_48.0C_dilution))
glyc_48.0C_se_lum <- numeric(length(glyc_48.0C_dilution))
glyc_48.0C_mean_conc <- numeric(length(glyc_48.0C_dilution))

for (i in 1:length(glyc_48.0C_dilution)) {
  df_val <- glyc_48.0C_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.0C_lum_values <- c(glyc_48.0C_luminescence[glyc_48.0C_dilution == df_val])
  glyc_48.0C_mean_lum[i] <- mean(glyc_48.0C_lum_values)
  glyc_48.0C_se_lum[i] <- sd(glyc_48.0C_lum_values) / sqrt(length(glyc_48.0C_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.0C_mean_conc[i] <- (glyc_48.0C_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.0C_data <- data.frame(
  glyc_48.0C_dilution_factor = glyc_48.0C_dilution,
  glyc_48.0C_mean_luminescence = glyc_48.0C_mean_lum,
  glyc_48.0C_se = glyc_48.0C_se_lum,
  glyc_48.0C_conc =  glyc_48.0C_mean_conc,
  label = paste0("48.0Cdf.", glyc_48.0C_dilution)
)
glyc_48.0C_mean_lum
glyc_48.0C_mean_conc
```

```{r add 48.0D** glycogen to standard curve, include=TRUE}
# Create sample data frame 48.0D**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.0D_mean_lum <- numeric(length(glyc_48.0D_dilution))
glyc_48.0D_se_lum <- numeric(length(glyc_48.0D_dilution))
glyc_48.0D_mean_conc <- numeric(length(glyc_48.0D_dilution))

for (i in 1:length(glyc_48.0D_dilution)) {
  df_val <- glyc_48.0D_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.0D_lum_values <- c(glyc_48.0D_luminescence[glyc_48.0D_dilution == df_val])
  glyc_48.0D_mean_lum[i] <- mean(glyc_48.0D_lum_values)
  glyc_48.0D_se_lum[i] <- sd(glyc_48.0D_lum_values) / sqrt(length(glyc_48.0D_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.0D_mean_conc[i] <- (glyc_48.0D_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.0D_data <- data.frame(
  glyc_48.0D_dilution_factor = glyc_48.0D_dilution,
  glyc_48.0D_mean_luminescence = glyc_48.0D_mean_lum,
  glyc_48.0D_se = glyc_48.0D_se_lum,
  glyc_48.0D_conc =  glyc_48.0D_mean_conc,
  label = paste0("48.0Ddf.", glyc_48.0D_dilution)
)
glyc_48.0D_mean_lum
glyc_48.0D_mean_conc
```

```{r add 48.0E** glycogen to standard curve, include=TRUE}
# Create sample data frame 48.0E**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.0E_mean_lum <- numeric(length(glyc_48.0E_dilution))
glyc_48.0E_se_lum <- numeric(length(glyc_48.0E_dilution))
glyc_48.0E_mean_conc <- numeric(length(glyc_48.0E_dilution))

for (i in 1:length(glyc_48.0E_dilution)) {
  df_val <- glyc_48.0E_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.0E_lum_values <- c(glyc_48.0E_luminescence[glyc_48.0E_dilution == df_val])
  glyc_48.0E_mean_lum[i] <- mean(glyc_48.0E_lum_values)
  glyc_48.0E_se_lum[i] <- sd(glyc_48.0E_lum_values) / sqrt(length(glyc_48.0E_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.0E_mean_conc[i] <- (glyc_48.0E_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.0E_data <- data.frame(
  glyc_48.0E_dilution_factor = glyc_48.0E_dilution,
  glyc_48.0E_mean_luminescence = glyc_48.0E_mean_lum,
  glyc_48.0E_se = glyc_48.0E_se_lum,
  glyc_48.0E_conc =  glyc_48.0E_mean_conc,
  label = paste0("48.0Edf.", glyc_48.0E_dilution)
)
glyc_48.0E_mean_lum
glyc_48.0E_mean_conc
```

```{r add 48.0F** glycogen to standard curve, include=TRUE}
# Create sample data frame 48.0F**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.0F_mean_lum <- numeric(length(glyc_48.0F_dilution))
glyc_48.0F_se_lum <- numeric(length(glyc_48.0F_dilution))
glyc_48.0F_mean_conc <- numeric(length(glyc_48.0F_dilution))

for (i in 1:length(glyc_48.0F_dilution)) {
  df_val <- glyc_48.0F_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.0F_lum_values <- c(glyc_48.0F_luminescence[glyc_48.0F_dilution == df_val])
  glyc_48.0F_mean_lum[i] <- mean(glyc_48.0F_lum_values)
  glyc_48.0F_se_lum[i] <- sd(glyc_48.0F_lum_values) / sqrt(length(glyc_48.0F_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.0F_mean_conc[i] <- (glyc_48.0F_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.0F_data <- data.frame(
  glyc_48.0F_dilution_factor = glyc_48.0F_dilution,
  glyc_48.0F_mean_luminescence = glyc_48.0F_mean_lum,
  glyc_48.0F_se = glyc_48.0F_se_lum,
  glyc_48.0F_conc =  glyc_48.0F_mean_conc,
  label = paste0("48.0Fdf.", glyc_48.0F_dilution)
)
glyc_48.0F_mean_lum
glyc_48.0F_mean_conc
```
```{r add 48.0G** glycogen to standard curve, include=TRUE}
# Create sample data frame 48.0G**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.0G_mean_lum <- numeric(length(glyc_48.0G_dilution))
glyc_48.0G_se_lum <- numeric(length(glyc_48.0G_dilution))
glyc_48.0G_mean_conc <- numeric(length(glyc_48.0G_dilution))

for (i in 1:length(glyc_48.0G_dilution)) {
  df_val <- glyc_48.0G_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.0G_lum_values <- c(glyc_48.0G_luminescence[glyc_48.0G_dilution == df_val])
  glyc_48.0G_mean_lum[i] <- mean(glyc_48.0G_lum_values)
  glyc_48.0G_se_lum[i] <- sd(glyc_48.0G_lum_values) / sqrt(length(glyc_48.0G_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.0G_mean_conc[i] <- (glyc_48.0G_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.0G_data <- data.frame(
  glyc_48.0G_dilution_factor = glyc_48.0G_dilution,
  glyc_48.0G_mean_luminescence = glyc_48.0G_mean_lum,
  glyc_48.0G_se = glyc_48.0G_se_lum,
  glyc_48.0G_conc =  glyc_48.0G_mean_conc,
  label = paste0("48.0Gdf.", glyc_48.0G_dilution)
)
glyc_48.0G_mean_lum
glyc_48.0G_mean_conc
```

```{r add 48.0H** glycogen to standard curve, include=TRUE}
# Create sample data frame 48.0H**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.0H_mean_lum <- numeric(length(glyc_48.0H_dilution))
glyc_48.0H_se_lum <- numeric(length(glyc_48.0H_dilution))
glyc_48.0H_mean_conc <- numeric(length(glyc_48.0H_dilution))

for (i in 1:length(glyc_48.0H_dilution)) {
  df_val <- glyc_48.0H_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.0H_lum_values <- c(glyc_48.0H_luminescence[glyc_48.0H_dilution == df_val])
  glyc_48.0H_mean_lum[i] <- mean(glyc_48.0H_lum_values)
  glyc_48.0H_se_lum[i] <- sd(glyc_48.0H_lum_values) / sqrt(length(glyc_48.0H_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.0H_mean_conc[i] <- (glyc_48.0H_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.0H_data <- data.frame(
  glyc_48.0H_dilution_factor = glyc_48.0H_dilution,
  glyc_48.0H_mean_luminescence = glyc_48.0H_mean_lum,
  glyc_48.0H_se = glyc_48.0H_se_lum,
  glyc_48.0H_conc =  glyc_48.0H_mean_conc,
  label = paste0("48.0Hdf.", glyc_48.0H_dilution)
)
glyc_48.0H_mean_lum
glyc_48.0H_mean_conc
```

```{r add 48.0I** glycogen to standard curve, include=TRUE}
# Create sample data frame 48.0I**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.0I_mean_lum <- numeric(length(glyc_48.0I_dilution))
glyc_48.0I_se_lum <- numeric(length(glyc_48.0I_dilution))
glyc_48.0I_mean_conc <- numeric(length(glyc_48.0I_dilution))

for (i in 1:length(glyc_48.0I_dilution)) {
  df_val <- glyc_48.0I_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.0I_lum_values <- c(glyc_48.0I_luminescence[glyc_48.0I_dilution == df_val])
  glyc_48.0I_mean_lum[i] <- mean(glyc_48.0I_lum_values)
  glyc_48.0I_se_lum[i] <- sd(glyc_48.0I_lum_values) / sqrt(length(glyc_48.0I_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.0I_mean_conc[i] <- (glyc_48.0I_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.0I_data <- data.frame(
  glyc_48.0I_dilution_factor = glyc_48.0I_dilution,
  glyc_48.0I_mean_luminescence = glyc_48.0I_mean_lum,
  glyc_48.0I_se = glyc_48.0I_se_lum,
  glyc_48.0I_conc =  glyc_48.0I_mean_conc,
  label = paste0("48.0Idf.", glyc_48.0I_dilution)
)
glyc_48.0I_mean_lum
glyc_48.0I_mean_conc
```

```{r add 48.0J** glycogen to standard curve, include=TRUE}
# Create sample data frame 48.0J**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.0J_mean_lum <- numeric(length(glyc_48.0J_dilution))
glyc_48.0J_se_lum <- numeric(length(glyc_48.0J_dilution))
glyc_48.0J_mean_conc <- numeric(length(glyc_48.0J_dilution))

for (i in 1:length(glyc_48.0J_dilution)) {
  df_val <- glyc_48.0J_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.0J_lum_values <- c(glyc_48.0J_luminescence[glyc_48.0J_dilution == df_val])
  glyc_48.0J_mean_lum[i] <- mean(glyc_48.0J_lum_values)
  glyc_48.0J_se_lum[i] <- sd(glyc_48.0J_lum_values) / sqrt(length(glyc_48.0J_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.0J_mean_conc[i] <- (glyc_48.0J_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.0J_data <- data.frame(
  glyc_48.0J_dilution_factor = glyc_48.0J_dilution,
  glyc_48.0J_mean_luminescence = glyc_48.0J_mean_lum,
  glyc_48.0J_se = glyc_48.0J_se_lum,
  glyc_48.0J_conc =  glyc_48.0J_mean_conc,
  label = paste0("48.0Jdf.", glyc_48.0J_dilution)
)
glyc_48.0J_mean_lum
glyc_48.0J_mean_conc
```

```{r add 48.2A** glycogen to standard curve, include=TRUE}
# Create sample data frame 48.2A**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.2A_mean_lum <- numeric(length(glyc_48.2A_dilution))
glyc_48.2A_se_lum <- numeric(length(glyc_48.2A_dilution))
glyc_48.2A_mean_conc <- numeric(length(glyc_48.2A_dilution))

for (i in 1:length(glyc_48.2A_dilution)) {
  df_val <- glyc_48.2A_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.2A_lum_values <- c(glyc_48.2A_luminescence[glyc_48.2A_dilution == df_val])
  glyc_48.2A_mean_lum[i] <- mean(glyc_48.2A_lum_values)
  glyc_48.2A_se_lum[i] <- sd(glyc_48.2A_lum_values) / sqrt(length(glyc_48.2A_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.2A_mean_conc[i] <- (glyc_48.2A_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.2A_data <- data.frame(
  glyc_48.2A_dilution_factor = glyc_48.2A_dilution,
  glyc_48.2A_mean_luminescence = glyc_48.2A_mean_lum,
  glyc_48.2A_se = glyc_48.2A_se_lum,
  glyc_48.2A_conc =  glyc_48.2A_mean_conc,
  label = paste0("48.2Adf.", glyc_48.2A_dilution)
)
glyc_48.2A_mean_lum
glyc_48.2A_mean_conc
```

```{r add 48.2B** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.2B**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.2B_mean_lum <- numeric(length(glyc_48.2B_dilution))
glyc_48.2B_se_lum <- numeric(length(glyc_48.2B_dilution))
glyc_48.2B_mean_conc <- numeric(length(glyc_48.2B_dilution))

for (i in 1:length(glyc_48.2B_dilution)) {
  df_val <- glyc_48.2B_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.2B_lum_values <- c(glyc_48.2B_luminescence[glyc_48.2B_dilution == df_val])
  glyc_48.2B_mean_lum[i] <- mean(glyc_48.2B_lum_values)
  glyc_48.2B_se_lum[i] <- sd(glyc_48.2B_lum_values) / sqrt(length(glyc_48.2B_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.2B_mean_conc[i] <- (glyc_48.2B_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.2B_data <- data.frame(
  glyc_48.2B_dilution_factor = glyc_48.2B_dilution,
  glyc_48.2B_mean_luminescence = glyc_48.2B_mean_lum,
  glyc_48.2B_se = glyc_48.2B_se_lum,
  glyc_48.2B_conc =  glyc_48.2B_mean_conc,
  label = paste0("48.2Bdf.", glyc_48.2B_dilution)
)
glyc_48.2B_mean_lum
glyc_48.2B_mean_conc
```

```{r add 48.2C** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.2C**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.2C_mean_lum <- numeric(length(glyc_48.2C_dilution))
glyc_48.2C_se_lum <- numeric(length(glyc_48.2C_dilution))
glyc_48.2C_mean_conc <- numeric(length(glyc_48.2C_dilution))

for (i in 1:length(glyc_48.2C_dilution)) {
  df_val <- glyc_48.2C_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.2C_lum_values <- c(glyc_48.2C_luminescence[glyc_48.2C_dilution == df_val])
  glyc_48.2C_mean_lum[i] <- mean(glyc_48.2C_lum_values)
  glyc_48.2C_se_lum[i] <- sd(glyc_48.2C_lum_values) / sqrt(length(glyc_48.2C_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.2C_mean_conc[i] <- (glyc_48.2C_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.2C_data <- data.frame(
  glyc_48.2C_dilution_factor = glyc_48.2C_dilution,
  glyc_48.2C_mean_luminescence = glyc_48.2C_mean_lum,
  glyc_48.2C_se = glyc_48.2C_se_lum,
  glyc_48.2C_conc =  glyc_48.2C_mean_conc,
  label = paste0("48.2Cdf.", glyc_48.2C_dilution)
)
glyc_48.2C_mean_lum
glyc_48.2C_mean_conc
```

```{r add 48.2D** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.2D**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.2D_mean_lum <- numeric(length(glyc_48.2D_dilution))
glyc_48.2D_se_lum <- numeric(length(glyc_48.2D_dilution))
glyc_48.2D_mean_conc <- numeric(length(glyc_48.2D_dilution))

for (i in 1:length(glyc_48.2D_dilution)) {
  df_val <- glyc_48.2D_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.2D_lum_values <- c(glyc_48.2D_luminescence[glyc_48.2D_dilution == df_val])
  glyc_48.2D_mean_lum[i] <- mean(glyc_48.2D_lum_values)
  glyc_48.2D_se_lum[i] <- sd(glyc_48.2D_lum_values) / sqrt(length(glyc_48.2D_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.2D_mean_conc[i] <- (glyc_48.2D_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.2D_data <- data.frame(
  glyc_48.2D_dilution_factor = glyc_48.2D_dilution,
  glyc_48.2D_mean_luminescence = glyc_48.2D_mean_lum,
  glyc_48.2D_se = glyc_48.2D_se_lum,
  glyc_48.2D_conc =  glyc_48.2D_mean_conc,
  label = paste0("48.2Ddf.", glyc_48.2D_dilution)
)
glyc_48.2D_mean_lum
glyc_48.2D_mean_conc
```

```{r add 48.2E** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.2E**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.2E_mean_lum <- numeric(length(glyc_48.2E_dilution))
glyc_48.2E_se_lum <- numeric(length(glyc_48.2E_dilution))
glyc_48.2E_mean_conc <- numeric(length(glyc_48.2E_dilution))

for (i in 1:length(glyc_48.2E_dilution)) {
  df_val <- glyc_48.2E_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.2E_lum_values <- c(glyc_48.2E_luminescence[glyc_48.2E_dilution == df_val])
  glyc_48.2E_mean_lum[i] <- mean(glyc_48.2E_lum_values)
  glyc_48.2E_se_lum[i] <- sd(glyc_48.2E_lum_values) / sqrt(length(glyc_48.2E_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.2E_mean_conc[i] <- (glyc_48.2E_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.2E_data <- data.frame(
  glyc_48.2E_dilution_factor = glyc_48.2E_dilution,
  glyc_48.2E_mean_luminescence = glyc_48.2E_mean_lum,
  glyc_48.2E_se = glyc_48.2E_se_lum,
  glyc_48.2E_conc =  glyc_48.2E_mean_conc,
  label = paste0("48.2Edf.", glyc_48.2E_dilution)
)
glyc_48.2E_mean_lum
glyc_48.2E_mean_conc
```

```{r add 48.2F** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.2F**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.2F_mean_lum <- numeric(length(glyc_48.2F_dilution))
glyc_48.2F_se_lum <- numeric(length(glyc_48.2F_dilution))
glyc_48.2F_mean_conc <- numeric(length(glyc_48.2F_dilution))

for (i in 1:length(glyc_48.2F_dilution)) {
  df_val <- glyc_48.2F_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.2F_lum_values <- c(glyc_48.2F_luminescence[glyc_48.2F_dilution == df_val])
  glyc_48.2F_mean_lum[i] <- mean(glyc_48.2F_lum_values)
  glyc_48.2F_se_lum[i] <- sd(glyc_48.2F_lum_values) / sqrt(length(glyc_48.2F_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.2F_mean_conc[i] <- (glyc_48.2F_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.2F_data <- data.frame(
  glyc_48.2F_dilution_factor = glyc_48.2F_dilution,
  glyc_48.2F_mean_luminescence = glyc_48.2F_mean_lum,
  glyc_48.2F_se = glyc_48.2F_se_lum,
  glyc_48.2F_conc =  glyc_48.2F_mean_conc,
  label = paste0("48.2Fdf.", glyc_48.2F_dilution)
)
glyc_48.2F_mean_lum
glyc_48.2F_mean_conc
```

```{r add 48.2G** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.2G**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.2G_mean_lum <- numeric(length(glyc_48.2G_dilution))
glyc_48.2G_se_lum <- numeric(length(glyc_48.2G_dilution))
glyc_48.2G_mean_conc <- numeric(length(glyc_48.2G_dilution))

for (i in 1:length(glyc_48.2G_dilution)) {
  df_val <- glyc_48.2G_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.2G_lum_values <- c(glyc_48.2G_luminescence[glyc_48.2G_dilution == df_val])
  glyc_48.2G_mean_lum[i] <- mean(glyc_48.2G_lum_values)
  glyc_48.2G_se_lum[i] <- sd(glyc_48.2G_lum_values) / sqrt(length(glyc_48.2G_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.2G_mean_conc[i] <- (glyc_48.2G_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.2G_data <- data.frame(
  glyc_48.2G_dilution_factor = glyc_48.2G_dilution,
  glyc_48.2G_mean_luminescence = glyc_48.2G_mean_lum,
  glyc_48.2G_se = glyc_48.2G_se_lum,
  glyc_48.2G_conc =  glyc_48.2G_mean_conc,
  label = paste0("48.2Gdf.", glyc_48.2G_dilution)
)
glyc_48.2G_mean_lum
glyc_48.2G_mean_conc
```

```{r add 48.2H** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.2H**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.2H_mean_lum <- numeric(length(glyc_48.2H_dilution))
glyc_48.2H_se_lum <- numeric(length(glyc_48.2H_dilution))
glyc_48.2H_mean_conc <- numeric(length(glyc_48.2H_dilution))

for (i in 1:length(glyc_48.2H_dilution)) {
  df_val <- glyc_48.2H_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.2H_lum_values <- c(glyc_48.2H_luminescence[glyc_48.2H_dilution == df_val])
  glyc_48.2H_mean_lum[i] <- mean(glyc_48.2H_lum_values)
  glyc_48.2H_se_lum[i] <- sd(glyc_48.2H_lum_values) / sqrt(length(glyc_48.2H_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.2H_mean_conc[i] <- (glyc_48.2H_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.2H_data <- data.frame(
  glyc_48.2H_dilution_factor = glyc_48.2H_dilution,
  glyc_48.2H_mean_luminescence = glyc_48.2H_mean_lum,
  glyc_48.2H_se = glyc_48.2H_se_lum,
  glyc_48.2H_conc =  glyc_48.2H_mean_conc,
  label = paste0("48.2Hdf.", glyc_48.2H_dilution)
)
glyc_48.2H_mean_lum
glyc_48.2H_mean_conc
```

```{r add 48.2I** glycogen to standard curve, include=TRUE}
# Create sample data frame for for 48.2I**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.2I_mean_lum <- numeric(length(glyc_48.2I_dilution))
glyc_48.2I_se_lum <- numeric(length(glyc_48.2I_dilution))
glyc_48.2I_mean_conc <- numeric(length(glyc_48.2I_dilution))

for (i in 1:length(glyc_48.2I_dilution)) {
  df_val <- glyc_48.2I_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.2I_lum_values <- c(glyc_48.2I_luminescence[glyc_48.2I_dilution == df_val])
  glyc_48.2I_mean_lum[i] <- mean(glyc_48.2I_lum_values)
  glyc_48.2I_se_lum[i] <- sd(glyc_48.2I_lum_values) / sqrt(length(glyc_48.2I_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.2I_mean_conc[i] <- (glyc_48.2I_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.2I_data <- data.frame(
  glyc_48.2I_dilution_factor = glyc_48.2I_dilution,
  glyc_48.2I_mean_luminescence = glyc_48.2I_mean_lum,
  glyc_48.2I_se = glyc_48.2I_se_lum,
  glyc_48.2I_conc =  glyc_48.2I_mean_conc,
  label = paste0("48.2Idf.", glyc_48.2I_dilution)
)
glyc_48.2I_mean_lum
glyc_48.2I_mean_conc
```

```{r add 48.2J** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.2J**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.2J_mean_lum <- numeric(length(glyc_48.2J_dilution))
glyc_48.2J_se_lum <- numeric(length(glyc_48.2J_dilution))
glyc_48.2J_mean_conc <- numeric(length(glyc_48.2J_dilution))

for (i in 1:length(glyc_48.2J_dilution)) {
  df_val <- glyc_48.2J_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.2J_lum_values <- c(glyc_48.2J_luminescence[glyc_48.2J_dilution == df_val])
  glyc_48.2J_mean_lum[i] <- mean(glyc_48.2J_lum_values)
  glyc_48.2J_se_lum[i] <- sd(glyc_48.2J_lum_values) / sqrt(length(glyc_48.2J_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.2J_mean_conc[i] <- (glyc_48.2J_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.2J_data <- data.frame(
  glyc_48.2J_dilution_factor = glyc_48.2J_dilution,
  glyc_48.2J_mean_luminescence = glyc_48.2J_mean_lum,
  glyc_48.2J_se = glyc_48.2J_se_lum,
  glyc_48.2J_conc =  glyc_48.2J_mean_conc,
  label = paste0("48.2Jdf.", glyc_48.2J_dilution)
)
glyc_48.2J_mean_lum
glyc_48.2J_mean_conc
```

```{r add 48.7A** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.7A**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.7A_mean_lum <- numeric(length(glyc_48.7A_dilution))
glyc_48.7A_se_lum <- numeric(length(glyc_48.7A_dilution))
glyc_48.7A_mean_conc <- numeric(length(glyc_48.7A_dilution))

for (i in 1:length(glyc_48.7A_dilution)) {
  df_val <- glyc_48.7A_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.7A_lum_values <- c(glyc_48.7A_luminescence[glyc_48.7A_dilution == df_val])
  glyc_48.7A_mean_lum[i] <- mean(glyc_48.7A_lum_values)
  glyc_48.7A_se_lum[i] <- sd(glyc_48.7A_lum_values) / sqrt(length(glyc_48.7A_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.7A_mean_conc[i] <- (glyc_48.7A_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.7A_data <- data.frame(
  glyc_48.7A_dilution_factor = glyc_48.7A_dilution,
  glyc_48.7A_mean_luminescence = glyc_48.7A_mean_lum,
  glyc_48.7A_se = glyc_48.7A_se_lum,
  glyc_48.7A_conc =  glyc_48.7A_mean_conc,
  label = paste0("48.7Adf.", glyc_48.7A_dilution)
)
glyc_48.7A_mean_lum
glyc_48.7A_mean_conc
```

```{r add 48.7B** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.7B**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.7B_mean_lum <- numeric(length(glyc_48.7B_dilution))
glyc_48.7B_se_lum <- numeric(length(glyc_48.7B_dilution))
glyc_48.7B_mean_conc <- numeric(length(glyc_48.7B_dilution))

for (i in 1:length(glyc_48.7B_dilution)) {
  df_val <- glyc_48.7B_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.7B_lum_values <- c(glyc_48.7B_luminescence[glyc_48.7B_dilution == df_val])
  glyc_48.7B_mean_lum[i] <- mean(glyc_48.7B_lum_values)
  glyc_48.7B_se_lum[i] <- sd(glyc_48.7B_lum_values) / sqrt(length(glyc_48.7B_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.7B_mean_conc[i] <- (glyc_48.7B_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.7B_data <- data.frame(
  glyc_48.7B_dilution_factor = glyc_48.7B_dilution,
  glyc_48.7B_mean_luminescence = glyc_48.7B_mean_lum,
  glyc_48.7B_se = glyc_48.7B_se_lum,
  glyc_48.7B_conc =  glyc_48.7B_mean_conc,
  label = paste0("48.7Bdf.", glyc_48.7B_dilution)
)
glyc_48.7B_mean_lum
glyc_48.7B_mean_conc
```

```{r add 48.7C** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.7C**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.7C_mean_lum <- numeric(length(glyc_48.7C_dilution))
glyc_48.7C_se_lum <- numeric(length(glyc_48.7C_dilution))
glyc_48.7C_mean_conc <- numeric(length(glyc_48.7C_dilution))

for (i in 1:length(glyc_48.7C_dilution)) {
  df_val <- glyc_48.7C_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.7C_lum_values <- c(glyc_48.7C_luminescence[glyc_48.7C_dilution == df_val])
  glyc_48.7C_mean_lum[i] <- mean(glyc_48.7C_lum_values)
  glyc_48.7C_se_lum[i] <- sd(glyc_48.7C_lum_values) / sqrt(length(glyc_48.7C_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.7C_mean_conc[i] <- (glyc_48.7C_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.7C_data <- data.frame(
  glyc_48.7C_dilution_factor = glyc_48.7C_dilution,
  glyc_48.7C_mean_luminescence = glyc_48.7C_mean_lum,
  glyc_48.7C_se = glyc_48.7C_se_lum,
  glyc_48.7C_conc =  glyc_48.7C_mean_conc,
  label = paste0("48.7Cdf.", glyc_48.7C_dilution)
)
glyc_48.7C_mean_lum
glyc_48.7C_mean_conc
```

```{r add 48.7D** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.7D**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.7D_mean_lum <- numeric(length(glyc_48.7D_dilution))
glyc_48.7D_se_lum <- numeric(length(glyc_48.7D_dilution))
glyc_48.7D_mean_conc <- numeric(length(glyc_48.7D_dilution))

for (i in 1:length(glyc_48.7D_dilution)) {
  df_val <- glyc_48.7D_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.7D_lum_values <- c(glyc_48.7D_luminescence[glyc_48.7D_dilution == df_val])
  glyc_48.7D_mean_lum[i] <- mean(glyc_48.7D_lum_values)
  glyc_48.7D_se_lum[i] <- sd(glyc_48.7D_lum_values) / sqrt(length(glyc_48.7D_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.7D_mean_conc[i] <- (glyc_48.7D_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.7D_data <- data.frame(
  glyc_48.7D_dilution_factor = glyc_48.7D_dilution,
  glyc_48.7D_mean_luminescence = glyc_48.7D_mean_lum,
  glyc_48.7D_se = glyc_48.7D_se_lum,
  glyc_48.7D_conc =  glyc_48.7D_mean_conc,
  label = paste0("48.7Ddf.", glyc_48.7D_dilution)
)
glyc_48.7D_mean_lum
glyc_48.7D_mean_conc
```

```{r add 48.7E** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.7E**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.7E_mean_lum <- numeric(length(glyc_48.7E_dilution))
glyc_48.7E_se_lum <- numeric(length(glyc_48.7E_dilution))
glyc_48.7E_mean_conc <- numeric(length(glyc_48.7E_dilution))

for (i in 1:length(glyc_48.7E_dilution)) {
  df_val <- glyc_48.7E_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.7E_lum_values <- c(glyc_48.7E_luminescence[glyc_48.7E_dilution == df_val])
  glyc_48.7E_mean_lum[i] <- mean(glyc_48.7E_lum_values)
  glyc_48.7E_se_lum[i] <- sd(glyc_48.7E_lum_values) / sqrt(length(glyc_48.7E_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.7E_mean_conc[i] <- (glyc_48.7E_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.7E_data <- data.frame(
  glyc_48.7E_dilution_factor = glyc_48.7E_dilution,
  glyc_48.7E_mean_luminescence = glyc_48.7E_mean_lum,
  glyc_48.7E_se = glyc_48.7E_se_lum,
  glyc_48.7E_conc =  glyc_48.7E_mean_conc,
  label = paste0("48.7Edf.", glyc_48.7E_dilution)
)
glyc_48.7E_mean_lum
glyc_48.7E_mean_conc
```

```{r add 48.7F** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.7F**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.7F_mean_lum <- numeric(length(glyc_48.7F_dilution))
glyc_48.7F_se_lum <- numeric(length(glyc_48.7F_dilution))
glyc_48.7F_mean_conc <- numeric(length(glyc_48.7F_dilution))

for (i in 1:length(glyc_48.7F_dilution)) {
  df_val <- glyc_48.7F_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.7F_lum_values <- c(glyc_48.7F_luminescence[glyc_48.7F_dilution == df_val])
  glyc_48.7F_mean_lum[i] <- mean(glyc_48.7F_lum_values)
  glyc_48.7F_se_lum[i] <- sd(glyc_48.7F_lum_values) / sqrt(length(glyc_48.7F_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.7F_mean_conc[i] <- (glyc_48.7F_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.7F_data <- data.frame(
  glyc_48.7F_dilution_factor = glyc_48.7F_dilution,
  glyc_48.7F_mean_luminescence = glyc_48.7F_mean_lum,
  glyc_48.7F_se = glyc_48.7F_se_lum,
  glyc_48.7F_conc =  glyc_48.7F_mean_conc,
  label = paste0("48.7Fdf.", glyc_48.7F_dilution)
)
glyc_48.7F_mean_lum
glyc_48.7F_mean_conc
```

```{r add 48.7G** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.7G**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.7G_mean_lum <- numeric(length(glyc_48.7G_dilution))
glyc_48.7G_se_lum <- numeric(length(glyc_48.7G_dilution))
glyc_48.7G_mean_conc <- numeric(length(glyc_48.7G_dilution))

for (i in 1:length(glyc_48.7G_dilution)) {
  df_val <- glyc_48.7G_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.7G_lum_values <- c(glyc_48.7G_luminescence[glyc_48.7G_dilution == df_val])
  glyc_48.7G_mean_lum[i] <- mean(glyc_48.7G_lum_values)
  glyc_48.7G_se_lum[i] <- sd(glyc_48.7G_lum_values) / sqrt(length(glyc_48.7G_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.7G_mean_conc[i] <- (glyc_48.7G_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.7G_data <- data.frame(
  glyc_48.7G_dilution_factor = glyc_48.7G_dilution,
  glyc_48.7G_mean_luminescence = glyc_48.7G_mean_lum,
  glyc_48.7G_se = glyc_48.7G_se_lum,
  glyc_48.7G_conc =  glyc_48.7G_mean_conc,
  label = paste0("48.7Gdf.", glyc_48.7G_dilution)
)
glyc_48.7G_mean_lum
glyc_48.7G_mean_conc
```

```{r add 48.7H** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.7H**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.7H_mean_lum <- numeric(length(glyc_48.7H_dilution))
glyc_48.7H_se_lum <- numeric(length(glyc_48.7H_dilution))
glyc_48.7H_mean_conc <- numeric(length(glyc_48.7H_dilution))

for (i in 1:length(glyc_48.7H_dilution)) {
  df_val <- glyc_48.7H_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.7H_lum_values <- c(glyc_48.7H_luminescence[glyc_48.7H_dilution == df_val])
  glyc_48.7H_mean_lum[i] <- mean(glyc_48.7H_lum_values)
  glyc_48.7H_se_lum[i] <- sd(glyc_48.7H_lum_values) / sqrt(length(glyc_48.7H_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.7H_mean_conc[i] <- (glyc_48.7H_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.7H_data <- data.frame(
  glyc_48.7H_dilution_factor = glyc_48.7H_dilution,
  glyc_48.7H_mean_luminescence = glyc_48.7H_mean_lum,
  glyc_48.7H_se = glyc_48.7H_se_lum,
  glyc_48.7H_conc =  glyc_48.7H_mean_conc,
  label = paste0("48.7Hdf.", glyc_48.7H_dilution)
)
glyc_48.7H_mean_lum
glyc_48.7H_mean_conc
```

```{r add 48.7I** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.7I**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.7I_mean_lum <- numeric(length(glyc_48.7I_dilution))
glyc_48.7I_se_lum <- numeric(length(glyc_48.7I_dilution))
glyc_48.7I_mean_conc <- numeric(length(glyc_48.7I_dilution))

for (i in 1:length(glyc_48.7I_dilution)) {
  df_val <- glyc_48.7I_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.7I_lum_values <- c(glyc_48.7I_luminescence[glyc_48.7I_dilution == df_val])
  glyc_48.7I_mean_lum[i] <- mean(glyc_48.7I_lum_values)
  glyc_48.7I_se_lum[i] <- sd(glyc_48.7I_lum_values) / sqrt(length(glyc_48.7I_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.7I_mean_conc[i] <- (glyc_48.7I_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.7I_data <- data.frame(
  glyc_48.7I_dilution_factor = glyc_48.7I_dilution,
  glyc_48.7I_mean_luminescence = glyc_48.7I_mean_lum,
  glyc_48.7I_se = glyc_48.7I_se_lum,
  glyc_48.7I_conc =  glyc_48.7I_mean_conc,
  label = paste0("48.7Idf.", glyc_48.7I_dilution)
)
glyc_48.7I_mean_lum
glyc_48.7I_mean_conc
```

```{r add 48.7J** glycogen to standard curve, include=TRUE}
# Create sample data frame for 48.7J**
# Calculate glycogen concentration from luminescence using the standard curve equation
# Rearranging y = mx + b to solve for x: x = (y - b) / m

# Calculate mean and SE for each dilution factor across replicates
glyc_48.7J_mean_lum <- numeric(length(glyc_48.7J_dilution))
glyc_48.7J_se_lum <- numeric(length(glyc_48.7J_dilution))
glyc_48.7J_mean_conc <- numeric(length(glyc_48.7J_dilution))

for (i in 1:length(glyc_48.7J_dilution)) {
  df_val <- glyc_48.7J_dilution[i]
  # Get luminescence values for this dilution factor from both rows
  glyc_48.7J_lum_values <- c(glyc_48.7J_luminescence[glyc_48.7J_dilution == df_val])
  glyc_48.7J_mean_lum[i] <- mean(glyc_48.7J_lum_values)
  glyc_48.7J_se_lum[i] <- sd(glyc_48.7J_lum_values) / sqrt(length(glyc_48.7J_lum_values))
  # Calculate concentration from mean luminescence
  glyc_48.7J_mean_conc[i] <- (glyc_48.7J_mean_lum[i] - glyc_intercept) / glyc_slope
}

glyc_48.7J_data <- data.frame(
  glyc_48.7J_dilution_factor = glyc_48.7J_dilution,
  glyc_48.7J_mean_luminescence = glyc_48.7J_mean_lum,
  glyc_48.7J_se = glyc_48.7J_se_lum,
  glyc_48.7J_conc =  glyc_48.7J_mean_conc,
  label = paste0("48.7Jdf.", glyc_48.7J_dilution)
)
glyc_48.7J_mean_lum
glyc_48.7J_mean_conc
```

### Plot glycogen standard curve, sample points

```{r plot glycogen standard curve with samples, include=TRUE}
# Create the plot
glyc_plot <- ggplot(glycogen_summary_data, aes(x = glyc_concentration, y = glyc_mean_luminescence)) +
  geom_smooth(aes(linetype = "Std Curve Best Fit Line"), method = "lm", se = FALSE, 
              color = "coral", linewidth = 1) +
  geom_errorbar(aes(ymin = glyc_mean_luminescence - glyc_se, ymax = glyc_mean_luminescence + glyc_se),
                width = 0.1, linewidth = 1, color = "darkblue") +
  geom_point(aes(color = "Standard Curve", shape = "Standard Curve"), size = 4) +
  
   geom_errorbar(data = glyc_0.0A_data, aes(x = glyc_0.0A_conc, y = glyc_0.0A_mean_luminescence,
                ymin = glyc_0.0A_mean_luminescence - glyc_0.0A_se, ymax = glyc_0.0A_mean_luminescence + glyc_0.0A_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_0.0A_data, aes(x = glyc_0.0A_conc, y = glyc_0.0A_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_0.0B_data, aes(x = glyc_0.0B_conc, y = glyc_0.0B_mean_luminescence,
                ymin = glyc_0.0B_mean_luminescence - glyc_0.0B_se, ymax = glyc_0.0B_mean_luminescence + glyc_0.0B_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_0.0B_data, aes(x = glyc_0.0B_conc, y = glyc_0.0B_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_0.0C_data, aes(x = glyc_0.0C_conc, y = glyc_0.0C_mean_luminescence,
                ymin = glyc_0.0C_mean_luminescence - glyc_0.0C_se, ymax = glyc_0.0C_mean_luminescence + glyc_0.0C_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_0.0C_data, aes(x = glyc_0.0C_conc, y = glyc_0.0C_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_0.0D_data, aes(x = glyc_0.0D_conc, y = glyc_0.0D_mean_luminescence,
                ymin = glyc_0.0D_mean_luminescence - glyc_0.0D_se, ymax = glyc_0.0D_mean_luminescence + glyc_0.0D_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_0.0D_data, aes(x = glyc_0.0D_conc, y = glyc_0.0D_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_0.0E_data, aes(x = glyc_0.0E_conc, y = glyc_0.0E_mean_luminescence,
                ymin = glyc_0.0E_mean_luminescence - glyc_0.0E_se, ymax = glyc_0.0E_mean_luminescence + glyc_0.0E_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_0.0E_data, aes(x = glyc_0.0E_conc, y = glyc_0.0E_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_0.0F_data, aes(x = glyc_0.0F_conc, y = glyc_0.0F_mean_luminescence,
                ymin = glyc_0.0F_mean_luminescence - glyc_0.0F_se, ymax = glyc_0.0F_mean_luminescence + glyc_0.0F_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_0.0F_data, aes(x = glyc_0.0F_conc, y = glyc_0.0F_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_0.0G_data, aes(x = glyc_0.0G_conc, y = glyc_0.0G_mean_luminescence,
                ymin = glyc_0.0G_mean_luminescence - glyc_0.0G_se, ymax = glyc_0.0G_mean_luminescence + glyc_0.0G_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_0.0G_data, aes(x = glyc_0.0G_conc, y = glyc_0.0G_mean_luminescence,
             color = label, shape = label), size = 4) +
  
    geom_errorbar(data = glyc_0.0H_data, aes(x = glyc_0.0H_conc, y = glyc_0.0H_mean_luminescence,
                ymin = glyc_0.0H_mean_luminescence - glyc_0.0H_se, ymax = glyc_0.0H_mean_luminescence + glyc_0.0H_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_0.0H_data, aes(x = glyc_0.0H_conc, y = glyc_0.0H_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_0.0I_data, aes(x = glyc_0.0I_conc, y = glyc_0.0I_mean_luminescence,
                ymin = glyc_0.0I_mean_luminescence - glyc_0.0I_se, ymax = glyc_0.0I_mean_luminescence + glyc_0.0I_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_0.0I_data, aes(x = glyc_0.0I_conc, y = glyc_0.0I_mean_luminescence,
             color = label, shape = label), size = 4) +
  
    geom_errorbar(data = glyc_0.0J_data, aes(x = glyc_0.0J_conc, y = glyc_0.0J_mean_luminescence,
                ymin = glyc_0.0J_mean_luminescence - glyc_0.0J_se, ymax = glyc_0.0J_mean_luminescence + glyc_0.0J_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_0.0J_data, aes(x = glyc_0.0J_conc, y = glyc_0.0J_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_48.0A_data, aes(x = glyc_48.0A_conc, y = glyc_48.0A_mean_luminescence,
                ymin = glyc_48.0A_mean_luminescence - glyc_48.0A_se, ymax = glyc_48.0A_mean_luminescence + glyc_48.0A_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.0A_data, aes(x = glyc_48.0A_conc, y = glyc_48.0A_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_48.0B_data, aes(x = glyc_48.0B_conc, y = glyc_48.0B_mean_luminescence,
                ymin = glyc_48.0B_mean_luminescence - glyc_48.0B_se, ymax = glyc_48.0B_mean_luminescence + glyc_48.0B_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.0B_data, aes(x = glyc_48.0B_conc, y = glyc_48.0B_mean_luminescence,
             color = label, shape = label), size = 4) +
  
    geom_errorbar(data = glyc_48.0C_data, aes(x = glyc_48.0C_conc, y = glyc_48.0C_mean_luminescence,
                ymin = glyc_48.0C_mean_luminescence - glyc_48.0C_se, ymax = glyc_48.0C_mean_luminescence + glyc_48.0C_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.0C_data, aes(x = glyc_48.0C_conc, y = glyc_48.0C_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.0D_data, aes(x = glyc_48.0D_conc, y = glyc_48.0D_mean_luminescence,
                ymin = glyc_48.0D_mean_luminescence - glyc_48.0D_se, ymax = glyc_48.0D_mean_luminescence + glyc_48.0D_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.0D_data, aes(x = glyc_48.0D_conc, y = glyc_48.0D_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.0E_data, aes(x = glyc_48.0E_conc, y = glyc_48.0E_mean_luminescence,
                ymin = glyc_48.0E_mean_luminescence - glyc_48.0E_se, ymax = glyc_48.0E_mean_luminescence + glyc_48.0E_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.0E_data, aes(x = glyc_48.0E_conc, y = glyc_48.0E_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.0F_data, aes(x = glyc_48.0F_conc, y = glyc_48.0F_mean_luminescence,
                ymin = glyc_48.0F_mean_luminescence - glyc_48.0F_se, ymax = glyc_48.0F_mean_luminescence + glyc_48.0F_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.0F_data, aes(x = glyc_48.0F_conc, y = glyc_48.0F_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.0G_data, aes(x = glyc_48.0G_conc, y = glyc_48.0G_mean_luminescence,
                ymin = glyc_48.0G_mean_luminescence - glyc_48.0G_se, ymax = glyc_48.0G_mean_luminescence + glyc_48.0G_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.0G_data, aes(x = glyc_48.0G_conc, y = glyc_48.0G_mean_luminescence,
             color = label, shape = label), size = 4) +
  
    geom_errorbar(data = glyc_48.0H_data, aes(x = glyc_48.0H_conc, y = glyc_48.0H_mean_luminescence,
                ymin = glyc_48.0H_mean_luminescence - glyc_48.0H_se, ymax = glyc_48.0H_mean_luminescence + glyc_48.0H_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.0H_data, aes(x = glyc_48.0H_conc, y = glyc_48.0H_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_48.0I_data, aes(x = glyc_48.0I_conc, y = glyc_48.0I_mean_luminescence,
                ymin = glyc_48.0I_mean_luminescence - glyc_48.0I_se, ymax = glyc_48.0I_mean_luminescence + glyc_48.0I_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.0I_data, aes(x = glyc_48.0I_conc, y = glyc_48.0I_mean_luminescence,
             color = label, shape = label), size = 4) +
  
    geom_errorbar(data = glyc_48.0J_data, aes(x = glyc_48.0J_conc, y = glyc_48.0J_mean_luminescence,
                ymin = glyc_48.0J_mean_luminescence - glyc_48.0J_se, ymax = glyc_48.0J_mean_luminescence + glyc_48.0J_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.0J_data, aes(x = glyc_48.0J_conc, y = glyc_48.0J_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_48.2A_data, aes(x = glyc_48.2A_conc, y = glyc_48.2A_mean_luminescence,
                ymin = glyc_48.2A_mean_luminescence - glyc_48.2A_se, ymax = glyc_48.2A_mean_luminescence + glyc_48.2A_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.2A_data, aes(x = glyc_48.2A_conc, y = glyc_48.2A_mean_luminescence,
             color = label, shape = label), size = 4) +
  
      geom_errorbar(data = glyc_48.2B_data, aes(x = glyc_48.2B_conc, y = glyc_48.2B_mean_luminescence,
                ymin = glyc_48.2B_mean_luminescence - glyc_48.2B_se, ymax = glyc_48.2B_mean_luminescence + glyc_48.2B_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.2B_data, aes(x = glyc_48.2B_conc, y = glyc_48.2B_mean_luminescence,
             color = label, shape = label), size = 4) +
  
    geom_errorbar(data = glyc_48.2C_data, aes(x = glyc_48.2C_conc, y = glyc_48.2C_mean_luminescence,
                ymin = glyc_48.2C_mean_luminescence - glyc_48.2C_se, ymax = glyc_48.2C_mean_luminescence + glyc_48.2C_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.2C_data, aes(x = glyc_48.2C_conc, y = glyc_48.2C_mean_luminescence,
             color = label, shape = label), size = 4) +

   geom_errorbar(data = glyc_48.2D_data, aes(x = glyc_48.2D_conc, y = glyc_48.2D_mean_luminescence,
                ymin = glyc_48.2D_mean_luminescence - glyc_48.2D_se, ymax = glyc_48.2D_mean_luminescence + glyc_48.2D_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.2D_data, aes(x = glyc_48.2D_conc, y = glyc_48.2D_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_48.2E_data, aes(x = glyc_48.2E_conc, y = glyc_48.2E_mean_luminescence,
                ymin = glyc_48.2E_mean_luminescence - glyc_48.2E_se, ymax = glyc_48.2E_mean_luminescence + glyc_48.2E_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.2E_data, aes(x = glyc_48.2E_conc, y = glyc_48.2E_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_48.2F_data, aes(x = glyc_48.2F_conc, y = glyc_48.2F_mean_luminescence,
                ymin = glyc_48.2F_mean_luminescence - glyc_48.2F_se, ymax = glyc_48.2F_mean_luminescence + glyc_48.2F_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.2F_data, aes(x = glyc_48.2F_conc, y = glyc_48.2F_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.2G_data, aes(x = glyc_48.2G_conc, y = glyc_48.2G_mean_luminescence,
                ymin = glyc_48.2G_mean_luminescence - glyc_48.2G_se, ymax = glyc_48.2G_mean_luminescence + glyc_48.2G_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.2G_data, aes(x = glyc_48.2G_conc, y = glyc_48.2G_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.2H_data, aes(x = glyc_48.2H_conc, y = glyc_48.2H_mean_luminescence,
                ymin = glyc_48.2H_mean_luminescence - glyc_48.2H_se, ymax = glyc_48.2H_mean_luminescence + glyc_48.2H_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.2H_data, aes(x = glyc_48.2H_conc, y = glyc_48.2H_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.2I_data, aes(x = glyc_48.2I_conc, y = glyc_48.2I_mean_luminescence,
                ymin = glyc_48.2I_mean_luminescence - glyc_48.2I_se, ymax = glyc_48.2I_mean_luminescence + glyc_48.2I_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.2I_data, aes(x = glyc_48.2I_conc, y = glyc_48.2I_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.2J_data, aes(x = glyc_48.2J_conc, y = glyc_48.2J_mean_luminescence,
                ymin = glyc_48.2J_mean_luminescence - glyc_48.2J_se, ymax = glyc_48.2J_mean_luminescence + glyc_48.2J_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.2J_data, aes(x = glyc_48.2J_conc, y = glyc_48.2J_mean_luminescence,
             color = label, shape = label), size = 4) +
  
    geom_errorbar(data = glyc_48.7A_data, aes(x = glyc_48.7A_conc, y = glyc_48.7A_mean_luminescence,
                ymin = glyc_48.7A_mean_luminescence - glyc_48.7A_se, ymax = glyc_48.7A_mean_luminescence + glyc_48.7A_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.7A_data, aes(x = glyc_48.7A_conc, y = glyc_48.7A_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_48.7B_data, aes(x = glyc_48.7B_conc, y = glyc_48.7B_mean_luminescence,
                ymin = glyc_48.7B_mean_luminescence - glyc_48.7B_se, ymax = glyc_48.7B_mean_luminescence + glyc_48.7B_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.7B_data, aes(x = glyc_48.7B_conc, y = glyc_48.7B_mean_luminescence,
             color = label, shape = label), size = 4) +
  
    geom_errorbar(data = glyc_48.7C_data, aes(x = glyc_48.7C_conc, y = glyc_48.7C_mean_luminescence,
                ymin = glyc_48.7C_mean_luminescence - glyc_48.7C_se, ymax = glyc_48.7C_mean_luminescence + glyc_48.7C_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.7C_data, aes(x = glyc_48.7C_conc, y = glyc_48.7C_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_48.7D_data, aes(x = glyc_48.7D_conc, y = glyc_48.7D_mean_luminescence,
                ymin = glyc_48.7D_mean_luminescence - glyc_48.7D_se, ymax = glyc_48.7D_mean_luminescence + glyc_48.7D_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.7D_data, aes(x = glyc_48.7D_conc, y = glyc_48.7D_mean_luminescence,
             color = label, shape = label), size = 4) +

    geom_errorbar(data = glyc_48.7E_data, aes(x = glyc_48.7E_conc, y = glyc_48.7E_mean_luminescence,
                ymin = glyc_48.7E_mean_luminescence - glyc_48.7E_se, ymax = glyc_48.7E_mean_luminescence + glyc_48.7E_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.7E_data, aes(x = glyc_48.7E_conc, y = glyc_48.7E_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  
    geom_errorbar(data = glyc_48.7F_data, aes(x = glyc_48.7F_conc, y = glyc_48.7F_mean_luminescence,
                ymin = glyc_48.7F_mean_luminescence - glyc_48.7F_se, ymax = glyc_48.7F_mean_luminescence + glyc_48.7F_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.7F_data, aes(x = glyc_48.7F_conc, y = glyc_48.7F_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.7G_data, aes(x = glyc_48.7G_conc, y = glyc_48.7G_mean_luminescence,
                ymin = glyc_48.7G_mean_luminescence - glyc_48.7G_se, ymax = glyc_48.7G_mean_luminescence + glyc_48.7G_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.7G_data, aes(x = glyc_48.7G_conc, y = glyc_48.7G_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.7H_data, aes(x = glyc_48.7H_conc, y = glyc_48.7H_mean_luminescence,
                ymin = glyc_48.7H_mean_luminescence - glyc_48.7H_se, ymax = glyc_48.7H_mean_luminescence + glyc_48.7H_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.7H_data, aes(x = glyc_48.7H_conc, y = glyc_48.7H_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.7I_data, aes(x = glyc_48.7I_conc, y = glyc_48.7I_mean_luminescence,
                ymin = glyc_48.7I_mean_luminescence - glyc_48.7I_se, ymax = glyc_48.7I_mean_luminescence + glyc_48.7I_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.7I_data, aes(x = glyc_48.7I_conc, y = glyc_48.7I_mean_luminescence,
             color = label, shape = label), size = 4) +
  
  geom_errorbar(data = glyc_48.7J_data, aes(x = glyc_48.7J_conc, y = glyc_48.7J_mean_luminescence,
                ymin = glyc_48.7J_mean_luminescence - glyc_48.7J_se, ymax = glyc_48.7J_mean_luminescence + glyc_48.7J_se,
                color = label), width = 0.1, linewidth = 1) +
  geom_point(data = glyc_48.7J_data, aes(x = glyc_48.7J_conc, y = glyc_48.7J_mean_luminescence,
             color = label, shape = label), size = 4) +

  scale_color_manual(name = "",
                     values = c("Standard Curve" = "steelblue",
                               "0.0Adf.20" = "darkorange",
                                "0.0Bdf.20" = "darkgreen",
                                "0.0Cdf.20" = "purple",
                                "0.0Ddf.20" = "brown",
                                "0.0Edf.20" = "green3",
                                "0.0Fdf.20" = "firebrick1",
                                "0.0Gdf.20" = "cyan",
                                "0.0Hdf.20" = "yellow3",
                                "0.0Idf.20" = "thistle3",
                                "0.0Jdf.20" = "azure4",
                                "48.0Adf.20" = "bisque4",
                                "48.0Bdf.20" = "chartreuse",
                                "48.0Cdf.20" = "chocolate4",
                                "48.0Ddf.20" = "darkslategray",
                                "48.0Edf.20" = "deeppink2",
                                "48.0Fdf.20" = "goldenrod3",
                                "48.0Gdf.20" = "hotpink3",
                                "48.0Hdf.20" = "mediumpurple2",
                                "48.0Idf.20" = "yellow",
                                "48.0Jdf.20" = "gray1",
                                "48.2Adf.20" = "darkcyan",
                                "48.2Bdf.20" = "darkorange",
                                "48.2Cdf.20" = "darkgreen",
                                "48.2Ddf.20" = "purple",
                                "48.2Edf.20" = "brown",
                                "48.2Fdf.20" = "green3",
                                "48.2Gdf.20" = "firebrick1",
                                "48.2Hdf.20" = "cyan",
                                "48.2Idf.20" = "yellow3",
                                "48.2Jdf.20" = "thistle3",
                                "48.7Adf.20" = "azure4",
                                "48.7Bdf.20" = "bisque4",
                                "48.7Cdf.20" = "chartreuse",
                                "48.7Ddf.20" = "chocolate4",
                                "48.7Edf.20" = "darkslategray",
                                "48.7Fdf.20" = "deeppink2",
                                "48.7Gdf.20" = "goldenrod3",
                                "48.7Hdf.20" = "hotpink3",
                                "48.7Idf.20" = "mediumpurple2",
                                "48.7Jdf.20" = "yellow"
                                )) +
  scale_shape_manual(name = "",
                     values = c("Standard Curve" = 16,
                                "0.0Adf.20" = 16,
                                "0.0Bdf.20" = 17,
                                "0.0Cdf.20" = 15,
                                "0.0Ddf.20" = 18,
                                "0.0Edf.20" = 8,
                                "0.0Fdf.20" = 4,
                                "0.0Gdf.20" = 5,
                                "0.0Hdf.20" = 0,
                                "0.0Idf.20" = 2,
                                "0.0Jdf.20" = 19,
                                "48.0Adf.20" = 20,
                                "48.0Bdf.20" = 17,
                                "48.0Cdf.20" = 15,
                                "48.0Ddf.20" = 18,
                                "48.0Edf.20" = 8,
                                "48.0Fdf.20" = 4,
                                "48.0Gdf.20" = 0,
                                "48.0Hdf.20" = 2,
                                "48.0Idf.20" = 19,
                                "48.0Jdf.20" = 20,
                                "48.2Adf.20" = 5, 
                                "48.2Bdf.20" = 0, 
                                "48.2Cdf.20" = 2, 
                                "48.2Ddf.20" = 19, 
                                "48.2Edf.20" = 20, 
                                "48.2Fdf.20" = 8, 
                                "48.2Gdf.20" = 4,
                                "48.2Hdf.20" = 5, 
                                "48.2Idf.20" = 17, 
                                "48.2Jdf.20" = 16, 
                                "48.7Adf.20" = 18, 
                                "48.7Bdf.20" = 5, 
                                "48.7Cdf.20" = 0, 
                                "48.7Ddf.20" = 19, 
                                "48.7Edf.20" = 20, 
                                "48.7Fdf.20" = 8, 
                                "48.7Gdf.20" = 4,
                                "48.7Hdf.20" = 2, 
                                "48.7Idf.20" = 16, 
                                "48.7Jdf.20" = 17
                                )) +
  scale_linetype_manual(name = "",
                        values = c("Std Curve Best Fit Line" = "dashed")) +
  annotate("label", x = max(glycogen_summary_data$glyc_concentration) * 0.75, 
           y = max(glycogen_summary_data$glyc_mean_luminescence) * 0.15,
           label = sprintf("y = %.2fx + %.2f\nR² = %.4f", glyc_slope, glyc_intercept, glyc_r_squared),
           size = 3.5, fontface = "bold", fill = "white", 
           color = "coral", label.padding = unit(0.3, "lines")) +
  labs(
    title = "glycogen Standard Curve",
    x = "glycogen Concentration (µg/µL)",
    y = "Luminescence",
    caption = "Error bars represent standard error of the mean (SEM)"
  ) +
  scale_x_continuous(breaks = glyc_concentrations) +
  theme_bw() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_line(linetype = "dashed", color = "grey70")
  )

# Display the plot
glyc_plot
glyc_0.0A_data
glyc_0.0B_data
glyc_0.0C_data
glyc_0.0D_data
glyc_0.0E_data
glyc_0.0F_data
glyc_0.0G_data
glyc_0.0H_data
glyc_0.0I_data
glyc_0.0J_data

glyc_48.0A_data
glyc_48.0B_data
glyc_48.0C_data
glyc_48.0D_data
glyc_48.0E_data
glyc_48.0F_data
glyc_48.0G_data
glyc_48.0H_data
glyc_48.0I_data
glyc_48.0J_data

glyc_48.2A_data
glyc_48.2B_data
glyc_48.2C_data
glyc_48.2D_data
glyc_48.2E_data
glyc_48.2F_data
glyc_48.2G_data
glyc_48.2H_data
glyc_48.2I_data
glyc_48.2J_data

glyc_48.7A_data
glyc_48.7B_data
glyc_48.7C_data
glyc_48.7D_data
glyc_48.7E_data
glyc_48.7F_data
glyc_48.7G_data
glyc_48.7H_data
glyc_48.7I_data
glyc_48.7J_data

# Print summary statistics
cat("glycogen Standard Curve Summary:\n")
cat(rep("=", 50), "\n", sep = "")
for (i in 1:nrow(glycogen_summary_data)) {
  cat(sprintf("Concentration: %g µg/µL\n", glycogen_summary_data$glyc_concentration[i]))
  cat(sprintf("  Mean Luminescence: %.2f\n",glycogen_summary_data$glyc_mean_luminescence[i]))
  cat(sprintf("  Standard Error: %.2f\n", glycogen_summary_data$glyc_se[i]))
  cat(sprintf("  CV%%: %.2f%%\n\n", glycogen_summary_data$glyc_cv[i]))
}
```

### Sample glycogen table

``` {r define tissue weights, include = TRUE}
#define tissue weights
glyc_0.0A_weight <- as.numeric(sample_weights[2,2])
glyc_0.0B_weight <- as.numeric(sample_weights[3,2])
glyc_0.0C_weight <- as.numeric(sample_weights[4,2])
glyc_0.0D_weight <- as.numeric(sample_weights[5,2])
glyc_0.0E_weight <- as.numeric(sample_weights[6,2])
glyc_0.0F_weight <- as.numeric(sample_weights[7,2])
glyc_0.0G_weight <- as.numeric(sample_weights[8,2])
glyc_0.0H_weight <- as.numeric(sample_weights[9,2])
glyc_0.0I_weight <- as.numeric(sample_weights[10,2])
glyc_0.0J_weight <- as.numeric(sample_weights[11,2])
glyc_48.0A_weight <- as.numeric(sample_weights[12,2])
glyc_48.0B_weight <- as.numeric(sample_weights[13,2])
glyc_48.0C_weight <- as.numeric(sample_weights[14,2])
glyc_48.0D_weight <- as.numeric(sample_weights[15,2])
glyc_48.0E_weight <- as.numeric(sample_weights[16,2])
glyc_48.0F_weight <- as.numeric(sample_weights[17,2])
glyc_48.0G_weight <- as.numeric(sample_weights[18,2])
glyc_48.0H_weight <- as.numeric(sample_weights[19,2])
glyc_48.0I_weight <- as.numeric(sample_weights[20,2])
glyc_48.0J_weight <- as.numeric(sample_weights[21,2])
glyc_48.2A_weight <- as.numeric(sample_weights[22,2])
glyc_48.2B_weight <- as.numeric(sample_weights[23,2])
glyc_48.2C_weight <- as.numeric(sample_weights[24,2])
glyc_48.2D_weight <- as.numeric(sample_weights[25,2])
glyc_48.2E_weight <- as.numeric(sample_weights[26,2])
glyc_48.2F_weight <- as.numeric(sample_weights[27,2])
#plate2
glyc_48.2G_weight <- as.numeric(sample_weights[28,2])
glyc_48.2H_weight <- as.numeric(sample_weights[29,2])
glyc_48.2I_weight <- as.numeric(sample_weights[30,2])
glyc_48.2J_weight <- as.numeric(sample_weights[31,2])
glyc_48.7A_weight <- as.numeric(sample_weights[32,2])
glyc_48.7B_weight <- as.numeric(sample_weights[33,2])
glyc_48.7C_weight <- as.numeric(sample_weights[34,2])
glyc_48.7D_weight <- as.numeric(sample_weights[35,2])
glyc_48.7E_weight <- as.numeric(sample_weights[36,2])
glyc_48.7F_weight <- as.numeric(sample_weights[37,2])
glyc_48.7G_weight <- as.numeric(sample_weights[38,2])
glyc_48.7H_weight <- as.numeric(sample_weights[39,2])
glyc_48.7I_weight <- as.numeric(sample_weights[40,2])
glyc_48.7J_weight <- as.numeric(sample_weights[41,2])
```

```{r, normalize calculated glycogen to sample weight, include = TRUE}
#normalize calculated glycogen to sample weight
#plate1
glyc_0.0A_mean_conc_normalized <- glyc_0.0A_mean_conc/glyc_0.0A_weight
glyc_0.0B_mean_conc_normalized <- glyc_0.0B_mean_conc/glyc_0.0B_weight
glyc_0.0C_mean_conc_normalized <- glyc_0.0C_mean_conc/glyc_0.0C_weight
glyc_0.0D_mean_conc_normalized <- glyc_0.0D_mean_conc/glyc_0.0D_weight
glyc_0.0E_mean_conc_normalized <- glyc_0.0E_mean_conc/glyc_0.0E_weight
glyc_0.0F_mean_conc_normalized <- glyc_0.0F_mean_conc/glyc_0.0F_weight
glyc_0.0G_mean_conc_normalized <- glyc_0.0G_mean_conc/glyc_0.0G_weight
glyc_0.0H_mean_conc_normalized <- glyc_0.0H_mean_conc/glyc_0.0H_weight
glyc_0.0I_mean_conc_normalized <- glyc_0.0I_mean_conc/glyc_0.0I_weight
glyc_0.0J_mean_conc_normalized <- glyc_0.0J_mean_conc/glyc_0.0J_weight
glyc_48.0A_mean_conc_normalized <- glyc_48.0A_mean_conc/glyc_48.0A_weight
glyc_48.0B_mean_conc_normalized <- glyc_48.0B_mean_conc/glyc_48.0B_weight
glyc_48.0C_mean_conc_normalized <- glyc_48.0C_mean_conc/glyc_48.0C_weight
glyc_48.0D_mean_conc_normalized <- glyc_48.0D_mean_conc/glyc_48.0D_weight
glyc_48.0E_mean_conc_normalized <- glyc_48.0E_mean_conc/glyc_48.0E_weight
glyc_48.0F_mean_conc_normalized <- glyc_48.0F_mean_conc/glyc_48.0F_weight
glyc_48.0G_mean_conc_normalized <- glyc_48.0G_mean_conc/glyc_48.0G_weight
glyc_48.0H_mean_conc_normalized <- glyc_48.0H_mean_conc/glyc_48.0H_weight
glyc_48.0I_mean_conc_normalized <- glyc_48.0I_mean_conc/glyc_48.0I_weight
glyc_48.0J_mean_conc_normalized <- glyc_48.0J_mean_conc/glyc_48.0J_weight
glyc_48.2A_mean_conc_normalized <- glyc_48.2A_mean_conc/glyc_48.2A_weight
glyc_48.2B_mean_conc_normalized <- glyc_48.2B_mean_conc/glyc_48.2B_weight
glyc_48.2C_mean_conc_normalized <- glyc_48.2C_mean_conc/glyc_48.2C_weight
glyc_48.2D_mean_conc_normalized <- glyc_48.2D_mean_conc/glyc_48.2D_weight
glyc_48.2E_mean_conc_normalized <- glyc_48.2E_mean_conc/glyc_48.2E_weight
glyc_48.2F_mean_conc_normalized <- glyc_48.2F_mean_conc/glyc_48.2F_weight
#plate2
glyc_48.2G_mean_conc_normalized <- glyc_48.2G_mean_conc/glyc_48.2G_weight
glyc_48.2H_mean_conc_normalized <- glyc_48.2H_mean_conc/glyc_48.2H_weight
glyc_48.2I_mean_conc_normalized <- glyc_48.2I_mean_conc/glyc_48.2I_weight
glyc_48.2J_mean_conc_normalized <- glyc_48.2J_mean_conc/glyc_48.2J_weight
glyc_48.7A_mean_conc_normalized <- glyc_48.7A_mean_conc/glyc_48.7A_weight
glyc_48.7B_mean_conc_normalized <- glyc_48.7B_mean_conc/glyc_48.7B_weight
glyc_48.7C_mean_conc_normalized <- glyc_48.7C_mean_conc/glyc_48.7C_weight
glyc_48.7D_mean_conc_normalized <- glyc_48.7D_mean_conc/glyc_48.7D_weight
glyc_48.7E_mean_conc_normalized <- glyc_48.7E_mean_conc/glyc_48.7E_weight
glyc_48.7F_mean_conc_normalized <- glyc_48.7F_mean_conc/glyc_48.7F_weight
glyc_48.7G_mean_conc_normalized <- glyc_48.7G_mean_conc/glyc_48.7G_weight
glyc_48.7H_mean_conc_normalized <- glyc_48.7H_mean_conc/glyc_48.7H_weight
glyc_48.7I_mean_conc_normalized <- glyc_48.7I_mean_conc/glyc_48.7I_weight
glyc_48.7J_mean_conc_normalized <- glyc_48.7J_mean_conc/glyc_48.7J_weight
```

Note: To normalize glycogen levels by dilution factor and tissue weight, the calculated glycogen is multiplied by (dilution factor/weight). This is reported as "Normalized glycogen (ug/uL/mg)"
```{r, sample glycogen table, include = TRUE}
#sample glycogen table
tab_glyc <- matrix(c(glyc_0.0A_mean_lum, glyc_0.0A_mean_conc, glyc_0.0A_dilution, glyc_0.0A_weight, glyc_0.0A_mean_conc*(glyc_0.0A_dilution/glyc_0.0A_weight), glyc_0.0B_mean_lum, glyc_0.0B_mean_conc, glyc_0.0B_dilution, glyc_0.0B_weight, glyc_0.0B_mean_conc*(glyc_0.0B_dilution/glyc_0.0B_weight), 
glyc_0.0C_mean_lum, glyc_0.0C_mean_conc, glyc_0.0C_dilution, glyc_0.0C_weight, glyc_0.0C_mean_conc*(glyc_0.0C_dilution/glyc_0.0C_weight), 
glyc_0.0D_mean_lum, glyc_0.0D_mean_conc, glyc_0.0D_dilution, glyc_0.0D_weight, glyc_0.0D_mean_conc*(glyc_0.0D_dilution/glyc_0.0D_weight), 
glyc_0.0E_mean_lum, glyc_0.0E_mean_conc, glyc_0.0E_dilution, glyc_0.0E_weight, glyc_0.0E_mean_conc*(glyc_0.0E_dilution/glyc_0.0E_weight), 
glyc_0.0F_mean_lum, glyc_0.0F_mean_conc, glyc_0.0F_dilution, glyc_0.0F_weight, glyc_0.0F_mean_conc*(glyc_0.0F_dilution/glyc_0.0F_weight), 
glyc_0.0G_mean_lum, glyc_0.0G_mean_conc, glyc_0.0G_dilution, glyc_0.0G_weight, glyc_0.0G_mean_conc*(glyc_0.0G_dilution/glyc_0.0G_weight), 
glyc_0.0H_mean_lum, glyc_0.0H_mean_conc, glyc_0.0H_dilution, glyc_0.0H_weight, glyc_0.0H_mean_conc*(glyc_0.0H_dilution/glyc_0.0H_weight), 
glyc_0.0I_mean_lum, glyc_0.0I_mean_conc, glyc_0.0I_dilution, glyc_0.0I_weight, glyc_0.0I_mean_conc*(glyc_0.0I_dilution/glyc_0.0I_weight), 
glyc_0.0J_mean_lum, glyc_0.0J_mean_conc, glyc_0.0J_dilution, glyc_0.0J_weight, glyc_0.0J_mean_conc*(glyc_0.0J_dilution/glyc_0.0J_weight), 

glyc_48.0A_mean_lum, glyc_48.0A_mean_conc, glyc_48.0A_dilution, glyc_48.0A_weight, glyc_48.0A_mean_conc*(glyc_48.0A_dilution/glyc_48.0A_weight), glyc_48.0B_mean_lum, glyc_48.0B_mean_conc, glyc_48.0B_dilution, glyc_48.0B_weight, glyc_48.0B_mean_conc*(glyc_48.0B_dilution/glyc_48.0B_weight), 
glyc_48.0C_mean_lum, glyc_48.0C_mean_conc, glyc_48.0C_dilution, glyc_48.0C_weight, glyc_48.0C_mean_conc*(glyc_48.0C_dilution/glyc_48.0C_weight), 
glyc_48.0D_mean_lum, glyc_48.0D_mean_conc, glyc_48.0D_dilution, glyc_48.0D_weight, glyc_48.0D_mean_conc*(glyc_48.0D_dilution/glyc_48.0D_weight), 
glyc_48.0E_mean_lum, glyc_48.0E_mean_conc, glyc_48.0E_dilution, glyc_48.0E_weight, glyc_48.0E_mean_conc*(glyc_48.0E_dilution/glyc_48.0E_weight), 
glyc_48.0F_mean_lum, glyc_48.0F_mean_conc, glyc_48.0F_dilution, glyc_48.0F_weight, glyc_48.0F_mean_conc*(glyc_48.0F_dilution/glyc_48.0F_weight), 
glyc_48.0G_mean_lum, glyc_48.0G_mean_conc, glyc_48.0G_dilution, glyc_48.0G_weight, glyc_48.0G_mean_conc*(glyc_48.0G_dilution/glyc_48.0G_weight), 
glyc_48.0H_mean_lum, glyc_48.0H_mean_conc, glyc_48.0H_dilution, glyc_48.0H_weight, glyc_48.0H_mean_conc*(glyc_48.0H_dilution/glyc_48.0H_weight), 
glyc_48.0I_mean_lum, glyc_48.0I_mean_conc, glyc_48.0I_dilution, glyc_48.0I_weight, glyc_48.0I_mean_conc*(glyc_48.0I_dilution/glyc_48.0I_weight), 
glyc_48.0J_mean_lum, glyc_48.0J_mean_conc, glyc_48.0J_dilution, glyc_48.0J_weight, glyc_48.0J_mean_conc*(glyc_48.0J_dilution/glyc_48.0J_weight), 

glyc_48.2A_mean_lum, glyc_48.2A_mean_conc, glyc_48.2A_dilution, glyc_48.2A_weight, glyc_48.2A_mean_conc*(glyc_48.2A_dilution/glyc_48.2A_weight), glyc_48.2B_mean_lum, glyc_48.2B_mean_conc, glyc_48.2B_dilution, glyc_48.2B_weight, glyc_48.2B_mean_conc*(glyc_48.2B_dilution/glyc_48.2B_weight), 
glyc_48.2C_mean_lum, glyc_48.2C_mean_conc, glyc_48.2C_dilution, glyc_48.2C_weight, glyc_48.2C_mean_conc*(glyc_48.2C_dilution/glyc_48.2C_weight), 
glyc_48.2D_mean_lum, glyc_48.2D_mean_conc, glyc_48.2D_dilution, glyc_48.2D_weight, glyc_48.2D_mean_conc*(glyc_48.2D_dilution/glyc_48.2D_weight), 
glyc_48.2E_mean_lum, glyc_48.2E_mean_conc, glyc_48.2E_dilution, glyc_48.2E_weight, glyc_48.2E_mean_conc*(glyc_48.2E_dilution/glyc_48.2E_weight), 
glyc_48.2F_mean_lum, glyc_48.2F_mean_conc, glyc_48.2F_dilution, glyc_48.2F_weight, glyc_48.2F_mean_conc*(glyc_48.2F_dilution/glyc_48.2F_weight), 
glyc_48.2G_mean_lum, glyc_48.2G_mean_conc, glyc_48.2G_dilution, glyc_48.2G_weight, glyc_48.2G_mean_conc*(glyc_48.2G_dilution/glyc_48.2G_weight), 
glyc_48.2H_mean_lum, glyc_48.2H_mean_conc, glyc_48.2H_dilution, glyc_48.2H_weight, glyc_48.2H_mean_conc*(glyc_48.2H_dilution/glyc_48.2H_weight), 
glyc_48.2I_mean_lum, glyc_48.2I_mean_conc, glyc_48.2I_dilution, glyc_48.2I_weight, glyc_48.2I_mean_conc*(glyc_48.2I_dilution/glyc_48.2I_weight), 
glyc_48.2J_mean_lum, glyc_48.2J_mean_conc, glyc_48.2J_dilution, glyc_48.2J_weight, glyc_48.2J_mean_conc*(glyc_48.2J_dilution/glyc_48.2J_weight), 

glyc_48.7A_mean_lum, glyc_48.7A_mean_conc, glyc_48.7A_dilution, glyc_48.7A_weight, glyc_48.7A_mean_conc*(glyc_48.7A_dilution/glyc_48.7A_weight), glyc_48.7B_mean_lum, glyc_48.7B_mean_conc, glyc_48.7B_dilution, glyc_48.7B_weight, glyc_48.7B_mean_conc*(glyc_48.7B_dilution/glyc_48.7B_weight), 
glyc_48.7C_mean_lum, glyc_48.7C_mean_conc, glyc_48.7C_dilution, glyc_48.7C_weight, glyc_48.7C_mean_conc*(glyc_48.7C_dilution/glyc_48.7C_weight), 
glyc_48.7D_mean_lum, glyc_48.7D_mean_conc, glyc_48.7D_dilution, glyc_48.7D_weight, glyc_48.7D_mean_conc*(glyc_48.7D_dilution/glyc_48.7D_weight), 
glyc_48.7E_mean_lum, glyc_48.7E_mean_conc, glyc_48.7E_dilution, glyc_48.7E_weight, glyc_48.7E_mean_conc*(glyc_48.7E_dilution/glyc_48.7E_weight), 
glyc_48.7F_mean_lum, glyc_48.7F_mean_conc, glyc_48.7F_dilution, glyc_48.7F_weight, glyc_48.7F_mean_conc*(glyc_48.7F_dilution/glyc_48.7F_weight), 
glyc_48.7G_mean_lum, glyc_48.7G_mean_conc, glyc_48.7G_dilution, glyc_48.7G_weight, glyc_48.7G_mean_conc*(glyc_48.7G_dilution/glyc_48.7G_weight), 
glyc_48.7H_mean_lum, glyc_48.7H_mean_conc, glyc_48.7H_dilution, glyc_48.7H_weight, glyc_48.7H_mean_conc*(glyc_48.7H_dilution/glyc_48.7H_weight), 
glyc_48.7I_mean_lum, glyc_48.7I_mean_conc, glyc_48.7I_dilution, glyc_48.7I_weight, glyc_48.7I_mean_conc*(glyc_48.7I_dilution/glyc_48.7I_weight), 
glyc_48.7J_mean_lum, glyc_48.7J_mean_conc, glyc_48.7J_dilution, glyc_48.7J_weight, glyc_48.7J_mean_conc*(glyc_48.7J_dilution/glyc_48.7J_weight)
), ncol=5, byrow=TRUE)
              
colnames(tab_glyc) <- c('Luminescence','Calculated Glycogen (ug/uL)','Dilution factor', 'Weight (mg)','Normalized glycogen (ug/uL/mg)')
rownames(tab_glyc) <- c('0.0A**','0.0B**','0.0C**','0.0D**','0.0E**','0.0F**','0.0G**','0.0H**','0.0I**','0.0J**','48.0A**','48.0B**','48.0C**','48.0D**','48.0E**','48.0F**','48.0G**','48.0H**','48.0I**','48.0J**','48.2A**','48.2B**','48.2C**','48.2D**','48.2E**','48.2F**','48.2G**','48.2H**','48.2I**','48.2J**','48.7A**','48.7B**','48.7C**','48.7D**','48.7E**','48.7F**','48.7G**','48.7H**','48.7I**','48.7J**')
tab_glyc <- as.table(tab_glyc)
tab_glyc
```