---
title: "Gen5-20260730-mgig-fams1_9-glycogen-glo"
author: "Sam White"
date: "2026-07-30"
output:
  github_document:
    toc: true
    number_sections: true
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
---

# BACKGROUND

Glycogen quantification of _Magallana gigas_ (Pacific oyster) ctenidia homogenates
using the [Glycogen-Glo Assay (Promega)](https://github.com/RobertsLab/resources/blob/master/protocols/Commercial_Protocols/Promega_Glycogen_Glo_Assay.pdf)
(GitHub; PDF), read on 2026-07-30.

The experiment compares two USDA oyster families (**family 1** and **family 9**)
across two temperature exposures (**ambient** and **36&deg;C**), with
**8 individuals per family per temperature** (32 samples total). Each sample was
run in **technical triplicate** at a **1:25 dilution**.

Two plates were read on the same instrument/protocol:

- **Plate 01**: all 32 samples, triplicate, filling all 96 wells (no standards).
- **Plate 02**: glycogen standard curve (20, 2, 0.2, 0.02, 0 &micro;g/mL) plus a
  buffer-only negative control, all in triplicate.

Because the standards live on a separate plate from the samples, a single standard
curve from plate 02 is applied to all samples on plate 01. Both plates were read
with the same protocol (`luminesence-96wells-1.0s-glycogen_glo.prt`, gain 135,
1.00 s integration, extended dynamic range) back-to-back, so this is reasonable,
but it does mean plate-to-plate signal drift is not controlled for. This is
flagged in the interpretation.

## Sample naming

Per [`../data/raw_luminescence/README.md`](../data/raw_luminescence/README.md),
plate layout entries follow:

- `<sample>-<assay_type>-<tissue_weight>-df.<dilution_factor>`

Here `<sample>` itself is composite:

- `<family>_<individual>_<temperature>`

E.g. `1_05_36C-glyc-25.7-df.25` is family 1, individual 05, 36&deg;C exposure,
glycogen assay, 25.7 mg of ctenidia tissue, diluted 1:25.

Standards follow `STD-<assay_type>-<concentration>`, e.g. `STD-glyc-20`
(20 &micro;g/mL glycogen); the buffer-only well is `NEG-glyc`.

## Important note(s)



1. **Units.** Promega's glycogen standards and linear range are specified in
   **&micro;g/mL** (linear to 20 &micro;g/mL), and the layout labels
   (`STD-glyc-20`) are on that scale. Earlier scripts in this repo labeled the
   same quantity "&micro;g/&micro;L"; that label is incorrect by a factor of
   1000 and is corrected here.

2. **Tissue weights** are taken from the plate layout labels rather than a
   separate weights CSV, since they are already encoded there for this experiment.

# SETUP

## Libraries
```{r setup, include=TRUE}
library(knitr)
library(ggplot2)
library(dplyr)
library(tidyr)

knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,         # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = "",        # Prevents appending '##' to beginning of lines in code output
  results = 'hold'     # Holds output so it's all printed together after code chunk
)
```


## Output directory
```{r output-dir}
# Output directory for this analysis (matches this file's name, per ../code/README.md)
output_dir <- "../output/Gen5-20260730-mgig-fams1_9-glycogen-glo"
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

cat("--- output_dir: destination for all figures and tables ---\n")
str(output_dir)
```

# DATA

Data are read from the local repo (`../data/raw_luminescence/`) so this document
renders before/after the files are pushed to GitHub. The commented-out URLs are
the remote equivalents.

```{r data, include=TRUE}
data_dir <- "../data/raw_luminescence"

# plate 1 - all 32 samples, triplicate, 1:25 dilution

plate_layout1 <- read.csv(file.path(data_dir, "layout-Gen5-20260730-mgig-fams1_9-plate-01.csv"),
                          header = FALSE, stringsAsFactors = FALSE)
raw_luminescence1 <- read.csv(file.path(data_dir, "raw_lum-Gen5-20260730-mgig-fams1_9-plate-01.csv"),
                              header = FALSE)

# plate 2 - glycogen standard curve + negative control, triplicates
plate_layout2 <- read.csv(file.path(data_dir, "layout-Gen5-20260730-mgig-fams1_9-plate-02.csv"),
                          header = FALSE, stringsAsFactors = FALSE)
raw_luminescence2 <- read.csv(file.path(data_dir, "raw_lum-Gen5-20260730-mgig-fams1_9-plate-02.csv"),
                              header = FALSE)

cat("Plate layout 1 (samples):\n\n")
str(plate_layout1)
cat("\nRaw luminescence 1 (samples):\n\n")
str(raw_luminescence1)
cat("\nPlate layout 2 (standards):\n\n")
str(plate_layout2)
cat("\nRaw luminescence 2 (standards):\n\n")
str(raw_luminescence2)
```

## Reshape plates to long format

Rather than defining one variable per sample, both plates are melted to a tidy
well-level data frame and the sample metadata is parsed out of the layout labels.
This keeps the code independent of how many samples are on the plate.

```{r plate-to-long, include=TRUE}
# Convert a paired layout/luminescence plate into one row per well
plate_to_long <- function(layout, luminescence, plate_label) {
  n_row <- 8
  n_col <- 12
  out <- expand.grid(plate_row_idx = 1:n_row, plate_col = 1:n_col)
  out$plate       <- plate_label
  out$plate_row   <- LETTERS[out$plate_row_idx]
  out$well        <- sprintf("%s%02d", out$plate_row, out$plate_col)
  out$label       <- trimws(as.character(mapply(function(i, j) layout[i, j],
                                                out$plate_row_idx, out$plate_col)))
  out$luminescence <- as.numeric(mapply(function(i, j) luminescence[i, j],
                                        out$plate_row_idx, out$plate_col))
  out <- out[order(out$plate_row_idx, out$plate_col), ]
  out[out$label != "" & !is.na(out$label), ]
}

plate1_long <- plate_to_long(plate_layout1, raw_luminescence1, "plate-01")
plate2_long <- plate_to_long(plate_layout2, raw_luminescence2, "plate-02")


cat("--- plate_to_long(): layout + luminescence -> one row per well ---\n\n")
str(args(plate_to_long))

cat("\n--- plate1_long: one row per occupied well, sample plate ---\n\n")
str(plate1_long)

cat("\n--- plate2_long: one row per occupied well, standards plate ---\n\n")
str(plate2_long)
```

# STANDARD CURVES

## Glycogen Standard Curve

### Extract luminescence data

Plate 02, row A holds the four non-zero glycogen standards
(20, 2, 0.2, 0.02 &micro;g/mL) in triplicate; row B columns 1-3 hold the
`STD-glyc-0` (zero-glycogen) standard and row C columns 1-3 hold the `NEG-glyc`
buffer-only negative control.

```{r glycogen standard data, include=TRUE}
standards <- plate2_long %>%
  filter(grepl("^STD-glyc-", label)) %>%
  mutate(glyc_concentration = as.numeric(sub("^STD-glyc-", "", label)))

negative_control <- plate2_long %>% filter(grepl("^NEG-glyc", label))

cat("Standard concentrations (ug/mL):",
    paste(sort(unique(standards$glyc_concentration)), collapse = ", "), "\n")
cat("Replicates per standard:",
    paste(unique(table(standards$glyc_concentration)), collapse = ", "), "\n\n")

cat("Negative control (buffer only) luminescence:",
    paste(negative_control$luminescence, collapse = ", "),
    "| mean =", round(mean(negative_control$luminescence), 1), "\n")
cat("Zero glycogen standard luminescence:  mean =",
    round(mean(standards$luminescence[standards$glyc_concentration == 0]), 1), "\n")

cat("\n--- standards: standard curve wells with parsed concentration ---\n")
str(standards)
cat("\n--- negative_control: buffer-only wells ---\n")
str(negative_control)
```

The buffer-only negative control and the zero-glycogen standard are
indistinguishable, as expected -- both report the reagent/instrument background.

### Glycogen standard curve summary statistics and linear regression

Following the convention used in previous analyses in this repo, the regression is
fit to the **mean** luminescence at each standard concentration. The fit to all
individual replicate wells is also reported, since it is the more honest estimate
of the curve's uncertainty.

```{r glycogen summary statistics and linear regression, include=TRUE}
glycogen_summary_data <- standards %>%
  group_by(glyc_concentration) %>%
  summarise(
    glyc_mean_luminescence = mean(luminescence),
    glyc_sd                = sd(luminescence),
    glyc_se                = sd(luminescence) / sqrt(n()),
    glyc_cv                = 100 * sd(luminescence) / mean(luminescence),
    glyc_n                 = n(),
    .groups = "drop"
  ) %>%
  arrange(glyc_concentration)

# Regression on the per-concentration means (repo convention)
lm_model      <- lm(glyc_mean_luminescence ~ glyc_concentration, data = glycogen_summary_data)
glyc_slope     <- coef(lm_model)[2]
glyc_intercept <- coef(lm_model)[1]
glyc_r_squared <- summary(lm_model)$r.squared

# Regression on all individual replicate wells (for comparison)
lm_model_reps    <- lm(luminescence ~ glyc_concentration, data = standards)
glyc_r2_reps     <- summary(lm_model_reps)$r.squared

glyc_conc_min_nonzero <- min(glycogen_summary_data$glyc_concentration[
  glycogen_summary_data$glyc_concentration > 0])
glyc_conc_max <- max(glycogen_summary_data$glyc_concentration)

kable(glycogen_summary_data,
      digits = c(2, 1, 1, 1, 2, 0),
      col.names = c("Glycogen (ug/mL)", "Mean luminescence", "SD", "SEM", "CV (%)", "n"),
      caption = "Glycogen standard curve summary statistics")

cat("\nFit to concentration means:      y =", sprintf("%.1f", glyc_slope), "x +",
    sprintf("%.1f", glyc_intercept), " R^2 =", sprintf("%.4f", glyc_r_squared), "\n")
cat("Fit to individual replicates:    R^2 =", sprintf("%.4f", glyc_r2_reps), "\n")
cat("Quantifiable range (standards):", glyc_conc_min_nonzero, "-", glyc_conc_max, "ug/mL\n")

cat("\n--- glycogen_summary_data: per-concentration summary statistics ---\n")
str(glycogen_summary_data)
cat("\n--- lm_model: regression on per-concentration means ---\n")
str(lm_model, max.level = 1, give.attr = FALSE)
cat("\n--- lm_model_reps: regression on all individual replicate wells ---\n")
str(lm_model_reps, max.level = 1, give.attr = FALSE)
```

### Extract sample data and calculate glycogen levels

Sample labels are parsed into family, individual, temperature, tissue weight, and
dilution factor. The delimiter inconsistencies noted above (`1_02-36C`,
`9-07-36C`, `9-08-36C`) are normalized here.

```{r glycogen samples, include=TRUE}
samples_wells <- plate1_long %>%
  # Normalize `-` to `_` in the <family>-<individual>-<temperature> prefix
  mutate(label_clean = sub("^([19])[-_]([0-9]{2})[-_](ambient|36C)-", "\\1_\\2_\\3-", label)) %>%
  mutate(
    sample_id  = sub("-glyc-.*$", "", label_clean),
    family     = sub("^([19])_.*$", "\\1", label_clean),
    individual = sub("^[19]_([0-9]{2})_.*$", "\\1", label_clean),
    temperature = sub("^[19]_[0-9]{2}_([^-]+)-.*$", "\\1", label_clean),
    weight_mg  = as.numeric(sub("^.*-glyc-([0-9.]+)-df\\..*$", "\\1", label_clean)),
    dilution   = as.numeric(sub("^.*-df\\.", "", label_clean))
  ) %>%
  mutate(
    family      = factor(paste("Family", family), levels = c("Family 1", "Family 9")),
    temperature = factor(temperature, levels = c("ambient", "36C"))
  )

# Fail loudly rather than silently dropping malformed labels
stopifnot(!any(is.na(samples_wells$weight_mg)),
          !any(is.na(samples_wells$dilution)),
          !any(is.na(samples_wells$temperature)))

cat("Wells parsed:", nrow(samples_wells),
    "| unique samples:", length(unique(samples_wells$sample_id)), "\n")
cat("Dilution factor(s) used:", paste(unique(samples_wells$dilution), collapse = ", "), "\n\n")
cat("Design (wells per group):\n")
print(table(samples_wells$family, samples_wells$temperature))
cat("\nTissue weight range (mg):",
    paste(range(samples_wells$weight_mg), collapse = " - "), "\n")

cat("\n--- samples_wells: one row per sample well, metadata parsed from label ---\n")
str(samples_wells)
```

Per-sample values are then computed by averaging the technical triplicate and
back-calculating through the standard curve:

1. **Well concentration** = `(mean luminescence - intercept) / slope` &rarr; &micro;g/mL in the assayed (diluted) well.
2. **Homogenate concentration** = well concentration &times; dilution factor (25) &rarr; &micro;g/mL in the undiluted homogenate.
3. **Normalized glycogen** = homogenate concentration / tissue weight &rarr; &micro;g/mL per mg tissue.

Note on step 3: because homogenization volume is not recorded in these data files,
normalized values are reported as &micro;g glycogen per mL homogenate per mg
tissue. If the standard 1.0 mL homogenization volume was used (750 &micro;L
PBS/0.3N HCl + 250 &micro;L Tris, per [`../glycogen-promega-notes.md`](../glycogen-promega-notes.md)),
these values are numerically equal to **&micro;g glycogen / mg tissue**. All
comparisons below are unaffected as long as the volume was constant across samples.

```{r calculate sample glycogen, include=TRUE}
sample_glycogen <- samples_wells %>%
  group_by(sample_id, family, individual, temperature, weight_mg, dilution) %>%
  summarise(
    n_reps        = n(),
    mean_lum      = mean(luminescence),
    sd_lum        = sd(luminescence),
    cv_lum        = 100 * sd(luminescence) / mean(luminescence),
    .groups = "drop"
  ) %>%
  mutate(
    well_conc_ug_mL       = (mean_lum - glyc_intercept) / glyc_slope,
    homogenate_conc_ug_mL = well_conc_ug_mL * dilution,
    norm_glycogen         = homogenate_conc_ug_mL / weight_mg,
    in_std_range          = well_conc_ug_mL >= glyc_conc_min_nonzero &
                            well_conc_ug_mL <= glyc_conc_max
  ) %>%
  arrange(family, temperature, individual)

cat("Samples quantified:", nrow(sample_glycogen), "\n")

cat("\n--- sample_glycogen: one row per individual, triplicate averaged and back-calculated ---\n")
str(sample_glycogen)
```

### Plot glycogen standard curve with sample points

```{r plot-glycogen-standard-curve, fig.width = 10, fig.height = 6, include=TRUE}
std_curve_plot <- ggplot(glycogen_summary_data,
                         aes(x = glyc_concentration, y = glyc_mean_luminescence)) +
  geom_smooth(aes(linetype = "Std curve best fit line"), method = "lm", se = FALSE,
              color = "coral", linewidth = 1) +
  geom_errorbar(aes(ymin = glyc_mean_luminescence - glyc_se,
                    ymax = glyc_mean_luminescence + glyc_se),
                width = 0.3, linewidth = 1, color = "darkblue") +
  geom_point(aes(color = "Glycogen standard"), shape = 16, size = 4) +
  geom_point(data = sample_glycogen,
             aes(x = well_conc_ug_mL, y = mean_lum, color = family),
             shape = 17, size = 2.5, alpha = 0.85) +
  geom_point(data = filter(sample_glycogen, !in_std_range),
             aes(x = well_conc_ug_mL, y = mean_lum, shape = "Extrapolated"),
             size = 4, stroke = 1.1, color = "red", fill = NA) +
  annotate("rect", xmin = glyc_conc_max, xmax = Inf, ymin = -Inf, ymax = Inf,
           fill = "red", alpha = 0.06) +
  scale_color_manual(name = "",
                     breaks = c("Glycogen standard", "Family 1", "Family 9"),
                     values = c("Glycogen standard" = "steelblue",
                                "Family 1"          = "#1b7837",
                                "Family 9"          = "#762a83")) +
  scale_shape_manual(name = "", values = c("Extrapolated" = 21)) +
  scale_linetype_manual(name = "", values = c("Std curve best fit line" = "dashed")) +
  guides(color = guide_legend(override.aes = list(shape = c(16, 17, 17), size = 3),
                              order = 1),
         shape = guide_legend(order = 2), linetype = guide_legend(order = 3)) +
  annotate("label",
           x = glyc_conc_max * 0.45, y = max(glycogen_summary_data$glyc_mean_luminescence) * 0.95,
           label = sprintf("y = %.1fx + %.1f\nR2 = %.4f", glyc_slope, glyc_intercept, glyc_r_squared),
           size = 3.5, fontface = "bold", fill = "white", color = "coral",
           label.padding = unit(0.3, "lines")) +
  labs(
    title = "Glycogen standard curve with sample well concentrations",
    subtitle = "Shaded region is above the top standard (20 ug/mL): samples there are extrapolated",
    x = "Glycogen concentration (ug/mL)",
    y = "Luminescence",
    caption = "Standard curve error bars are SEM of triplicate wells (n = 3)"
  ) +
  theme_bw() +
  theme(
    plot.title  = element_text(size = 14, face = "bold"),
    axis.title  = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_line(linetype = "dashed", color = "grey70")
  )

cat("--- std_curve_plot: ggplot object structure ---\n")
summary(std_curve_plot)

ggsave(file.path(output_dir, "glycogen_standard_curve.png"), std_curve_plot,
       width = 10, height = 6, dpi = 300)

std_curve_plot
```

The linear fit is anchored almost entirely by the 20 &micro;g/mL standard. On a
log-log scale it is clear that the three lowest standards (0.02, 0.2, 2
&micro;g/mL) sit essentially on the baseline of this linear model, i.e. the curve
has little resolving power at the bottom of the range:

```{r plot-glycogen-standard-curve-log, fig.width = 8, fig.height = 5.5, include=TRUE}
std_curve_log_plot <- ggplot(filter(glycogen_summary_data, glyc_concentration > 0),
                             aes(x = glyc_concentration, y = glyc_mean_luminescence)) +
  geom_hline(aes(yintercept = mean(negative_control$luminescence),
                 linetype = "Background (buffer only)"), color = "grey40") +
  geom_line(color = "steelblue", linewidth = 0.8) +
  geom_errorbar(aes(ymin = glyc_mean_luminescence - glyc_se,
                    ymax = glyc_mean_luminescence + glyc_se),
                width = 0.08, color = "darkblue") +
  geom_point(color = "steelblue", size = 3.5) +
  scale_x_log10() +
  scale_y_log10() +
  scale_linetype_manual(name = "", values = c("Background (buffer only)" = "dotted")) +
  labs(
    title = "Glycogen standard curve (log-log)",
    x = "Glycogen concentration (ug/mL, log scale)",
    y = "Luminescence (log scale)"
  ) +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 11, face = "bold"))

cat("--- std_curve_log_plot: ggplot object structure ---\n")
summary(std_curve_log_plot)

ggsave(file.path(output_dir, "glycogen_standard_curve_loglog.png"), std_curve_log_plot,
       width = 8, height = 5.5, dpi = 300)

std_curve_log_plot
```

# QUALITY CONTROL

## Technical replicate variability

```{r replicate-cv, include=TRUE}
cat("Coefficient of variation across technical triplicates (%):\n")
print(round(summary(sample_glycogen$cv_lum), 2))

cat("\nSamples with CV > 15%:\n")
high_cv <- sample_glycogen %>% filter(cv_lum > 15) %>%
  select(sample_id, mean_lum, sd_lum, cv_lum)
if (nrow(high_cv) == 0) {
  cat("  none\n")
} else {
  print(as.data.frame(high_cv), row.names = FALSE, digits = 4)
}

cat("\n--- high_cv: samples exceeding the 15% triplicate CV threshold ---\n\n")
str(high_cv)
```

## Samples outside the standard curve range

```{r out-of-range, include=TRUE}
oor <- sample_glycogen %>%
  filter(!in_std_range) %>%
  select(sample_id, family, temperature, mean_lum, well_conc_ug_mL,
         weight_mg, norm_glycogen) %>%
  arrange(desc(well_conc_ug_mL))

cat("Samples with well concentrations outside", glyc_conc_min_nonzero, "-",
    glyc_conc_max, "ug/mL:", nrow(oor), "of", nrow(sample_glycogen), "\n\n")
kable(oor, digits = c(0, 0, 0, 0, 2, 1, 2),
      col.names = c("Sample", "Family", "Temp", "Mean lum.", "Well conc. (ug/mL)",
                    "Tissue (mg)", "Normalized glycogen"),
      caption = "Samples requiring extrapolation beyond the top standard")

cat("\n--- Samples outside the standard curve range ---\n\n")
str(oor)
```

All four are **above** the 20 &micro;g/mL top standard, so their concentrations
are extrapolated rather than interpolated. Notably, `1_08_ambient` reads
~201,000 RLU (3.3&times; the top standard) with a very tight triplicate CV
(0.2%), which is the signature of a well near the top of the detector's usable
range. The other three (`1_06_ambient`, `1_05_36C`, `9_05_ambient`) are more
modestly above range (~70,000-97,000 RLU, roughly 1.5-1.6&times; the top
standard). These samples should be re-run at a higher dilution (e.g. 1:50-1:100
for `1_08_ambient`) before the values are treated as quantitative. Group
comparisons below are therefore repeated with these samples excluded as a
sensitivity check.

## Plate 01 luminescence map

A well-position map checks for edge effects or systematic gradients that would
confound the family/temperature contrasts. In this design, family and
temperature are both encoded as two-row blocks (rows A/B = family 1 ambient,
C/D = family 1 36&deg;C, E/F = family 9 ambient, G/H = family 9 36&deg;C), while
column position encodes only the individual (with each individual's triplicate
occupying 3 adjacent columns). Any row-wise artifact would therefore alias
directly onto the family/temperature contrasts of interest; a column-wise
artifact would not, since individuals are distributed independently of family
and temperature group.

```{r plate-heatmap, fig.width = 10, fig.height = 5, include=TRUE}
plate_map_plot <- ggplot(samples_wells,
                         aes(x = factor(plate_col), y = factor(plate_row, levels = rev(LETTERS[1:8])))) +
  geom_tile(aes(fill = log10(luminescence)), color = "white", linewidth = 0.5) +
  geom_text(aes(label = format(luminescence, big.mark = ",")), size = 2.4, color = "grey15") +
  scale_fill_viridis_c(option = "magma", name = "log10(RLU)") +
  labs(
    title = "Plate 01 raw luminescence by well",
    subtitle = "Row-pairs A/B, C/D, E/F, G/H = family1-amb, family1-36C, family9-amb, family9-36C; columns = individual (triplicate)",
    x = "Plate column", y = "Plate row"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        panel.grid = element_blank())

cat("--- plate_map_plot: ggplot object structure ---\n\n")
summary(plate_map_plot)

ggsave(file.path(output_dir, "plate01_luminescence_map.png"), plate_map_plot,
       width = 10, height = 5, dpi = 300)

plate_map_plot
```


## Relationship between tissue weight and signal

If normalization by tissue weight is working, normalized glycogen should show no
strong trend against input tissue weight.

```{r weight-vs-signal, fig.width = 10, fig.height = 4.5, include=TRUE}
weight_check <- sample_glycogen %>%
  select(sample_id, family, temperature, weight_mg,
         homogenate_conc_ug_mL, norm_glycogen) %>%
  pivot_longer(c(homogenate_conc_ug_mL, norm_glycogen),
               names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(metric,
                         homogenate_conc_ug_mL = "Homogenate glycogen (ug/mL)",
                         norm_glycogen         = "Normalized glycogen (ug/mL/mg tissue)"))

weight_plot <- ggplot(weight_check, aes(x = weight_mg, y = value)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey50", linetype = "dashed") +
  geom_point(aes(color = family, shape = temperature), size = 2.8) +
  facet_wrap(~ metric, scales = "free_y") +
  scale_color_manual(values = c("Family 1" = "#1b7837", "Family 9" = "#762a83"),
                     name = "") +
  scale_shape_manual(values = c("ambient" = 16, "36C" = 17), name = "") +
  scale_y_log10() +
  labs(title = "Input tissue weight vs. glycogen signal",
       x = "Ctenidia tissue weight (mg)", y = NULL) +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        strip.text = element_text(face = "bold"))

cat("--- weight_check: long-format data for the two-panel weight check ---\n\n")
str(weight_check)
cat("\n--- weight_plot: ggplot object structure ---\n\n")
summary(weight_plot)

ggsave(file.path(output_dir, "tissue_weight_vs_glycogen.png"), weight_plot,
       width = 10, height = 4.5, dpi = 300)

weight_plot

cat("Spearman correlation, tissue weight vs. homogenate glycogen concentration:\n\n")
print(cor.test(sample_glycogen$weight_mg, sample_glycogen$homogenate_conc_ug_mL,
               method = "spearman"))
cat("\nSpearman correlation, tissue weight vs. weight-normalized glycogen:\n\n")
print(cor.test(sample_glycogen$weight_mg, sample_glycogen$norm_glycogen,
               method = "spearman"))
```

**Interpretation.** Homogenate glycogen concentration (before dividing by
weight) is strongly and significantly correlated with input tissue weight
(rho = 0.66, p < 0.0001). This is expected, not a sign of a problem: every
sample was homogenized and diluted by the same fixed protocol regardless of
starting tissue mass, so a heavier piece of ctenidia simply contributes
proportionally more total glycogen into the same assay volume, mechanically
inflating homogenate-level concentration. It reflects a feature of the
protocol (fixed homogenization/dilution volume, variable input mass), not
biological variation in glycogen content per se.

Dividing by tissue weight is meant to remove exactly this scaling artifact.
After normalization, the correlation with weight drops sharply, to a weak and
non-significant residual (rho = 0.30, p = 0.10) -- roughly a threefold
reduction in the strength of association, and no longer distinguishable from
no relationship at this sample size (n = 32). This is the expected signature
of normalization working as intended: most, though perhaps not quite all, of
the weight-driven scaling has been removed, leaving normalized glycogen
approximately independent of how much tissue went into each homogenate. The
residual rho = 0.30 is small enough that it does not indicate normalization
failure, but it is also not exactly zero, so a mild weight-related trend in
the normalized values cannot be fully excluded from this dataset alone
(possible contributors include measurement noise at low tissue weights, where
small absolute pipetting/weighing errors have larger proportional effects, or
a true small biological effect such as larger individuals tending to have
either richer or more dilute ctenidia glycogen stores). Because none of the
family or temperature comparisons above showed a significant effect, this
residual weight association is not currently a confound for the reported
group comparisons, but it is worth re-checking if a future dataset shows a
significant family or temperature effect that also happens to coincide with a
weight imbalance between groups.

# SAMPLE GLYCOGEN TABLE

```{r sample glycogen table, include=TRUE}
sample_table <- sample_glycogen %>%
  transmute(
    Sample        = sample_id,
    Family        = family,
    Temperature   = temperature,
    Individual    = individual,
    `Tissue (mg)` = weight_mg,
    `Dilution`    = dilution,
    `Mean lum.`   = round(mean_lum, 0),
    `CV (%)`      = round(cv_lum, 1),
    `Well glycogen (ug/mL)`        = round(well_conc_ug_mL, 3),
    `Homogenate glycogen (ug/mL)`  = round(homogenate_conc_ug_mL, 2),
    `Normalized glycogen (ug/mL/mg)` = round(norm_glycogen, 2),
    `In std range` = ifelse(in_std_range, "yes", "NO - extrapolated")
  )

kable(sample_table, caption = "Per-sample glycogen quantification (mean of technical triplicate)")

write.csv(sample_table, file.path(output_dir, "sample_glycogen_results.csv"), row.names = FALSE)

cat("\n--- sample_table: formatted per-sample results written to CSV ---\n\n")
str(sample_table)
```

# FAMILY AND TEMPERATURE COMPARISONS

## Group summary statistics

```{r group-summary, include=TRUE}
group_summary <- sample_glycogen %>%
  group_by(Family = family, Temperature = temperature) %>%
  summarise(
    n      = n(),
    mean   = mean(norm_glycogen),
    sd     = sd(norm_glycogen),
    se     = sd(norm_glycogen) / sqrt(n()),
    median = median(norm_glycogen),
    min    = min(norm_glycogen),
    max    = max(norm_glycogen),
    .groups = "drop"
  )

kable(group_summary, digits = 2,
      caption = "Normalized glycogen (ug/mL per mg tissue) by family and temperature")

write.csv(group_summary, file.path(output_dir, "group_summary_stats.csv"), row.names = FALSE)

cat("\nSame summary excluding the 4 extrapolated samples:\n\n")
group_summary_inrange <- sample_glycogen %>%
  filter(in_std_range) %>%
  group_by(Family = family, Temperature = temperature) %>%
  summarise(n = n(), mean = mean(norm_glycogen), sd = sd(norm_glycogen),
            median = median(norm_glycogen), .groups = "drop")
print(as.data.frame(group_summary_inrange), row.names = FALSE, digits = 4)

cat("\n--- group_summary: family x temperature summary, all 32 samples ---\n\n")
str(group_summary)
cat("\n--- group_summary_inrange: same summary, extrapolated samples excluded ---\n\n")
str(group_summary_inrange)
```

## Plot: glycogen by family and temperature

```{r plot-family-temp, fig.width = 9, fig.height = 6, include=TRUE}
group_plot <- ggplot(sample_glycogen,
                     aes(x = temperature, y = norm_glycogen, fill = family)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.35, width = 0.6,
               position = position_dodge(width = 0.75)) +
  geom_point(aes(color = family, shape = in_std_range, group = family),
             position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.75,
                                             seed = 42),
             size = 2.8) +
  scale_fill_manual(values = c("Family 1" = "#1b7837", "Family 9" = "#762a83"), name = "") +
  scale_color_manual(values = c("Family 1" = "#1b7837", "Family 9" = "#762a83"), name = "") +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 4),
                     labels = c("TRUE" = "within standard curve",
                                "FALSE" = "extrapolated"),
                     name = "") +
  labs(
    title = "Ctenidia glycogen content by USDA family and temperature exposure",
    subtitle = "Each point is one individual (mean of technical triplicate); n = 8 per group",
    x = "Temperature exposure",
    y = "Normalized glycogen (ug/mL per mg tissue)",
    caption = "x = value extrapolated above the 20 ug/mL top standard"
  ) +
  theme_bw() +
  theme(
    plot.title  = element_text(size = 14, face = "bold"),
    axis.title  = element_text(size = 12, face = "bold"),
    legend.position = "bottom"
  )

cat("--- group_plot: ggplot object structure ---\n\n")
summary(group_plot)

ggsave(file.path(output_dir, "glycogen_by_family_temperature.png"), group_plot,
       width = 9, height = 6, dpi = 300)

group_plot
```

Note: this boxplot uses a linear y-axis. Given the ~30-fold range of values
and the four extrapolated high points, the linear scale visually compresses
the lower/mid-range groups toward the bottom of the plot; a log-scale version
was reviewed and rejected in favor of this one, but keep the compression
in mind when comparing box heights/positions across groups by eye -- the
statistical tests below are unaffected either way, since they are run on
log10-transformed values regardless of which scale is plotted.

## Plot: family x temperature interaction

The boxplot above shows the full per-individual distribution, which makes it
hard to visually judge whether the family gap widens, narrows, or reverses
with temperature -- exactly the question an interaction test asks. The plot
below isolates that comparison: group means +/- SEM only, connected by family,
so that the two line slopes are a visual proxy for the interaction effect size
being tested in the ANOVA (Family x Temperature, p = 0.268 on
log10-transformed values; see Statistical tests below). Near-parallel lines
indicate no interaction; converging, diverging, or crossing lines would
indicate the families respond differently to temperature.

```{r plot-interaction, fig.width = 8, fig.height = 5.5, include=TRUE}
interaction_data <- sample_glycogen %>%
  group_by(family, temperature) %>%
  summarise(mean_glyc = mean(norm_glycogen),
            se_glyc   = sd(norm_glycogen) / sqrt(n()),
            .groups = "drop")

interaction_plot <- ggplot(interaction_data,
                           aes(x = temperature, y = mean_glyc,
                               color = family, group = family)) +
  geom_line(linewidth = 1.1) +
  geom_errorbar(aes(ymin = mean_glyc - se_glyc, ymax = mean_glyc + se_glyc),
                width = 0.08, linewidth = 0.9) +
  geom_point(size = 4) +
  scale_color_manual(values = c("Family 1" = "#1b7837", "Family 9" = "#762a83"), name = "") +
  labs(
    title = "Family x temperature interaction",
    subtitle = "Group means +/- SEM, linear scale (n = 8 per group)",
    x = "Temperature exposure",
    y = "Mean normalized glycogen (ug/mL/mg tissue)"
  ) +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"),
        axis.title = element_text(size = 11, face = "bold"))

cat("--- interaction_data: group means and SEM plotted below ---\n")
str(interaction_data)
cat("\n--- interaction_plot: ggplot object structure ---\n")
summary(interaction_plot)

ggsave(file.path(output_dir, "glycogen_family_temperature_interaction.png"), interaction_plot,
       width = 8, height = 5.5, dpi = 300)

interaction_plot
```

## Statistical tests

Normalized glycogen spans roughly two orders of magnitude and is strongly
right-skewed, so residuals from a two-way ANOVA on the raw scale are non-normal.
A log10 transform fixes this, so the log-scale ANOVA is the primary test and the
raw-scale ANOVA plus distribution-free rank tests are reported alongside it.

```{r assumption-checks, include=TRUE}
aov_raw <- aov(norm_glycogen ~ family * temperature, data = sample_glycogen)
aov_log <- aov(log10(norm_glycogen) ~ family * temperature, data = sample_glycogen)

cat("Shapiro-Wilk on ANOVA residuals:\n")
cat("  raw scale:   p =", signif(shapiro.test(residuals(aov_raw))$p.value, 4), "\n")
cat("  log10 scale: p =", signif(shapiro.test(residuals(aov_log))$p.value, 4), "\n")
cat("\nBartlett test of homogeneity of variance (log10 scale): p =",
    signif(bartlett.test(log10(norm_glycogen) ~ interaction(family, temperature),
                         data = sample_glycogen)$p.value, 4), "\n")

cat("\n--- aov_raw: two-way ANOVA model, raw scale ---\n")
str(aov_raw, max.level = 1, give.attr = FALSE)
cat("\n--- aov_log: two-way ANOVA model, log10 scale (primary test) ---\n")
str(aov_log, max.level = 1, give.attr = FALSE)
```

```{r two-way-anova, include=TRUE}
cat("=== Two-way ANOVA on log10(normalized glycogen) [primary] ===\n")
print(summary(aov_log))

cat("\n=== Two-way ANOVA on raw normalized glycogen [for reference] ===\n")
print(summary(aov_raw))
```

```{r nonparametric-tests, include=TRUE}
cat("=== Wilcoxon rank-sum: family 1 vs family 9 (temperatures pooled) ===\n")
print(wilcox.test(norm_glycogen ~ family, data = sample_glycogen))

cat("=== Wilcoxon rank-sum: ambient vs 36C (families pooled) ===\n")
print(wilcox.test(norm_glycogen ~ temperature, data = sample_glycogen))

cat("=== Stratified comparisons ===\n")
strat <- rbind(
  data.frame(comparison = "ambient vs 36C, within family 1",
             p = wilcox.test(norm_glycogen ~ temperature,
                             data = filter(sample_glycogen, family == "Family 1"))$p.value),
  data.frame(comparison = "ambient vs 36C, within family 9",
             p = wilcox.test(norm_glycogen ~ temperature,
                             data = filter(sample_glycogen, family == "Family 9"))$p.value),
  data.frame(comparison = "family 1 vs 9, within ambient",
             p = wilcox.test(norm_glycogen ~ family,
                             data = filter(sample_glycogen, temperature == "ambient"))$p.value),
  data.frame(comparison = "family 1 vs 9, within 36C",
             p = wilcox.test(norm_glycogen ~ family,
                             data = filter(sample_glycogen, temperature == "36C"))$p.value)
)
strat$p_holm <- p.adjust(strat$p, method = "holm")
kable(strat, digits = 4,
      col.names = c("Comparison", "p (Wilcoxon)", "p (Holm-adjusted)"),
      caption = "Stratified pairwise comparisons of normalized glycogen")

cat("\n--- strat: stratified Wilcoxon comparisons with Holm-adjusted p-values ---\n")
str(strat)
```

### Sensitivity analysis: excluding extrapolated samples

Three of the four samples that read above the top standard are family 1
(`1_08_ambient`, `1_06_ambient`, `1_05_36C`) and one is family 9
(`9_05_ambient`), so it is important to confirm they are not driving any
apparent family or temperature effect.

```{r sensitivity-analysis, include=TRUE}
in_range_only <- filter(sample_glycogen, in_std_range)

cat("n retained:", nrow(in_range_only), "of", nrow(sample_glycogen), "\n\n")
cat("=== Two-way ANOVA on log10(normalized glycogen), in-range samples only ===\n")
print(summary(aov(log10(norm_glycogen) ~ family * temperature, data = in_range_only)))

cat("\n=== Wilcoxon, family 1 vs 9 (in-range only, temperatures pooled) ===\n")
print(wilcox.test(norm_glycogen ~ family, data = in_range_only))

cat("\n--- in_range_only: sample_glycogen with extrapolated samples removed ---\n")
str(in_range_only)
```

```{r save-workspace-summary, include=TRUE}
# Persist the tidy well-level data alongside the summaries for downstream reuse
write.csv(samples_wells %>%
            select(plate, well, sample_id, family, temperature, individual,
                   weight_mg, dilution, luminescence),
          file.path(output_dir, "well_level_luminescence.csv"), row.names = FALSE)

write.csv(glycogen_summary_data,
          file.path(output_dir, "standard_curve_summary.csv"), row.names = FALSE)

cat("Files written to", output_dir, ":\n")
cat(paste(" -", list.files(output_dir)), sep = "\n")
```

# SUMMARY

**Standard curve.** The four-point glycogen curve plus zero standard is linear
across the full range when fit to concentration means (R&sup2; = 0.9999), but the fit
is dominated by the single 20 &micro;g/mL point; fit to individual replicate wells
R&sup2; drops to 0.986, and the 20 &micro;g/mL standard itself has a 12.6% CV. The
three lowest standards (0.02-2 &micro;g/mL) are barely resolved above the
buffer-only background (~910 RLU). Adding intermediate standards (e.g. 5 and 10
&micro;g/mL) and tightening replicate pipetting at the top standard would make
back-calculation in the mid-range considerably more trustworthy.

**Technical quality.** Triplicate CVs are good overall (median 4.9%, mean 5.5%);
only `1_04_36C` exceeds 15% (30.1%, driven by one low replicate). The plate map
shows no obvious row-wise gradient across the family/temperature row-blocks
(A/B, C/D, E/F, G/H), so the row-block design does not appear to be confounded
with plate position.

**Range.** Four samples read above the top standard and are extrapolated:
`1_08_ambient`, `1_06_ambient`, `1_05_36C`, `9_05_ambient`. `1_08_ambient` in
particular (~201,000 RLU, 3.3&times; the top standard) should be re-run at a
higher dilution; the other three are more modestly above range (~94,000-97,000
RLU). The extrapolated samples are not concentrated in a single family.

Recommended re-assay dilutions for these four samples, targeting a well-level
concentration near the middle of the quantifiable curve (~10 &micro;g/mL, i.e.
roughly the midpoint between the 2 and 20 &micro;g/mL standards, corresponding
to ~10 &micro;g/mL/mg tissue-normalized once re-scaled by weight) rather than
just clearing the 20 &micro;g/mL ceiling:

| Sample | Mean luminescence (RLU, current 1:25) | Current dilution | Recommended total dilution | Approx. fold increase |
|---|---:|---:|---:|---:|
| `1_08_ambient` | ~201,000 | 1:25 | ~1:150 | ~6&times; |
| `1_06_ambient` | ~97,300  | 1:25 | ~1:80  | ~3&times; |
| `1_05_36C`     | ~94,100  | 1:25 | ~1:75  | ~3&times; |
| `9_05_ambient` | ~70,200  | 1:25 | ~1:60  | ~2&times; |

These are rough targets based on each sample's current homogenate-level
glycogen estimate (extrapolated off the standard curve) and are intended as a
starting point for the next dilution series, not a precise prescription --
actual optimal dilution should be confirmed empirically, since the
extrapolated homogenate concentrations that these estimates are based on are
themselves the values in question.

**Family effect.** Pooling temperatures, group means are 27.1 (family 1) vs
20.9 (family 9) &micro;g/mL/mg tissue. The effect is not significant in the
log-scale two-way ANOVA (F = 0.72,
p = 0.404) or by Wilcoxon (W = 150, p = 0.423). None of the four stratified
family/temperature contrasts reach significance either before or after Holm
correction (smallest unadjusted p = 0.195, within ambient).

**Temperature effect.** No detectable effect of 36&deg;C exposure. Neither the
main effect (log-scale ANOVA p = 0.936; Wilcoxon p = 0.867) nor the family
&times; temperature interaction (p = 0.268) reaches significance. Group means
by family and temperature are: family 1 ambient 31.7, family 1 36&deg;C 22.5,
family 9 ambient 19.8, family 9 36&deg;C 22.0 &micro;g/mL/mg tissue -- no
consistent within-family shift with temperature, and no interaction pattern
suggestive of family-specific heat response.

**Weight relationship.** Homogenate glycogen concentration correlates with
input tissue weight (Spearman rho = 0.66, p < 0.0001), as expected since larger
samples were run at the same fixed dilution rather than volume-adjusted --
this is a protocol artifact of fixed homogenization volume, not a biological
signal. Weight-normalized glycogen shows a much weaker, non-significant
residual correlation with weight (rho = 0.30, p = 0.10), consistent with the
normalization step removing most, though perhaps not all, of that scaling
effect. The residual association is small enough not to indicate a
normalization failure, and since neither family nor temperature showed a
significant effect in this dataset, it is not currently confounding those
comparisons -- but it would be worth revisiting if a future run shows a
significant group difference alongside a weight imbalance between groups.

**Caveats.**

1. None of the family, temperature, or family &times; temperature effects
   are statistically significant at n = 8 per group -- excluding the four
   extrapolated samples changes this picture only slightly (log-scale ANOVA
   family p = 0.652, temperature p = 0.467, interaction p = 0.292, n = 28),
   so the qualitative conclusion (no detectable family or temperature effect)
   is robust to that exclusion.
2. Standards were on a separate plate from the samples, so plate-to-plate signal
   offset is uncorrected. Including a short standard curve on each sample plate
   would remove this assumption.
3. Homogenization volume is not recorded in these files; normalized values are
   therefore per mL of homogenate per mg tissue. Group comparisons are unaffected
   provided the volume was constant, but absolute glycogen per mg tissue cannot be
   stated without it.
