---
title: "Gen5-20260804-mgig-fams1_9-glycogen-glo-combined"
author: "Sam White"
date: "2026-08-04"
output:
  github_document:
    toc: true
    number_sections: true
  bookdown::html_document2:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    code_download: true
---

# BACKGROUND

Combined glycogen analysis of *Magallana gigas* (Pacific oyster) ctenidia
samples from USDA families 1 and 9, comparing ambient vs. 36&deg;C temperature
exposure, using the [Glycogen-Glo Assay (Promega)](https://github.com/RobertsLab/resources/blob/master/protocols/Commercial_Protocols/Promega_Glycogen_Glo_Assay.pdf)
(GitHub; PDF).

This document merges results from two assay dates:

- **2026-07-30** ([`Gen5-20260730-mgig-fams1_9-glycogen-glo`](Gen5-20260730-mgig-fams1_9-glycogen-glo.md)):
  the full 32-sample experiment (all family &times; temperature combinations),
  samples run at 1:25 dilution on one plate, standard curve + negative
  control run on a separate plate.

- **2026-08-03** ([`Gen5-20260803-mgig-fams1_9-glycogen-glo`](Gen5-20260803-mgig-fams1_9-glycogen-glo.md)):
  a targeted re-assay of 5 samples at 1:200 dilution (~8&times; higher),
  run alongside a fresh, same-plate standard curve. Four of these five
  (`1_08_ambient`, `1_06_ambient`, `1_05_36C`, `9_05_ambient`) had read above
  the top standard (20 &micro;g/mL) in the original run and required
  extrapolation; the fifth (`1_04_36C`) was in range originally but had an
  elevated technical-replicate CV (30.1%).

**Reconciliation rule applied here:** for each of the 32 samples, this
document selects **whichever available measurement falls within its own
run's standard-curve range** (interpolated rather than extrapolated),
preferring the more recent run as a tiebreaker if more than one candidate
qualifies. In practice this means:

- `1_08_ambient`, `1_06_ambient`, `1_05_36C`, `9_05_ambient`: the
  **2026-08-03 (1:200)** value is used, since it is in-range and the
  2026-07-30 value was extrapolated.
- `1_04_36C`: the **2026-07-30 (1:25)** value is used per instruction --
  its 2026-08-03 (1:200) re-assay read below the standard curve's floor
  (negative back-calculated concentration, unusable), so the original
  in-range-but-high-CV value is retained here rather than the unusable
  re-assay value.
- All other 28 samples: only the 2026-07-30 value exists and is used as-is.

Full detail on the re-assay rationale, the standard-curve QC checks, and the
outlier/replicate-exclusion analysis for `1_04_36C` (a Grubbs test found no
statistical basis for dropping a replicate, and a general "exclude any
replicate outside 1 SD of the triplicate mean" rule was tested and rejected
as a mathematical inevitability for n=3, not a data-quality signal) are
documented in the two source analyses linked above and are not repeated here.

## Sample naming

Per [`../data/raw_luminescence/README.md`](../data/raw_luminescence/README.md),
plate layout entries follow `<sample>-<assay_type>-<tissue_weight>-df.<dilution_factor>`,
with sample tokens formatted as `<family>_<individual>_<temperature>`, e.g.
`1_06_ambient-glyc-11.0-df.25`. Standards are labeled `STD-glyc-<concentration>`
and the negative (buffer-only) control `NEG-glyc`.

# SETUP

## Libraries
```{r setup, include=TRUE}
library(knitr)
library(ggplot2)
library(dplyr)
library(tidyr)

knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = TRUE,         # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  comment = "",        # Prevents appending '##' to beginning of lines in code output
  results = 'hold'     # Holds output so it's all printed together after code chunk
)
```

## Set output directory
```{r output-dir}
output_dir <- "../output/Gen5-20260804-mgig-fams1_9-glycogen-glo-combined"
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
```

# DATA IMPORT

Raw plate-reader exports (comma-separated, no header, one row per plate row
`A`-`H` and one column per plate column `1`-`12`) are read separately for
each plate, for both assay dates.

```{r read-raw-data}
data_dir <- "../data/raw_luminescence"

# 2026-07-30: plate-01 (32 samples, triplicate wells), plate-02 (standard curve + neg. control)
layout_0730_plate01 <- read.csv(file.path(data_dir, "layout-Gen5-20260730-mgig-fams1_9-plate-01.csv"),
                                header = FALSE, stringsAsFactors = FALSE)
raw_0730_plate01    <- read.csv(file.path(data_dir, "raw_lum-Gen5-20260730-mgig-fams1_9-plate-01.csv"),
                                header = FALSE)
layout_0730_plate02 <- read.csv(file.path(data_dir, "layout-Gen5-20260730-mgig-fams1_9-plate-02.csv"),
                                header = FALSE, stringsAsFactors = FALSE)
raw_0730_plate02    <- read.csv(file.path(data_dir, "raw_lum-Gen5-20260730-mgig-fams1_9-plate-02.csv"),
                                header = FALSE)

# 2026-08-03: single plate (5 re-assayed samples + fresh standard curve + neg. control)
layout_0803_plate01 <- read.csv(file.path(data_dir, "layout-Gen5-20260803-mgig-fams1_9-plate-01.csv"),
                                header = FALSE, stringsAsFactors = FALSE)
raw_0803_plate01    <- read.csv(file.path(data_dir, "raw_lum-Gen5-20260803-mgig-fams1_9-plate-01.csv"),
                                header = FALSE)

cat("2026-07-30 plate-01 dims:", dim(layout_0730_plate01), "\n")
cat("2026-07-30 plate-02 dims:", dim(layout_0730_plate02), "\n")
cat("2026-08-03 plate-01 dims:", dim(layout_0803_plate01), "\n")
```

# PLATE RESHAPING

Converts each plate's row/column layout into long format (one row per well),
dropping any well without a sample/standard label (empty wells).

```{r plate-to-long-function}
plate_to_long <- function(layout, luminescence, plate_label) {
  n_row <- 8
  n_col <- 12
  out <- expand.grid(plate_row_idx = 1:n_row, plate_col = 1:n_col)
  out$plate        <- plate_label
  out$plate_row    <- LETTERS[out$plate_row_idx]
  out$well         <- sprintf("%s%02d", out$plate_row, out$plate_col)
  out$label        <- trimws(as.character(mapply(function(i, j) layout[i, j],
                                                  out$plate_row_idx, out$plate_col)))
  out$luminescence <- as.numeric(mapply(function(i, j) luminescence[i, j],
                                        out$plate_row_idx, out$plate_col))
  out <- out[order(out$plate_row_idx, out$plate_col), ]
  out[out$label != "" & !is.na(out$label), ]
}
```

```{r reshape-plates}
plate_0730_01_long <- plate_to_long(layout_0730_plate01, raw_0730_plate01, "0730-plate-01")
plate_0730_02_long <- plate_to_long(layout_0730_plate02, raw_0730_plate02, "0730-plate-02")
plate_0803_01_long <- plate_to_long(layout_0803_plate01, raw_0803_plate01, "0803-plate-01")

cat("--- plate_0730_01_long ---\n")
str(plate_0730_01_long)
cat("\n--- plate_0730_02_long ---\n")
str(plate_0730_02_long)
cat("\n--- plate_0803_01_long ---\n")
str(plate_0803_01_long)
```

# STANDARD CURVES

Each assay date has its own standard curve (glycogen standards run alongside
the samples), fit independently -- samples are always back-calculated against
the curve from *their own* assay date/plate, never a curve from a different
run.

```{r extract-standards}
standards_0730 <- plate_0730_02_long %>%
  filter(grepl("^STD-glyc-", label)) %>%
  mutate(glyc_concentration = as.numeric(sub("^STD-glyc-", "", label)),
         run = "2026-07-30")

negative_0730 <- plate_0730_02_long %>%
  filter(grepl("^NEG-glyc", label)) %>%
  mutate(run = "2026-07-30")

standards_0803 <- plate_0803_01_long %>%
  filter(grepl("^STD-glyc-", label)) %>%
  mutate(glyc_concentration = as.numeric(sub("^STD-glyc-", "", label)),
         run = "2026-08-03")

negative_0803 <- plate_0803_01_long %>%
  filter(grepl("^NEG-glyc", label)) %>%
  mutate(run = "2026-08-03")

cat("--- standards_0730 ---\n")
str(standards_0730)
cat("\n--- standards_0803 ---\n")
str(standards_0803)

cat("\nNegative control, 2026-07-30 (n =", nrow(negative_0730), "): mean luminescence =",
    round(mean(negative_0730$luminescence), 1), "\n")
cat("Negative control, 2026-08-03 (n =", nrow(negative_0803), "): mean luminescence =",
    round(mean(negative_0803$luminescence), 1), "\n")
```

```{r fit-standard-curves}
fit_curve <- function(standards) {
  summ <- standards %>%
    group_by(glyc_concentration) %>%
    summarise(
      glyc_mean_luminescence = mean(luminescence),
      glyc_sd                = sd(luminescence),
      glyc_se                = sd(luminescence) / sqrt(n()),
      glyc_cv                = 100 * sd(luminescence) / mean(luminescence),
      glyc_n                 = n(),
      .groups = "drop"
    ) %>%
    arrange(glyc_concentration)

  lm_model <- lm(glyc_mean_luminescence ~ glyc_concentration, data = summ)

  list(
    summary   = summ,
    model     = lm_model,
    slope     = unname(coef(lm_model)[2]),
    intercept = unname(coef(lm_model)[1]),
    r2        = summary(lm_model)$r.squared,
    conc_min  = min(summ$glyc_concentration[summ$glyc_concentration > 0]),
    conc_max  = max(summ$glyc_concentration),
    fit_min   = min(summ$glyc_concentration),
    fit_max   = max(summ$glyc_concentration)
  )
}

curve_0730 <- fit_curve(standards_0730)
curve_0803 <- fit_curve(standards_0803)

cat("--- 2026-07-30 standard curve ---\n")
str(curve_0730$summary)
cat("slope:", curve_0730$slope, " intercept:", curve_0730$intercept,
    " R2:", round(curve_0730$r2, 5), "\n")
cat("quantifiable range:", curve_0730$conc_min, "-", curve_0730$conc_max, "ug/mL\n\n")

cat("--- 2026-08-03 standard curve ---\n")
str(curve_0803$summary)
cat("slope:", curve_0803$slope, " intercept:", curve_0803$intercept,
    " R2:", round(curve_0803$r2, 5), "\n")
cat("quantifiable range:", curve_0803$conc_min, "-", curve_0803$conc_max, "ug/mL\n")
```

## Plot standard curves
```{r plot-standard-curves, fig.width=10, fig.height=5}
curve_plot_data <- bind_rows(
  curve_0730$summary %>% mutate(run = "2026-07-30"),
  curve_0803$summary %>% mutate(run = "2026-08-03")
)
curve_fit_lines <- bind_rows(
  data.frame(run = "2026-07-30", slope = curve_0730$slope, intercept = curve_0730$intercept,
             fit_min = curve_0730$fit_min, fit_max = curve_0730$fit_max, r2 = curve_0730$r2),
  data.frame(run = "2026-08-03", slope = curve_0803$slope, intercept = curve_0803$intercept,
             fit_min = curve_0803$fit_min, fit_max = curve_0803$fit_max, r2 = curve_0803$r2)
) %>%
  mutate(
    y_start   = intercept + slope * fit_min,
    y_end     = intercept + slope * fit_max,
    r2_label  = paste0("R\u00b2 = ", sprintf("%.4f", r2))
  )

cat("--- curve_fit_lines: trend-line endpoints + R2 per run ---\n")
str(curve_fit_lines)

standard_curve_plot <- ggplot(curve_plot_data, aes(x = glyc_concentration, y = glyc_mean_luminescence)) +
  geom_segment(data = curve_fit_lines,
               aes(x = fit_min, xend = fit_max, y = y_start, yend = y_end),
               color = "steelblue", linewidth = 0.6, inherit.aes = FALSE) +
  geom_errorbar(aes(ymin = glyc_mean_luminescence - glyc_se, ymax = glyc_mean_luminescence + glyc_se),
                width = 0.3, color = "grey40") +
  geom_point(size = 2.4, color = "steelblue") +
  geom_text(data = curve_fit_lines, aes(x = -Inf, y = Inf, label = r2_label),
            hjust = -0.15, vjust = 1.8, inherit.aes = FALSE, size = 3.6, fontface = "bold") +
  facet_wrap(~ run, scales = "free_y") +
  labs(x = "Glycogen standard concentration (\u00b5g/mL)",
       y = "Mean luminescence (RLU)",
       title = "Standard curves by assay date") +
  theme_bw(base_size = 12) +
  theme(strip.text = element_text(face = "bold"))

print(standard_curve_plot)
ggsave(file.path(output_dir, "standard_curves_both_runs.png"), standard_curve_plot,
       width = 10, height = 5, dpi = 300)
```

# SAMPLE LABEL PARSING

Sample labels are parsed into family, individual, temperature, tissue
weight, and dilution factor. Both runs' sample wells (excluding standards and
the negative control) use the same label grammar.

```{r parse-samples-function}
parse_samples <- function(plate_long) {
  plate_long %>%
    filter(!grepl("^STD-glyc-|^NEG-glyc", label)) %>%
    mutate(label_clean = sub("^([19])[-_]([0-9]{2})[-_](ambient|36C)-", "\\1_\\2_\\3-", label)) %>%
    mutate(
      sample_id   = sub("-glyc-.*$", "", label_clean),
      family      = sub("^([19])_.*$", "\\1", label_clean),
      individual  = sub("^[19]_([0-9]{2})_.*$", "\\1", label_clean),
      temperature = sub("^[19]_[0-9]{2}_([^-]+)-.*$", "\\1", label_clean),
      weight_mg   = as.numeric(sub("^.*-glyc-([0-9.]+)-df\\..*$", "\\1", label_clean)),
      dilution    = as.numeric(sub("^.*-df\\.", "", label_clean))
    ) %>%
    mutate(
      family      = factor(paste("Family", family), levels = c("Family 1", "Family 9")),
      temperature = factor(temperature, levels = c("ambient", "36C"))
    )
}
```

```{r parse-samples}
samples_0730 <- parse_samples(plate_0730_01_long)
samples_0803 <- parse_samples(plate_0803_01_long)

cat("--- samples_0730 ---\n")
str(samples_0730)
cat("\n--- samples_0803 ---\n")
str(samples_0803)

cat("\n2026-07-30: ", length(unique(samples_0730$sample_id)), "unique samples,",
    nrow(samples_0730), "wells (triplicate).\n")
cat("2026-08-03: ", length(unique(samples_0803$sample_id)), "unique samples,",
    nrow(samples_0803), "wells (triplicate).\n")
```

# PER-SAMPLE GLYCOGEN QUANTIFICATION

For each sample, technical-replicate wells are averaged, back-calculated
against that assay date's own standard curve, corrected for dilution factor
and tissue weight, and flagged for whether the well-level luminescence falls
within the quantifiable range of that date's standard curve.

```{r calc-glycogen-function}
calc_glycogen <- function(samples_wells, curve, run_label) {
  samples_wells %>%
    group_by(sample_id, family, individual, temperature, weight_mg, dilution) %>%
    summarise(
      n_reps   = n(),
      mean_lum = mean(luminescence),
      sd_lum   = sd(luminescence),
      cv_lum   = 100 * sd(luminescence) / mean(luminescence),
      .groups  = "drop"
    ) %>%
    mutate(
      run                   = run_label,
      well_conc_ug_mL       = (mean_lum - curve$intercept) / curve$slope,
      homogenate_conc_ug_mL = well_conc_ug_mL * dilution,
      norm_glycogen         = homogenate_conc_ug_mL / weight_mg,
      in_std_range          = well_conc_ug_mL >= curve$conc_min & well_conc_ug_mL <= curve$conc_max
    ) %>%
    arrange(family, temperature, individual)
}
```

```{r calc-glycogen}
glycogen_0730 <- calc_glycogen(samples_0730, curve_0730, "2026-07-30")
glycogen_0803 <- calc_glycogen(samples_0803, curve_0803, "2026-08-03")

cat("--- glycogen_0730 ---\n")
str(glycogen_0730)
cat("\n--- glycogen_0803 ---\n")
str(glycogen_0803)
```

# COEFFICIENT OF VARIATION (CV) CHECK

Technical-replicate CV is computed for every sample in both runs. Samples
exceeding 15% CV are **flagged, not excluded** -- per the analysis
requirements, high CV is a data-quality note carried alongside the result,
not a reason to drop a sample.

```{r cv-flagging}
all_glycogen_by_run <- bind_rows(glycogen_0730, glycogen_0803) %>%
  mutate(high_cv = cv_lum > 15)

cat("--- all_glycogen_by_run: every measurement from both runs, CV-flagged ---\n")
str(all_glycogen_by_run)

high_cv_samples <- all_glycogen_by_run %>%
  filter(high_cv) %>%
  select(sample_id, run, n_reps, mean_lum, cv_lum, norm_glycogen, in_std_range)

cat("\nSamples exceeding 15% technical-replicate CV (flagged, not excluded):\n")
kable(high_cv_samples, digits = 2)
```

# STANDARD CURVES WITH SAMPLES OVERLAID

Each run's samples are plotted against that run's own standard curve, on the
well-concentration scale, so it's visually clear which samples fall inside
vs. outside the curve's quantifiable range.

## Plot 20260730 Standard Curves with Samples

```{r plot-standards-with-samples-0730, fig.width=8, fig.height=5.5}
r2_label_0730 <- paste0("R\u00b2 = ", sprintf("%.4f", curve_0730$r2))

std_samples_0730 <- ggplot() +
  geom_ribbon(data = data.frame(x = c(curve_0730$conc_min, curve_0730$conc_max)),
              aes(x = x, ymin = -Inf, ymax = Inf), fill = "grey85", alpha = 0.4,
              inherit.aes = FALSE) +
  geom_segment(aes(x = curve_0730$fit_min, xend = curve_0730$fit_max,
                    y = curve_0730$intercept + curve_0730$slope * curve_0730$fit_min,
                    yend = curve_0730$intercept + curve_0730$slope * curve_0730$fit_max),
               color = "steelblue", linewidth = 0.6) +
  geom_point(data = curve_0730$summary, aes(x = glyc_concentration, y = glyc_mean_luminescence),
             size = 2.4, color = "steelblue") +
  geom_point(data = glycogen_0730, aes(x = well_conc_ug_mL, y = mean_lum, shape = in_std_range),
             size = 2.2, color = "firebrick", alpha = 0.75) +
  scale_shape_manual(values = c(`TRUE` = 16, `FALSE` = 4),
                      labels = c(`TRUE` = "in range", `FALSE` = "out of range"),
                      name = "Sample status") +
  annotate("text", x = -Inf, y = Inf, label = r2_label_0730,
           hjust = -0.15, vjust = 1.8, size = 3.6, fontface = "bold") +
  labs(x = "Glycogen concentration (\u00b5g/mL, well-level)", y = "Luminescence (RLU)",
       title = "2026-07-30: standard curve with samples overlaid",
       subtitle = "Shaded band = quantifiable range of the standard curve") +
  theme_bw(base_size = 12)

print(std_samples_0730)
ggsave(file.path(output_dir, "standard_curve_with_samples_20260730.png"), std_samples_0730,
       width = 8, height = 5.5, dpi = 300)
```


## Plot 20260803 Standard Curves with Samples
```{r plot-standards-with-samples-0803, fig.width=8, fig.height=5.5}
r2_label_0803 <- paste0("R\u00b2 = ", sprintf("%.4f", curve_0803$r2))

std_samples_0803 <- ggplot() +
  geom_ribbon(data = data.frame(x = c(curve_0803$conc_min, curve_0803$conc_max)),
              aes(x = x, ymin = -Inf, ymax = Inf), fill = "grey85", alpha = 0.4,
              inherit.aes = FALSE) +
  geom_segment(aes(x = curve_0803$fit_min, xend = curve_0803$fit_max,
                    y = curve_0803$intercept + curve_0803$slope * curve_0803$fit_min,
                    yend = curve_0803$intercept + curve_0803$slope * curve_0803$fit_max),
               color = "steelblue", linewidth = 0.6) +
  geom_point(data = curve_0803$summary, aes(x = glyc_concentration, y = glyc_mean_luminescence),
             size = 2.4, color = "steelblue") +
  geom_point(data = glycogen_0803, aes(x = well_conc_ug_mL, y = mean_lum, shape = in_std_range),
             size = 2.2, color = "firebrick", alpha = 0.75) +
  scale_shape_manual(values = c(`TRUE` = 16, `FALSE` = 4),
                      labels = c(`TRUE` = "in range", `FALSE` = "out of range"),
                      name = "Sample status") +
  annotate("text", x = -Inf, y = Inf, label = r2_label_0803,
           hjust = -0.15, vjust = 1.8, size = 3.6, fontface = "bold") +
  labs(x = "Glycogen concentration (\u00b5g/mL, well-level)", y = "Luminescence (RLU)",
       title = "2026-08-03: standard curve with re-assayed samples overlaid",
       subtitle = "Shaded band = quantifiable range of the standard curve") +
  theme_bw(base_size = 12)

print(std_samples_0803)
ggsave(file.path(output_dir, "standard_curve_with_samples_20260803.png"), std_samples_0803,
       width = 8, height = 5.5, dpi = 300)
```

# MERGING RUNS INTO A SINGLE FINAL DATASET

For the 5 samples with a measurement from both dates, the value used going
forward is the one that falls within its own run's standard-curve range
(interpolated), with the more recent run used as a tiebreaker if both
qualify. All other 27 samples have only a single (2026-07-30) measurement and
pass through unchanged. This yields the reconciliation described in
Background: the 4 originally-extrapolated samples take their 2026-08-03
(1:200) value, while `1_04_36C` takes its 2026-07-30 (1:25) value, since its
2026-08-03 re-assay fell below the standard curve's floor (negative
back-calculated concentration).

```{r merge-runs}
combined_measurements <- bind_rows(glycogen_0730, glycogen_0803) %>%
  mutate(high_cv = cv_lum > 15) %>%
  arrange(sample_id, run)

cat("--- combined_measurements: every available measurement, both runs, before reconciliation ---\n")
str(combined_measurements)

final_glycogen <- combined_measurements %>%
  group_by(sample_id) %>%
  arrange(desc(in_std_range), desc(run), .by_group = TRUE) %>%
  mutate(n_runs_available = n()) %>%
  slice(1) %>%
  ungroup() %>%
  arrange(family, temperature, individual)

cat("\n--- final_glycogen: one reconciled measurement per sample ---\n")
str(final_glycogen)

cat("\nSamples for which the 2026-08-03 re-assay superseded the 2026-07-30 value:\n")
print(as.data.frame(final_glycogen %>%
  filter(n_runs_available > 1, run == "2026-08-03") %>%
  select(sample_id, run, in_std_range, cv_lum, norm_glycogen)))

cat("\nSamples for which the 2026-07-30 value was retained despite a 2026-08-03 re-assay existing:\n")
print(as.data.frame(final_glycogen %>%
  filter(n_runs_available > 1, run == "2026-07-30") %>%
  select(sample_id, run, in_std_range, cv_lum, norm_glycogen)))

cat("\nTotal samples in final dataset:", nrow(final_glycogen), "\n")
cat("Samples still out-of-range after reconciliation:",
    sum(!final_glycogen$in_std_range), "\n")
```

## Excluding out-of-range samples from statistical comparisons

Per the analysis requirements, any sample still falling outside its
standard curve's quantifiable range after reconciliation is **excluded from
the statistical comparisons** below (family/temperature ANOVA and pairwise
tests), though it remains visible in the full results table and raw
distribution plots.

```{r exclude-out-of-range}
stats_data <- final_glycogen %>% filter(in_std_range)

cat("Samples excluded from statistical comparisons (out of range after reconciliation):\n")
print(as.data.frame(final_glycogen %>% filter(!in_std_range) %>%
  select(sample_id, run, in_std_range, norm_glycogen)))

cat("\nn samples entering statistical comparisons:", nrow(stats_data), "of", nrow(final_glycogen), "\n")
str(stats_data)
```

# RESULTS TABLE

```{r results-table}
results_table <- final_glycogen %>%
  transmute(
    Sample = sample_id,
    Family = family,
    Temperature = temperature,
    Individual = individual,
    `Tissue (mg)` = weight_mg,
    Dilution = dilution,
    `Run used` = run,
    `Mean lum.` = round(mean_lum, 1),
    `CV (%)` = round(cv_lum, 1),
    `High CV (>15%)` = ifelse(high_cv, "yes", "no"),
    `Well glycogen (ug/mL)` = round(well_conc_ug_mL, 3),
    `Normalized glycogen (ug/mL/mg)` = round(norm_glycogen, 2),
    `In std range` = ifelse(in_std_range, "yes", "NO - excluded from stats")
  )

kable(results_table)

write.csv(results_table, file.path(output_dir, "sample_glycogen_results_combined.csv"), row.names = FALSE)
```

# GROUP SUMMARY STATISTICS

```{r group-summary}
group_summary_stats <- stats_data %>%
  group_by(Family = family, Temperature = temperature) %>%
  summarise(
    n      = n(),
    mean   = mean(norm_glycogen),
    sd     = sd(norm_glycogen),
    se     = sd(norm_glycogen) / sqrt(n()),
    median = median(norm_glycogen),
    .groups = "drop"
  )

cat("--- group_summary_stats ---\n")
str(group_summary_stats)
kable(group_summary_stats, digits = 2)

write.csv(group_summary_stats, file.path(output_dir, "group_summary_stats_combined.csv"), row.names = FALSE)
```

# STATISTICAL ANALYSIS

Statistical analysis is performed **before** the box plot and interaction
plot below, so that the plots can be annotated with the significance calls
made here. Normalized glycogen values span roughly an order of magnitude and
are right-skewed, so the primary test uses a **log10 transform** to better
meet ANOVA's normality/homoscedasticity assumptions; the raw-scale ANOVA and
a non-parametric (Wilcoxon rank-sum) alternative are reported alongside it as
robustness checks. Out-of-range samples (if any remain after reconciliation)
are excluded from all tests below (`stats_data`, `r nrow(stats_data)`
samples).

## Two-way ANOVA (family x temperature)

```{r anova-log}
aov_log <- aov(log10(norm_glycogen) ~ family * temperature, data = stats_data)
cat("--- Two-way ANOVA, log10(normalized glycogen) ---\n")
print(summary(aov_log))
```

```{r anova-raw}
aov_raw <- aov(norm_glycogen ~ family * temperature, data = stats_data)
cat("--- Two-way ANOVA, raw-scale normalized glycogen (robustness check) ---\n")
print(summary(aov_raw))
```

## Overall family and temperature comparisons (Wilcoxon rank-sum)

```{r wilcoxon-overall}
wilcox_family <- wilcox.test(norm_glycogen ~ family, data = stats_data)
wilcox_temp   <- wilcox.test(norm_glycogen ~ temperature, data = stats_data)

cat("Family 1 vs. Family 9 (pooled across temperature):\n")
print(wilcox_family)
cat("\nAmbient vs. 36C (pooled across family):\n")
print(wilcox_temp)
```

## Reusable significance-annotation helpers

These three helpers are plot-agnostic: they take a `family`/`temperature`-style
pairwise-comparison table (built in the next section) and a data frame, and
compute/draw asterisk-annotated brackets. They generalize directly to other
analyses (different family counts, additional temperature levels, or
different grouping factors) as long as the same long-format comparison table
shape (`comparison_type`, `fixed_level`, `group1`, `group2`, `p_value`) is
provided. Defined here, ahead of the comparisons themselves, since the
comparison-building chunk below calls `sig_stars()` directly.

```{r sig-annotation-helpers}
# sig_stars(): map p-values to conventional significance stars.
# Returns NA (i.e. "no annotation") for non-significant results, so callers
# can filter on !is.na(sig) to find only what needs to be drawn.
sig_stars <- function(p, thresholds = c(0.001, 0.01, 0.05), symbols = c("***", "**", "*")) {
  ord <- order(thresholds)
  thresholds <- thresholds[ord]; symbols <- symbols[ord]
  out <- rep(NA_character_, length(p))
  for (i in seq_along(thresholds)) out[is.na(out) & p < thresholds[i]] <- symbols[i]
  out
}

# dodge_offset(): x-axis offset of the k-th level of an n-level dodged group,
# matching ggplot2's position_dodge(width)/position_jitterdodge(dodge.width = width).
dodge_offset <- function(group_index, n_groups, width = 0.75) {
  -width / 2 + width / n_groups * (group_index - 0.5)
}

# build_sig_brackets(): compute bracket endpoints + stacked heights for every
# significant (non-NA sig) row of `comparisons`, for a two-factor plot with
# `x_var` on the x-axis and `dodge_var` as the fill/color/dodge factor.
#   - comparison_type == "across_x"     : fixed dodge_var level, compare two
#                                         x_var levels (horizontal bracket
#                                         spanning x positions).
#   - comparison_type == "across_dodge" : fixed x_var level, compare two
#                                         dodge_var levels (horizontal bracket
#                                         spanning dodge offsets within one
#                                         x position; dodge_width = 0 collapses
#                                         this to a zero-width case, so only
#                                         use "across_dodge" when dodge_width > 0).
# Returns list(brackets, tip) where brackets is empty (0 rows) when nothing
# is significant -- callers should add zero layers in that case, which
# geom_sig_brackets() does automatically.
build_sig_brackets <- function(data, y_var, x_var, dodge_var, comparisons,
                                dodge_width = 0.75, step = NULL, tip = NULL) {
  y_vals  <- data[[y_var]]
  y_range <- diff(range(y_vals, na.rm = TRUE))
  if (is.null(step)) step <- 0.09 * y_range
  if (is.null(tip))  tip  <- 0.02 * y_range

  x_levels     <- levels(data[[x_var]])
  dodge_levels <- levels(data[[dodge_var]])
  n_dodge      <- length(dodge_levels)

  offset_for <- function(lvl) dodge_offset(match(lvl, dodge_levels), n_dodge, dodge_width)
  xnum_for   <- function(lvl) match(lvl, x_levels)

  sig_rows <- comparisons %>% filter(!is.na(sig))
  if (nrow(sig_rows) == 0) return(list(brackets = sig_rows, tip = tip))

  base_y <- max(y_vals, na.rm = TRUE) + step

  brackets <- sig_rows %>%
    rowwise() %>%
    mutate(
      x1 = if (comparison_type == "across_x") xnum_for(group1) + offset_for(fixed_level)
           else                               xnum_for(fixed_level) + offset_for(group1),
      x2 = if (comparison_type == "across_x") xnum_for(group2) + offset_for(fixed_level)
           else                               xnum_for(fixed_level) + offset_for(group2)
    ) %>%
    ungroup() %>%
    filter(x1 != x2) %>%   # degenerate (zero-width) brackets are not drawable
    arrange(pmin(x1, x2), abs(x2 - x1)) %>%
    mutate(y = base_y + (row_number() - 1) * step, x_mid = (x1 + x2) / 2)

  list(brackets = brackets, tip = tip)
}

# geom_sig_brackets(): ggplot layer list drawing a horizontal bracket + star
# for every row of a build_sig_brackets() result. Adds nothing when there are
# no significant comparisons, so it is always safe to add to a plot.
geom_sig_brackets <- function(bracket_result) {
  b <- bracket_result$brackets
  if (nrow(b) == 0) return(list())
  tip <- bracket_result$tip
  list(
    geom_segment(data = b, aes(x = x1, xend = x2, y = y, yend = y),
                 inherit.aes = FALSE, linewidth = 0.5),
    geom_segment(data = b, aes(x = x1, xend = x1, y = y, yend = y - tip),
                 inherit.aes = FALSE, linewidth = 0.5),
    geom_segment(data = b, aes(x = x2, xend = x2, y = y, yend = y - tip),
                 inherit.aes = FALSE, linewidth = 0.5),
    geom_text(data = b, aes(x = x_mid, y = y, label = sig),
              inherit.aes = FALSE, vjust = -0.3, size = 4.2, fontface = "bold")
  )
}

cat("Defined: sig_stars(), dodge_offset(), build_sig_brackets(), geom_sig_brackets()\n")
```

## Within-family temperature comparisons, and within-temperature family comparisons

Four stratified pairwise comparisons, each via Wilcoxon rank-sum, with Holm
correction applied across the family of four tests. This is written
generically over the levels of `family` and `temperature` present in
`stats_data`, so it extends unchanged to analyses with additional families
or temperature levels: every same-family pair across temperature levels, and
every same-temperature pair across families, is tested and Holm-corrected as
one family of comparisons.

```{r stratified-comparisons}
family_levels <- levels(stats_data$family)
temp_levels   <- levels(stats_data$temperature)

# Every pairwise comparison within a fixed family, across temperature levels.
# comparison_type = "across_x" -- generic tag consumed by build_sig_brackets()
# below: it means "x_var (temperature) varies, dodge_var (family) is fixed".
across_temp_tests <- bind_rows(lapply(family_levels, function(fam) {
  d <- filter(stats_data, family == fam)
  combn(temp_levels, 2, simplify = FALSE) %>%
    lapply(function(pair) {
      p <- wilcox.test(norm_glycogen ~ temperature,
                        data = filter(d, temperature %in% pair))$p.value
      data.frame(comparison_type = "across_x", fixed_level = fam,
                 group1 = pair[1], group2 = pair[2], p_value = p,
                 label_prefix = "within-family", stringsAsFactors = FALSE)
    }) %>% bind_rows()
}))

# Every pairwise comparison within a fixed temperature, across family levels.
# comparison_type = "across_dodge" -- "dodge_var (family) varies, x_var
# (temperature) is fixed".
across_family_tests <- bind_rows(lapply(temp_levels, function(tp) {
  d <- filter(stats_data, temperature == tp)
  combn(family_levels, 2, simplify = FALSE) %>%
    lapply(function(pair) {
      p <- wilcox.test(norm_glycogen ~ family,
                        data = filter(d, family %in% pair))$p.value
      data.frame(comparison_type = "across_dodge", fixed_level = tp,
                 group1 = pair[1], group2 = pair[2], p_value = p,
                 label_prefix = "within-temperature", stringsAsFactors = FALSE)
    }) %>% bind_rows()
}))

stratified_tests <- bind_rows(across_temp_tests, across_family_tests) %>%
  mutate(
    p_holm = p.adjust(p_value, method = "holm"),
    sig    = sig_stars(p_holm),
    label  = paste0(label_prefix, " (", fixed_level, "): ", group1, " vs. ", group2)
  )

cat("--- stratified_tests ---\n")
str(stratified_tests)
kable(stratified_tests %>% select(label, comparison_type, p_value, p_holm, sig), digits = 4)
cat("\nSignificant (Holm-corrected) stratified comparisons in this dataset:",
    sum(!is.na(stratified_tests$sig)), "of", nrow(stratified_tests), "\n")
```

# PLOTTING

The box plot is annotated with brackets for every statistically significant
(Holm-corrected p < 0.05) pairwise comparison from `stratified_tests` above,
via the reusable `build_sig_brackets()`/`geom_sig_brackets()` helpers. With
the current dataset none of the four stratified comparisons are significant,
so no brackets are drawn -- the annotation layer is present but empty. If a
future re-analysis (e.g. additional families, or a stronger temperature
effect) produces a significant pairwise result, a bracket and asterisk(s)
will appear automatically without any code changes.


## Box plots
```{r plot-family-temp, fig.width=8, fig.height=5.5}
box_brackets <- build_sig_brackets(stats_data, y_var = "norm_glycogen",
                                    x_var = "temperature", dodge_var = "family",
                                    comparisons = stratified_tests, dodge_width = 0.75)

cat("--- box_brackets$brackets (significant comparisons drawn on the box plot) ---\n")
str(box_brackets$brackets)

group_plot <- ggplot(stats_data, aes(x = temperature, y = norm_glycogen, fill = family)) +
  geom_boxplot(position = position_dodge(0.75), alpha = 0.35, outlier.shape = NA) +
  geom_point(aes(shape = family), position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.75),
             size = 2.8, alpha = 0.8) +
  geom_sig_brackets(box_brackets) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.12))) +
  labs(x = "Temperature exposure", y = "Normalized glycogen (\u00b5g/mL/mg tissue)",
       title = "Glycogen content by family and temperature",
       subtitle = "Combined dataset (2026-07-30 + 2026-08-03 reconciled); boxplot + individual points",
       fill = "Family", shape = "Family") +
  scale_fill_manual(values = c("Family 1" = "#1b9e77", "Family 9" = "#d95f02")) +
  theme_bw(base_size = 13) +
  theme(axis.title = element_text(size = 11, face = "bold"),
        legend.position = "bottom")

print(group_plot)
ggsave(file.path(output_dir, "glycogen_by_family_temperature_combined.png"), group_plot,
       width = 8, height = 5.5, dpi = 300)
```

## Family x temperature interaction

The interaction plot is annotated using the same `build_sig_brackets()`
helper, restricted to the `"across_temperature"` (within-family) comparisons
and drawn with `dodge_width = 0` since family is encoded by line color here
rather than by horizontal dodging. Within-temperature family-vs-family
comparisons are already shown as brackets on the box plot above rather than
duplicated here, since family is not spatially separated at a given
temperature in this plot.

```{r plot-interaction, fig.width=7, fig.height=5.5}
interaction_data <- stats_data %>%
  group_by(family, temperature) %>%
  summarise(mean_glycogen = mean(norm_glycogen),
            se_glycogen   = sd(norm_glycogen) / sqrt(n()),
            .groups = "drop")

cat("--- interaction_data ---\n")
str(interaction_data)

interaction_brackets <- build_sig_brackets(
  interaction_data %>% mutate(norm_glycogen = mean_glycogen + se_glycogen),
  y_var = "norm_glycogen", x_var = "temperature", dodge_var = "family",
  comparisons = filter(stratified_tests, comparison_type == "across_x"),
  dodge_width = 0
)

cat("--- interaction_brackets$brackets (significant within-family comparisons drawn) ---\n")
str(interaction_brackets$brackets)

interaction_plot <- ggplot(interaction_data, aes(x = temperature, y = mean_glycogen,
                                                  color = family, group = family)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_glycogen - se_glycogen, ymax = mean_glycogen + se_glycogen),
                width = 0.1) +
  geom_sig_brackets(interaction_brackets) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.12))) +
  scale_color_manual(values = c("Family 1" = "#1b9e77", "Family 9" = "#d95f02")) +
  labs(x = "Temperature exposure", y = "Mean normalized glycogen (\u00b5g/mL/mg, \u00b1 SE)",
       title = "Family x temperature interaction",
       subtitle = "Combined dataset (2026-07-30 + 2026-08-03 reconciled); group means \u00b1 SE",
       color = "Family") +
  theme_bw(base_size = 13) +
  theme(axis.title = element_text(size = 11, face = "bold"),
        legend.position = "bottom")

print(interaction_plot)
ggsave(file.path(output_dir, "glycogen_family_temperature_interaction_combined.png"), interaction_plot,
       width = 7, height = 5.5, dpi = 300)
```

# SUMMARY

- **Combined dataset:** `r nrow(final_glycogen)` samples (families 1 and 9,
  ambient and 36&deg;C), reconciling 2026-07-30 and 2026-08-03 measurements
  per the rule in Background. `r sum(final_glycogen$n_runs_available > 1)`
  samples had a measurement from both dates; for 4 of these
  (`1_08_ambient`, `1_06_ambient`, `1_05_36C`, `9_05_ambient`) the
  2026-08-03 (1:200) in-range value superseded the original extrapolated
  value, while `1_04_36C` retained its 2026-07-30 (1:25) value since its
  2026-08-03 re-assay fell below the standard curve's floor.

- **Out-of-range samples:** `r sum(!final_glycogen$in_std_range)` sample(s)
  remain outside their standard curve's quantifiable range after
  reconciliation and were excluded from all statistical comparisons above
  (see "Excluding out-of-range samples" for detail).

- **High-CV samples:** samples exceeding 15% technical-replicate CV were
  flagged (see "Coefficient of Variation (CV) Check") but retained in all
  analyses, per the analysis requirements.

- **Family effect:** no statistically significant difference between
  Family 1 and Family 9 in normalized glycogen content, either pooled
  across temperature (two-way ANOVA `family` term, and the overall
  Wilcoxon test) or within either temperature stratum (Holm-corrected
  pairwise comparisons).

- **Temperature effect:** no statistically significant difference between
  ambient and 36&deg;C exposure, either pooled across family or within
  either family stratum.

- **Family x temperature interaction:** no statistically significant
  interaction term in the two-way ANOVA. The interaction plot shows Family
  1's mean glycogen trending down from ambient to 36&deg;C while Family 9's
  trends up, but neither the interaction term nor the stratified
  within-family comparisons reach significance at this sample size --
  this pattern should be interpreted as a visual trend, not a confirmed
  effect.


