---
title: "01-multi-assay-family-survival-prediction"
author: "Steven Roberts"
date: "2026-10-02"
output:
  github_document:
    toc: true
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
set.seed(20261002)

out_dir <- "../outputs/01-multi-assay-family-survival-prediction"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
```

# Goal

Integrate family-level resazurin, Na<sup>+</sup>/K<sup>+</sup>-ATPase and citrate
synthase data to find the combination of measures that best predicts family
survival.

**Survival target:** `composite_score` from the cross-experiment survival
summary (mean within-experiment survival percentile, 0-100, higher = hardier).

**The main constraint is sample size.** Survival is a family-level phenotype and
there are at most 9 families, against ~180 candidate predictors. Picking the
best combination from that many candidates will always find something that fits
9 points well, so this notebook separates two things:

1. **Apparent fit**: the best combinations found by exhaustive search, scored by
   leave-one-family-out (LOFO) cross-validation. This is optimistic because the
   same data chose the features.
2. **Honest performance**: *nested* LOFO cross-validation, where feature
   selection is repeated inside each training fold, plus a permutation test that
   runs the whole selection procedure on shuffled survival scores.

# Data

## Survival

```{r survival}
survival <- read_csv(
  "../../heat-survivorship/outputs/09-mgig-survivorship-cross-experiment-summary/family_ranking.csv",
  col_types = cols(family = col_character())
) %>%
  select(family, composite_score, mean_surv_prop, n_experiments)

knitr::kable(survival)
```

## Resazurin

Family-mean curve features from
`Resazurin/code/04-resazurin-family-phenotype-prediction.Rmd`, for each stress
context (`all`, `sw_heat`, `fw_heat`, `freshwater_rt`) and value metric
(`corrected_fc`, `metabolism_per_area_mm2_measurement`). Bookkeeping columns
(`cup_id_group`, `round_group`) are dropped.

```{r resazurin}
resazurin_long <- read_csv(
  "../../Resazurin/outputs/04-resazurin-family-phenotype-prediction/family_feature_matrix.csv",
  col_types = cols(family = col_character())
) %>%
  filter(!feature %in% c("cup_id_group", "round_group")) %>%
  mutate(metric = recode(value_metric,
                         corrected_fc = "fc",
                         metabolism_per_area_mm2_measurement = "area"),
         predictor = paste("rz", context, metric, feature, sep = "."))

resazurin <- resazurin_long %>%
  select(family, predictor, family_mean) %>%
  pivot_wider(names_from = predictor, values_from = family_mean)

resazurin_long %>%
  group_by(context) %>%
  summarise(families = paste(sort(as.integer(unique(family))), collapse = ","),
            n_predictors = n_distinct(predictor), .groups = "drop") %>%
  knitr::kable()
```

The freshwater contexts have no family 2.

## Na/K-ATPase

Family medians of gill ATPase activity from the June 2026 sampling (see
`NaK-ATPase/code/02-NaK-ATPase-vs-family-survival.Rmd`).

```{r atpase}
atpase_raw <- read_tsv("../../NaK-ATPase/data/202606-sampling.tsv") %>%
  rename(ATPase = `ATPase (umol ADP/mg protein/hr)`) %>%
  left_join(read_csv("../../sampling-event-metadata/june-2026/sampling-log.csv") %>%
              select(`Tube ID`, Condition, `Oyster Family ID`),
            by = "Tube ID") %>%
  mutate(family = str_remove(`Oyster Family ID`, "Family "))

stopifnot(!any(is.na(atpase_raw$Condition)))

atpase <- atpase_raw %>%
  group_by(family) %>%
  summarise(
    atp.ambient   = median(ATPase[Condition == "Ambient"]),
    atp.heat      = median(ATPase[Condition == "36C"]),
    atp.log2ratio = log2(atp.heat / atp.ambient),
    .groups = "drop"
  )

knitr::kable(atpase, digits = 3)
```

## Citrate synthase

Citrate synthase has only been assayed in families 5 and 7, so it cannot enter
a family-level model. It is summarized descriptively at the end.

```{r cs}
cs <- read_csv(
  "../../Citrate_synthase/outputs/Gen5-20260825-mgig-sormi-citrate_synthase-F05-F07-temperature-comparison/cs_activity_all_families_no_background.csv"
) %>%
  filter(QC == "usable") %>%
  mutate(family = str_remove(Family, "^F0?"))

count(cs, family, Temperature)
```

## Combined family × predictor matrix

```{r combine}
fam_data <- survival %>%
  inner_join(resazurin, by = "family") %>%
  inner_join(atpase, by = "family")

predictor_cols <- setdiff(names(fam_data), names(survival))

# drop predictors that are constant, and exact duplicates (e.g. timing features
# are identical for the fc and per-area metrics)
predictor_cols <- predictor_cols[map_lgl(predictor_cols, function(p) {
  x <- fam_data[[p]]
  sum(!is.na(x)) >= 8 && sd(x, na.rm = TRUE) > 0
})]
dup <- duplicated(as.list(fam_data[predictor_cols]))
predictor_cols <- predictor_cols[!dup]

tibble(source = if_else(str_starts(predictor_cols, "atp"), "ATPase", "resazurin")) %>%
  count(source) %>%
  knitr::kable(caption = "Candidate predictors after removing constant and duplicate columns")
```

Two family sets are analysed:

- **Primary (8 families):** families 1, 3, 5, 6, 7, 8, 9, 10, which have every
  predictor, including the freshwater contexts.
- **Sensitivity (9 families):** adds family 2, restricted to predictors measured
  in all 9 families (resazurin `all` and `sw_heat` contexts, plus ATPase).

```{r family-sets}
complete_cols <- predictor_cols[map_lgl(predictor_cols, ~ !anyNA(fam_data[[.x]]))]

primary <- fam_data %>%
  filter(if_all(all_of(predictor_cols), ~ !is.na(.x)))
sensitivity <- fam_data

c(primary_families = nrow(primary), primary_predictors = length(predictor_cols),
  sensitivity_families = nrow(sensitivity), sensitivity_predictors = length(complete_cols))
```

# Methods

All models are ordinary linear regressions of `composite_score` on 1-3
predictors. Leave-one-out prediction errors are computed exactly from the hat
matrix, so every candidate model can be scored without refitting.

```{r functions}
# Leave-one-out residuals for lm(y ~ X) via the hat matrix
loo_resid <- function(X, y) {
  X <- cbind(1, X)
  q <- qr(X)
  if (q$rank < ncol(X)) return(rep(NA_real_, length(y)))
  e <- qr.resid(q, y)
  h <- rowSums(qr.Q(q)^2)
  if (any(h > 1 - 1e-8)) return(rep(NA_real_, length(y)))
  e / (1 - h)
}

loo_rmse <- function(X, y) sqrt(mean(loo_resid(X, y)^2))

# Forward selection of up to k_max predictors, each step adding the predictor
# that minimises LOO RMSE; final size chosen by LOO RMSE.
select_model <- function(Xall, y, k_max = 3) {
  chosen <- character(0)
  path <- list()
  for (k in seq_len(k_max)) {
    cand <- setdiff(colnames(Xall), chosen)
    scores <- vapply(cand, function(p) loo_rmse(Xall[, c(chosen, p), drop = FALSE], y), numeric(1))
    if (all(is.na(scores))) break
    chosen <- c(chosen, cand[which.min(scores)])
    path[[k]] <- list(vars = chosen, rmse = min(scores, na.rm = TRUE))
  }
  path[[which.min(map_dbl(path, "rmse"))]]$vars
}

# Nested LOFO CV: selection is redone within each training fold
nested_cv <- function(Xall, y, k_max = 3) {
  n <- length(y)
  sel <- vector("list", n)
  pred <- numeric(n)
  for (i in seq_len(n)) {
    vars <- select_model(Xall[-i, , drop = FALSE], y[-i], k_max)
    fit <- lm.fit(cbind(1, Xall[-i, vars, drop = FALSE]), y[-i])
    pred[i] <- sum(c(1, Xall[i, vars]) * fit$coefficients)
    sel[[i]] <- vars
  }
  list(pred = pred,
       rho = cor(pred, y, method = "spearman"),
       rmse = sqrt(mean((pred - y)^2)),
       selected = sel)
}

# Permutation test of the whole nested procedure
perm_test <- function(Xall, y, observed_rho, n_perm = 200, k_max = 3) {
  null <- replicate(n_perm, nested_cv(Xall, sample(y), k_max)$rho)
  list(null = null, p = (sum(null >= observed_rho) + 1) / (n_perm + 1))
}

# Exhaustive search over all combinations of size k
exhaustive <- function(Xall, y, k) {
  combos <- combn(ncol(Xall), k)
  scores <- apply(combos, 2, function(v) {
    r <- loo_resid(Xall[, v, drop = FALSE], y)
    if (anyNA(r)) return(c(NA_real_, NA_real_))
    c(sqrt(mean(r^2)), cor(y - r, y, method = "spearman"))
  })
  tibble(k = k,
         predictors = apply(combos, 2, function(v) paste(colnames(Xall)[v], collapse = " + ")),
         loo_rmse = scores[1, ],
         loo_rho = scores[2, ]) %>%
    filter(!is.na(loo_rmse)) %>%
    arrange(loo_rmse)
}

as_matrix <- function(df, cols) {
  m <- scale(as.matrix(df[cols]))
  rownames(m) <- df$family
  m
}
```

# Apparent fit: best combinations by exhaustive search

Primary 8-family set. Predictors are z-scored, so the null model (predicting
the training mean) has LOO RMSE of about
`r round(sqrt(mean(((primary$composite_score - mean(primary$composite_score)) * nrow(primary) / (nrow(primary) - 1))^2)), 1)`.

```{r exhaustive}
X_primary <- as_matrix(primary, predictor_cols)
y_primary <- primary$composite_score

search <- bind_rows(
  exhaustive(X_primary, y_primary, 1),
  exhaustive(X_primary, y_primary, 2),
  exhaustive(X_primary, y_primary, 3)
) %>%
  mutate(includes_atpase = str_detect(predictors, "atp\\."),
         includes_resazurin = str_detect(predictors, "rz\\."))

write_csv(search %>% group_by(k) %>% slice_min(loo_rmse, n = 100),
          file.path(out_dir, "exhaustive_search_top100_per_k.csv"))

search %>%
  group_by(k) %>%
  slice_min(loo_rmse, n = 5) %>%
  ungroup() %>%
  select(k, predictors, loo_rmse, loo_rho) %>%
  knitr::kable(digits = 3, caption = "Top 5 combinations of each size (LOFO CV, optimistic)")
```

How often each predictor appears among the 50 best 2- and 3-predictor
combinations shows which measures recur, rather than relying on a single winner.

```{r recurring}
recurring <- search %>%
  filter(k > 1) %>%
  group_by(k) %>%
  slice_min(loo_rmse, n = 50) %>%
  ungroup() %>%
  separate_rows(predictors, sep = " \\+ ") %>%
  count(predictors, name = "n_top_models", sort = TRUE)

write_csv(recurring, file.path(out_dir, "recurring_predictors.csv"))
knitr::kable(head(recurring, 15))
```

```{r best-by-source}
search %>%
  mutate(source = case_when(
    includes_atpase & includes_resazurin ~ "resazurin + ATPase",
    includes_atpase ~ "ATPase only",
    TRUE ~ "resazurin only")) %>%
  group_by(source, k) %>%
  slice_min(loo_rmse, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(source, k, predictors, loo_rmse, loo_rho) %>%
  arrange(k, loo_rmse) %>%
  knitr::kable(digits = 3, caption = "Best combination by data source and size")
```

# Honest performance: nested cross-validation

For each predictor pool, forward selection (up to 3 predictors) is rerun inside
every training fold, the held-out family is predicted, and the procedure is
repeated on 200 permutations of the survival scores.

```{r nested}
pools <- list(
  "ATPase only"        = str_subset(predictor_cols, "^atp\\."),
  "resazurin only"     = str_subset(predictor_cols, "^rz\\."),
  "resazurin + ATPase" = predictor_cols
)

nested <- imap_dfr(pools, function(cols, pool) {
  X <- as_matrix(primary, cols)
  cv <- nested_cv(X, y_primary)
  pt <- perm_test(X, y_primary, cv$rho)
  sel <- table(unlist(cv$selected))
  tibble(pool = pool, n_predictors = length(cols),
         nested_rho = cv$rho, nested_rmse = cv$rmse, perm_p = pt$p,
         most_selected = paste(names(sort(sel, decreasing = TRUE))[1:min(3, length(sel))],
                               collapse = "; "),
         pred = list(tibble(family = primary$family, observed = y_primary, predicted = cv$pred)),
         null = list(pt$null))
})

write_csv(nested %>% select(-pred, -null), file.path(out_dir, "nested_cv_performance.csv"))

nested %>%
  select(pool, n_predictors, nested_rho, nested_rmse, perm_p, most_selected) %>%
  knitr::kable(digits = 3, caption = "Nested LOFO CV, primary 8-family set")
```

```{r nested-plot, fig.width=10, fig.height=4}
nested %>%
  select(pool, pred) %>%
  unnest(pred) %>%
  ggplot(aes(x = predicted, y = observed)) +
  geom_abline(linetype = "dashed", colour = "grey60") +
  geom_point(size = 2.5) +
  geom_text(aes(label = family), nudge_y = 2, size = 3) +
  facet_wrap(~ pool) +
  labs(x = "Predicted composite score (nested LOFO)",
       y = "Observed composite score",
       title = "Held-out predictions after in-fold feature selection") +
  theme_bw()
```

```{r null-plot, fig.width=10, fig.height=3.5}
nested %>%
  select(pool, nested_rho, null) %>%
  unnest(null) %>%
  ggplot(aes(x = null)) +
  geom_histogram(bins = 25, fill = "grey75") +
  geom_vline(aes(xintercept = nested_rho), colour = "#E45756", linewidth = 1) +
  facet_wrap(~ pool) +
  labs(x = "Nested CV Spearman rho on permuted survival scores",
       y = "Permutations",
       title = "Observed nested CV rho (red) vs permutation null") +
  theme_bw()
```

# Sensitivity: 9 families

Adds family 2 and restricts to predictors measured in all 9 families.

```{r sensitivity}
pools9 <- map(pools, ~ intersect(.x, complete_cols))
y9 <- sensitivity$composite_score

nested9 <- imap_dfr(pools9, function(cols, pool) {
  X <- as_matrix(sensitivity, cols)
  cv <- nested_cv(X, y9)
  pt <- perm_test(X, y9, cv$rho)
  sel <- table(unlist(cv$selected))
  tibble(pool = pool, n_predictors = length(cols),
         nested_rho = cv$rho, nested_rmse = cv$rmse, perm_p = pt$p,
         most_selected = paste(names(sort(sel, decreasing = TRUE))[1:min(3, length(sel))],
                               collapse = "; "))
})

write_csv(nested9, file.path(out_dir, "nested_cv_performance_9_families.csv"))
knitr::kable(nested9, digits = 3, caption = "Nested LOFO CV, 9-family set")
```

# Citrate synthase (families 5 and 7 only)

```{r cs-summary}
cs_summary <- cs %>%
  group_by(family, Temperature) %>%
  summarise(median_activity = median(`Activity, no BG (mU/mg)`), n = n(), .groups = "drop") %>%
  pivot_wider(names_from = Temperature, values_from = c(median_activity, n)) %>%
  mutate(log2_36C_vs_ambient = log2(median_activity_36C / median_activity_ambient)) %>%
  left_join(survival %>% select(family, composite_score), by = "family")

knitr::kable(cs_summary, digits = 2)
```

```{r cs-plot, fig.width=6, fig.height=4}
cs %>%
  left_join(survival, by = "family") %>%
  mutate(family_label = sprintf("Family %s\n(survival score %.0f)", family, composite_score),
         Temperature = factor(Temperature, levels = c("ambient", "36C"))) %>%
  ggplot(aes(x = Temperature, y = `Activity, no BG (mU/mg)`)) +
  geom_boxplot(outlier.shape = NA, fill = "grey90") +
  geom_jitter(width = 0.1, size = 1.5) +
  facet_wrap(~ family_label) +
  labs(y = "Citrate synthase activity (mU / mg protein)", x = NULL) +
  theme_bw()
```

# Conclusions

- **Apparent fits are perfect, and that is the warning sign.** Exhaustive search
  finds 2-predictor resazurin combinations with LOFO Spearman rho = 1.0 and
  3-predictor combinations with LOFO RMSE under 1 point. With 162 resazurin
  predictors and 8 families, that many combinations will always contain
  near-perfect fits by chance.
- **Under nested cross-validation the combinations do not hold up.** When
  selection is redone inside each fold (8 families), held-out predictions are
  essentially uncorrelated with survival (rho about 0.1 for every pool), RMSE
  (17-18) is worse than predicting the mean (about 13.5), and none beats the
  permutation null (p 0.27-0.40). The selected predictors change from fold to
  fold.
- **Adding ATPase does not help.** ATPase alone does not predict survival, and
  resazurin + ATPase performs the same as resazurin alone; ATPase rarely enters
  the selected models.
- **The 9-family sensitivity analysis is more encouraging for resazurin.** With
  family 2 included and only `all`/`sw_heat` resazurin features, nested rho is
  0.68 (permutation p = 0.045), driven mainly by `rz.all.area.trough_value`
  (family-mean minimum per-area resazurin signal), but RMSE (12.4) is only
  slightly better than the mean. This is the most promising lead, not a
  validated predictor.
- **Recurring resazurin timing features** (`fw_heat` `time_to_min_slope`,
  `all` `inflection_time`) appear most often among top combinations and are
  worth targeting in future experiments.
- **Citrate synthase** is the most striking contrast but covers only two
  families: the hardiest family (5) cuts CS activity by about 70% at 36C
  (log2 change -1.67), while the least hardy (7) does not change. Assaying CS in
  the remaining families is the clearest next step for testing whether
  metabolic down-regulation under heat predicts survival.

Bottom line: no combination of the current resazurin and ATPase measures
predicts family survival robustly with 8-9 families. More families, or
measurements and survival on the same individuals, are needed before a
multi-assay predictor can be selected and trusted.
